### 2022-06-28 MCQMC 2020 stuff/MCQMC_QMCSoftwareTut.tex
Update from Office Computer

%Talk given virtually for SIAM UQ 2020 March
\documentclass[11pt,compress,xcolor={usenames,dvipsnames},aspectratio=169]{beamer}
%\documentclass[xcolor={usenames,dvipsnames},aspectratio=169]{beamer} %slides and 
%notes
\usepackage{amsmath,
	amssymb,
	datetime,
	mathtools,
	bbm,
	%mathabx,
	array,
	booktabs,
	xspace,
	multirow,
	calc,
	colortbl,
	siunitx,
 	graphicx}
\usepackage[usenames]{xcolor}
\usepackage[giveninits=false,backend=biber,style=nature, maxcitenames =10, mincitenames=9]{biblatex}
\addbibresource{FJHown23.bib}
\addbibresource{FJH23.bib}
\usepackage{newpxtext}
\usepackage[euler-digits,euler-hat-accent]{eulervm}
\usepackage{media9}
\usepackage[autolinebreaks]{mcode}
\usepackage[tikz]{mdframed}

\usetheme{FJHSlimNoFoot169}
\setlength{\parskip}{2ex}
\setlength{\arraycolsep}{0.5ex}

\DeclareMathOperator{\sol}{SOL}
\DeclareMathOperator{\app}{APP}
\DeclareMathOperator{\alg}{ALG}
\DeclareMathOperator{\ACQ}{ACQ}
\DeclareMathOperator{\ERR}{ERR}
\DeclareMathOperator{\COST}{COST}
\DeclareMathOperator{\COMP}{COMP}
\newcommand{\dataN}{\bigl(\hf(\vk_i)\bigr)_{i=1}^n}
\newcommand{\dataNj}{\bigl(\hf(\vk_i)\bigr)_{i=1}^{n_j}}
\newcommand{\dataNjd}{\bigl(\hf(\vk_i)\bigr)_{i=1}^{n_{j^\dagger}}}
\newcommand{\ERRN}{\ERR\bigl(\dataN,n\bigr)}

\newcommand{\Sapp}{S_{\textup{app}}}
\newcommand{\LambdaStd}{\Lambda^{\textup{std}}}
\newcommand{\LambdaSer}{\Lambda^{\textup{ser}}}
\newcommand{\LambdaAll}{\Lambda^{\textup{all}}}
\newcommand{\oton}{1\!:\!n}
\newcommand{\talert}[1]{\alert{\text{#1}}}
\DeclareMathOperator{\init}{init}
\DeclareMathOperator{\GP}{\cg\cp}
\newcommand{\MLE}{\textup{EB}}
\newcommand{\mCtheta}{{\mathsf{C}_{\vtheta}}}
\newcommand{\mCInv}{\mathsf{C}^{-1}}

%\DeclareMathOperator{\app}{app}

\providecommand{\HickernellFJ}{H.\xspace}

\iffalse

\fi

\renewcommand{\OffTitleLength}{-7ex}
\setlength{\FJHThankYouMessageOffset}{-8ex}
\title{Quasi-Monte Carlo Software}
\author[]{Fred J. Hickernell}
\institute{Department of Applied Mathematics \&
	Center for Interdisciplinary Scientific Computation \\  Illinois Institute of Technology \quad
	\href{mailto:hickernell@iit.edu}{\url{hickernell@iit.edu}} \quad
	\href{http://mypages.iit.edu/~hickernell}{\url{mypages.iit.edu/~hickernell}}}

\thanksnote{with Sou-Cheng Choi, Mike McCourt, Jagadeeswaran Rathinavel, Aleksei Sorokin \\ 
and the rest of the GAIL an QMCPy teams \\
	partially supported by SigOpt and NSF-DMS-1522687 \\[2ex]
	Thanks to the organizers, especially during these unusual times \\
	Slides at  \href{https://speakerdeck.com/fjhickernell/quasi-monte-carlo-software}{\nolinkurl{speakerdeck.com/fjhickernell/quasi-monte-carlo-software}}\\
	Google Colaboratory notebook at \href{https://tinyurl.com/QMCPyTutorial}{\nolinkurl{tinyurl.com/QMCPyTutorial}}\\
	Blog at \href{https://qmcpy.wordpress.com/}{\nolinkurl{qmcpy.wordpress.com/}}\\
	Recording of this talk at \href{https://media.ed.ac.uk/playlist/dedicated/51612401/1_0z0wec2z/1_2k12mwiw}{\nolinkurl{media.ed.ac.uk/playlist/dedicated/51612401/1_0z0wec2z/1_2k12mwiw}}

}
\event{MCQMC 2020}
\date[]{August 10, 2020}

\input FJHDef.tex

\newlength{\figwidth}
\setlength{\figwidth}{0.25\textwidth}

\newlength{\figwidthSmall}
\setlength{\figwidthSmall}{0.2\textwidth}

\newcommand{\financePict}{\href{http://i2.cdn.turner.com/money/dam/assets/130611131918-chicago-board-options-exchange-1024x576.jpg}{\includegraphics[width
		= 3cm]{ProgramsImages/130611131918-chicago-board-options-exchange-1024x576.jpg}}}

	\newcommand{\scoop}[1]{\parbox{#1}{\includegraphics[width=#1]{IceCreamScoop.eps}}\xspace}
	\newcommand{\smallscoop}{\scoop{1cm}}
	\newcommand{\medscoop}{\scoop{1.8cm}}
	\newcommand{\largescoop}{\scoop{3cm}}
	\newcommand{\ICcone}[1]{\parbox{#1}{\includegraphics[width=#1,angle=270]{MediumWaffleCone.eps}}\xspace}
	\newcommand{\medcone}{\ICcone{1.2cm}}
	\newcommand{\largercone}{\parbox{2.2cm}{\vspace*{-0.2cm}\includegraphics[width=1cm,angle=270]{MediumWaffleCone.eps}}\xspace}
	\newcommand{\largecone}{\ICcone{1.8cm}}
	\newcommand{\smallcone}{\parbox{1.1cm}{\includegraphics[width=0.5cm,angle=270]{MediumWaffleCone.eps}}\xspace}

\newcommand{\northeaststuff}[3]{
	\begin{tikzpicture}[remember picture, overlay]
	\node [shift={(-#1 cm,-#2 cm)}]  at (current page.north east){#3};
	\end{tikzpicture}}

\begin{document}
	\tikzstyle{every picture}+=[remember picture]
	\everymath{\displaystyle}

\frame{\titlepage}

\section{Background}

\begin{frame}{You have heard about QMC.  How do you try it?}

	\vspace{-5ex}
\begin{itemize}
\setlength{\itemsep}{0cm}
    \item<1-> \emph{QMC will give you 100 times the accuracy in the same amount of time as simple MC} \\
     \uncover<2->{\alert{Often}}
    \item<1-> \emph{Just replace your IID random points with low discrepancy (LD) points}\\
    \uncover<2->{\alert{Sometimes}}

    \vspace{4ex}

    \item<3-> \emph{Where can I get accurate, efficient, easy to use QMC software to try for my problem?}\\
     \item<3-> \emph{How can I make my great QMC software available for others?}\\
    \uncover<4->{\alert{Let's try to help}}

\end{itemize}
\end{frame}

\begin{frame}{Quasi-Monte Carlo (QMC) uses low discrepancy (LD) sequences}
\vspace{-3ex}
\begin{tabular}{>{\centering}p{0.47\textwidth}@{\quad}>{\centering}p{0.47\textwidth}}
%Independent \& Identically Distributed (IID) &
%Low Discrepancy (LD) \tabularnewline
\includegraphics[height=5cm]{ProgramsImages/IIDPoints.eps} &
\includegraphics[height=5cm]{ProgramsImages/SSobolPoints.eps}
\tabularnewline
$\vT_i$ are random &
$\vX_i$ may be deterministic \alert{or} random 
\tabularnewline
$\vT_1, \vT_2 \cdots \alert{\IIDsim} F$ &
$\vX_1, \vX_2 \cdots \alert{\LDsim} F$ 
\tabularnewline
$\vT_i$ do not know about one another &
$\{\vX_i\}_{i=1}^n$ represent $F$ well
\tabularnewline
$F_{n}(\vt_1, \ldots, \vt_n) = F(\vt_1) \cdots F(\vt_n)$ &
$F_{\{\vX_i\}_{i=1}^n}(\vx) \approx F(\vx)$
\end{tabular}
\end{frame}

\begin{frame}{Where is the software?  (apologies to those I missed)}
			\vspace{-3ex}

	\renewcommand{\arraystretch}{1.15}
	\begin{tabular}{>{\centering}m{0.47\textwidth}@{\qquad}>{\centering}m{0.47\textwidth}}
		\alert{LD Sequence Generators} & \alert{Multi-Level, Stopping Criteria, Applications}
		\tabularnewline \toprule
		\uncover<1>{\href{https://www.mathworks.com}{\alert{MATLAB Statistics Toolbox}}---\newline Sobol' and Halton} &
		\href{https://people.maths.ox.ac.uk/gilesm/mlmc/}{\alert{Mike Giles}}---Multi-Level (Quasi-)Monte Carlo  \uncover<1>{in C++, MATLAB, Python, and R}
		\tabularnewline
		\href{https://cran.r-project.org/web/packages/qrng/qrng.pdf}{\alert{Marius Hofert \& Christiane Lemieux }}---\texttt{qrng} \uncover<1>{R package,} Sobol' and Halton &
	  \href{http://gailgithub.github.io/GAIL_Dev/}{\alert{Guaranteed Automatic Integration Library (GAIL)}}---Stopping criteria  \uncover<1>{in MATLAB}
		\tabularnewline
		\uncover<1>{\href{http://www.broda.co.uk}{\alert{BRODA, Sergei Kucherenko}}---Sobol' in C, MATLAB, and Excel}& 
		\uncover<1>{\href{https://www.uqlab.com}{\alert{UQLab}}---Framework for Uncertainty Quantification in MATLAB}
		\tabularnewline
		\href{http://statweb.stanford.edu/~owen/code/}{\alert{Art Owen}}---randomized Halton\uncover<1>{ in R}&
		\multirow{3}{0.47\textwidth}{\uncover<1>{\centering \href{http://www.openturns.org}{\alert{OpenTURNS}---An Open source initiative for the Treatment of Uncertainties, Risks 'N Statistics in Python}}}
		\tabularnewline
		\uncover<1>{\href{https://github.com/PieterjanRobbe/QMC.jl}{\alert{Pieterjan Robbe}---LD sequences in Julia}}
		\tabularnewline
		\href{https://pytorch.org/}{\alert{PyTorch}---Sobol' \uncover<1>{in Python}}
		\tabularnewline
		\multicolumn{2}{>{\centering}m{0.96\textwidth}}{\href{http://simul.iro.umontreal.ca}{\alert{Pierre L'Ecuyer}---Lattice Builder \uncover<1>{and  Stochastic Simulation in C/C++ and Java}}}
		\tabularnewline
		\multicolumn{2}{>{\centering}m{0.96\textwidth}}{\href{https://people.cs.kuleuven.be/~dirk.nuyens/}{\alert{Dirk Nuyens}}---Magic Point Shop \uncover<1>{and QMC4PDE in MATLAB, Python, and C++}}
\tabularnewline
		\multicolumn{2}{>{\centering}m{0.96\textwidth}}{\uncover<1>{\href{http://people.sc.fsu.edu/~jburkardt/}{\alert{John Burkhardt}}---variety in C++, Fortran, MATLAB, \& Python}}
\tabularnewline
		\multicolumn{2}{>{\centering}m{0.96\textwidth}}{\href{https://qmcsoftware.github.io/QMCSoftware/}{\alert{QMCPy}}---Python package \alert<2->{incorporating and connecting} the work of different groups}
\tabularnewline
	\end{tabular}

\renewcommand{\arraystretch}{1}

\end{frame}

\begin{frame}{Why LD is better than IID}
	\vspace{-4ex}
	\begin{description}
		\setlength{\itemsep}{0.5cm}
		\item[Integration/Expectation]  Arising in finance, uncertainty quantification, Bayesian inference, \ldots
		\begin{equation*}
		\mu := \Ex[f(\vX)] = \int_\cx f(\vx) \, \varrho(\vx) \, \dif \vx \approx \frac 1n \sum_{i=1}^n f(\vX_i) =:  \hmu_n
		\end{equation*}
		LD points give faster convergence than IID

		\item[Design of Computer Experiments] LD can be more space filling (even than Latin hypercube sampling) for use in constructing surrogate models and uncertainty quantification

		\item[Global Optimization]  LD points are more space filling and find good starting points for local methods

		\end{description}

\end{frame}

\begin{frame}{What software components do we need?}

	\vspace{-6ex}

	\[
	 \uncover<2->{\mu :=  \int_{\ct} \alert<3>{g(\vt)} \, \alert<2>{\lambda(\vt) \, \dif \vt} = \cdots = } \only<1>{\mu :=} 	\underbrace{\Ex[f(\vX)]}_{\text{expectation}} = \underbrace{\int_\cx \alert<3>{f(\vx)} \, \varrho(\vx) \, \dif \vx}_{\text{integration}} \approx  \frac 1{\alert<4>{n}} \sum_{i=1}^{\alert<4>{n}} f(\alert<1>{\vX_i}) =: \hmu_{\alert<4>{n}}
	\]

		\vspace{-3ex}

	\begin{description}[<+->]
				\setlength{\itemsep}{0.5cm}

		\item[LD Generator] producing $\{\vX_1, \vX_2, \dots \}$ that mimics the distribution with PDF $\varrho$, e.g., uniform

		\item[True Measure] that defines the original integral, e.g., Lebesgue

		\item[Integrand] $g$, which defines the original integral, plus the transformed version, $f$, to fit the LD generator

		\item[Stopping Criterion] that determines how large $n$ should be to ensure that $\abs{\mu - \hmu_n} \le \varepsilon$
	\end{description}
\end{frame}

\section{QMCPy on Google Colaboratory}
\begin{frame}{QMCPy in a Jupyter Notebooks on Google Colaboratory}
	\vspace{-4ex}

	\alert{Prerequisites}---No Python, Jupyter, or QMC knowledge assumed

	\vspace{-1ex}

	\alert{Goals}

		\vspace{-3ex}
		\begin{itemize}
		\item Show you how QMC software works
		\item Interest you in using/contributing
	\end{itemize}

	\vspace{-1ex}

	\alert{Directions}

	\vspace{-3ex}
	\begin{itemize}
		\item Point your browser to \href{https://tinyurl.com/QMCPyTutorial}{\nolinkurl{https://tinyurl.com/QMCPyTutorial}}
		\item Open the file in Google Colaboratory (may need to push a button at the top of your browser)
		\item Make a copy of this file onto your own Google drive account \\
		\texttt{File} $\rightarrow$ \texttt{Save a copy in Drive}
	\end{itemize}

\alert{Pause for questions}
\end{frame}

\section{Why collaborate?}

\begin{frame}
	\frametitle{Acknowledgments for QMCPy}

	\begin{itemize}
		\item Coded almost entirely by Aleksei Sorokin (BS/MS '21)

		\item Relies on code from several groups (see above)

		\item Documentation and testing by Sou-Cheng Choi and Jagadeeswaran R.

		\item Funded and encouraged by Mike McCourt at SigOpt
	\end{itemize}
\end{frame}

\begin{frame}
	{The Guaranteed Automatic Integration Library (GAIL) and QMCPy teams}

	\vspace{-2ex}
	\includegraphics[angle = 180, origin = c, width = 0.32\textwidth]{ProgramsImages/GAIL2014RE.jpeg} \
	\includegraphics[width = 0.32\textwidth]{ProgramsImages/GAILatSIAM2018Hi.jpeg} \ 
	\includegraphics[width = 0.32\textwidth]{ProgramsImages/GAILatChinatown2018.jpg}

	\vspace{-4ex}

	{\small 
		\hspace{-4ex}\begin{tabular}{p{0.545\textwidth}p{0.44\textwidth}}

		\begin{itemize}
			\setlength{\itemsep}{0ex}

			\item Sou-Cheng Choi (Chief Data Scientist, Kamakura)

			\item Yuhan Ding (IIT PhD '15, Lecturer, IIT)

			\item Lan Jiang  (IIT PhD '16, Compass)

			\item Llu\'is Antoni Jim\'enez Rugama (IIT PhD '17, UBS)

			\item Mike McCourt (IIT BS '07, Cornell PhD '12, \\ Head of Research, SigOpt)

			\item Jagadeeswaran Rathinavel (IIT PhD '19, Wi-Tronix)

		\end{itemize}

		&

		\begin{itemize}

			\setlength{\itemsep}{0ex}

			\item Aleksei Sorokin (IIT BS \& MAS '21 exp.)

			\item  Xin Tong (IIT MS, UIC PhD '20 exp.)

			\item Kan Zhang (IIT PhD '20 exp.)

			\item Yizhi Zhang (IIT PhD '18, Jamran Int'l)

			\item Xuan Zhou (IIT PhD '15, JP Morgan)

			\item and others

		\end{itemize}

	\end{tabular}
}

\end{frame}

\begin{frame}{The argument for community software}

	\vspace{-4ex}

	\begin{itemize}
	\item<+-> Our research groups are typically expert at only part of the whole picture:
			\begin{itemize}
			\item LD sequence generators

			\item Increasing efficiency, e.g., MLMC, MDM

			\item Stopping criteria

			\item Use cases
		\end{itemize}

		A community library lets us all take advantage of the best

	\item<+-> Provides a consistent interface for pieces from different places

	\item<+-> Enables reproducible computational research

	\item<+-> Tedious stuff only done once

		\item<+-> Many eyes find and correct errors
		\begin{itemize}
			\item MATLAB's Sobol' generator's  scrambling corrected in MATLAB 2017a after Tony Jim\'enez Rugama noticed the problem

			\item PyTorch's Sobol' generator found to be wrong unless double precision is proactively specified; also missing the first point; reported at \url{https://github.com/pytorch/pytorch/issues/32047}
		\end{itemize}

\end{itemize}
\end{frame}

\begin{frame}{How you can contribute}

		\vspace{-3ex}
	Try out  QMCPy and then

			\vspace{-1ex}

	\begin{description}
		\item[Easy] Submit your bugs and feature requests as issues to \url{https://github.com/QMCSoftware/QMCSoftware/issues}

		\item[Moderately Difficult]  Ask your students or collaborators to try QMCPy themselves and submit their bugs and feature requests\\[1ex]

		Email us your blog post to add to \href{https://qmcpy.wordpress.com/}{\nolinkurl{qmcpy.wordpress.com/}}

		\item[Heroic] Add a feature or use case and make a pull request at \url{https://github.com/QMCSoftware/QMCSoftware/pulls} \\ so that we can included it in our next release
	\end{description}

		\vspace{-1ex}

	Questions or suggestions?  Email us at \href{mailto:qmc-software@googlegroups.com}{\nolinkurl{qmc-software@googlegroups.com}}

			\vspace{-1ex}

	\uncover<2>{If you are wedded to another language, think of designing your software so that others can add to it easily.}

\end{frame}

\finalthanksnote{These slides are  available at \\  \href{https://speakerdeck.com/fjhickernell/quasi-monte-carlo-software}{\nolinkurl{speakerdeck.com/fjhickernell/quasi-monte-carlo-software}}\\
Google Colaboratory notebook at \href{https://tinyurl.com/QMCPyTutorial}{\nolinkurl{tinyurl.com/QMCPyTutorial}}\\
Blog at \href{https://qmcpy.wordpress.com/}{\nolinkurl{qmcpy.wordpress.com/}}}

\thankyouframe

\end{document}

### 2022-06-28 MCQMC 2020 stuff/QMCSoftwareArticleNoRho.tex
Update from Office Computer

%
% sample root file for your "contribution" to a contributed volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[graybox,footinfo]{svmult}

\smartqed
\usepackage{mathptmx}       % selects Times Roman as basic font
\usepackage{helvet}         % selects Helvetica as sans-serif font
\usepackage{courier}        % selects Courier as typewriter font
\usepackage{type1cm}        % activate if the above 3 fonts are
% not available on your system
\usepackage{graphicx}       % standard LaTeX graphics tool
% when including figure files

\usepackage{array,colortbl}
\usepackage{amsmath,amsfonts,amssymb,bm} % no amsthm, Springer defines Theorem, Lemma, etc themselves
%\usepackage[mathx]{mathabx}
\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{
	<5> <6> <7> <8> <9> <10>
	<10.95> <12> <14.4> <17.28> <20.74> <24.88>
	mathx10
}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareFontSubstitution{U}{mathx}{m}{n}
\DeclareMathAccent{\widecheck}{0}{mathx}{"71}

% Note that Springer defines the following already:
%
% \D upright d for differential d
% \I upright i for imaginary unit
% \E upright e for exponential function
% \tens depicts tensors as sans serif upright
% \vec depicts vectors as boldface characters instead of the arrow accent
%
% Additionally we throw in the following common used macro's:
\input{macros}

% Macros below are now included in macros.tex from MCQMC 2016 web site
% This spot formerly included macros that are now in macros.tex

% indicator boldface 1:
\DeclareSymbolFont{bbold}{U}{bbold}{m}{n}
\DeclareSymbolFontAlphabet{\mathbbold}{bbold}
%\newcommand{\ind}{\mathbbold{1}}

\usepackage{microtype} % good font tricks

\usepackage[colorlinks=true,linkcolor=black,citecolor=black,urlcolor=blue]{hyperref}
\urlstyle{same}
\usepackage{bookmark}
\pdfstringdefDisableCommands{\def\and{, }}
\makeatletter % to avoid hyperref warnings:
\providecommand*{\toclevel@author}{999}
\providecommand*{\toclevel@title}{0}
\makeatother

\usepackage{bbm,mathtools,array,longtable,booktabs,graphicx,color,enumitem,xspace}
%\input FJHDef.tex

\newcommand{\QMCPYnorm}[2][{}]{\ensuremath{\left \lVert #2 \right \rVert}_{#1}}
\newcommand{\QMCPYnormnorm}[2][{}]{\ensuremath{\lVert #2 \rVert}_{#1}}
\newcommand{\QMCPYbignorm}[2][{}]{\ensuremath{\bigl \lVert #2 \bigr \rVert}_{#1}}
\newcommand{\QMCPYBignorm}[2][{}]{\ensuremath{\Bigl \lVert #2 \Bigr \rVert}_{#1}}
\newcommand{\QMCPYabs}[1]{\ensuremath{{\left \lvert #1 \right \rvert}}}
\newcommand{\QMCPYbigabs}[1]{\ensuremath{{\bigl \lvert #1 \bigr \rvert}}}
\newcommand{\QMCPYBigabs}[1]{\ensuremath{{\Bigl \lvert #1 \Bigr \rvert}}}
\newcommand{\QMCPYbiggabs}[1]{\ensuremath{{\biggl \lvert #1 \biggr \rvert}}}

\providecommand{\HickernellFJ}{Hickernell}

\definecolor{orange}{rgb}{1.0,0.3,0.0}
\definecolor{violet}{rgb}{0.75,0,1}
\newcommand{\frednote}[1]{  {\textcolor{red}  {\mbox{**Fred:} #1}}}
\newcommand{\scnote}[1]{ {\textcolor{blue}  {\mbox{**SC:} #1}}}
\newcommand{\tonynote}[1]{ {\textcolor{orange}  {\mbox{**Tony:} #1}}}

%\journal{Journal of Complexity}

\allowdisplaybreaks[4]

%   python styling
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{listings}
\usepackage{xcolor}
\definecolor{darkgreen}{rgb}{0,0.6,0}
\lstdefinestyle{Python}{
    showstringspaces=false,
    language        = Python,
    basicstyle      = \small\ttfamily,
    morekeywords = {as},
    keywordstyle    = \color{blue},
    stringstyle     = \color{darkgreen},
    commentstyle    = \color{darkgreen}\ttfamily,
	breaklines = true,
	postbreak=\text{$\hookrightarrow$\space},
	% style >>> and ... 
	%   see: https://tex.stackexchange.com/questions/326655/make-a-keyword-in-listings-enviorment
	alsoletter = {>,.} ,
    morekeywords = [2]{>>>,...},
    keywordstyle = [2]\color{cyan}\bfseries}

\newcommand{\AGSComment}[1]{{\color{cyan} Aleksei: #1}}
\newcommand{\FJHComment}[1]{{\color{magenta} Fred: #1}}
\newcommand{\JRComment}[1]{{\color{violet} Jag: #1}}

\newcommand{\hmu}{\widehat{\mu}}
\newcommand{\IID}{\textup{IID}}
\newcommand{\LD}{\textup{LD}}
\newcommand{\IMP}{\textup{IS}}
\newcommand{\unif}{\textup{unif}}
\newcommand{\IIDsim}{\overset{\IID}{\sim}}
\newcommand{\LDsim}{\overset{\LD}{\sim}}
\newcommand{\Msim}{\overset{\textup{M}}{\sim}}
\newcommand{\dig}{\textup{dig}}
\newcommand{\mA}{\mathsf{A}}
\newcommand{\mSigma}{\mathsf{\Sigma}}
\newcommand{\mLambda}{\mathsf{\Lambda}}
\newcommand{\mI}{\mathsf{I}}
\newcommand{\cube}{[0,1]^d}
\newcommand{\Kum}{\textup{Kum}}

\iffalse
Here are some of the main suggestions, other grammatical / wording suggestions are in the reports

* General ~ Sou-Cheng DONE
Expand section 2 and talk more about how QMCPy combines the best of available QMC softwares
Fix LatNet Builder reference Fred
Add tensorflow

* Discrete Distribution ~ Fred
Discuss difference between stratification and RQMC
Comment on the name DiscreteDistribution, the referee and Dirk Nuyens found this confusing Fred done
List criterion used to construct Lattice sequences Aleksei done
Clarify that the better uniformity is for embedded Lattices and is not true in general Fred done

* Code / Choice of Parameters ~ Aleksei, Jags check
Show how the user can customize Sobol' and Lattice generative vectors
Make parameters more explicit e.g. Sobol(2) --> Sobol(dimension=2), and explain more about what the code does for non-python / QMCPy users

* Integrand / Stopping Criterion ~ Jags done, Sou-Cheng check
Explain why we use the trapezoidal rule for AsianOption and how we chose the drift for the importance sampling example? 
Better explanation of CubQMCSobolG algorithm

* Importance Sampling ~ Aleksei done, Fred check
Provide better discussion of importance sampling and justification for the example
Motivation for successive change of variables
Explain the Kumaraswamy distribution

\fi

\begin{document}

\title*{Quasi-Monte Carlo Software}
\authorrunning{S.-C.\ T.\ Choi et al.}
\author{Sou-Cheng T. Choi \and Fred J. Hickernell \and R. Jagadeeswaran \and Michael J. McCourt \and Aleksei G. Sorokin}
\institute{Sou-Cheng T. Choi \at Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616; and Kamakura Corporation, 2222 Kalakaua Ave, Suite 1400, Honolulu, HI 96815 \email{schoi32@iit.edu}
\and
Fred J. Hickernell \at Center for Interdisciplinary Scientific Computation and \\
Department of Applied Mathematics, Illinois Institute of Technology \\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{hickernell@iit.edu}
\and
R. Jagadeeswaran \at
Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{jrathin1@iit.edu}; and \\Wi-Tronix LLC, 631 E Boughton Rd, Suite 240, Bolingbrook, IL 60440
\and
Michael J. McCourt \at SigOpt, an Intel company, \\
100 Bush St., Suite 1100, San Francisco, CA 94104
\email{mccourt@sigopt.com}
\and 
Aleksei G. Sorokin \at
Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{asorokin@hawk.iit.edu}}

\maketitle

\abstract{Practitioners wishing to experience the efficiency gains from using low discrepancy sequences need correct, robust, well-written software.  This article, based on our MCQMC 2020 tutorial, describes some of the better quasi-Monte Carlo (QMC) software available.  We highlight the key software components required  by QMC to approximate multivariate integrals or expectations of functions of vector random variables.  We have combined these components in QMCPy, a Python open-source library, which we hope will draw the support of the QMC community.  Here we introduce QMCPy.}

\section{Introduction} \label{sec:intro}

Quasi-Monte Carlo (QMC) methods promise great efficiency gains over independent and identically distributed (IID) Monte Carlo (MC) methods.  In some cases, QMC  achieves one hundredth of the error of IID MC in the same amount of time (see Figure \ref{fig:sc_comp}). Often, these efficiency gains are obtained simply by replacing IID sampling with low discrepancy (LD) sampling, which is the heart of QMC.

Practitioners might wish to test whether QMC would speed up their computation; doing so would be made easier with access the best QMC algorithms available.  Theoreticians or algorithm developers might want to demonstrate their ideas on various use cases to show their practical value.

This tutorial points to some of the best QMC software available.  Then we describe  QMCPy \cite{QMCPy2020a}\footnote{QMCPy is in active development. This article is based on version 1.0.1 on PyPI.}, which is crafted to be a community-owned Python library that combines the best QMC algorithms and interesting use cases from various authors under a common user interface.

The model problem for QMC is approximating a multivariate integral,
\begin{equation} \label{eq:integral}
	\mu := \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst,
\end{equation}
where $g$ is the integrand, and $\lambda$ is a non-negative weight.  If $\lambda$ is a probability distribution (PDF) for the random variable $\bsT$, then $\mu$ is the mean of $g(\bsT)$.  Regardless, we perform a suitable variable transformation to interpret this integral as the  mean of a function of a multivariate, standard uniform random variable:
\begin{equation} \label{eq:fintegral}
	\mu = \bbE[f(\bsX)] =  \int_{\cube}  f(\bsx) \,  \D \bsx , \qquad \bsX \sim \calU\cube.
\end{equation}

QMC approximates the population mean, $\mu$,  by a sample mean,
\begin{equation} \label{eq:samplemean}
	\hmu := \frac 1n \sum_{i=0}^{n-1} f(\bsX_i), \qquad \bsX_0, \bsX_1, \ldots \Msim \calU\cube.
\end{equation}
The choice of the sequence $\{\bsX_i\}_{i=0}^\infty$ and the choice of $n$ to satisfy  the prescribed error requirement,
\begin{equation} \label{eq:err_req}
	\QMCPYabs{\mu - \hmu} \le \varepsilon \qquad \text{absolutely or with high probability},
\end{equation} 
are important decisions, which  QMC software helps the user make.

Here, the notation $\Msim$ means that the sequence mimics the specified, target distribution, but not necessarily in a probabilistic way.  We  use this notation in two forms:  $\IIDsim$ and $\LDsim$.

IID sequences must be random. The position of any point is not influenced by any other, so clusters and gaps occur.  A randomly chosen subsequence of an IID sequence is also IID.  When we say that $\bsX_0, \bsX_1, \ldots \IIDsim F$ for some distribution $F$, we mean that for any positive integer $n$, the  multivariate probability distribution of $\bsX_0, \ldots, \bsX_{n-1}$ is the product of the marginals, specifically,
\begin{equation*}
	F_{n}(\bsx_0, \ldots, \bsx_{n-1}) = F(\bsx_0) \cdots  F(\bsx_{n-1}).
\end{equation*}
When IID points are used to approximate $\mu$ by the sample mean, the root mean squared error is $\calO(n^{-1/2})$.  Figure \ref{fig:comparePts} displays IID uniform points, $\bsX^{\IID}_0, \bsX^{\IID}_1, \ldots \IIDsim \calU[0,1]^2$, i.e.,  the target distribution is $F_\unif(\bsx) = x_1 x_2$.

\begin{figure}[t]
	\includegraphics[height=5.8cm]{QMCSoftwareArticle/vfigs/dd_iid_uniform_pts.eps}
	\quad
	\includegraphics[height=5.8cm]{QMCSoftwareArticle/vfigs/dd_sobol_pts.eps}
	\caption{IID points (left) contrasted with LD points (right).  The LD points cover the square more evenly.}
	\label{fig:comparePts}
\end{figure}

LD sequences may be deterministic or random, but each point is carefully coordinated with the others so that they fill the domain well.  Subsequences of LD sequences are generally not LD.  When we say that $\bsX_0, \bsX_1, \ldots \LDsim \calU\cube$, we mean that for any positive integer $n$,  the \emph{empirical distribution} of $\bsX_0, \ldots, \bsX_{n-1}$, denoted $F_{\{\bsX_i\}_{i=0}^{n-1}}$,  approximates the uniform distribution, $F_\unif$, well (relative to $n$).  (The empirical distribution of a set assigns equal probability to each point.)

A measure of the difference between the empirical distribution of a set of points and the uniform distribution is called a \emph{discrepancy} and is denoted $D(\{\bsX_i\}_{i=0}^{n-1})$ \cite{DicEtal14a,Hic97a,Hic99a,Nie92}.  This is the origin of the term ``low discrepancy'' points or sequences.  LD points by definition have a smaller discrepancy than IID points.  Figure \ref{fig:comparePts} contrasts IID uniform points with LD points, $\bsX^{\LD}_0, \bsX^{\LD}_1 \ldots \LDsim \calU[0,1]^2$, in this case, linearly scrambled and digitally shifted Sobol' points.

The error in using the sample mean to approximate the integral can be bounded according to the Koksma-Hlawka inequality and its extensions \cite{DicEtal14a,Hic97a,Hic99a,Nie92} as the product of the discrepancy of the sampling sequence and the variation of the integrand, denoted $V(\cdot)$:
\begin{equation}
	\QMCPYabs{\mu - \hmu} = \QMCPYbiggabs{\int_{\cube} f(\bsx) \, \D \bigl(F_\unif - F_{\{\bsX_i\}_{i=0}^{n-1}}\bigr) (\bsx)} \le D\bigl(\{\bsX_i\}_{i=0}^{n-1}\bigr) V(f).
\end{equation} 
The variation is a (semi-) norm of the integrand in a suitable Banach space.  The discrepancy corresponds to the norm of the error functional for that Banach space.  For typical Banach spaces, the discrepancy of LD points is $\calO(n^{-1+\epsilon})$, which is a higher convergence order than for IID points.  For details, the reader is pointed to the references.

Here, we expect the reader to see in Figure \ref{fig:comparePts} that the LD points cover the integration domain more evenly than IID points.  LD sampling can be thought of as a more even distribution of the sampling sites than IID.  LD sampling is similar to stratified sampling.  In the examples below, the reader will see the demonstrably smaller cubature errors arising from using LD points.

In the sections that follow, we first overview available QMC software.  We next describe an architecture for good QMC software, i.e., what the key components are and how they should interact.  We then describe how we have implemented this architecture in QMCPy.  Finally, we summarize further directions that we hope QMCPy and other QMC software projects will take.  Those interested in following or contributing to the development of QMCPy are urged to visit the GitHub repository at \href{https://github.com/QMCSoftware/QMCSoftware}{\nolinkurl{https://github.com/QMCSoftware/QMCSoftware}}.

We have endeavored to be as accurate as possible at the time of writing this article.  We hope that progress in QMC software development will make this article obsolete in a good way in the coming years.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Available Software for QMC} \label{sec:available} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
QMC software spans  LD sequence generators, cubatures, and applications.  We review the better-known software collections, recognizing that some software overlaps multiple categories. Whenever applicable, we state each library's accessibility in QMCPy, or contrast its functionalities with QMCPy's --- where we lag behind in QMCPy, we strive to catch up in the near future.

Software focusing on generating high-quality LD sequences and their generators includes the following, listed in alphabetical order:
\begin{description}[format=\textup,format=\textbf]
% https://tex.stackexchange.com/questions/74279/how-to-add-bullets-to-description-lists
	\item[BRODA] Commerical and non-commercial software developed jointly with I.M.~Sobol' in C++, MATLAB, and Excel~\cite{BRODA20a}. BRODA can generate Sobol' sequences up to 65,536 dimensions. In comparison, QMCPy supports Sobol' sequences up to 21,201 dimensions.	
	\item[Burkhardt] Various QMC software for generating van der Corput, Faure, Halton, Hammersley,  Niederreiter, or Sobol' sequences in C, C++, Fortran, MATLAB, or Python~\cite{Bur20a}. In QMCPy, we have implemented digital net, lattice, and Halton generators.
	\item[LatNet Builder] The successor to \emph{Lattice Builder}~\cite{LEcMun14a}, LatNet Builder is a C++ library with Python and Java interfaces (in SSJ below) for generating vectors/matrices for lattices and digital nets~\cite{LatNet,LEcEtal22a}. QMCPy contains a module to parse the generators from LatNet Builder for compatibility with our LD point generators. %contains a \href{https://qmcpy.readthedocs.io/en/latest/demo_rst/latnetbuilder.html}{demo for LatNet Builder}. 
	\item[MATLAB] Commercial software \cite{MAT9.10}, which contains Sobol' and Halton sequences in the Statistics and Machine Learning Toolbox. Both generators can be applied in conjunction with MATLAB's Parallel Computing Toolbox to accelerate their execution speed. The maximum dimension of the Sobol' sequences is restricted to 1,111, which is relatively small, yet sufficient for most applications. 
	\item[MPS] Magic Point Shop contains lattices and Sobol' sequences in Python, Matlab, and C++ \cite{Nuy17a}. We have used MPS as a starting point for developing LD generators in QMCPy.
	\item[Owen] Randomized Halton sequences with dimensions up to 1,000 \cite{Owe20a} and scrambled Sobol' sequences with dimensions up to 21,021  \cite{owen1998scrambling, owen2021r}  in R. QMCPy currently supports Owen's Halton randomization method, and we plan to implement Owen's nested uniform scrambling for digital nets in the near future.
	\item[PyTorch] Open-source Python library for deep learning, with scrambled or unscrambled Sobol' sequences \cite{NEURIPS2019_9015,PyTorch}. PyTorch enables seamless utilization of Graphics Processing Units (GPUs) or Field Programmable Gate Arrays (FPGAs).
	\item[QMC.jl] LD Sequences in Julia \cite{Rob20a}. Julia~\cite{bezanson2012julia} is an interpreted language similar to Python and R in terms of ease of use, but is designed to run much faster.
	\item [qrng] Randomized Sobol', Halton, and Korobov sequences in R~\cite{QRNG2020}. The default Halton randomization in QMCPy utilizes the methods from qrng. 
	%translated qrng to Python; see  \url{https://github.com/QMCSoftware/qrngpy}. The Halton randomization from qrng is also optionally available in QMCPy. \AGSComment{We don't really use qrng besides the QRNG Halton randomization.} In addition, the C source files in qrng can be compiled and accessed in QMCPy as one of the backends. In some cases, our version of qrng with its compiled code in QMCPy is more efficient than the original software in R \scnote{Why or how? By taking advantage of MRG63k3a?}; see the demo, \href{https://qmcpy.readthedocs.io/en/latest/demo_rst/quasirandom_generators.html}{\emph{Quasi-Random Sequence Generator Comparison}}, in the QMCPy documentation~\cite{QMCPyDocs}. 
	\item[SciPy] Scientific computing library in Python with Latin hypercube,  Halton, and Sobol' generators \cite{2020SciPy-NMeth}.
	\item[TF Quant Finance] Google's Tensorflow  deep-learning library~\cite{tensorflow2015-whitepaper} specialized for financial modeling \cite{tfqf21}. The software contains lattice and Sobol' generators alongside algorithms accelerated useing GPUs, FGPAs, or Tensor Processing Units (TPUs).% as opposed to the Central Processing Units (CPUs).		
\end{description}
Software focusing on QMC cubatures and applications includes the following:
\begin{description}[format=\textup,format=\textbf]
	\item[GAIL] Guaranteed Automatic Integration Library  contains automatic (Q)MC stopping criteria in MATLAB \cite{ChoEtal21a,HCJJ14}. These are iterative procedures for one- or high-dimensional integration that take a user's input error tolerance(s) and determine the number of (Q)MC sampling points necessary to achieve the desired accuracy (almost surely).  Most of GAIL's (Q)MC functions, some with enhancements, are implemented in Python in QMCPy.
	\item[ML(Q)MC] Multi-Level (Q)MC routines in C, C++, MATLAB, Python, and R~\cite{GilesSoft}. We have ported ML(Q)MC functions to QMCPy.
	\item[MultilevelEstimators.jl] ML(Q)MC methods in Julia \cite{Rob21}. The author, Pieterjan Robbe, has contributed cubature algorithms and use cases to QMCPy.
	\item[OpenTURNS] Open source initiative for the Treatment of Uncertainties, Risks'N Statistics \cite{OpenTURNS} written in C++ and Python, leveraging R statistical packages, as well as LAPACK and BLAS for numerical linear algebra.
	\item[QMC4PDE] QMC for elliptic PDEs with random diffusion coefficients in Python \cite{KuoNuy16a}.
	\item[SSJ] Stochastic Simulation with the \texttt{hups} package in Java \cite{SSJ}.
	\item[UQLab] Framework for Uncertainty Quantification in MATLAB \cite{UQLab2014}.  The core of UQLab is closed source, but a large portion of the library is open source. Recently,  UQ[py]Lab, the beta release of UQLab with Python bindings, is  available as Software as a Service (SaaS) via UQCloud~\cite{lataniotis2021uncertainty}.

\end{description}

The sections that follow describe QMCPy \cite{QMCPy2020a}, which is our attempt to establish a framework for QMC software and to combine the best of the above software under a common user interface written in Python 3.  The choice of language was determined by the desire to make QMC software accessible to a broad audience, especially the technology industry.

\section{Components of QMC Software}
QMC cubature can be summarized as follows.  We want to approximate the expectation, $\mu$, well by the sample mean, $\hmu$, where \eqref{eq:integral}, \eqref{eq:fintegral}, and \eqref{eq:samplemean} combine to give
\begin{multline} \label{eq:cubSummary}
	\mu : = \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  = \bbE[f(\bsX)] = \int_{\cube} f(\bsx) \, \D \bsx \approx \frac 1n \sum_{i=0}^{n-1} f(\bsX_i) =: \hmu, \\
	 \bsX \sim \calU\cube, \ \bsX_0, \bsX_1, \ldots \Msim \calU\cube.
\end{multline}
Moreover, we want to satisfy the error requirement in \eqref{eq:err_req}.
This requires four components, which we implement as QMCPy classes.

\begin{description}[format=\textup,format=\textbf]

	\item[Discrete Distribution]  produces the sequence $\bsX_0, \bsX_1, \dots$ that mimics $\calU[0,1]^d$;

	\item[True Measure] $\bst \mapsto \lambda (\bst) \D \bst$  defines the original measure, e.g., Gaussian or Lebesgue;

	\item[Integrand] $g$  defines the original integrand, and $f$ defines the transformed version to fit the \texttt{DiscreteDistribution}; and

	\item[Stopping Criterion] determines how large $n$ should be to ensure that $\QMCPYabs{\mu - \hmu} \le \varepsilon$ as in (\ref{eq:err_req}).
\end{description}

The software libraries referenced in Section \ref{sec:available} provide one or more of these components. QMCPy combines multiple examples of all these components under an object-oriented framework. Each example is implemented as a concrete class that realizes the properties and methods required by the abstract class for that component. The following sections detail descriptions and specific examples for each component.

Thorough documentation of all QMCPy classes is available in \cite{QMCPyDocs}. Demonstrations of how QMCPy works are given in Google Colab notebooks \cite{QMCPyTutColab2020,QMCPyTutColab2020_paper}. The project may be installed from PyPI into a Python 3 environment via the command \texttt{pip install qmcpy}. In the code snippets that follow, we assume QMCPy has been imported alongside NumPy \cite{numpy} via the following commands in a Python Console:
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/import.txt}

\section{Discrete Distributions}

LD sequences typically mimic $\calU[0,1]^d$.  Good sequences mimicking other distributions are obtained by transformations as described in the next section.  We denote by \texttt{DiscreteDistribution} the abstract class containing LD sequence generators.  In most cases that have been implemented, the points $\bsX_0, \ldots, \bsX_{n-1}$ have an empirical (discrete) distribution that closely approximates the uniform distribution, say, in the sense of discrepancy.  We also envision future possible \texttt{DiscreteDistribution} objects that assign unequal weights to the sampling points.  In any case, the term ``discrete'' refers to the fact that these are sequences of points (and weights), not continuous distributions.

QMCPy implements \emph{extensible} LD sequences, i.e., those that allow practitioners to obtain and use $\bsX_n, \bsX_{n+1}, \ldots $ without discarding $\bsX_0, \ldots, \bsX_{n-1}$.  Halton sequences do not have preferred sample sizes $n$, but extensible integration lattices and digital sequences in base $b$ prefer $n$ to be a power of $b$.  For integration lattices and digital sequences, we have focused on base $b=2$ since this is a popular choice and for convenience in generating extensible sequences.

Integration lattices and digital sequences in base $2$ have an elegant group structure, which we summarize in Table \ref{tab:GroupProp}.  The addition operator  is $\oplus$\footnote{The operator $\oplus$ is commonly used to denote exclusive-or, does correspond to its meaning for digital sequences in base $2$.  However, we are using it here in a more general sense.}, and its inverse is $\ominus$.  The unshifted sequence is $\bsZ_0, \bsZ_1, \ldots$ and the randomly shifted sequence is $\bsX_0, \bsX_1, \ldots$

\begin{table}[h]
	\centering
	\caption{Properties of lattices and digital net sequences.  Note that they share group properties but also have distinctives.} \label{tab:GroupProp}
\[
	\renewcommand{\arraystretch}{1.3}
\begin{array}{c@{\qquad}c}
	\toprule
	\multicolumn{2}{c}{\text{Define \ldots}} \\
	\multicolumn{2}{c}{\bsZ_1, \bsZ_2, \bsZ_4, \ldots \in [0,1)^d \text{ chosen well} } \\
	\multicolumn{2}{c}{
	\bsZ_{i} := i_0  \bsZ_1 \oplus i_1 \bsZ_{2} \oplus i_2  \bsZ_{4} \oplus  i_3  \bsZ_{8} \oplus \cdots 
	\quad
	\text{for }i = i_0 +i_1 2 + i_2 4 + i_3 8 + \cdots, \; i_\ell \in \{0,1\}} \\
    \multicolumn{2}{c}{\bsX_i := \bsZ_i \oplus \bsDelta, \qquad \text{where }\bsDelta \IIDsim [0,1)^d} \\  \hline
	\text{Rank-1 Integration Lattices} & \text{Digital Nets} \\
		\bst \oplus \bsx : = (\bst + \bsx) \bmod \bsone & \bst \oplus \bsx := \text{binary digitwise addition}, \oplus_{\dig} \\ 
		\text{require } \bsZ_{2^{m}} \oplus \bsZ_{2^{m}} = \bsZ_{\lfloor2^{m-1}\rfloor}
		\quad \forall m \in \bbN_0 
%		 \\ \text{require } \begin{array}{l} \bsZ_1 = (1/2, \ldots, 1/2) \\
%		 	\bsZ_{2^{m}} \oplus \bsZ_{2^{m}} = \bsZ_{2^{m-1}} \quad \forall m \in \bbN 
%	 	\end{array}
		\\[.5ex]
\toprule
\multicolumn{2}{c}{\text{Then it follows that \ldots}} \\
	\multicolumn{2}{c}{\left . \begin{array}{r}
			\calP_m := \{\bsZ_0, \ldots, \bsZ_{2^m-1}\}, \quad
			\bsZ_i \oplus \bsZ_j \in \calP_m \\
			\calP_{\bsDelta,m} := \{\bsX_0, \ldots, \bsX_{2^m-1}\}, \quad
			\bsX_i \oplus \bsX_j \ominus \bsX_k \in \calP_{\bsDelta,m}
	\end{array} \right \}\quad \begin{array}{l}\forall  i,j,k \in \{0, \ldots, 2^{m} -1\} \\ \forall m \in \bbN_0\end{array}} \\
\bottomrule
\end{array}
\]
\end{table}

We illustrate lattice and Sobol' sequences using QMCPy. First, we create an instance of a $d=2$ dimensional \texttt{Lattice} object of the  \texttt{DiscreteDistribution} abstract class. Then we generate the first eight (non-randomized) points in this lattice. 
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/dd_lattice.txt}
The first three generators for this lattice are $\bsZ_1 = (0.5, 0.5)$, $\bsZ_2 = (0.25, 0.75)$, and $\bsZ_4 = (0.125, 0.375)$.  One can check that $(\bsZ_2 + \bsZ_4) \bmod \bsone = (0.375, 0.125) = \bsZ_6$, as Table \ref{tab:GroupProp} specifies.

The random shift has been turned off above to illuminate the group structure.  We normally include the randomization to ensure that there are no points on the boundary of $[0,1]^d$.  Then, when points are transformed to mimic distributions such as the Gaussian, no LD points will be transformed to infinity.  Turning off the randomization generates a warning when the \texttt{gen\_samples} method is called.

Now, we generate Sobol' points using a similar process as we did for lattice points.  Sobol' sequences are one of the most popular example of digital sequences.
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/dd_sobol.txt}
Here, $\bsZ_4$ differs from that for lattices, but more importantly, addition for digital sequences differs from that for lattices.  Using digitwise addition for digital sequences, we can confirm that according to Table \ref{tab:GroupProp},
\begin{align*}
\bsZ_2 \oplus_{\dig} \bsZ_4 = (0.25,0.75)  \oplus_{\dig} (0.125,0.625) \\
=  ({}_20.010,{}_20.110)  \oplus_{\dig} ({}_20.001,{}_20.101) &= ({}_20.011,{}_20.011) \\
& = (0.375,0.375) = \bsZ_6.
\end{align*}

By contrast, if we construct a digital sequence using the generators for the lattice above with $\bsZ_2 = (0.25, 0.75)$, and $\bsZ_4 = (0.125, 0.375)$, we would obtain
\begin{multline*}
\bsZ_6 = \bsZ_2 \oplus_{\dig} \bsZ_4   = ({}_20.010,{}_20.110)  \oplus_{\dig} ({}_20.001,{}_20.011)  \\
= ({}_20.011,{}_20.101) = (0.375, 0.625),
\end{multline*}
which differs from the $\bsZ_6=(0.375, 0.125)$ constructed for lattices.  To emphasize, lattices and digital sequences are different, even if they share the same generators, $\bsZ_1, \bsZ_2, \bsZ_4, \ldots$.

The examples of \texttt{qp.Lattice} and \texttt{qp.Sobol} illustrate how QMCPy LD generators share a common user interface.  The dimension is specified when the instance is constructed, and the number of points is specified when the \texttt{gen\_samples} method is called.  Following Python practice, parameters can be input without specifying their names if they are input in the prescribed order.  QMCPy also includes Halton sequences and IID sequences, again deferring details to the QMCPy documentation \cite{QMCPyDocs}.

A crucial difference between IID generators and LD generators is reflected in the behavior of generating $n$ points.  For an IID generator, asking for $n$ points repeatedly gives different points each time because they are meant to be random and independent.
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/dd_iidone.txt}
Your output may look different depending on the seed used to generate these random numbers.

On the other hand, for an LD generator, asking for $n$ points repeatedly gives \emph{the same} points each time because they are meant to be the first $n$ points of a specific LD sequence.  
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/dd_latticeone.txt}
Here we allow the randomization so that the first point in the sequence is not the origin.  To obtain the \emph{next} $n$ points, one may specify the start and ending indices of the sequence.
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/dd_latticenextone.txt}

Figure \ref{fig:increase_n} shows how increasing the number of lattice and Sobol' LD points through powers of two fills in the gaps in an even way.

\begin{figure}[t]
	\includegraphics[width=1\textwidth]{QMCSoftwareArticle/vfigs/dd_lattice_successive.eps}
	\qquad
	\includegraphics[width=1\textwidth]{QMCSoftwareArticle/vfigs/dd_sobol_successive.eps}
	\caption{Randomized lattice and Sobol' points mimicking a $\calU[0,1]^2$ measure for $n = 64, 128,$ and 256. Note how increasing the number of points evenly fills in the gaps between the points.}
	\label{fig:increase_n}
\end{figure}

\section{True Measures}

The LD sequences implemented as \texttt{DiscreteDistribution} objects mimic the $\calU[0,1]^d$ distribution.  However, we may need sequences to mimic other distributions.  This is implemented via variable transformations, $\bsPsi$.  In general, if $\bsX \Msim \calU[0,1]^d$, then
\begin{subequations} \label{eq:exampleVarTrans}
\begin{gather}
\bsT = \bsPsi(\bsX) := \bsa  + (\bsb - \bsa) \odot \bsX \Msim  \calU[\bsa,\bsb], \\
\label{eq:exampleVarTransGauss}
\bsT = \bsPsi(\bsX) := \bsa + \mA \bsPhi^{-1}(\bsX)  \Msim \calN(\bsa, \mSigma), \\
\nonumber  \text{where }  \bsPhi^{-1}(\bsX) : = \begin{pmatrix} \Phi^{-1}(X_1) \\ \vdots \\ \Phi^{-1}(X_d)\end{pmatrix}, \qquad \mSigma = \mA \mA^T,
\end{gather}
\end{subequations}
and $\odot$ denotes term-by-term (Hadamard) multiplication.  Here, $\bsa$ and $\bsb$ are assumed to be finite, and $\Phi$ is the standard Gaussian distribution function.  Again we use $\Msim$ to denote mimicry, not necessarily in a probabilistic sense.

Figure \ref{fig:tm_ug} displays LD sequences transformed as described above to mimic a uniform and a Gaussian distribution.  The code to generate these points takes the following form for uniform points based on a Halton sequence: 
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/tm_uniform.txt}
whereas for Gaussian points based on a lattice sequence, we have:
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/tm_gaussian.txt}
Here the covariance decomposition $\mSigma = \mA \mA^T$ is done using principal component analysis. The Cholesky decomposition is also available.

\begin{figure}[t]
	\includegraphics[width=.45\textwidth]{QMCSoftwareArticle/vfigs/tm_uniform.eps} 
	%	\quad
	\includegraphics[width=.45\textwidth]{QMCSoftwareArticle/vfigs/tm_gaussian.eps}
	\caption{Halton samples transformed to mimic a uniform $\calU\left( \begin{bmatrix} -2 \\ 0 \end{bmatrix},\begin{bmatrix} 2 \\ 4 \end{bmatrix} \right)$ 
	%\frednote{This is not correct, we need the upper and lower endponts of a two dimensional interval.  Also, are these Sobol' or Halton as in the code?} 
	distribution (left) and lattice samples transformed to mimic a Gaussian $\calN\left(\begin{bmatrix} 3 \\ 2 \end{bmatrix}, \begin{bmatrix} 9 & 5 \\ 5 & 4 \end{bmatrix} \right)$ distribution (right).}
	\label{fig:tm_ug}
\end{figure}

The Brownian motion distribution arises often in financial risk applications.  Here the $d$ components of the variable $\bsT$ correspond to the discretized Brownian motion at times $\tau/d, 2\tau/d, \ldots, \tau$, where $\tau$ is the time horizon.  The distribution is a special case of the Gaussian with covariance 
\begin{equation} \label{eq:BMcov}
	\mSigma = (\tau/d) \bigl (\min(j,k) \bigr)_{j,k=1}^d
\end{equation}
and mean $\bsa$, which  is proportional to the times $(\tau/d)(1, 2, \ldots, d)^T$. The code for generating a Brownian motion is
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/tm_brownian_motion.txt}
Figure \ref{fig:tm_bm} displays a Brownian motion based on Sobol' sequence with and without a drift.

\begin{figure}[t]
	\includegraphics[width=1\textwidth]{QMCSoftwareArticle/vfigs/tm_bm.eps} 
	\caption{Sobol' samples transformed to mimic a 52-dimensional Brownian Motion without drift (left) and with drift coefficient 2 (right).}
	\label{fig:tm_bm}
\end{figure}

\section{Integrands}

Let's return  to the integration problem in \eqref{eq:integral}, which we must rewrite as \eqref{eq:fintegral}.  We choose a transformation of variables defined as $\bst = \bsPsi(\bsx)$ where $\bsPsi:\cube \to \calT$.  This leads to 
\begin{align}
	\nonumber 
 \mu &= \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  = \int_{\cube} g\bigl(\bsPsi(\bsx)\bigr) \, \lambda\bigl(\bsPsi(\bsx)\bigr) \,\QMCPYabs{\bsPsi'(\bsx)}  \, \D \bsx =  \int_{\cube}f(\bsx) \, \D \bsx,  \\
 \label{eq:transVar}
  & \qquad \qquad \text{where } f(\bsx)  = g\bigl(\bsPsi(\bsx)\bigr)  \, \lambda\bigl(\bsPsi(\bsx)\bigr) \,\QMCPYabs{\bsPsi'(\bsx)} ,
\end{align}
and $\QMCPYabs{\bsPsi'(\bsx)} := \QMCPYabs{\partial \bsPsi/\partial \bsx}$ represents the Jacobian of the variable transformation.  The abstract class \texttt{Integrand} provides $f$ based on the user's input of $g$ and the \texttt{TrueMeasure} instance, which defines $\lambda$ and the transformation $\bsPsi$. Different choices of $\bsPsi$ lead to different $f$, which may give different rates of convergence of the cubature, $\hmu$ to $\mu$.

We illustrate the \texttt{Integrand} class via an example of Keister \cite{Kei96}:
\begin{equation} \label{eq:KeisterIntegral}
	\mu 
	= \int_{\bbR^d} \cos(\lVert \bst \rVert) \exp(-\bst^T \bst) \, \D \bst \\ 
	= \int_{\bbR^d} \underbrace{\pi^{d/2} \cos(\lVert \bst \rVert)}_{g(\bst)}\, \underbrace{\pi^{-d/2} \exp(-\bst^T \bst) }_{\lambda(\bst)} \, \D \bst.
\end{equation}
Since $\lambda$ is the density for $\calN(\bszero,\mI/2)$, it is natural to choose $\bsPsi$ according to \eqref{eq:exampleVarTransGauss} with $\mA = \sqrt{1/2} \, \mI$, in which case $\lambda(\bsPsi(\bsx)) \QMCPYabs{\bsPsi'(\bsx)}  = 1$, and so 
\[
\mu = \int_{[0,1]^d} \underbrace{\pi^{d/2} \cos(\lVert \bsPsi(\bsx) \rVert)}_{f(\bsx)} \, \D \bsx, \qquad 
\bsPsi(\bsx) := \sqrt{1/2} \,\bsPhi^{-1}(\bsx).
\]

The code below sets up an \texttt{Integrand} instance using QMCPy's \texttt{CustomFun} wrapper to tie a user-defined function $g$ into the QMCPy framework.  Then we evaluate the sample mean of $n=1000$ $f$ values obtained by sampling at transformed Halton points. Notice how a two-dimensional Halton generator is used to construct a Gaussian true measure, which is applied alongside the \texttt{my\_Keister} function to instantiate a customized, QMCPy-compatible integrand for this problem.
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/i_keister.txt}
We have no indication yet of how accurate our approximation is.  That topic is treated in the next section.  Figure \ref{fig:ikc} visualizes sampling on the original integrand, $g$, and sampling on the transformed integrand, $f$.

\begin{figure}[t]
	\includegraphics[height=5cm]{QMCSoftwareArticle/vfigs/i_keister_contours.eps}
	\caption{Right: Sampling the transformed Keister integrand $f$ at Halton points $\bsX_i \LDsim \calU[0,1]^2$. Left: Sampling the original Keister integrand $g$ at $\bsT_i = \bsPsi(\bsX_i) \Msim \calN(\bszero,\mI/2)$ where $\bsPsi$ is defined in \eqref{eq:exampleVarTransGauss}.  } \label{fig:ikc}
\end{figure}

Another way to approximate the Keister integral in \eqref{eq:KeisterIntegral} is to write it as an integral with respect to the Lebesgue measure:
\begin{align*} 
	\mu 
	& = \int_{\bbR^d} \underbrace{\cos(\lVert \bst \rVert) \exp(-\bst^T \bst) }_{g(\bst)} \, \underbrace{1}_{\lambda(\bst)} \,\D \bst \\
	& = \int_{[0,1]^d} \underbrace{\cos(\lVert \bsPsi(\bsx) \rVert) \exp(-\bsPsi^T\!\!(\bsx) \bsPsi(\bsx)) \QMCPYabs{\bsPsi'(\bsx)}  }_{f(\bsx)} \, \D \bsx,
\end{align*}
where $\bsPsi$ is any transformation from $[0,1]^d$ to $\bbR^d$. Now $\lambda$ is \emph{not} a PDF.  QMCPy can perform the cubature this way as well.
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/i_keisterLebesgue.txt}
The $\bsPsi$ chosen when transforming uniform sequences on the unit cube to fill $\bbR^d$ is  given by \eqref{eq:exampleVarTransGauss} with $\mA = \mI$.

In the examples above, one must input the correct $g$ into \texttt{CustomFun} along with the correct \texttt{TrueMeasure} $\lambda$ to define the integration problem. The \texttt{Keister} integrand included in the QMCPy library takes a more flexible approach to defining the integration problem $\mu$ in \eqref{eq:KeisterIntegral}. Selecting a different \texttt{sampler} $\bsPsi$ performs  \emph{importance sampling}, which leaves $\mu$ unchanged.  
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/i_keisterBuiltIn.txt}

In the first case above, the $\lambda$ in \eqref{eq:transVar} corresponds to the Gaussian density  with mean zero and variance $1/2$ by default, and the corresponding variable transformation, $\bsPsi$, is chosen to make $\lambda\bigl(\bsPsi(\bsx)\bigr) \QMCPYabs{\bsPsi'(\bsx)}  = 1$ and $f(\bsx) = g\left(\bsPsi(\bsx) \right)$.  In the second case,  we choose an importance sampling density $\lambda_{\IMP}$, corresponding  to standard  Gaussian, and the variable transformation $\bsPsi_{\IMP}$ makes $\lambda_\IMP\bigl(\bsPsi_{\IMP}(\bsx)\bigr) \QMCPYabs{\bsPsi'_{\IMP}(\bsx)}  = 1$.  Then
\begin{align}
	\nonumber 
	\mu &= \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  = \int_\calT g(\bst) \, \frac{\lambda(\bst)}{\lambda_\IMP(\bst) } \, \lambda_\IMP(\bst) \D \bst  \\ 
	\nonumber
	& =  \int_{\cube} g\bigl(\bsPsi_\IMP(\bsx)\bigr) \, \frac{\lambda\bigl(\bsPsi_\IMP(\bsx)\bigr)} {\lambda_\IMP\bigl(\bsPsi_\IMP(\bsx)\bigr) } \lambda_\IMP\bigl(\bsPsi_\IMP(\bsx)\bigr) \,\QMCPYabs{\bsPsi'_{\IMP}(\bsx)}  \, \D \bsx \\
	\nonumber
	& =  \int_{\cube} f_\IMP(\bsx) \, \D \bsx  \\
	\label{eq:transVarImp}
	& \qquad \qquad \text{where } f_\IMP(\bsx)  = g\bigl(\bsPsi_\IMP(\bsx)\bigr)  \,  \frac{\lambda\bigl(\bsPsi_\IMP(\bsx)\bigr)} {\lambda_\IMP\bigl(\bsPsi_\IMP(\bsx)\bigr) }.
\end{align}
Because LD samples mimic $\calU\cube$, choosing a different \texttt{sampler} is equivalent to choosing a different variable transform.

\section{Stopping Criteria} \label{sec:stopping_crit}

The \texttt{StoppingCriterion} object determines the number of samples $n$ that are required for the sample mean approximation $\hmu$ to be within error tolerance $\varepsilon$ of the true mean $\mu$.  Several QMC stopping criteria have been implemented in QMCPy, including replications, stopping criteria that track the decay of the Fourier complex exponential or Walsh coefficients of the integrand \cite{HicJim16a,HicEtal17a,JimHic16a}, and stopping criteria based on
Bayesian credible intervals \cite{RatHic19a,JagHic22a}.

The \texttt{CubQMCSobolG} stopping criterion used in the example below assumes the Walsh-Fourier coefficients of the integrand are absolutely convergent. The algorithm iteratively doubles the number of samples used in the integration and estimates the error using the decay of the Walsh coefficients \cite{HicJim16a}. When the estimated error is below the user-specified tolerance, it finishes the computation and returns the estimated integral.

Let us return to the Keister example from the previous section.  After setting up  a default \texttt{Keister} instance via a Sobol' \texttt{DiscreteDistribution}, we choose a \texttt{StoppingCriterion} object that matches the \texttt{DiscreteDistribution} and input our desired tolerance.  Calling the  \texttt{integrate} method returns the approximate integral plus some useful information about the computation.
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/sc_keister_qmc.txt}
The second output of the stopping criterion provides helpful diagnostic information.  This computation requires $n=2^{13}$ Sobol' points and $0.009$ seconds to complete.  The error bound is $0.000637$, which falls below the absolute tolerance.

QMC, which uses LD sequences, is touted as providing substantially greater computational efficiency compared to IID MC.
Figure \ref{fig:sc_comp} compares the time and sample sizes needed to compute the $5$-dimensional Keister integral \eqref{eq:KeisterIntegral} using IID sequences and LD lattice sequences. Consistent with what is stated in Section \ref{sec:intro}, the error of IID MC is $\calO(n^{-1/2})$, which means that the time and sample size to obtain an absolute error tolerance of $\varepsilon$ is $\calO(\varepsilon^{-2})$.  By contrast, the  error of QMC using LD sequences is $\calO(n^{-1+\epsilon})$, which implies $\calO(\varepsilon^{-1-\epsilon})$ times and sample sizes.  We see that QMC methods often require orders of magnitude fewer samples than MC methods to achieve the same error tolerance.

\begin{figure}[t]
	\includegraphics[height=6cm]{QMCSoftwareArticle/vfigs/sc_comp.eps}
	\caption{Comparison of run times and sample sizes for computing the $5$-dimensional Keister integral \eqref{eq:KeisterIntegral} using IID and LD lattice sequences for a variety of absolute error tolerances. The respective stopping criteria are  \texttt{qp.CubMCG} \cite{HicEtal14a} and  \texttt{qp.CubQMCLatticeG} \cite{JimHic16a}. The LD sequences provide the desired answer much more efficiently.}
	\label{fig:sc_comp}
\end{figure}

For another illustration of QMC cubature, we turn to pricing an Asian arithmetic mean call option. The (continuous-time) Asian option is defined in terms of the average of the stock price, which is written in terms of an integral. The payoff of this option is the positive difference between the strike price, $K$, averaged over the time horizon: 
$$
\text{payoff}(\bsS) = \max\left(\frac{1}{2d}\sum_{j=1}^d (S_{j-1}+S_j)-K, \; 0\right), \quad \bsS = (S_0, \ldots, S_d).
$$
Here $S_j$ denotes the asset price at time $\tau j/d$, and a trapezoidal rule is used for discrete approximation of the integral in time that defines the average. 
The trapezoidal rule is a more accurate approximation to the integral than a rectangle rule. % Comment from Fred
A basic model for asset prices is a geometric Brownian motion, 
\[
S_j(\bsT) = S_0 \exp((r - \sigma^2/2) \tau j/d + \sigma T_j),   \;  j = 1, \ldots, d, \; \bsT = (T_1, \ldots, T_d)\sim \calN(\bszero,\mSigma),
\]
where $\mSigma$ is defined in \eqref{eq:BMcov}, $r$ is the interest rate, $\sigma$ is the volatility, and $S_0$ is the initial asset price.  The fair price of the option is then the expected value of the discounted payoff, namely,
\begin{equation*}
	\text{price} = \mu = \bbE[g(\bsT)], \quad \text{where } g(\bst) = \text{payoff}\bigl(\bsS(\bst) \bigr) \exp(-r \tau).
\end{equation*}

The following code utilizes QMCPy's Asian option \texttt{Integrand} object to approximate the value of an Asian call option for a particular choice of the parameters.
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/sc_aco.txt}
Because this \texttt{Integrand} object has the built-in Brownian motion \texttt{TrueMeasure}, one only need provide the LD sampler.

Out of the money option price calculations can be sped up by adding an upward drift to the Brownian motion.  The upward drift produces more in the money paths and also reduces the variation or variance of the final integrand, $f$.  This is a form of importance sampling.  Using a Brownian motion without drift we get
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/wo_imp_samp_aco.txt}
Adding the upward drift gives us the answer faster:
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/w_imp_samp_aco.txt}
The choice of a good drift is an art.

The improvement in time is less than that in $n$ because the integrand is more expensive to compute when the drift is employed.  Referring to \eqref{eq:transVarImp}, in the case of no drift, the $\lambda$ corresponds to the density for the discrete Brownian motion, and the variable transformation $\bsPsi$ is chosen so that $f(\bsx) = g\left(\bsPsi(\bsx) \right)$.  However, in the case of a drift, the integrand becomes  $f_\IMP(\bsx)  = g\bigl(\bsPsi_\IMP(\bsx)\bigr)  \lambda\bigl(\bsPsi_\IMP(\bsx)\bigr)/\lambda_\IMP\bigl(\bsPsi_\IMP(\bsx)\bigr)$, which requires more computation time per integrand value.

\section{Under the Hood}

In this section, we look at the inner workings of QMCPy and point out features we hope will benefit the community of QMC researchers and practitioners. We also highlight important nuances of QMC methods and how QMCPy addresses these challenges. For details, readers should refer to the QMCPy documentation \cite{QMCPyDocs}.

\subsection{LD Sequences}

LD sequences are the backbone of QMC methods. QMCPy provides generators that combine research from across the QMC community to enable advanced features and customization options.

Two popular LD sequences are integration lattices and digital nets which we previously outlined in Table \ref{tab:GroupProp}. These LD generators are comprised of two parts: the static generating vectors $\bsZ_1,\bsZ_2,\bsZ_4, \ldots \in [0,1)^d$ and the callable generator function. By default, QMCPy provides a number of high-quality generating vectors for users to choose from. For instance, the default ordinary lattice vector was constructed by Cools, Kuo, and Nuyens \cite{doi:10.1137/06065074X} using component-by-component search, is extensible, has order-2 weights, and supports up to 3600 dimensions and $2^{20}$ samples. 
%https://web.maths.unsw.edu.au/~fkuo/lattice/index.html
%https://web.maths.unsw.edu.au/~fkuo/sobol/
However, users who require more samples but fewer dimensions may switch to a generating vector constructed using LatNet Builder \cite{LatNet,LEcEtal22a} to support 750 dimensions and $2^{24}$ samples. Moreover, the \texttt{qp.Lattice} and \texttt{qp.DigitalNet} objects allow users to input their own generating vectors to produce highly customized sequences. To find such vectors, we recommend using LatNet Builder's construction routines as the results can be easily parsed into a QMCPy-compatible format.

Along with the selection of a generating vector, QMCPy's low discrepancy sequence routines expose a number of other customization parameters. For instance, the lattice generator extends the Magic Point Shop \cite{Nuy17a} to support either linear or natural ordering. Digital sequences permit either standard or Gray code ordering and may be randomized via a digital shift optionally combined with a linear scrambling \cite{Mat98}. Halton sequences may be randomized via the routines of either Owen \cite{Owe20a} or Hofert and Lemieux \cite{QRNG2020}.

\subsection{A Word of Caution When Using LD Sequences}

Although QMCPy's \texttt{DiscreteDistribution}s have many of the same parameters and methods, users should be careful when swapping IID sequences with LD sequences. While IID node-sets have no preferred sample size, LD sequences often require special sampling ranges to ensure optimal discrepancy. As mentioned  earlier, base-2 digital nets and extensible integration lattices show better evenness for sample sizes that are powers of 2. On the other hand, the preferred sample sizes for $d$-dimensional Halton sequences are $n = \prod_{j=1}^d p^{m_j}_j$ where $p_j$ is the $j^{\text{th}}$ prime number and $m_j \in \mathbb{N}_0$ for $j=1,\dots,d$. Due to the infrequency of such values, Halton sequences are often regarded as not having a preferred sample size.

Users may also run into trouble when trying to generate too many points. Since QMCPy's generators construct sequences in 32-bit precision, generating greater than $2^{32}$ consecutive samples will cause the sequence to repeat. In the future, we plan to expand our generators to support optional 64-bit precision at the cost of greater computational overhead.

Another subtlety arises when transforming LD sequences to mimic different distributions. As mentioned earlier, unrandomized lattice and digital sequences include the origin, making transformations such as \eqref{eq:exampleVarTransGauss} produce infinite values.

Some popular implementations of LD sequences drop the first point, which is the origin in the absence of randomization.  The rationale is to avoid the transformation of the origin to infinity when mimicking a Gaussian or other distribution with an infinite sample space.  Unfortunately, dropping the first point destroys some of the nice properties of the first $n = 2^m$ points of LD sequences, which can degrade the order of convergence for QMC cubature. A careful  discussion of this matter is given by~\cite{owen2020dropping}.

\subsection{Transformations}

The transformation $\bsPsi$  connects a \texttt{DiscreteDistribution} and \texttt{TrueMeasure}. So far, we have assumed the \texttt{DiscreteDistribution} mimics a $\calU[0,1]^d$ distribution with PDF $\varrho(\bsx)=1$. However, it may be advantageous to utilize a \texttt{DiscreteDistribution} that mimics a different distribution. 
%For instance, QMCPy includes a \texttt{DiscreteDistribution} corresponding to an IID standard Gaussian sequence.  It may be transformed to mimic a general Gaussian distribution \texttt{TrueMeasure} while avoiding the inverse CDF transform required for nodes mimicking $\calU[0,1]^d$.
% the IID standard Gaussian samples still use inverse CDF, so we don't actually save anything aside from maybe a minor speed up from numpy's faster code

Suppose we have a \texttt{DiscreteDistribution} mimicking  density $\varrho$ supported on $\calX$.  Then using the variable transformation $\bsPsi$, 
\begin{equation*}
	\mu = \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  =  \int_{\calX}f(\bsx) \, \varrho(\bsx) \D \bsx \quad \text{for }
   f(\bsx)  = g\bigl(\bsPsi(\bsx)\bigr)  \, \frac{\lambda\bigl(\bsPsi(\bsx)\bigr)}{\varrho(\bsx)} \, \lvert \bsPsi'(\bsx)\rvert,
\end{equation*}
which generalizes \eqref{eq:transVar}.
QMCPy also includes support for successive changes of measures so users may build complex variable transformations in an intuitive manner. Suppose that the variable transformation is a composition of several transformations: $\bsPsi=\widehat{\bsPsi}_L = \bsPsi_L \circ \bsPsi_{L-1} \circ \dots \circ \bsPsi_1$ as in \eqref{eq:transVarImp}. Here, $\bsPsi_l:\calX_{l-1} \to \calX_l$,  $\calX_0=\calX$, and $\calX_L=\calT$ so that the transformations are compatible with the \texttt{DiscreteDistribution} and \texttt{TrueMeasure}. Let $\widehat{\bsPsi}_l=\bsPsi_l \circ \bsPsi_{l-1} \circ \dots \circ \bsPsi_1$ denote the composition of the first $l$ transforms and assume that $\widehat{\bsPsi}_0(\bsx)=\bsx$, the identity transform. Then we may write $\mu =  \int_{\calX}f(\bsx) \, \varrho(\bsx) \D \bsx$ for 
\begin{equation*}
    f(\bsx) 
 = g\bigl(\widehat{\bsPsi}_L(\bsx)\bigr)\frac{\lambda\bigl(\widehat{\bsPsi}_L(\bsx)\bigr)}{\varrho(\bsx)}\prod_{l=1}^L \bigl\lvert \bsPsi_l'\bigl(\widehat{\bsPsi}_{l-1}(\bsx)\bigr) \bigr\rvert .
\end{equation*}
It is often the case that $\bsPsi_l$ is chosen such that $\bsPsi_l(\bsX)$ is stochastically equivalent to a random variable with density $\lambda_l$ on sample space $\calX_l$ when $\bsX$ is a random variable with density $\varrho_l$ on sample space $\calX_{l-1}$.  This implies  $\varrho_l(\bsx) = \lambda_l(\bsPsi_l(\bsx))\lvert  \bsPsi'_l (\bsx) \rvert$ so that
\begin{equation*}
f(\bsx) 
 = g\bigl(\widehat{\bsPsi}_L(\bsx)\bigr)\frac{\lambda\bigl(\widehat{\bsPsi}_L(\bsx)\bigr)}{\varrho(\bsx)}\prod_{l=1}^L \frac{\varrho_l(\widehat{\bsPsi}_{l-1}(\bsx))}{\lambda_l(\widehat{\bsPsi}_l(\bsx))} .
\end{equation*}

For an example, we return to the Keister integral \eqref{eq:KeisterIntegral}. The following code constructs three \texttt{Keister} instances: one without importance sampling, one importance sampled by a Gaussian distribution, and one importance sampled by the composition of a Gaussian distribution with a Kumaraswamy distribution \cite{Kumaraswamy}. All \texttt{Integrand}s use a Sobol' \texttt{DiscreteDistribution}, making $\varrho(\bsx)=1$ and $\calX = [0,1]^d$. The \texttt{TrueMeasure} is $\calN(\bszero,\mI/2)$ making $\lambda(\bst) = \pi^{-d/2} \exp(-\bst^T \bst)$ and $\calT = \bbR^d$.

The table below displays the variable transformations and the measures for these three cases.  In all cases $\varrho_1(\bsx) = \cdots = \varrho_L(\bsx) = 1$ because the $\bsPsi_l$ utilize inverse cumulative distributions.

\[\begin{array}{ccccccc}
	\texttt{Integrand} & L & \lambda_1 & \bsPsi_1 & \lambda_2 & \bsPsi_2 & f \tabularnewline
	\toprule
	\texttt{K} & 1 & \calN(\bszero, \mI/2) & \eqref{eq:exampleVarTransGauss}& && g(\bsPsi_1(\cdot)) \tabularnewline [1ex]
	\texttt{K\_gauss} & 1 & \calN(\bszero,3 \mI/4) & \eqref{eq:exampleVarTransGauss} & && 
	\displaystyle g(\bsPsi_1(\cdot)) \frac{\lambda(\bsPsi_1(\cdot))}{\lambda_1(\bsPsi_1(\cdot))} \tabularnewline [2.5ex]
	\texttt{K\_gauss\_kuma} & 2 & \Kum & \bsF^{-1}_\Kum & \calN(\bszero,\mI) &  \eqref{eq:exampleVarTransGauss} &
	\displaystyle \frac{g(\bsPsi_2(\bsPsi_1(\cdot))) \lambda(\bsPsi_2(\bsPsi_1(\cdot))) }{\lambda_1(\bsPsi_1(\cdot)) \lambda_2(\bsPsi_2(\bsPsi_1(\cdot)))} 
   \\ \bottomrule
\end{array}
\]

Here $\Kum$ denotes the multivariate Kumaraswamy distribution with independent marginals, and $ \boldsymbol{F}^{-1}_\Kum$ denotes the element-wise inverse cumulative distribution function.  The following code evaluates the Keister integral \eqref{eq:KeisterIntegral} for $d=1$ and error tolerance $\varepsilon = 5 \times 10^{-8}$.  The timings for each of these different integrands are displayed.

\lstinputlisting[style=Python]{QMCSoftwareArticle/python/composed_IS.txt}

%Successful importance sampling places more points where the original integrand, $g$, varies more.  Equivalently, 
Successful importance sampling makes the transformed integrand, $f$, more flat. The shorter cubature times correspond to flatter integrands, as  illustrated in Figure \ref{fig:mIS}. The above example uses $d=1$ to facilitate the plot in Figure \ref{fig:mIS}; however, the same example works for arbitrary dimensions.
\begin{figure}[t]
    \centering
	\includegraphics[width=.8\textwidth]{QMCSoftwareArticle/vfigs/multiple_IS.eps}
	\caption{Keister functions with and without importance sampling (\texttt{IS}). Note that the Keister functions using importance sampling are generally less variable and therefore easier to integrate, as evidenced by the faster integration times.} \label{fig:mIS}
\end{figure}

Automatic Bayesian stopping criteria choose $n$ needed to satisfy the error tolerance via a Bayesian-credible interval for $\mu$ \cite{RatHic19a,JagHic22a}.  These are also options in QMCPy.

\section{Further Work} \label{sec:further}

QMCPy is ripe for growth and development in several areas.  We hope that the QMC community will join us in making this a reality.

Multi-level (quasi-)Monte Carlo (ML(Q)MC) methods make possible the computation of expectations of functionals of stochastic differential equations and partial differential equations with random coefficients.  Such use cases appear in quantitative finance and geophysical applications.  QMCPy's ML(Q)MC's capability is rudimentary, but under active development.

We hope to add a greater variety of use cases and are engaging collaborators to help.  Sobol' indices, partial differential equations with random coefficients, expected improvement measures for Bayesian optimization, and multivariate probabilities are some of those on our radar.

Recently, several QMC experts have focused on developing LD generators for Python. 
Well-established packages such as SciPy \cite{SCIPY} and PyTorch \cite{PyTorch} are have developed QMC modules that support numerous LD sequences and related functionalities. We plan to integrate the routines as optional backends for QMCPy's LD generators. Creating ties to these other packages will allow users to call their preferred generators from within the QMCPy framework.  Moreover, as features in QMCPy become more common and prove their value, we will try to incorporate them into SciPy and other popular, general-purpose packages.

We also plan to expand our library of digital net generating matrices. We wish to incorporate interlaced digital nets, polynomial lattices, and Niederreiter sequences, among others. By including high-quality defaults in QMCPy, we hope to make these sequences more readily available to the public.

Our \texttt{DiscreteDistrution} places equal weights on each support point, $\bsX_i$.  In the future, we might generalize this to unequal weights.

QMCPy already includes importance sampling, but the choice of sampling distribution must be chosen a priori.  We would like to see an automatic, adaptive choice following the developments of \cite{AsmGly2007a,approx_zero_variance_simulation,OweZho00a}.

Control variates can be useful for QMC as well as for IID MC \cite{HicEtal03}.  These should be incorporated into QMCPy in a seamless way.

We close with an invitation.  Try QMCPy.  If you find bugs or missing features, please submit an issue to \href{https://github.com/QMCSoftware/QMCSoftware/issues}{\nolinkurl{https://github.com/QMCSoftware/QMCSoftware/issues}}.  If you wish to add your great algorithm or use case, please submit a pull request to our GitHub repository at \href{https://github.com/QMCSoftware/QMCSoftware/pulls}{\nolinkurl{https://github.com/QMCSoftware/QMCSoftware/pulls}}.  We hope that the community will embrace QMCPy.

\begin{acknowledgement}
The authors would like to thank the organizers for a wonderful MCQMC 2020. 
We also thank the referees for their many helpful suggestions.  This work is supported in part by SigOpt and National Science Foundation grant DMS-1522687.
\end{acknowledgement}

%\section*{References}
%\nocite{*}
\bibliographystyle{spmpsci.bst}
\bibliography{FJH23,FJHown23,QMCSoftwareArticle,sc}

\iffalse
\AGSComment{

\section{Extra Derivations}

\subsection{Importance sampling with Gaussian}

$$ \bsPsi(\bsx) = A\Phi^{-1}(\bsx) + \mu \quad \text{where} \quad \Sigma = AA^T$$
$$ \left\lvert \frac{\D \bsPsi}{\D \bsx} \right\rvert = \det(J) \quad \text{for} \quad J_{ij} = \frac{\D \Psi_i}{\D x_j} = \frac{A_{ij}}{\phi(\Phi^{-1}(x_j))}$$
$$ \det(J) = \det\left(A . \text{diag}\left(\frac{1}{\phi(\Phi^{-1}(\bsx))}\right)\right) = \frac{\det(A)}{\prod_{j=1}^d \phi(\Phi^{-1}(x_j))}$$}

\FJHComment{Given $g$ and $\lambda$ to define your integral, suppose that $\lambda(\bst)$ is some multiple of some multivariate normal distribution:
	\begin{equation*}
		\lambda(\bst)  = \frac{c \exp\bigl(-(\bst -\bsb)^T \mLambda^{-1}(\bst -\bsb)/2\bigr)}{\sqrt{ (2\pi)^d \det(\mLambda)}} 
	\end{equation*}
Suppose that we use the variable transformation $\bsPsi$ defined in \eqref{eq:exampleVarTransGauss}.  By what you have derived above:
\begin{align*}
	\QMCPYabs{\frac{\partial \bsPsi}{\partial \bsx}} 
	& = \frac{\det(\mA^T)}{\prod_{j=1}^d \phi(\Phi^{-1}(x_j))} = \sqrt{(2\pi)^d \det(\mSigma)} \exp\bigl( \QMCPYnormnorm[2]{\bsPhi^{-1}(\bsx)}^2/2\bigr) \\
	\MoveEqLeft[2]{\lambda\bigl(\bsPsi(\bsx)\bigr) \QMCPYabs{\frac{\partial \bsPsi}{\partial \bsx}}} \\
	& =  c \sqrt{\frac{\det(\mSigma)}{\det(\mLambda)}} \exp \bigl( [ \bsPhi^{-1}(\bsx)^T \bsPhi^{-1}(\bsx) \\
	& \qquad \qquad -   (\mA\bsPhi^{-1}(\bsx) +\bsa -\bsb)^T\mLambda^{-1}(\mA\bsPhi^{-1}(\bsx) +\bsa -\bsb)        ] /2   \bigr) \\
	& =  c \sqrt{\frac{\det(\mSigma)}{\det(\mLambda)}} \exp \bigl( [ \bsPhi^{-1}(\bsx)^T \bsPhi^{-1}(\bsx) 
	-   \bsPhi^{-1}(\bsx)^T\mA^T\mLambda^{-1}\mA \bsPhi^{-1}(\bsx)\\
	& \qquad \qquad  - 2(\bsa -\bsb)^T \mLambda^{-1}\mA\bsPhi^{-1}(\bsx) - (\bsa -\bsb)^T \mLambda^{-1} (\bsa -\bsb)] /2   \bigr) 
\end{align*}
If $\mLambda = \mSigma$, then this simplifies to
\begin{equation*}
		\lambda\bigl(\bsPsi(\bsx)\bigr) \QMCPYabs{\frac{\partial \bsPsi}{\partial \bsx}}
	 =  c \exp \bigl( - (\bsa -\bsb)^T \mA^{-T} [
	2 \bsPhi^{-1}(\bsx) + \mA^{-1}(\bsa -\bsb)] /2   \bigr) .
\end{equation*}
If $\mLambda = \lambda^2 \mI$ and  $\mA = \sigma \mI$, then 
\begin{align*}
	\lambda\bigl(\bsPsi(\bsx)\bigr) \QMCPYabs{\frac{\partial \bsPsi}{\partial \bsx}}
	& = \frac{c\sigma}{\lambda} 
	\exp \bigl( [ (\lambda^2 -\sigma^2)\bsPhi^{-1}(\bsx)^T\bsPhi^{-1}(\bsx) 
	\\
	& -  (\bsa -\bsb)^T [
	2\sigma \bsPhi^{-1}(\bsx) + (\bsa -\bsb)] /(2\lambda^2)   \bigr) .
\end{align*}
}

\AGSComment{

\subsection{Chaining Transforms}
With $k$ transforms, let $\widehat{\bsPsi}_j(\bsx) = (\bsPsi_j \circ \bsPsi_{j-1} \circ \dots \circ \bsPsi_1)(\bsx)$ denote the first $j$ compositions and $\bst = \widehat{\bsPsi}_k(\bsx)$ be the complete transformation. Then, 
$$f(\bsx) = g(\widehat{\bsPsi}_k(\bsx)) \frac{\lambda(\widehat{\bsPsi}_k(\bsx))}{\varrho(\bsx)} \prod_{i=1}^{k} \left\lvert \frac{\D \, \bsPsi_i}{\D \, \widehat{\bsPsi}_{i-1}(\bsx)} \right\rvert$$
where $\widehat{\bsPsi}_0(\bsx) = \bsx$.
}

\FJHComment{

\subsection{Software Framework}

\newcommand{\dd}{\texttt{DiscreteDistribution}\xspace}
\newcommand{\tm}{\texttt{TrueMeasure}\xspace}

\begin{itemize}
\item \dd $\rightarrow$ \tm gives an instance that can generate a sequence that mimics the measure specified by \tm if \tm is a probability measure; it also has a \texttt{gen\_samples} method.

\item if \tm is not a probability measure, then \dd $\rightarrow$ \tm sets up a transformation from $\calX$ to $\calT$.

\item \texttt{CustomFun} takes a user-defined function $g$ and a \tm instance and uses the $\lambda$ inferred from the \tm to form $f$.

\item Built-in \texttt{Integrand} classes know what $g$ and $f$ are.

\begin{itemize}
    \item One way is to have these \texttt{Integrand} classes assume a \tm and only need a \dd input. If you do not like the default \tm, then you need to input your own.

    \item Another way is to have these \texttt{Integrand} classes take a \tm input and use that, plus their knowledge of $g$ and $\lambda$ to find $f$, perhaps by what could be considered importance sampling.

\end{itemize}
\end{itemize}
}

\AGSComment{
\subsection{\texttt{print}ing  QMCPy Instances}

QMCPy provides the a \texttt{print} command to display information about instances of QMCPy classes. In the sample code in Section \ref{sec:stopping_crit} this command was used show the output of a \texttt{StoppingCriteroion}'s \texttt{integrate} method.

faulted for any given component. This feature was already used to print the data object returned by the stopping criterion's \texttt{integrate} method in Section \ref{sec:stopping_crit}. The following code prints a drifted \texttt{BrownianMotion} instance and displays the connection to an underlying Gaussian measure. 
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/print_bm.txt}}

\AGSComment{Do we want the implementation to follow the first or second equation? QMCPy current follows first equation (directly computes Jacobians), which may allow for transforms that do not correspond to PDFs. However, the second equation would be easier for QMCPy contributors who then do not need to compute Jacobians for their transforms. May also be computationally cheaper i.e. Gaussian Jacobian requires PDF and inverse CDF computations, where the second equation would save the inverse CDF computation.}

\fi

\end{document}

### 2022-06-28 MCQMC 2020 stuff/referee_reports/author_followup_report1.tex
Update from Office Computer

\documentclass{amsart}
\usepackage{mathtools,upref,siunitx,upquote,fancyvrb,color}
\usepackage[hyphens]{url}
\usepackage[utf8]{inputenc}

\textwidth 6.5in
\hoffset -0.7in
\allowdisplaybreaks %allow equations to span multiple pages

\input{../FJHDef}

\newcommand{\scnote}[1]{ {\textcolor{blue}  {\mbox{\bf Response:} #1}}}

\begin{document}
\title{QMCPy Authors' Feedback for Referee Report 1}
%\author{}

\maketitle

Report on the paper "Quasi-Monte Carlo Software", by Sou-Cheng T. Choi, Fred J. Hickernell, R. Jagadeeswaran, Michael J. McCourt, Aleksei G. Sorokin

This work is devoted to discuss quasi-MonteCarlo (QMC) software available for practitioners and researchers. The main focus is the QMCPy Python project, developed by the authors of the paper.

The introduction of the paper discusses the general framework of QMC methods, their main properties, and notation. Section 2 lists a number of QMC software available (in some cases in commercial packages), with a very short reference to their contents. Section 3 conceptualizes four main components of QMC software: discrete distribution; true measure; integrand; stopping criterion. The following sections (4 thru 7) discuss in more detail how these components are included/represented/used in the QMCPy library. Section 8 go into further depth for some additional features of the library. Section 9 presents conclusions and future work.

The paper seems devoid of obvious mistakes or errors. The  availability of QMC libraries is important for more widespread use of the existing methods, as well as for trying out new ideas.

Nevertheless, I feel that Section 2 of the paper is very short and gives almost no information on the existing software, the strengths and weaknesses, etc.  Also, while conceptualizing the main four components of a QMC library can be useful, the authors do not map these components for existing software, except their own QMCPy library. They state that ``The sections that follow describe QMCPy [3], which is our attempt to combine some of the best of the above software under a common user interface written in Python 3.''; but they do not explain which are the features of ``the best of the above software'' that they have included in their library.

In this sense, the paper is very focused on this particularly library, and the title, and the abstract do not quite represent the actual contents.

My recommendation is a minor revision, so that the authors can expand Section 2, or let more clear that the paper will be focused specifically on their library.

%\bibliographystyle{amsplain}
%\bibliography{FJH23,FJHown23}

\vspace{10ex}

\scnote{ 
We thank the reviewer for the valid and helpful feedback especially for Section 2 of the manuscript. We have expanded the section from half a page to almost two pages with the following changes: 
\begin{enumerate}
 \item added a few packages (SciPy, TF Quant Finance, and MultilevelEstimators.jl) that have come to our attention recently; 
 \item for each software, 
    \begin{enumerate}
        \item described in more detail some of its best features;
        \item specified if some or all of the best features are accessible and, if applicable, enhanced through QMCPy by various mechanisms (e.g., direct calls, wrapper functions, code compilation, or direct translation).
    \end{enumerate}
\end{enumerate}
}

\end{document}

### 2022-06-28 MCQMC 2020 stuff/referee_reports/author_followup_report2.tex
Update from Office Computer

\documentclass{amsart}
\usepackage{mathtools,upref,siunitx,upquote,fancyvrb,color}
\usepackage[hyphens]{url}
\usepackage[utf8]{inputenc}

\textwidth 6.5in
\hoffset -0.7in
\allowdisplaybreaks %allow equations to span multiple pages

\input{FJHDef}

\begin{document}
\title{QMCPy Authors' Feedback for Referee Report 2}
%\author{}

\newcommand{\AGSComment}[1]{{\color{blue} #1}}
\newcommand{\FJHComment}[1]{{\color{blue} #1}}

\maketitle

\section{Overall Evaluation}

This paper proposes and discusses a software tool named QMCPy, in the form of a Python library, for estimating integrals via quasi-Monte Carlo (QMC). The library offers different types of QMC point sets and randomizations. The point sets are extensible in the number of points, using successive powers of 2. The main function takes an extensible point set, an integrand, and a target maximal error (or variance), and computes an RQMC estimator by doubling the number of points successively until the error bound is below the target. Several examples are given, some with a change of measure and/or a change of variables, and the corresponding Python code.

This is certainly a very useful tool, designed by knowledgeable people, which is likely to grow and spread. The paper is reasonably well written and should have its place in the MCQMC 2020 book. However, there are several pieces that would need clarification, as well as some additional explanations to make the paper more self-contained. I understand that one can also study the tutorial and the Colab notebooks to understand how things work, but it would be good if the paper can also be understood easily just by itself.

\bigskip

\FJHComment{We thank the referee for his or her careful reading of our manuscript and many helpful comments, including pointers to some of the literature that we were unfamiliar with.  We have tried to address them all.}

\section{Details}

Page 3: ``The measure of the difference...'' should be ``A measure of the difference...'' (not
unique).

\AGSComment{Fixed.}
\vspace{1cm}

Page 4, Eq.(5): Replace the comma by a period.

\AGSComment{Fixed.}
\vspace{1cm}

Page 4, ``highly stratified sampling'': A key difference between RQMC and stratification is that in the latter, once the partition is defined, the positions of the points in the different pieces are conditionally independent. Because of that, the variance can never increase compared with Monte Carlo, which is not true for RQMC. See Section 2 of [2], for example. This independence property can make it easier to prove convergence rates for the variance; see
[3], Proposition 6, for an example of this.

\FJHComment{Thanks to the referee for pointing this out.  We have modified the wording to say that LD sampling is similar to stratified sampling without going into the details.}

\vspace{1cm}

Page 5, SSJ: You may add “in the hups package”.

\AGSComment{Added.}
\vspace{1cm}

The authors call an RQMC point set a discrete distribution, which I find a bit confusing. The
RQMC points do not really follow a discrete distribution and there is no sampling from a discrete distribution. After they are randomized, conditional on their values, the n points form a deterministic set and one can think of a discrete uniform distribution over these n points, but one never generate points from this discrete distribution. A DiscreteDistribution object has a gen samples method that returns all the n points of the RQMC point set in a
large two-dimensional array. It is probably too late to change the name, but a comment on this could be added.

\FJHComment{The intention in the name is that the finite points have an empirical distribution that has discrete support.  We might in the future add the possibility of unequal weights although we have not done so yet.  A comment added in the first paragraph of Section 4.}

\vspace{1cm}

Page 6: The $\oplus$ sign is commonly used to denote the exclusive-or operation, so its use for
something more general is a bit confusing.

\FJHComment{An explanation has been added in the footnote.}

\vspace{1cm}

``lattices and digital sequences prefer $n$ to be a power of 2'' should be better qualified. This is true for digital nets in base 2. For digital nets in prime base $b > 2$, we would prefer a power of $b$. For lattice rules, a prime $n$ is sometimes better than a power of 2; see Figure 6 of [4], for example. I think the real reason for using a power of 2 for lattices in this paper is to have a sequence of embedded lattices. This should be clarified.

\FJHComment{Explanation and clarification added early in Section 4.  You are correct that we want embedded lattices or nets.}

\vspace{1cm}

Bottom of table 1: the 	$\ominus$ operator is not defined. What is it?

\FJHComment{The definition is given in the text where we discuss Table 1.}

\vspace{1cm}

From what we see in Section 4, it seems that the user cannot select the point sets (direction numbers, generating matrices, generating vectors for the lattices). A comment on this would be good.

\AGSComment{We have added a discussion in Section 8, ``Under the Hood'',  regarding QMCPy's support for custom generating vectors for Lattice sequences and custom generating matrices for digital nets in base 2.}
\vspace{1cm}

In several places, a reader who is not already familiar with both Python and QMCPy (like me) may struggle to understand what goes on when reading the Python code. For example, on page 8, it would be good to recall quickly in the text what the parameters mean in \texttt{IIDStdUniform(2)}, \texttt{gen\_samples(1)}, and \texttt{gen\_samples(n min=1,n max=2)}.

\AGSComment{We have added some discussion/keywords to try and make input parameters more transparent.}
\vspace{1cm}

On page 12, it would also be good to explain the code a bit. What is \texttt{CustomFun} doing? What are \texttt{Halton(2)} and \texttt{Gaussian} doing? Do we have \texttt{d = 2} in this case?

\AGSComment{We have added some discussion/keywords to try and make input parameters more transparent.}
\vspace{1cm}

The different ways of rewriting the integral shown on pages 12 and 13 are only remotely connected to the software itself. I suppose the software can take any reformulation as an integral over the unit hypercube. Of course, the MC and RQMC variances will depend on the reformulation. We also miss some insight about the specific choices of reformulations given here. For example, the optimal change of measure for the importance sampling is the one that makes the integrand completely flat and gives a variance of zero [1]. This zero variance IS can often be well approximated [6]. Basically, we want $\lambda_\text{IS}$ to be approximately proportional to $g \cdot \lambda$. But the choice made in the paper seems arbitrary and is not really justified.

\FJHComment{You are correct that importance sampling can be difficult to do well.  It is an art.  QMCPy is not yet developed enough to do sophisticated importance sampling.  This is future work.}

\vspace{1cm}

The code on page 14: Say what \texttt{CubQMCSobolG} is doing exactly. How is the error or variance estimated? What randomization is used for the Sobol points?

\FJHComment{\texttt{CubQMCSobolG} implements the stopping criterion of Hickernell and Jim\'enez Rugama in reference [18] as now explained in the text.  A more detailed description of the stopping criterion would be too long.  The randomization is linear matrix scrambling, but it is immaterial.  We are not using replications.}

\vspace{1cm}

Page 16: Why use a trapezoidal rule? I think Asian option payoffs in practice are already defined by a finite arithmetic average, no?

\FJHComment{The continuous time Asian option is defined as an integral.  The trapezoidal rule gives a better approximation to the integral.}

\vspace{1cm}

``Integrand object to approximate the Asian call option''? You mean to ``approximate the value of ... ''?

\AGSComment{Yes. This has been reworded.}
\vspace{1cm}

Bottom of page 16: How did you select the upward drift? An optimal choice that would provide zero variance would have to be dynamic, as explained in [6] for the Markov chain
setting.

\FJHComment{Again, dynamic choice is beyond the scope of what we can do.}

\vspace{1cm}

Section 8.1: What criteria were used to construct these lattice sequences?

\AGSComment{We added a bit of discussion on how the default lattice generating vector was found and where it was sourced from.}
\vspace{1cm}

Section 8.2: “better evenness for ... powers of 2” is because of the specific construction of embedded lattices. It is not always true in general. Maybe this should be clarified.

\AGSComment{We added a bit of discussion throughout that tries to clarify and reinforce this.}
\vspace{1cm}

Section 8.3, first paragraph: How are the IID Gaussian points obtained if it is not by inversion?

\AGSComment{Good point, we are still using inversion for IID Standard Gaussian, so this does not avoid the computational overhead of the inverse normal.}
\vspace{1cm}

Second paragraph: Here you consider successive changes of measures and/or successive changes of variables. It is not clear why one would want to do that. In the end, the combined changes of variables corresponds to a single change of variable, no? We miss a motivation for this.

\AGSComment{We have built-in support for successive changes of variables to help users easily construct complex variable transformations without having to work out the math. This discussion has been added to the paper.}
\vspace{1cm}

What is the Kumaraswamy distribution?

\AGSComment{It is similar to the Beta distribution, but with closed-form CDF and quantile functions. We have added a reference for this.}
\vspace{1cm}

Page 20: ``Successful importance sampling places more points where the original integrand $g$ varies more'' is not quite correct. Again, see [1] and [6] for the optimal choice.

\AGSComment{We have replaced this sentence with ``Successful importance sampling makes the transformed integrand, $f$, more flat.''}
\vspace{1cm}

Page 22: Automatic adaptive choice of IS: How?

\AGSComment{Not quite sure yet. Perhaps we can do some distribution-hyperparameter optimization? We need to think through this more, but initial computer experiments have shown some promise.}

\vspace{1cm}

Reference 27 is not correct. There is the journal paper [5] on Lattice Builder. Then there is the LatNet Builder software at https://github.com/umontreal-simul/latnetbuilder, which has more contributors: L’Ecuyer, Munger, Marion, Godin, Puchhammer.

\FJHComment{The references to Lattice Builder and LatNet Builder have been updated.}

\vspace{1cm}

%\bibliographystyle{amsplain}
%\bibliography{FJH23,FJHown23}

\end{document}

### 2022-06-28 MCQMC_2022_ChallengesDevelopSoftware.tex
Update from Office Computer

--- 2022-06-28 MCQMC 2020 stuff/MCQMC_QMCSoftwareTut.tex
@@
-%Talk given virtually for SIAM UQ 2020 March
+%Talk given at MCQMC 2022 July
 \documentclass[11pt,compress,xcolor={usenames,dvipsnames},aspectratio=169]{beamer}
@@
 
+Challenges in Developing Great MCQMC Software
+Fred J. Hickernell, Illinois Institute of Technology, Chicago, IL USA
+
+The process of translating new Monte Carlo and Quasi-Monte Carlo (MCQMC) algorithms into software libraries faces several challenges.
+Great software should be easy to use with reasonable default options for the novice and advanced features for the developer or experienced user.
+The library architecture must allow for growth.
+Ensuring connectivity with other software libraries will facilitate a larger user base of the MCQMC library.
+Coding algorithms the right way may significantly improve their runtime or portability.
+Since development team members will come and go, the shared wisdom of the development team must be documented and transmitted to succeeding generations.
+Software developers must keep abreast of the newest computing environments to ensure peak performance.
+This talk will highlight some of these challenges and ways to address them.
+
 \fi
@@
 \setlength{\FJHThankYouMessageOffset}{-8ex}
-\title{Quasi-Monte Carlo Software}
+\title{Challenges in Developing Great MCQMC Software}
 \author[]{Fred J. Hickernell}
@@
 
-\thanksnote{with Sou-Cheng Choi, Mike McCourt, Jagadeeswaran Rathinavel, Aleksei Sorokin \\ 
-and the rest of the GAIL an QMCPy teams \\
-	partially supported by SigOpt and NSF-DMS-1522687 \\[2ex]
-	Thanks to the organizers, especially during these unusual times \\
-	Slides at  \href{https://speakerdeck.com/fjhickernell/quasi-monte-carlo-software}{\nolinkurl{speakerdeck.com/fjhickernell/quasi-monte-carlo-software}}\\
-	Google Colaboratory notebook at \href{https://tinyurl.com/QMCPyTutorial}{\nolinkurl{tinyurl.com/QMCPyTutorial}}\\
-	Blog at \href{https://qmcpy.wordpress.com/}{\nolinkurl{qmcpy.wordpress.com/}}\\
-	Recording of this talk at \href{https://media.ed.ac.uk/playlist/dedicated/51612401/1_0z0wec2z/1_2k12mwiw}{\nolinkurl{media.ed.ac.uk/playlist/dedicated/51612401/1_0z0wec2z/1_2k12mwiw}}
-
+\thanksnote{with Mark Klinchin and the rest of the GAIL an QMCPy teams \\
+	partially supported by SigOpt, an Intel company \\[2ex]
+	Thanks to the organizers as we meet again in person in still challenging times\\
+	Slides at  \href{https://speakerdeck.com/fjhickernell/quasi-monte-carlo-software}{\nolinkurl{speakerdeck.com/fjhickernell/quasi-monte-carlo-software}}
 }
-\event{MCQMC 2020}
-\date[]{August 10, 2020}
+\event{MCQMC 2022}
+\date[]{July 18, 2022}
 

### 2022-07-07 MCQMC_2022_ChallengesDevelopSoftware.tex
Updates from Overleaf
Office Computer Updates

@@
 \usepackage[usenames]{xcolor}
-\usepackage[giveninits=false,backend=biber,style=nature, maxcitenames =10, mincitenames=9]{biblatex}
+\usepackage[giveninits=false,backend=biber,style=nature,style=authoryear,maxcitenames =10, mincitenames=9]{biblatex}
 \addbibresource{FJHown23.bib}
@@
 \usepackage[tikz]{mdframed}
+
+\usepackage[T1]{fontenc}
+\usepackage{tgadventor} %Font found at https://tug.org/FontCatalogue/
+%\usepackage{newpxtext}
+\usepackage[euler-digits,euler-hat-accent]{eulervm}
 
@@
 \section{Background}
+
+\begin{frame}{My Definition of Great MCQMC Software}
+
+	\vspace{-5ex}
+	\begin{itemize}
+		\item \alert{Correct}, e.g., not omit the zeroth point in a low discrepancy sequence \parencite{Owe22a, scipySobol2020a}
+
+		\item
+
+		\item
+
+		\item
+
+		\item
+
+	\end{itemize}
+\end{frame}
+
+\printbibliography
+\end{document}
+
+\begin{frame}{Quasi-Monte Carlo (QMC) uses low discrepancy (LD) sequences}
+	\vspace{-3ex}
+	\begin{tabular}{>{\centering}p{0.47\textwidth}@{\quad}>{\centering}p{0.47\textwidth}}
+		%Independent \& Identically Distributed (IID) &
+		%Low Discrepancy (LD) \tabularnewline
+		\includegraphics[height=5cm]{ProgramsImages/IIDPoints.eps} &
+		\includegraphics[height=5cm]{ProgramsImages/SSobolPoints.eps}
+		\tabularnewline
+		$\vT_i$ are random &
+		$\vX_i$ may be deterministic \alert{or} random 
+		\tabularnewline
+		$\vT_1, \vT_2 \cdots \alert{\IIDsim} F$ &
+		$\vX_1, \vX_2 \cdots \alert{\LDsim} F$ 
+		\tabularnewline
+		$\vT_i$ do not know about one another &
+		$\{\vX_i\}_{i=1}^n$ represent $F$ well
+		\tabularnewline
+		$F_{n}(\vt_1, \ldots, \vt_n) = F(\vt_1) \cdots F(\vt_n)$ &
+		$F_{\{\vX_i\}_{i=1}^n}(\vx) \approx F(\vx)$
+	\end{tabular}
+\end{frame}
 

### 2022-07-09 MCQMC_2022_ChallengesDevelopSoftware.tex
Office Computer Updates

@@
 
-		\item
+		\item  \alert{Complete}---contain the components or have easy access to components in other libraries to solve real, complex problems
 
-		\item
+		\item  \alert{Accessible}---tutorials, demos, discussion forums, etc.\ for (new) users; written in a language that potential users speak; provide a consistent user interface
 
-		\item
+		\item \alert{Efficient}---in terms of compute time and memory
 
-		\item
+		\item \alert{Current}---include the latest and best algorithms
+
+		\item  \alert{Sustainable}--have a sufficient user base and developer community for updates and maintenance
+
+		\item \alert{Scalable}--take advantage of latest computer architectures for speed and to tackle large problems
 

### 2022-07-13 MCQMC_2022_ChallengesDevelopSoftware.tex
Updates from Overleaf
Office Computer Updates

@@
 
-\section{Background}
+\section{Introduction}
 
@@
 		\item \alert{Scalable}--take advantage of latest computer architectures for speed and to tackle large problems
+
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Why Do We Need Software?}
+\end{frame}
+
+\begin{frame}{Selective History of MC\textbf{QMC} Software}
+	\vspace{-6ex}
+	\begin{itemize}
+		\item \textcite{LEc2017a} provides a history of \alert{random number generation}
+		\item Notorious \alert{\texttt{randu}} from the 1960s \parencite{RANDU}, which fails the spectral test
+		\item \alert{\texttt{VEGAS}} \parencite{Lep78a, Lep21a} importance sampling Monte Carlo algorithm favored by physicists
+		\item \alert{ACM} low discrepancy point generators \parencite{BraFox88,BraFoxNie92,HonHic00a}
+		\item FinDer \parencite{PasTra95,FinDer} and BRODA \parencite{BRODA20a} targeting quantitative finance
+		\item Korobov cubature and scrambled Sobol' generators in \alert{NAG} \parencite{NAG27} for decades		
+		\item Scrambled Sobol' and Halton generators in \textcite{MAT9.13} since 2008
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Selective History of MC\textbf{QMC} Software}
+	\vspace{-6ex}
+	\begin{itemize}
+		\item Scrambled Sobol' and Halton generators in \textcite{MAT9.13} since 2008
+		\item Scrambled Sobol' in SciPy \parencite{virtanen2020scipy} and PyTorch \parencite{paszke2019pytorch}
 

### 2022-07-13 MCQMC_2022_ChallengesDevelopSoftware.tex
Updates from Overleaf

--- 2022-07-09 MCQMC_2022_ChallengesDevelopSoftware.tex
@@
 
-\thanksnote{with Mark Klinchin and the rest of the GAIL an QMCPy teams \\
+\thanksnote{with input from Mark Klinchin, the GAIL an QMCPy teams, and 
+several friends \\
 	partially supported by SigOpt, an Intel company \\[2ex]
@@
 	\end{itemize}
+\end{frame}
+
+\begin{frame}{Why Do We Need Software?}
+\end{frame}
+
+\begin{frame}{Brief History of MC\textbf{QMC} Software}
 \end{frame}

### 2022-07-13 MCQMC_2022_ChallengesDevelopSoftware.tex
Office Computer Updates

--- 2022-07-13 MCQMC_2022_ChallengesDevelopSoftware.tex
@@
 \usepackage[giveninits=false,backend=biber,style=nature,style=authoryear,maxcitenames =10, mincitenames=9]{biblatex}
+\AtBeginBibliography{\small}
 \addbibresource{FJHown23.bib}
@@
 
-\renewcommand{\OffTitleLength}{-7ex}
+\renewcommand{\OffTitleLength}{-10ex}
 \setlength{\FJHThankYouMessageOffset}{-8ex}
@@
 
-\thanksnote{with Mark Klinchin and the rest of the GAIL an QMCPy teams \\
+\thanksnote{with input from Mark Klinchin, the GAIL an QMCPy teams, and  friends \\
 	partially supported by SigOpt, an Intel company \\[2ex]
-	Thanks to the organizers as we meet again in person in still challenging times\\
-	Slides at  \href{https://speakerdeck.com/fjhickernell/quasi-monte-carlo-software}{\nolinkurl{speakerdeck.com/fjhickernell/quasi-monte-carlo-software}}
+	Please join us Tuesday, July 19, for lunch at ??? \\ to discuss the future of MCQMC software \\[2ex]
+	Thanks to the organizers as we meet again in person\\
+	Slides at  \href{??}{\nolinkurl{speakerdeck.com/fjhickernell/quasi-monte-carlo-software}}
 }
-\event{MCQMC 2022}
+\event{MCQMC 2022, Linz}
 \date[]{July 18, 2022}
@@
 		\item Korobov cubature and scrambled Sobol' generators in \alert{NAG} \parencite{NAG27} for decades		
-		\item Scrambled Sobol' and Halton generators in \textcite{MAT9.13} since 2008
+		\item Scrambled Sobol' and Halton generators in MATLAB \textcite{MAT9.13} since 2008
 	\end{itemize}
@@
 	\begin{itemize}
-		\item Scrambled Sobol' and Halton generators in \textcite{MAT9.13} since 2008
+		\item Keller
+		\item L'Ecuyer
+		\item Nuyens
+		\item Hickernell
+		\item TensorFlow
 		\item Scrambled Sobol' in SciPy \parencite{virtanen2020scipy} and PyTorch \parencite{paszke2019pytorch}
@@
 
+\begin{frame}{Software Library Architecture}
+	content...
+\end{frame}
+
+\begin{frame}{Choice of Language}
+	content...
+\end{frame}
+
+\begin{frame}{Advanced Architectures}
+	content...
+\end{frame}
+
+\begin{frame}{Introductory Demos}
+	content...
+\end{frame}
+
+\begin{frame}{Connecting Demos}
+	content...
+\end{frame}
+
+\begin{frame}{Building User \& Developer Ownership}
+	content...
+\end{frame}
+
+\begin{frame}{Next Steps}
+	content...
+\end{frame}
+
+\begin{frame}[allowframebreaks]{References}
 	\printbibliography
+\end{frame}
+
 \end{document}

### 2022-07-13 MCQMC_2022_ChallengesDevelopSoftware.tex
Office Computer Updates

@@
 		\item \alert{ACM} low discrepancy point generators \parencite{BraFox88,BraFoxNie92,HonHic00a}
-		\item FinDer \parencite{PasTra95,FinDer} and BRODA \parencite{BRODA20a} targeting quantitative finance
+		\item \alert{FinDer} \parencite{PasTra95,FinDer} and \alert{BRODA} \parencite{BRODA20a} targeting quantitative finance
 		\item Korobov cubature and scrambled Sobol' generators in \alert{NAG} \parencite{NAG27} for decades		
-		\item Scrambled Sobol' and Halton generators in MATLAB \textcite{MAT9.13} since 2008
+		\item Scrambled Sobol' and Halton generators in MATLAB \parencite{MAT9.13} since 2008, and fixed a few years later
 	\end{itemize}
@@
 	\begin{itemize}
+		\item Stan \cite{STAN} for 
 		\item Keller
@@
 \begin{frame}{Building User \& Developer Ownership}
+	content...
+\end{frame}
+
+\begin{frame}{From Private to Public}
 	content...

### 2022-07-13 MCQMC_2022_ChallengesDevelopSoftware.tex
Updates from Overleaf

@@
 
-\begin{frame}{Why Do We Need Software?}
-\end{frame}
-
 \begin{frame}{Selective History of MC\textbf{QMC} Software}
-	\vspace{-6ex}
+	\vspace{-4ex}
 	\begin{itemize}
 		\item \textcite{LEc2017a} provides a history of \alert{random number generation}
-		\item Notorious \alert{\texttt{randu}} from the 1960s \parencite{RANDU}, which fails the spectral test
-		\item \alert{\texttt{VEGAS}} \parencite{Lep78a, Lep21a} importance sampling Monte Carlo algorithm favored by physicists
+		\item Notorious \alert{randu} from the 1960s \parencite{RANDU}, which fails the spectral test
+		\item \alert{VEGAS} \parencite{Lep78a, Lep21a} importance sampling Monte Carlo algorithm favored by physicists
 		\item \alert{ACM} low discrepancy point generators \parencite{BraFox88,BraFoxNie92,HonHic00a}
@@
 	\begin{itemize}
-		\item Stan \cite{STAN} for 
+		\item \alert{Stan} \parencite{STAN} for Markov Chain Monte Carlo (MCMC)
 		\item Keller
@@
 \end{frame}
+
+\begin{frame}{Why Do We Need Software?}
+\end{frame}
+
+\section{Structure}
 
@@
 
+\section{Supporting Documentation}
 \begin{frame}{Introductory Demos}
@@
 
+\section{People}
 \begin{frame}{Building User \& Developer Ownership}
@@
 \end{frame}
+
+\section{Conclusion}
 

### 2022-07-13 MCQMC_2022_ChallengesDevelopSoftware.tex
Office Computer Updates

@@
 
-		\item  \alert{Complete}---contain the components or have easy access to components in other libraries to solve real, complex problems
+		\item  \alert{Complete}---contain the components or  easily access components in other libraries to solve real, complex problems
 
@@
 
-		\item  \alert{Sustainable}--have a sufficient user base and developer community for updates and maintenance
+		\item  \alert{Sustainable}---have a sufficient user base and developer community for updates and maintenance
 
-		\item \alert{Scalable}--take advantage of latest computer architectures for speed and to tackle large problems
+		\item \alert{Scalable}---take advantage of advanced computer architectures for speed and to solve large problems
 

### 2022-07-14 MCQMC_2022_ChallengesDevelopSoftware.tex
Office Computer Updates

@@
 	partially supported by SigOpt, an Intel company \\[2ex]
-	Please join us Tuesday, July 19, for lunch at ??? \\ to discuss the future of MCQMC software \\[2ex]
+	Please join us Tuesday, July 19, at 12:45 PM for lunch at Bella Casa \\ to discuss the future of MCQMC software \\[2ex]
 	Thanks to the organizers as we meet again in person\\
@@
 \begin{frame}{Next Steps}
-	content...
+	\includegraphics[width=0.5\textwidth]{ProgramsImages/PizeriaBellaCasa.png}
 \end{frame}

### 2022-07-14 MCQMC_2022_ChallengesDevelopSoftware.tex
Office Computer Updates

@@
 
-\begin{frame}{Selective History of MC\textbf{QMC} Software}
-	\vspace{-4ex}
+\begin{frame}[allowframebreaks]{Selective History of MC\textbf{QMC} Software}
 	\begin{itemize}
@@
 		\item Korobov cubature and scrambled Sobol' generators in \alert{NAG} \parencite{NAG27} for decades		
-		\item Scrambled Sobol' and Halton generators in MATLAB \parencite{MAT9.13} since 2008, and fixed a few years later
-	\end{itemize}
-\end{frame}
-
-\begin{frame}{Selective History of MC\textbf{QMC} Software}
-	\vspace{-6ex}
-	\begin{itemize}
-		\item \alert{Stan} \parencite{STAN} for Markov Chain Monte Carlo (MCMC)
-		\item Keller
+		\item Scrambled Sobol' and Halton generators in \alert{MATLAB} \parencite{MAT9.13} since 2008, and fixed a few years later
+		\item \alert{BUGS} \parencite{BUGSBook, BUGSweb} and \alert{Stan} \parencite{STAN} for Markov Chain Monte Carlo (MCMC)
+		\item Low discrepancy generators \parencite{FriKel02,FriKelweb}, \alert{SamplePack} \parencite{SamplePack}, \textcite{Gru},  and \alert{MatBuilder} \parencite{paulin2022} by Alex Keller and collaborators
 		\item L'Ecuyer
 		\item Nuyens
+		\item Giles
 		\item Hickernell

### 2022-07-14 MCQMC_2022_ChallengesDevelopSoftware.tex
Updates from Overleaf

@@
 \usepackage[usenames]{xcolor}
-\usepackage[giveninits=false,backend=biber,style=nature,style=authoryear,maxcitenames =10, mincitenames=9]{biblatex}
+\usepackage[giveninits=false,
+    backend=biber,
+    giveninits=true,
+    style=nature,
+    style=authoryear,
+    maxcitenames =2,
+    mincitenames=1]{biblatex}
 \AtBeginBibliography{\small}
@@
 \begin{frame}[allowframebreaks]{Selective History of MC\textbf{QMC} Software}
+\vspace{-2ex}
 	\begin{itemize}
@@
 		\item \alert{BUGS} \parencite{BUGSBook, BUGSweb} and \alert{Stan} \parencite{STAN} for Markov Chain Monte Carlo (MCMC)
-		\item Low discrepancy generators \parencite{FriKel02,FriKelweb}, \alert{SamplePack} \parencite{SamplePack}, \textcite{Gru},  and \alert{MatBuilder} \parencite{paulin2022} by Alex Keller and collaborators
+		\item Low discrepancy generators \parencite{FriKel02,FriKelweb}, \alert{SamplePack} \parencite{SamplePack}, \textcite{GruWeb},  and \alert{MatBuilder} \parencite{paulin2022} by Alex Keller and collaborators
 		\item L'Ecuyer
@@
 		\item Hickernell
-		\item TensorFlow
-		\item Scrambled Sobol' in SciPy \parencite{virtanen2020scipy} and PyTorch \parencite{paszke2019pytorch}
+		\item \alert{TensorFlow}
+		\item Scrambled Sobol' in \alert{SciPy} \parencite{virtanen2020scipy} and \alert{PyTorch} \parencite{paszke2019pytorch} since several years ago
 

### 2022-07-14 ProgramsImages/PlotRandomCoefficient.m
Office Computer Updates

@@ -1,0 +1,19 @@
+%% Plot of our random coefficient
+
+n = 2^10;
+d = 6;
+
+p = scramble(sobolset(d),"MatousekAffineOwen");
+wpts = 2*net(p,n)-1;
+xpts = (0:0.002:1)';
+chebpoly = cos(acos(2*xpts-1)*(1:d))./((1:d).^2);
+
+for beta = [0.6 0.8 0.9]
+    amat = 1 + beta*chebpoly*wpts';
+    plot(xpts,amat)
+    xlabel('\(x\)')
+    ylabel('\(a(x,\mathbf{W})\)')
+    axis([0 1 -0.2 2.2])
+    title(['\(\beta = ' num2str(beta) '\)'])
+    print('-depsc',['axwbeta' num2str(beta) '.eps'])
+end

### 2022-07-14 MCQMC_2022_ChallengesDevelopSoftware.tex
Office Computer Updates

@@
     style=authoryear,
-    maxcitenames =2,
+    maxcitenames=2,
     mincitenames=1]{biblatex}
@@
 		\item Low discrepancy generators \parencite{FriKel02,FriKelweb}, \alert{SamplePack} \parencite{SamplePack}, \textcite{GruWeb},  and \alert{MatBuilder} \parencite{paulin2022} by Alex Keller and collaborators
-		\item L'Ecuyer
+		\item \alert{RNGStreams} \parencite{LEcEtal02a},  L'Ecuyer
 		\item Nuyens
@@
 	content...
+\end{frame}
+
+\begin{frame}{Connecting QMCPy with FEniCS/Dolphin}
+	\begin{itemize}
+		\item Want to demonstrate how our QMC software could work with a popular DE solver to solve DEs with random coefficients
+		\item Partial success
+	\end{itemize}
+\begin{columns}
+ \begin{column}{0.5\textwidth}
+ 	\begin{gather*}
+ 		-\frac{\dif }{\dif x}\biggl(a(x,\vW) \frac{\dif}{\dif x} u(x,\vW) \biggr) = 2x - 1 \\
+ 		 0 \leq x \leq 1, \qquad
+ 		u(0) = u(1) = 0 \\
+ 		a(x,\vW) = 1 + \beta \sum_{k=1}^d \frac{W_k T_{k}(2x-1)}{k^2} \\
+ 		 \vW \sim \cu[-1,1]^d \\
+ 	\end{gather*}
+ \end{column}
+ \begin{column}{0.5\textwidth}
+ 	\centering
+ 	\includegraphics[width=\textwidth]{ProgramsImages/timing-2.eps}
+\end{column}
+\end{columns}
 \end{frame}

### 2022-07-14 ProgramsImages/PlotRandomCoefficient.m

@@ -1,6 +1,7 @@
 %% Plot of our random coefficient
 
-n = 2^10;
+close all
+n = 2^8;
 d = 6;
 
 p = scramble(sobolset(d),"MatousekAffineOwen");
@@ -9,11 +10,16 @@ xpts = (0:0.002:1)';
 chebpoly = cos(acos(2*xpts-1)*(1:d))./((1:d).^2);
 
 for beta = [0.6 0.8 0.9]
+    figure
     amat = 1 + beta*chebpoly*wpts';
     plot(xpts,amat)
     xlabel('\(x\)')
     ylabel('\(a(x,\mathbf{W})\)')
+    if beta == 0.6
+        axis([0 1 0 2])
+        print('-depsc',['axwtrue' num2str(beta) '.eps'])
+    end
     axis([0 1 -0.2 2.2])
     title(['\(\beta = ' num2str(beta) '\)'])
-    print('-depsc',['axwbeta' num2str(beta) '.eps'])
+    print('-depsc',['axwbeta' num2str(beta) '.eps']) 
 end

### 2022-07-15 MCQMC_2022_ChallengesDevelopSoftware.tex
Office Computer Updates
Updates from Overleaf

--- 2022-07-14 MCQMC_2022_ChallengesDevelopSoftware.tex
@@
 	content...
+\end{frame}
+
+\begin{frame}{Connecting QMCPy with FEniCS/Dolfin}
+	\begin{itemize}
+		\item Want to demonstrate how our QMC software could work with a popular DE solver for DEs with random coefficients
+		\item Partial success
+	\end{itemize}
+ 	\vspace{-3ex}
+ \begin{minipage}{0.48\textwidth}
+ 	\begin{gather*}
+ 		-\frac{\dif }{\dif x}\biggl(a(x,\vW) \frac{\dif}{\dif x} u(x,\vW) \biggr) = 200( 2x - 1)\\
+ 		 0 \leq x \leq 1, \qquad
+ 		u(0) = u(1) = 0 \\
+ 		a(x,\vW) = 1 + 0.6 \sum_{k=1}^d \frac{W_k T_{k}(2x-1)}{k^2} \\
+ 		 \vW \sim \cu[-1,1]^d \\
+ 		 \Ex[u(0.25,\vW)] = \, ?
+ 	\end{gather*}
+ \end{minipage}%
+\hfill
+ \begin{minipage}{0.48\textwidth}
+ 	\centering
+ 	\includegraphics<1>[width=\textwidth]{ProgramsImages/axwtrue0.6.eps}
+ 	\includegraphics<2->[width=\textwidth]{ProgramsImages/timing-4.eps}
+\end{minipage}
+
 \end{frame}

### 2022-07-15 MCQMC_2022_ChallengesDevelopSoftware.tex
Office Computer Updates

@@
 		\item \alert{BUGS} \parencite{BUGSBook, BUGSweb} and \alert{Stan} \parencite{STAN} for Markov Chain Monte Carlo (MCMC)
-		\item Low discrepancy generators \parencite{FriKel02,FriKelweb}, \alert{SamplePack} \parencite{SamplePack}, \textcite{GruWeb},  and \alert{MatBuilder} \parencite{paulin2022} by Alex Keller and collaborators
-		\item \alert{RNGStreams} \parencite{LEcEtal02a},  L'Ecuyer
-		\item Nuyens
-		\item Giles
-		\item Hickernell
+		\item \alert{LatMRG} \parencite{LEcCou97}, \alert{RNGStreams} \parencite{LEcEtal02a},  \alert{SSJ} \parencite{LEc2002a,SSJ}, and \alert{LatNetBuilder} \parencite{LatNet}
+		\item Low discrepancy generators \parencite{FriKel02,FriKelweb}, \alert{SamplePack} \parencite{SamplePack}, \textcite{GruWeb},  and \alert{MatBuilder} \parencite{paulin2022}
+		\item \alert{Fast CBC}, \alert{Magic Point Shop}, \alert{QMC4PDE} and other code since 2004  at \textcite{NuyWeb}
+		\item Multi-level software \parencite{GilesSoft,GilesQSoft} 
+		\item Data-driven error bounds and stopping criteria in \alert{GAIL} \parencite{ChoEtal21a} and \alert{QMCPy} \parencite{QMCPy2020a, ChoEtal22a} 
+		\framebreak
 		\item \alert{TensorFlow}

### 2022-07-15 MCQMC_2022_ChallengesDevelopSoftware.tex
Office Computer Updates

@@
 
-\begin{frame}[allowframebreaks]{Selective History of MC\textbf{QMC} Software}
+\begin{frame}[allowframebreaks]{Selective History of Publicly Available MC\textbf{QMC} Software}
 \vspace{-2ex}
@@
 		\item \textcite{LEc2017a} provides a history of \alert{random number generation}
-		\item Notorious \alert{randu} from the 1960s \parencite{RANDU}, which fails the spectral test
+		\item Notorious \alert{RANDU} from the 1960s \parencite{RANDU}, which fails the spectral test
 		\item \alert{VEGAS} \parencite{Lep78a, Lep21a} importance sampling Monte Carlo algorithm favored by physicists
@@
 		\item \alert{BUGS} \parencite{BUGSBook, BUGSweb} and \alert{Stan} \parencite{STAN} for Markov Chain Monte Carlo (MCMC)
-		\item \alert{LatMRG} \parencite{LEcCou97}, \alert{RNGStreams} \parencite{LEcEtal02a},  \alert{SSJ} \parencite{LEc2002a,SSJ}, and \alert{LatNetBuilder} \parencite{LatNet}
+		\item \alert{LatMRG} \parencite{LEcCou97}, \alert{RNGStreams} \parencite{LEcEtal02},  \alert{SSJ} \parencite{LEc2002a,SSJ}, and \alert{LatNetBuilder} \parencite{LatNet}
 		\item Low discrepancy generators \parencite{FriKel02,FriKelweb}, \alert{SamplePack} \parencite{SamplePack}, \textcite{GruWeb},  and \alert{MatBuilder} \parencite{paulin2022}
@@
 		\item Data-driven error bounds and stopping criteria in \alert{GAIL} \parencite{ChoEtal21a} and \alert{QMCPy} \parencite{QMCPy2020a, ChoEtal22a} 
-		\framebreak
-		\item \alert{TensorFlow}
+		\item Uncertainty quantification packages \alert{Dakota} \parencite{Dakota}, \alert{UQTk} \parencite{DebEtal04,UQTk}, and \alert{MUQ} \parencite{MUQ} have some basic level low discrepancy
+	    \item QMC framework in Julia since 2019 \parencite{QMCJulia} by Robbe and others
 		\item Scrambled Sobol' in \alert{SciPy} \parencite{virtanen2020scipy} and \alert{PyTorch} \parencite{paszke2019pytorch} since several years ago
+		\item \alert{TensorFlow} QMC framework \parencite{tfqfQMC2021a} since a year or so ago
 
@@
 \begin{frame}{Why Do We Need Software?}
+	\begin{itemize}
+		\item \alert{Theory} 
+		\begin{itemize}
+			\item Explains and justifies \alert{algorithms}
+			\item Assumptions$\implies$success should be viewed as failure$\implies$why
+		\end{itemize}
+		\item \alert{Software} 
+		\begin{itemize}
+			\item Makes \alert{algorithms} practical
+			\item Solves (new) \alert{applications}
+		\end{itemize}
+		\item New \alert{applications} inspire new \alert{theory}
+	\end{itemize}
 \end{frame}
@@
 \begin{frame}{Building User \& Developer Ownership}
-	content...
+ Research Software Engineers International 	\url{https://researchsoftware.org}, ``Research Software Engineers are people who combine professional software expertise with an understanding of research.
+They go by various job titles but the term Research Software Engineer (RSE) is fast gaining international recognition.''
+  \url{https://us-rse.org}
 \end{frame}

### 2022-07-15 MCQMC_2022_ChallengesDevelopSoftware.tex
Office Computer Updates

@@
 		\item Data-driven error bounds and stopping criteria in \alert{GAIL} \parencite{ChoEtal21a} and \alert{QMCPy} \parencite{QMCPy2020a, ChoEtal22a} 
-		\item Uncertainty quantification packages \alert{Dakota} \parencite{Dakota}, \alert{UQTk} \parencite{DebEtal04,UQTk}, and \alert{MUQ} \parencite{MUQ} have some basic level low discrepancy
+		\item Uncertainty quantification packages \alert{Dakota} \parencite{DakotaUsersManual}, \alert{UQTk} \parencite{DebEtal04,UQTk}, and \alert{MUQ} \parencite{MUQ} have some basic level low discrepancy sampling
 	    \item QMC framework in Julia since 2019 \parencite{QMCJulia} by Robbe and others
@@
 			\item Solves (new) \alert{applications}
+			\item Eliminates do-it-yourself
 		\end{itemize}
@@
 \begin{frame}{Software Library Architecture}
-	content...
+	\vspace{-7ex}
+
+\[
+\mu :=  \int_{\ct} g(\vt) \, \lambda(\vt) \, \dif \vt = \cdots = 	
+\underbrace{\Ex[f(\vX)]}_{\text{expectation}} = \underbrace{\int_{[0,1]^d}  f(\vx) \, \dif \vx}_{\text{integration}} \approx  \frac 1{n} \sum_{i=1}^{n} f(\vX_i) =: \hmu_{n}
+\]
+
+\vspace{-3ex}
+
+\begin{description}
+	%\setlength{\itemsep}{0.2cm}
+
+	\item[LD Generator] producing $\{\vX_1, \vX_2, \dots \}$ that mimics the distribution with PDF $\varrho$, e.g., uniform
+
+	\item[True Measure] that defines the original integral, e.g., Lebesgue; embodies the transformation $\vt = \Psi(\vx)$
+
+	\item[Integrand] $g$, which defines the original integral, plus the transformed version, $f$, to fit the LD generator
+
+	\item[Stopping Criterion] based on a data-driven error bound, whic determines how large $n$ should be to ensure that $\abs{\mu - \hmu_n} \le \varepsilon$
+\end{description}
+
+\vspace{-1ex}
+\uncover<2->{May we agree on a \alert{common definition} for these these objects like we have for floating point numbers or basic mathematical functions?}
 \end{frame}
 
-\begin{frame}{Choice of Language}
+\begin{frame}{Choice of Language or Package}
 	content...

### 2022-07-15 MCQMC_2022_ChallengesDevelopSoftware.tex
Updates from Overleaf

@@
 		\item Data-driven error bounds and stopping criteria in \alert{GAIL} \parencite{ChoEtal21a} and \alert{QMCPy} \parencite{QMCPy2020a, ChoEtal22a} 
-		\item Uncertainty quantification packages \alert{Dakota} \parencite{DakotaUsersManual}, \alert{UQTk} \parencite{DebEtal04,UQTk}, and \alert{MUQ} \parencite{MUQ} have some basic level low discrepancy sampling
+		\item Uncertainty quantification libraries \alert{Dakota} \parencite{DakotaUsersManual}, \alert{UQTk} \parencite{DebEtal04,UQTk}, and \alert{MUQ} \parencite{MUQ} have some basic level low discrepancy sampling
 	    \item QMC framework in Julia since 2019 \parencite{QMCJulia} by Robbe and others
@@
 		\item \alert{TensorFlow} QMC framework \parencite{tfqfQMC2021a} since a year or so ago
+	\end{itemize}
 
-	\end{itemize}
+In summary, we have stand-alone algorithms, small libraries, and pieces of larger libraries
 \end{frame}
@@
 			\item Solves (new) \alert{applications}
-			\item Eliminates do-it-yourself
+			\item Eliminates \alert{do-it-yourself}
 		\end{itemize}
-		\item New \alert{applications} inspire new \alert{theory}
+		\item \alert{Applications} 
+  		\begin{itemize}
+			\item Make \alert{societal} impact
+			\item Inspire new \alert{theory}
+		\end{itemize}
+
 	\end{itemize}
@@
 
-\begin{frame}{Choice of Language or Package}
-	content...
+\begin{frame}{Choice of Language, Library, or Environment}
+	\begin{itemize}
+    \item Language/library choices of the users are driven by familiarity and speed
+
+    \item Several favorites aligned with various communities, but no one dominates all
+
+    \item Diversity of multi-processor environments complicates software development; where does parallelism really help?
+
+    \item<2-> Recommendations
+    \begin{itemize}
+    \item Contribute your new idea as a demo in your favorite library
+    \item If your contribution is substantial, add as a feature to other libraries
+    \item Move basic features of your library to larger library
+    \item Write wrappers or demos to connect other libraries to your library
+
+    \end{itemize}
+ \end{itemize}
 \end{frame}
 
-\begin{frame}{Advanced Architectures}
-	content...
+\begin{frame}{Ex.\ Connecting QMCPy with FEniCS/Dolfin}
+\vspace{-4ex}
+	\begin{itemize}
+		\item Want to demonstrate how our QMC software could work with a popular DE solver for DEs with random coefficients
+		\item Partial success \only<3->{\alert{Why?}}
+	\end{itemize}
+ 	\vspace{-2ex}
+ \begin{minipage}{0.48\textwidth}
+ 	\only<1-2>{\begin{gather*}
+ 		-\frac{\dif }{\dif x}\biggl(a(x,\vW) \frac{\dif}{\dif x} u(x,\vW) \biggr) = 200( 2x - 1)\\
+ 		 0 \leq x \leq 1, \qquad
+ 		u(0) = u(1) = 0 \\
+ 		a(x,\vW) = 1 + 0.6 \sum_{k=1}^d \frac{W_k T_{k}(2x-1)}{k^2} \\
+ 		 \vW \sim \cu[-1,1]^d \\
+ 		 \Ex[u(0.25,\vW)] = \only<1>{\, ?}\only<2>{ -1.7462}
+ 	\end{gather*}}
+  \only<3->{
+\begin{itemize}
+\item Even with FEniCS/Dolfin's extensive documentation we had \alert{difficulties}
+\begin{itemize}
+    \item Learning how to solve a simple ODE, so reverted to an older version
+    \item Learning how to change the random instance of coefficient without tiggering the JIT compile
+    \item Do not yet know how to express the random coefficient in terms of the covariance kernel
+\end{itemize}
+\end{itemize}
+  }
+ \end{minipage}%
+\hfill
+ \begin{minipage}{0.48\textwidth}
+ 	\only<1-2>{\centering
+ 	\includegraphics<1>[width=\textwidth]{ProgramsImages/axwtrue0.6.eps}
+ 	\includegraphics<2->[width=\textwidth]{ProgramsImages/timing-4.eps}}%
+  \only<3->{
+\begin{itemize}
+\item But
+\begin{itemize}
+    \item It does work
+    \item If we can overcome the challenges, will post a blog and demo at \textcite{QMCBlog}
+\end{itemize}
+\end{itemize}
+
+  }
+\end{minipage}
+
 \end{frame}
 
-\section{Supporting Documentation}
-\begin{frame}{Introductory Demos}
+\section{Support}
+\begin{frame}{Encouraging Users Requires}
 	content...
@@
 
-\begin{frame}{Connecting QMCPy with FEniCS/Dolfin}
-	\begin{itemize}
-		\item Want to demonstrate how our QMC software could work with a popular DE solver for DEs with random coefficients
-		\item Partial success
-	\end{itemize}
- 	\vspace{-3ex}
- \begin{minipage}{0.48\textwidth}
- 	\begin{gather*}
- 		-\frac{\dif }{\dif x}\biggl(a(x,\vW) \frac{\dif}{\dif x} u(x,\vW) \biggr) = 200( 2x - 1)\\
- 		 0 \leq x \leq 1, \qquad
- 		u(0) = u(1) = 0 \\
- 		a(x,\vW) = 1 + 0.6 \sum_{k=1}^d \frac{W_k T_{k}(2x-1)}{k^2} \\
- 		 \vW \sim \cu[-1,1]^d \\
- 		 \Ex[u(0.25,\vW)] = \, ?
- 	\end{gather*}
- \end{minipage}%
-\hfill
- \begin{minipage}{0.48\textwidth}
- 	\centering
- 	\includegraphics<1>[width=\textwidth]{ProgramsImages/axwtrue0.6.eps}
- 	\includegraphics<2->[width=\textwidth]{ProgramsImages/timing-4.eps}
-\end{minipage}
-
-\end{frame}
-
-\section{People}
 \begin{frame}{Building User \& Developer Ownership}

### 2022-07-17 MCQMC_2022_ChallengesDevelopSoftware.tex
Office Computer Updates

@@
 \setlength{\arraycolsep}{0.5ex}
+\setbeamersize{text margin top=10mm}
 
@@
 
-	\vspace{-5ex}
+	%\vspace{-5ex}
 	\begin{itemize}
@@
 \begin{frame}{Why Do We Need Software?}
+	\vspace{-5ex}
 	\begin{itemize}
@@
 	\end{itemize}
+
+\uncover<2->{The United States Department of Energy is investing in studying how to develop great scientific software \parencite{Her19a,ASCR-SSSDU}}
 \end{frame}
@@
 
-	\item[Stopping Criterion] based on a data-driven error bound, whic determines how large $n$ should be to ensure that $\abs{\mu - \hmu_n} \le \varepsilon$
+	\item[Stopping Criterion] based on a data-driven error bound, which determines how large $n$ should be to ensure that $\abs{\mu - \hmu_n} \le \varepsilon$
 \end{description}
@@
 
-    \item Diversity of multi-processor environments complicates software development; where does parallelism really help?
+    \item Growth of multi-processor environments 
+    \begin{itemize}
+    	\item Provides opportunity
+    	\item Complicates software development; where does parallelism really help?
+    \end{itemize}
 
@@
 
-\begin{frame}{Ex.\ Connecting QMCPy with FEniCS/Dolfin}
+\begin{frame}{Ex.\ Connecting QMCPy with FEniCS/Dolfin \parencite{}}
 \vspace{-4ex}
@@
 \begin{frame}{Encouraging Users Requires}
-	content...
+	\begin{itemize}
+		\item A repository where
+		\begin{itemize}
+		\item  Your library can be easily downloaded
+		\item Issues or queries can be posted
+		\item There are updates
+		\end{itemize}
+		\item (Elementary) \alert{demos} on how to use the software and highlighting  its advantages
+		\item Demos \alert{connecting}  your library with others
+		\item Recommending your students and postdocs to \alert{use} your library
+		\item \alert{Welcoming} instructions on how to contribute an algorithm or a demo
+	\end{itemize}
+
 \end{frame}
 
-\begin{frame}{Connecting Demos}
-	content...
-\end{frame}
+\begin{frame}{Building Developer Team}
 
-\begin{frame}{Building User \& Developer Ownership}
- Research Software Engineers International 	\url{https://researchsoftware.org}, ``Research Software Engineers are people who combine professional software expertise with an understanding of research.
-They go by various job titles but the term Research Software Engineer (RSE) is fast gaining international recognition.''
-  \url{https://us-rse.org}
-\end{frame}
+\begin{itemize}
 
-\begin{frame}{From Private to Public}
-	content...
+\item Many senior researchers \alert{don't code}, and rely on transient team members (students, postdocs)
+\begin{itemize}
+	\item Encourage a sense of ownership that lasts beyond the time your team members are part of your lab
+	\item Encourage 
+\end{itemize}
+
+\item Recognize good software as \alert{valuable} scholarly output
+
+\item Recognize \alert{research software engineer} as a valued vocation; look at \alert{Research Software Engineers International} 	\url{https://researchsoftware.org}, \emph{Research Software Engineers are people who combine professional software expertise with an understanding of research.
+They go by various job titles but the term Research Software Engineer (RSE) is fast gaining international recognition.}
+  \end{itemize}
 \end{frame}

### 2022-07-17 MCQMC_2022_ChallengesDevelopSoftware.tex
Office Computer Updates

@@
 \setlength{\arraycolsep}{0.5ex}
-\setbeamersize{text margin top=10mm}
+%\setbeamersize{text margin top=10mm}
 
@@
 
-\begin{frame}{My Definition of Great MCQMC Software}
+\begin{frame}{My Definition of a Great MCQMC Software Ecosystem}
 
@@
 		\item Low discrepancy generators \parencite{FriKel02,FriKelweb}, \alert{SamplePack} \parencite{SamplePack}, \textcite{GruWeb},  and \alert{MatBuilder} \parencite{paulin2022}
+		\item Sobol' \alert{direction numbers} \parencite{JoeKuo03,JoeKuo08,SobolDirection}
 		\item \alert{Fast CBC}, \alert{Magic Point Shop}, \alert{QMC4PDE} and other code since 2004  at \textcite{NuyWeb}
-		\item Multi-level software \parencite{GilesSoft,GilesQSoft} 
+		\item \alert{Multi-level} software \parencite{GilesSoft,GilesQSoft} 
 		\item Data-driven error bounds and stopping criteria in \alert{GAIL} \parencite{ChoEtal21a} and \alert{QMCPy} \parencite{QMCPy2020a, ChoEtal22a} 
@@
 
-In summary, we have stand-alone algorithms, small libraries, and pieces of larger libraries
+We have stand-alone algorithms, small libraries, and pieces of larger libraries
 \end{frame}
@@
 
-\uncover<2->{The United States Department of Energy is investing in studying how to develop great scientific software \parencite{Her19a,ASCR-SSSDU}}
+\uncover<2->{The US Department of Energy is investing in studying how to develop great scientific software \parencite{Her19a,ASCR-SSSDU}}
 \end{frame}
@@
 
-\begin{frame}{Software Library Architecture}
+\begin{frame}{Software Architecture}
 	\vspace{-7ex}
@@
 \vspace{-1ex}
-\uncover<2->{May we agree on a \alert{common definition} for these these objects like we have for floating point numbers or basic mathematical functions?}
+\uncover<2->{May we agree on a \alert{common software definition} for these these objects like we have for floating point numbers or basic mathematical functions?}
 \end{frame}
@@
 \begin{frame}{Choice of Language, Library, or Environment}
+	\vspace{-5ex}
 	\begin{itemize}
-    \item Language/library choices of the users are driven by familiarity and speed
+    \item Language/library choices of the users are driven by \alert{familiarity} and \alert{speed}
 
-    \item Several favorites aligned with various communities, but no one dominates all
+  	\item  Favorites aligned with various \alert{communities}, but no one dominates all
+
+	\item Difference between \alert{prototyping} environments and ninja written \alert{production} code
 
@@
 
-\begin{frame}{Ex.\ Connecting QMCPy with FEniCS/Dolfin \parencite{}}
+\begin{frame}{Ex.\ QMCPy $+$ FEniCS/Dolfin \parencite{LoggEtal_10_2012}}
 \vspace{-4ex}

### 2022-07-17 MCQMC_2022_ChallengesDevelopSoftware.tex
Office Computer Updates

@@
 \begin{frame}{Ex.\ QMCPy $+$ FEniCS/Dolfin \parencite{LoggEtal_10_2012}}
-\vspace{-4ex}
+\vspace{-3ex}
 	\begin{itemize}
@@
 	\end{itemize}
- 	\vspace{-2ex}
+ 	\only<1>{\vspace{-3ex}}\only<2>{\vspace{-2ex}}\only<3>{\vspace{0ex}}
  \begin{minipage}{0.48\textwidth}
@@
 \begin{itemize}
-    \item Learning how to solve a simple ODE, so reverted to an older version
-    \item Learning how to change the random instance of coefficient without tiggering the JIT compile
+    \item Did not understand the latest version, so reverted to an older version
+    \item Had to learn how to change the random instance of coefficient without tiggering the JIT compile
     \item Do not yet know how to express the random coefficient in terms of the covariance kernel
@@
 \begin{itemize}
-    \item It does work
+    \item It  works
     \item If we can overcome the challenges, will post a blog and demo at \textcite{QMCBlog}
@@
 	\item Encourage a sense of ownership that lasts beyond the time your team members are part of your lab
-	\item Encourage 
+	\item Encourage those outside your own group to contribute
 \end{itemize}
@@
 \begin{frame}{Next Steps}
-	\includegraphics[width=0.5\textwidth]{ProgramsImages/PizeriaBellaCasa.png}
+	\begin{minipage}{0.48\textwidth}
+		content...
+	\end{minipage}
+	\begin{minipage}{0.48\textwidth}
+		Join us for a lunch discussion at Bella Casa tomorrow, July 19, at 12:45 PM
+
+		\vspace{2ex}
+
+		\includegraphics[width=\textwidth]{ProgramsImages/PizeriaBellaCasa.png}
+\end{minipage}
+
 \end{frame}

### 2022-07-18 MCQMC_2022_ChallengesDevelopSoftware.tex
Office Computer Updates

@@
 \setlength{\arraycolsep}{0.5ex}
-%\setbeamersize{text margin top=10mm}
 
@@
 	Thanks to the organizers as we meet again in person\\
-	Slides at  \href{??}{\nolinkurl{speakerdeck.com/fjhickernell/quasi-monte-carlo-software}}
+	Slides at  \href{https://speakerdeck.com/fjhickernell/challenges-of-developing-mcqmc-software}%
+	{\nolinkurl{speakerdeck.com/fjhickernell/challenges-of-developing-mcqmc-software}}
 }
@@
 
+	\newcommand{\talktime}[1]{{\color{Purple} \textbf{Talk #1}}}
+
 \newcommand{\northeaststuff}[3]{
@@
 
-\begin{frame}{My Definition of a Great MCQMC Software Ecosystem}
+\begin{frame}{A Great MCQMC Software Ecosystem Should Be}
 
-	%\vspace{-5ex}
+	\vspace{-3ex}
 	\begin{itemize}
@@
 
-		\item  \alert{Accessible}---tutorials, demos, discussion forums, etc.\ for (new) users; written in a language that potential users speak; provide a consistent user interface
+		\item  \alert{Accessible}---downloadable; tutorials, demos, discussion forums, etc.\ for (new) users; written in a language that potential users speak; provide a consistent user interface
 
@@
 		\item \alert{BUGS} \parencite{BUGSBook, BUGSweb} and \alert{Stan} \parencite{STAN} for Markov Chain Monte Carlo (MCMC)
-		\item \alert{LatMRG} \parencite{LEcCou97}, \alert{RNGStreams} \parencite{LEcEtal02},  \alert{SSJ} \parencite{LEc2002a,SSJ}, and \alert{LatNetBuilder} \parencite{LatNet}
-		\item Low discrepancy generators \parencite{FriKel02,FriKelweb}, \alert{SamplePack} \parencite{SamplePack}, \textcite{GruWeb},  and \alert{MatBuilder} \parencite{paulin2022}
+		\item \alert{LatMRG} \parencite{LEcCou97}, \alert{RNGStreams} \parencite{LEcEtal02},  \alert{SSJ} \parencite{LEc2002a,SSJ}, and \alert{LatNetBuilder} \parencite{LatNet} \talktime{M 11}
+		\item Low discrepancy generators \parencite{FriKel02,FriKelweb}, \alert{SamplePack} \parencite{SamplePack}, \textcite{GruWeb},  and \alert{MatBuilder} \parencite{paulin2022} \talktime{M 12, Tu 12}
 		\item Sobol' \alert{direction numbers} \parencite{JoeKuo03,JoeKuo08,SobolDirection}
-		\item \alert{Fast CBC}, \alert{Magic Point Shop}, \alert{QMC4PDE} and other code since 2004  at \textcite{NuyWeb}
-		\item \alert{Multi-level} software \parencite{GilesSoft,GilesQSoft} 
-		\item Data-driven error bounds and stopping criteria in \alert{GAIL} \parencite{ChoEtal21a} and \alert{QMCPy} \parencite{QMCPy2020a, ChoEtal22a} 
-		\item Uncertainty quantification libraries \alert{Dakota} \parencite{DakotaUsersManual}, \alert{UQTk} \parencite{DebEtal04,UQTk}, and \alert{MUQ} \parencite{MUQ} have some basic level low discrepancy sampling
-	    \item QMC framework in Julia since 2019 \parencite{QMCJulia} by Robbe and others
+		\item \alert{Fast CBC}, \alert{Magic Point Shop}, \alert{QMC4PDE} and other code since 2004  at \textcite{NuyWeb} \talktime{Tu 10:30}
+		\item \alert{Multi-level} software \parencite{GilesSoft,GilesQSoft} \talktime{M 10:30}
+		\item Data-driven error bounds and stopping criteria in \alert{GAIL} \parencite{ChoEtal21a} and \alert{QMCPy} \parencite{QMCPy2020a, ChoEtal22a}  \talktime{Tu 11}
+		\item Uncertainty quantification libraries \alert{Dakota} \parencite{DakotaUsersManual}, \alert{UQTk} \parencite{DebEtal04,UQTk}, and \alert{MUQ} \parencite{MUQ} \talktime{Tu 11:30} have some basic level low discrepancy sampling
+	    \item QMC framework in Julia since 2019 \parencite{QMCJulia} by Robbe and others \talktime{M 17}
 		\item Scrambled Sobol' in \alert{SciPy} \parencite{virtanen2020scipy} and \alert{PyTorch} \parencite{paszke2019pytorch} since several years ago
@@
 \begin{frame}{Why Do We Need Software?}
-	\vspace{-5ex}
+	\vspace{-4ex}
 	\begin{itemize}
@@
 		\begin{itemize}
-			\item Explains and justifies \alert{algorithms}
-			\item Assumptions$\implies$success should be viewed as failure$\implies$why
+			\item Explains, justifies, and leads to \alert{algorithms}
+			\item Assumptions$\implies$success can be viewed as failure$\implies$what went wrong
 		\end{itemize}
@@
 
-\uncover<2->{The US Department of Energy is investing in studying how to develop great scientific software \parencite{Her19a,ASCR-SSSDU}}
+\uncover<2->{The US Department of Energy is investing in studying how to develop great scientific software, including a recent \emph{Workshop on the Science of Scientific-Software Development and Use} \parencite{Her19a,ASCR-SSSDU}}
 \end{frame}
@@
 \begin{frame}{Software Architecture}
-	\vspace{-7ex}
+	\vspace{-6ex}
 
@@
 
-\begin{description}
-	%\setlength{\itemsep}{0.2cm}
+\begin{description}[\labelwidth=5ex,\leftmargin=10ex]
+	%\setlength{\labelwidth}{5ex}
 
-	\item[LD Generator] producing $\{\vX_1, \vX_2, \dots \}$ that mimics the distribution with PDF $\varrho$, e.g., uniform
+	\item[Low Discrepancy Generator] producing $\{\vX_1, \vX_2, \dots \}$ that mimics the uniform distribution
 
-	\item[True Measure] that defines the original integral, e.g., Lebesgue; embodies the transformation $\vt = \Psi(\vx)$
+	\item[True Measure] that defines the original integral, e.g., Lebesgue; embodies the transformation $\vt = \vPsi(\vx)$
 
@@
 
-\begin{frame}{Choice of Language, Library, or Environment}
+\begin{frame}{Choosing a Language, Library, or Environment}
 	\vspace{-5ex}
@@
 
-    \item Growth of multi-processor environments 
+    \item Growth of \alert{multi-processor} environments 
     \begin{itemize}
@@
     \begin{itemize}
-    \item Contribute your new idea as a demo in your favorite library
-    \item If your contribution is substantial, add as a feature to other libraries
-    \item Move basic features of your library to larger library
-    \item Write wrappers or demos to connect other libraries to your library
+    \item Contribute your new idea as a \alert{demo} in your favorite library
+    \item If your contribution is substantial, add as a \alert{feature} to other libraries
+    \item Move basic features of your library to \alert{larger} library
+    \item Write wrappers or demos to \alert{connect} other libraries to your library
 
@@
     \item Did not understand the latest version, so reverted to an older version
-    \item Had to learn how to change the random instance of coefficient without tiggering the JIT compile
+    \item Had to learn how to change the random instance of the coefficient without tiggering the JIT compiler
     \item Do not yet know how to express the random coefficient in terms of the covariance kernel
@@
 \section{Support}
-\begin{frame}{Encouraging Users Requires}
+\begin{frame}{Encouraging Users of Your Library Requires}
+
+	\vspace{-4ex}
 	\begin{itemize}
@@
 		\item  Your library can be easily downloaded
+		\item There is documentation
 		\item Issues or queries can be posted
 		\item There are updates
+		\item Tests are run regularly to mitigate against bugs
 		\end{itemize}
@@
 
-\begin{frame}{Building Developer Team}
+\begin{frame}{Building a Developer Team for Your Library}
 
+	\vspace{-4ex}
 \begin{itemize}
@@
 \begin{itemize}
-	\item Encourage a sense of ownership that lasts beyond the time your team members are part of your lab
-	\item Encourage those outside your own group to contribute
+	\item Encourage a sense of \alert{ownership} that lasts beyond the time your team members are part of your lab
+	\item Encourage those \alert{outside} your own group to contribute
 \end{itemize}
+
+\item Those of us who understand theory and appreciate software need to encourage and steer popular software to ensure that it is implemented well.
 
@@
 
-\item Recognize \alert{research software engineer} as a valued vocation; look at \alert{Research Software Engineers International} 	\url{https://researchsoftware.org}, \emph{Research Software Engineers are people who combine professional software expertise with an understanding of research.
+\item Recognize \alert{research software engineer} as a valued vocation; see \alert{Research Software Engineers International} 	\url{https://researchsoftware.org}, \emph{Research Software Engineers are people who combine professional software expertise with an understanding of research.
 They go by various job titles but the term Research Software Engineer (RSE) is fast gaining international recognition.}
@@
 
-\section{Conclusion}
+\section{Next Steps}
 
 \begin{frame}{Next Steps}
+
+	\vspace{-2ex}
 	\begin{minipage}{0.48\textwidth}
-		content...
+		\begin{itemize}
+			\item Which exemplary \alert{theoretical developments} are not yet implemented in readily available software?
+			\item What potential MCQMC user communities are \alert{ill-served} by existing  software libraries?
+			\item Should those of us interested in software gather regularly for \alert{show-and-tell} to foster more collaboration?
+		\end{itemize}
 	\end{minipage}
 	\begin{minipage}{0.48\textwidth}
-		Join us for a lunch discussion at Bella Casa tomorrow, July 19, at 12:45 PM
+		Join us for a \alert{lunch} discussion at Bella Casa tomorrow, July 19, at 12:45 PM
 

### 2022-07-24 MCQMC_2022_ChallengesDevelopSoftware.tex
Office Computer Updates

@@
 \thanksnote{with input from Mark Klinchin, the GAIL an QMCPy teams, and  friends \\
-	partially supported by SigOpt, an Intel company \\[2ex]
-	Please join us Tuesday, July 19, at 12:45 PM for lunch at Bella Casa \\ to discuss the future of MCQMC software \\[2ex]
+	partially supported by SigOpt (Intel) and an Illinois  Tech alum\\[2ex]
+	Join us Tuesday, July 19, at 12:45 PM for lunch at Bella Casa \\ to discuss the future of MCQMC software \\[2ex]
 	Thanks to the organizers as we meet again in person\\
@@
 
-\begin{description}[\labelwidth=5ex,\leftmargin=10ex]
+\begin{description}[alabel]
 	%\setlength{\labelwidth}{5ex}
@@
 
-    \item<2-> Recommendations
+    \item<2-> Recommendations if you want to contribute software
     \begin{itemize}
-    \item Contribute your new idea as a \alert{demo} in your favorite library
-    \item If your contribution is substantial, add as a \alert{feature} to other libraries
+    \item Realize your new idea as a \alert{demo} in your favorite library
+    \item Add a significant idea as a \alert{feature} to other libraries
     \item Move basic features of your library to \alert{larger} library

### 2022-10-31 MCQMC2022QMCSoftwareArticle.tex
Office Computer Updates

--- 2022-06-28 MCQMC 2020 stuff/QMCSoftwareArticleNoRho.tex
@@
+% MCQMC 2022 article on the Challenges of Great MCQMC Software
 %
@@
 
-\usepackage[colorlinks=true,linkcolor=black,citecolor=black,urlcolor=blue]{hyperref}
+\usepackage[colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black]{hyperref}
 \urlstyle{same}
@@
 \newcommand{\QMCPYbigabs}[1]{\ensuremath{{\bigl \lvert #1 \bigr \rvert}}}
-\newcommand{\QMCPYBigabs}[1]{\ensuremath{{\Bigl \lvert #1 \Bigr \rvert}}}
-\newcommand{\QMCPYbiggabs}[1]{\ensuremath{{\biggl \lvert #1 \biggr \rvert}}}
 
@@
 \newcommand{\frednote}[1]{  {\textcolor{red}  {\mbox{**Fred:} #1}}}
-\newcommand{\scnote}[1]{ {\textcolor{blue}  {\mbox{**SC:} #1}}}
+\newcommand{\yuhannote}[1]{ {\textcolor{violet}  {\mbox{**Yuhan:} #1}}}
 \newcommand{\tonynote}[1]{ {\textcolor{orange}  {\mbox{**Tony:} #1}}}
@@
 \newcommand{\FJHComment}[1]{{\color{magenta} Fred: #1}}
-\newcommand{\JRComment}[1]{{\color{violet} Jag: #1}}
 
@@
 \newcommand{\LD}{\textup{LD}}
-\newcommand{\IMP}{\textup{IS}}
+\newcommand{\IMP}{\textup{imp}}
 \newcommand{\unif}{\textup{unif}}
@@
 \newcommand{\LDsim}{\overset{\LD}{\sim}}
-\newcommand{\Msim}{\overset{\textup{M}}{\sim}}
 \newcommand{\dig}{\textup{dig}}
@@
 \newcommand{\mI}{\mathsf{I}}
-\newcommand{\cube}{[0,1]^d}
-\newcommand{\Kum}{\textup{Kum}}
-
-\iffalse
-Here are some of the main suggestions, other grammatical / wording suggestions are in the reports
-
-* General ~ Sou-Cheng DONE
-Expand section 2 and talk more about how QMCPy combines the best of available QMC softwares
-Fix LatNet Builder reference Fred
-Add tensorflow
-
-* Discrete Distribution ~ Fred
-Discuss difference between stratification and RQMC
-Comment on the name DiscreteDistribution, the referee and Dirk Nuyens found this confusing Fred done
-List criterion used to construct Lattice sequences Aleksei done
-Clarify that the better uniformity is for embedded Lattices and is not true in general Fred done
-
-* Code / Choice of Parameters ~ Aleksei, Jags check
-Show how the user can customize Sobol' and Lattice generative vectors
-Make parameters more explicit e.g. Sobol(2) --> Sobol(dimension=2), and explain more about what the code does for non-python / QMCPy users
-
-* Integrand / Stopping Criterion ~ Jags done, Sou-Cheng check
-Explain why we use the trapezoidal rule for AsianOption and how we chose the drift for the importance sampling example? 
-Better explanation of CubQMCSobolG algorithm
-
-* Importance Sampling ~ Aleksei done, Fred check
-Provide better discussion of importance sampling and justification for the example
-Motivation for successive change of variables
-Explain the Kumaraswamy distribution
-
-\fi
 
@@
 \title*{Quasi-Monte Carlo Software}
-\authorrunning{S.-C.\ T.\ Choi et al.}
-\author{Sou-Cheng T. Choi \and Fred J. Hickernell \and R. Jagadeeswaran \and Michael J. McCourt \and Aleksei G. Sorokin}
-\institute{Sou-Cheng T.
-Choi \at Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616; and Kamakura Corporation, 2222 Kalakaua Ave, Suite 1400, Honolulu, HI 96815 \email{schoi32@iit.edu}
+\authorrunning{S.-C.\ T.\ Choi et al}
+\author{Sou-Cheng Terrya Choi \and Fred J. Hickernell \and R. Jagadeeswaran \and Michael J. McCourt \and Aleksei G. Sorokin}
+\institute{Sou-Cheng Terrya Choi \at Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{schoi32@hawk.iit.edu}
 \and
@@
 R. Jagadeeswaran \at
-Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{jrathin1@iit.edu}; and \\Wi-Tronix LLC, 631 E Boughton Rd, Suite 240, Bolingbrook, IL 60440
 \and
-Michael J. McCourt \at SigOpt, an Intel company, \\
-100 Bush St., Suite 1100, San Francisco, CA 94104
-\email{mccourt@sigopt.com}
+Michael J. McCourt \at ???
 \and 
@@
 
-\abstract{Practitioners wishing to experience the efficiency gains from using low discrepancy sequences need correct, robust, well-written software.
-This article, based on our MCQMC 2020 tutorial, describes some of the better quasi-Monte Carlo (QMC) software available.
-We highlight the key software components required  by QMC to approximate multivariate integrals or expectations of functions of vector random variables.
-We have combined these components in QMCPy, a Python open-source library, which we hope will draw the support of the QMC community.
-Here we introduce QMCPy.}
+\abstract{This is the article based on my MCQMC 2020 Tutorial.
+\AGSComment{Latest code demo at: \href{https://colab.research.google.com/drive/1wijwt0ugsJ9chUHtzSb3lQpc3-N8mCDs?usp=sharing}{https://colab.research.google.com/drive/1wijwt0ugsJ9chUHtzSb3lQpc3-N8mCDs?usp=sharing}}.}
 
 \section{Introduction} \label{sec:intro}
+Quasi-Monte Carlo (QMC) methods promise great efficiency gains over independent and identically distributed (IID) Monte Carlo (MC) methods.
+In some cases QMC  achieves one hundredth of the error of IID MC in the same amount of time.
+Often, these efficiency gains are obtained simply by  replacing IID sampling by low discrepancy (LD) sampling, which is the heart of QMC.
 
-Quasi-Monte Carlo (QMC) methods promise great efficiency gains over independent and identically distributed (IID) Monte Carlo (MC) methods.
-In some cases, QMC  achieves one hundredth of the error of IID MC in the same amount of time (see Figure \ref{fig:sc_comp}).
-Often, these efficiency gains are obtained simply by replacing IID sampling with low discrepancy (LD) sampling, which is the heart of QMC.
-
-Practitioners might wish to test whether QMC would speed up their computation; doing so would be made easier with access the best QMC algorithms available.
-Theoreticians or algorithm developers might want to demonstrate their ideas on various use cases to show their practical value.
+As a practitioner, you might wish to test whether QMC would speed your computation.
+You would like easy access to the best QMC algorithms available.
+As a theoretician or an algorithm developer, you would want to demonstrate your best ideas on  various  use cases to show their practical value.
 
 This tutorial points to some of the best QMC software available.
-Then we describe  QMCPy \cite{QMCPy2020a}\footnote{QMCPy is in active development.
-This article is based on version 1.0.1 on PyPI.}, which is crafted to be a community-owned Python library that combines the best QMC algorithms and interesting use cases from various authors under a common user interface.
+Moreover, we describe QMCPy \cite{QMCPy2020a}, which is designed to be a community owned Python library that combines the best QMC algorithms from various authors under a common user interface.
 
-The model problem for QMC is approximating a multivariate integral,
+The model problem for QMC is approximating an integral,
 \begin{equation} \label{eq:integral}
@@
 where $g$ is the integrand, and $\lambda$ is a non-negative weight.
-If $\lambda$ is a probability distribution (PDF) for the random variable $\bsT$, then $\mu$ is the mean of $g(\bsT)$.
-Regardless, we perform a suitable variable transformation to interpret this integral as the  mean of a function of a multivariate, standard uniform random variable:
+We use $\mu$ to denote the value of this integral because we interpret it as the population mean of a random variable after a suitable variable transformation:
 \begin{equation} \label{eq:fintegral}
-	\mu = \bbE[f(\bsX)] =  \int_{\cube}  f(\bsx) \,  \D \bsx , \qquad \bsX \sim \calU\cube.
+	\mu = \bbE[f(\bsX)] = \int_\calX f(\bsx) \, \varrho(\bsx) \, \D \bsx =  \int_\calX f(\bsx) \,  \D F(\bsx) ,
 \end{equation}
+where $\varrho$ is a probability density with corresponding probability distribution $F$.
+In practice, $\calX$ often corresponds to the unit cube, $[0,1]^d$, and $F$ corresponds to the uniform distribution.
 
-QMC approximates the population mean, $\mu$,  by a sample mean,
+QMC approximates this population mean by a sample mean,
 \begin{equation} \label{eq:samplemean}
-	\hmu := \frac 1n \sum_{i=0}^{n-1} f(\bsX_i), \qquad \bsX_0, \bsX_1, \ldots \Msim \calU\cube.
+	\hmu := \frac 1n \sum_{i=0}^{n-1} f(\bsX_i), \qquad \bsX_0, \bsX_1, \ldots \sim F.
 \end{equation}
-The choice of the sequence $\{\bsX_i\}_{i=0}^\infty$ and the choice of $n$ to satisfy  the prescribed error requirement,
-\begin{equation} \label{eq:err_req}
-	\QMCPYabs{\mu - \hmu} \le \varepsilon \qquad \text{absolutely or with high probability},
-\end{equation} 
-are important decisions, which  QMC software helps the user make.
+The choice of this sequence and the choice of $n$ to satisfy  the prescribed error tolerance are important decisions, which  QMC software helps the user make.
 
-Here, the notation $\Msim$ means that the sequence mimics the specified, target distribution, but not necessarily in a probabilistic way.  We  use this notation in two forms:  $\IIDsim$ and $\LDsim$.
+Here, the notation ``$\sim$'' means that the sequence mimics the specified, target distribution, but not necessarily in a probabilistic way.
+We  use this notation in two forms:  $\IIDsim$ and $\LDsim$.
 
@@
 The position of any point is not influenced by any other, so clusters and gaps occur.
-A randomly chosen subsequence of an IID sequence is also IID.
-When we say that $\bsX_0, \bsX_1, \ldots \IIDsim F$ for some distribution $F$, we mean that for any positive integer $n$, the  multivariate probability distribution of $\bsX_0, \ldots, \bsX_{n-1}$ is the product of the marginals, specifically,
+A subsequence of IID points chosen randomly is also IID.
+When we say that $\bsX_0, \bsX_1, \ldots \IIDsim F$, we mean that for any positive integer $n$, the  multivariate probability distribution of $\bsX_0, \ldots, \bsX_{n-1}$ is the product of the marginals, specifically,
 \begin{equation*}
@@
 \end{equation*}
-When IID points are used to approximate $\mu$ by the sample mean, the root mean squared error is $\calO(n^{-1/2})$.
-Figure \ref{fig:comparePts} displays IID uniform points, $\bsX^{\IID}_0, \bsX^{\IID}_1, \ldots \IIDsim \calU[0,1]^2$, i.e.,  the target distribution is $F_\unif(\bsx) = x_1 x_2$.
+When IID points are used to approximate $\mu$ by the sample mean, the error is $\calO(n^{-1/2})$.
+Figure \ref{fig:comparePts} displays IID uniform points, $\bsX^{\IID}_0, \bsX^{\IID}_1, \ldots \IIDsim \calU[0,1]^2$, where the target distribution is $F(\bsx) = x_1 x_2$.
 
-\begin{figure}[t]
-	\includegraphics[height=5.8cm]{QMCSoftwareArticle/vfigs/dd_iid_uniform_pts.eps}
-	\quad
-	\includegraphics[height=5.8cm]{QMCSoftwareArticle/vfigs/dd_sobol_pts.eps}
-	\caption{IID points (left) contrasted with LD points (right).  The LD points cover the square more evenly.}
+\begin{figure}
+	\includegraphics[height=5cm]{QMCSoftwareArticle/figs/dd_iid_uniform_pts.png}
+	\qquad
+	\includegraphics[height=5cm]{QMCSoftwareArticle/figs/dd_sobol_pts.png}
+	\caption{IID points contrasted with LD points.  The LD points cover the square more evenly.}
 	\label{fig:comparePts}
@@
 
-LD sequences may be deterministic or random, but each point is carefully coordinated with the others so that they fill the domain well.
-Subsequences of LD sequences are generally not LD.
-When we say that $\bsX_0, \bsX_1, \ldots \LDsim \calU\cube$, we mean that for any positive integer $n$,  the \emph{empirical distribution} of $\bsX_0, \ldots, \bsX_{n-1}$, denoted $F_{\{\bsX_i\}_{i=0}^{n-1}}$,  approximates the uniform distribution, $F_\unif$, well (relative to $n$).
-(The empirical distribution of a set assigns equal probability to each point.)
-
-A measure of the difference between the empirical distribution of a set of points and the uniform distribution is called a \emph{discrepancy} and is denoted $D(\{\bsX_i\}_{i=0}^{n-1})$ \cite{DicEtal14a,Hic97a,Hic99a,Nie92}.
+LD sequences may be deterministic or random, but each point is carefully coordinated with the other so that they fill the domain well.
+When we say that $\bsX_0, \bsX_1, \ldots \LDsim F$, we mean that for any positive integer $n$,  the \emph{empirical distribution} of $\bsX_0, \ldots, \bsX_{n-1}$, denoted $F_{\{\bsX_i\}_{i=0}^{n-1}}$,  approximates the target distribution, $F$, well (relative to $n$).
+(The empirical distribution of a set assigns equal probability to each point.)  The measure of the difference between the empirical distribution of a set of points and the target distribution is called a \emph{discrepancy} and is denoted $D(\{\bsX_i\}_{i=0}^{n-1}, F)$.
 This is the origin of the term ``low discrepancy'' points or sequences.
 LD points by definition have a smaller discrepancy than IID points.
-Figure \ref{fig:comparePts} contrasts IID uniform points with LD points, $\bsX^{\LD}_0, \bsX^{\LD}_1 \ldots \LDsim \calU[0,1]^2$, in this case, linearly scrambled and digitally shifted Sobol' points.
+Figure \ref{fig:comparePts} contrasts IID uniform points with LD points, $\bsX^{\LD}_0, \bsX^{\LD}_1 \ldots \LDsim \calU[0,1]^2$, in this case linearly scrambled Sobol' points.
+For nearly all LD sequences, the target distribution is $\calU[0,1]^d$.
 
-The error in using the sample mean to approximate the integral can be bounded according to the Koksma-Hlawka inequality and its extensions \cite{DicEtal14a,Hic97a,Hic99a,Nie92} as the product of the discrepancy of the sampling sequence and the variation of the integrand, denoted $V(\cdot)$:
+The error in using the sample mean to approximate the integral can be bounded according to the Koksma-Hlawka inequality and its extensions as the product of the discrepancy of the sampling sequence and the variation of the integrand, denoted $V(\cdot)$:
 \begin{equation}
-	\QMCPYabs{\mu - \hmu} = \QMCPYbiggabs{\int_{\cube} f(\bsx) \, \D \bigl(F_\unif - F_{\{\bsX_i\}_{i=0}^{n-1}}\bigr) (\bsx)} \le D\bigl(\{\bsX_i\}_{i=0}^{n-1}\bigr) V(f).
+	\QMCPYabs{\mu - \hmu} = \QMCPYabs{\int_{\calX} f(\bsx) \, \D (F - F_{\{\bsX_i\}_{i=0}^{n-1}}) (\bsx)} \le D(\{\bsX_i\}_{i=0}^{n-1}, F) V(f),
 \end{equation} 
@@
 For typical Banach spaces, the discrepancy of LD points is $\calO(n^{-1+\epsilon})$, which is a higher convergence order than for IID points.
-For details, the reader is pointed to the references.
+For details, the reader is referred to the references.
+For our purposes, we expect the reader to see in Figure \ref{fig:comparePts} that the LD points cover the integration domain more evenly than IID points.
+In the examples below the reader will see the demonstrably smaller cubature errors arising from using LD points.
 
-Here, we expect the reader to see in Figure \ref{fig:comparePts} that the LD points cover the integration domain more evenly than IID points.
-LD sampling can be thought of as a more even distribution of the sampling sites than IID.
-LD sampling is similar to stratified sampling.
-In the examples below, the reader will see the demonstrably smaller cubature errors arising from using LD points.
+In the sections that follow we first overview available QMC software.
+We next describe an architecture for good QMC software, i.e., what are the key components and how should they interact.
+We then describe how we have implemented this architecture in QMCPy.
+Finally, we summarize further directions that we hope QMCPy and related software projects will take.
+Those interested in following the development of QMCPy or even contributing to that development are urged to visit the GitHub repository at \href{https://qmcsoftware.github.io/QMCSoftware/}{\nolinkurl{https://qmcsoftware.github.io/QMCSoftware/}}.
 
-In the sections that follow, we first overview available QMC software.
-We next describe an architecture for good QMC software, i.e., what the key components are and how they should interact.
-We then describe how we have implemented this architecture in QMCPy.
-Finally, we summarize further directions that we hope QMCPy and other QMC software projects will take.
-Those interested in following or contributing to the development of QMCPy are urged to visit the GitHub repository at \href{https://github.com/QMCSoftware/QMCSoftware}{\nolinkurl{https://github.com/QMCSoftware/QMCSoftware}}.
-
-We have endeavored to be as accurate as possible at the time of writing this article.
-We hope that progress in QMC software development will make this article obsolete in a good way in the coming years.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Available Software for QMC} \label{sec:available} 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-QMC software spans  LD sequence generators, cubatures, and applications.
-We review the better-known software collections, recognizing that some software overlaps multiple categories.
-Whenever applicable, we state each library's accessibility in QMCPy, or contrast its functionalities with QMCPy's --- where we lag behind in QMCPy, we strive to catch up in the near future.
-
-Software focusing on generating high-quality LD sequences and their generators includes the following, listed in alphabetical order:
+QMC software spans  LD sequence generators, algorithms, and applications.  We review the better known software collections, recognizing that some software overlaps multiple categories.
+Software focusing on generating high quality LD sequences  or their generators includes
 \begin{description}[format=\textup,format=\textbf]
 % https://tex.stackexchange.com/questions/74279/how-to-add-bullets-to-description-lists
-	\item[BRODA] Commerical and non-commercial software developed jointly with I.M.~Sobol' in C++, MATLAB, and Excel~\cite{BRODA20a}.
-BRODA can generate Sobol' sequences up to 65,536 dimensions.
-In comparison, QMCPy supports Sobol' sequences up to 21,201 dimensions.
-	\item[Burkhardt] Various QMC software for generating van der Corput, Faure, Halton, Hammersley,  Niederreiter, or Sobol' sequences in C, C++, Fortran, MATLAB, or Python~\cite{Bur20a}.
-In QMCPy, we have implemented digital net, lattice, and Halton generators.
-	\item[LatNet Builder] The successor to \emph{Lattice Builder}~\cite{LEcMun14a}, LatNet Builder is a C++ library with Python and Java interfaces (in SSJ below) for generating vectors/matrices for lattices and digital nets~\cite{LatNet,LEcEtal22a}.
-QMCPy contains a module to parse the generators from LatNet Builder for compatibility with our LD point generators.
-%contains a \href{https://qmcpy.readthedocs.io/en/latest/demo_rst/latnetbuilder.html}{demo for LatNet Builder}.
-	\item[MATLAB] Commercial software \cite{MAT9.10}, which contains Sobol' and Halton sequences in the Statistics and Machine Learning Toolbox.
-Both generators can be applied in conjunction with MATLAB's Parallel Computing Toolbox to accelerate their execution speed.
-The maximum dimension of the Sobol' sequences is restricted to 1,111, which is relatively small, yet sufficient for most applications.
-	\item[MPS] Magic Point Shop contains lattices and Sobol' sequences in Python, Matlab, and C++ \cite{Nuy17a}. We have used MPS as a starting point for developing LD generators in QMCPy.
-	\item[Owen] Randomized Halton sequences with dimensions up to 1,000 \cite{Owe20a} and scrambled Sobol' sequences with dimensions up to 21,021  \cite{owen1998scrambling, owen2021r}  in R.
-QMCPy currently supports Owen's Halton randomization method, and we plan to implement Owen's nested uniform scrambling for digital nets in the near future.
-	\item[PyTorch] Open-source Python library for deep learning, with scrambled or unscrambled Sobol' sequences \cite{NEURIPS2019_9015,PyTorch}.
-PyTorch enables seamless utilization of Graphics Processing Units (GPUs) or Field Programmable Gate Arrays (FPGAs).
-	\item[QMC.jl] LD Sequences in Julia \cite{Rob20a}. Julia~\cite{bezanson2012julia} is an interpreted language similar to Python and R in terms of ease of use, but is designed to run much faster.
-	\item [qrng] Randomized Sobol', Halton, and Korobov sequences in R~\cite{QRNG2020}. The default Halton randomization in QMCPy utilizes the methods from qrng. 
-	%translated qrng to Python; see  \url{https://github.com/QMCSoftware/qrngpy}.
-The Halton randomization from qrng is also optionally available in QMCPy.
-\AGSComment{We don't really use qrng besides the QRNG Halton randomization.} In addition, the C source files in qrng can be compiled and accessed in QMCPy as one of the backends.
-In some cases, our version of qrng with its compiled code in QMCPy is more efficient than the original software in R \scnote{Why or how?
-By taking advantage of MRG63k3a?}; see the demo, \href{https://qmcpy.readthedocs.io/en/latest/demo_rst/quasirandom_generators.html}{\emph{Quasi-Random Sequence Generator Comparison}}, in the QMCPy documentation~\cite{QMCPyDocs}.
-	\item[SciPy] Scientific computing library in Python with Latin hypercube,  Halton, and Sobol' generators \cite{2020SciPy-NMeth}.
-	\item[TF Quant Finance] Google's Tensorflow  deep-learning library~\cite{tensorflow2015-whitepaper} specialized for financial modeling \cite{tfqf21}.
-The software contains lattice and Sobol' generators alongside algorithms accelerated useing GPUs, FGPAs, or Tensor Processing Units (TPUs).% as opposed to the Central Processing Units (CPUs).
+	\item[BRODA] Sobol' sequences in C, MATLAB, and Excel \cite{BRODA20a},
+	\item[Burkhardt] various QMC software in C++, Fortran, MATLAB, \& Python \cite{Bur20a},
+	\item[LatNet Builder] Generating vectors/matrices for lattices and digital nets \cite{LEcEtal22a,LatNet},
+	\item[MATLAB] Sobol' and Halton sequences, commercial \cite{MAT9.9},
+	\item[MPS] Magic Point Shop, lattices and Sobol' sequences \cite{Nuy17a},
+	\item[Owen] Randomized Halton sequences in R \cite{Owe20a},
+	\item[PyTorch] Scrambled Sobol' sequences \cite{PyTorch},
+	\item[QMC.jl] LD Sequences in Julia \cite{Rob20a}, and
+	\item [qrng]  Sobol', Halton, and Korobov sequences in R \cite{QRNG2020}.
 \end{description}
-Software focusing on QMC cubatures and applications includes the following:
+Software focusing on QMC algorithms and applications includes
 \begin{description}[format=\textup,format=\textbf]
-	\item[GAIL] Guaranteed Automatic Integration Library  contains automatic (Q)MC stopping criteria in MATLAB \cite{ChoEtal21a,HCJJ14}.
-These are iterative procedures for one- or high-dimensional integration that take a user's input error tolerance(s) and determine the number of (Q)MC sampling points necessary to achieve the desired accuracy (almost surely).
-Most of GAIL's (Q)MC functions, some with enhancements, are implemented in Python in QMCPy.
-	\item[ML(Q)MC] Multi-Level (Q)MC routines in C, C++, MATLAB, Python, and R~\cite{GilesSoft}. We have ported ML(Q)MC functions to QMCPy.
-	\item[MultilevelEstimators.jl] ML(Q)MC methods in Julia \cite{Rob21}. The author, Pieterjan Robbe, has contributed cubature algorithms and use cases to QMCPy.
-	\item[OpenTURNS] Open source initiative for the Treatment of Uncertainties, Risks'N Statistics \cite{OpenTURNS} written in C++ and Python, leveraging R statistical packages, as well as LAPACK and BLAS for numerical linear algebra.
-	\item[QMC4PDE] QMC for elliptic PDEs with random diffusion coefficients in Python \cite{KuoNuy16a}.
-	\item[SSJ] Stochastic Simulation with the \texttt{hups} package in Java \cite{SSJ}.
+	\item[GAIL] Automatic (Q)MC stopping criteria in MATLAB \cite{ChoEtal20a},
+	\item[ML(Q)MC] Multi-Level (Quasi-)Monte Carlo routines in C++, MATLAB, Python, and R \cite{GilesSoft},
+	\item[OpenTURNS] Open source initiative for the Treatment of Uncertainties, Risks 'N Statistics in Python \cite{OpenTURNS},
+	\item[QMC4PDE] QMC for elliptic PDEs with random diffusion coefficients \cite{KuoNuy16a},
+	\item[SSJ] Stochastic Simulation in Java \cite{SSJ}, and
 	\item[UQLab] Framework for Uncertainty Quantification in MATLAB \cite{UQLab2014}.
-The core of UQLab is closed source, but a large portion of the library is open source.
-Recently,  UQ[py]Lab, the beta release of UQLab with Python bindings, is  available as Software as a Service (SaaS) via UQCloud~\cite{lataniotis2021uncertainty}.
-
 \end{description}
 
-The sections that follow describe QMCPy \cite{QMCPy2020a}, which is our attempt to establish a framework for QMC software and to combine the best of the above software under a common user interface written in Python 3.
-The choice of language was determined by the desire to make QMC software accessible to a broad audience, especially the technology industry.
+The sections that follow describe QMCPy \cite{QMCPy2020a}, which is our attempt to combine the best of the above software under a common user interface written in Python 3.
+The choice of language was determined by the desire to make QMC software accessible to a broad audience, especially the tech industry.
 
@@
 QMC cubature can be summarized as follows.
-We want to approximate the expectation, $\mu$, well by the sample mean, $\hmu$, where \eqref{eq:integral}, \eqref{eq:fintegral}, and \eqref{eq:samplemean} combine to give
+We want to approximate the integral oe expectation, $\mu$, well by the sample mean, $\hmu$, where \eqref{eq:integral}, \eqref{eq:fintegral}, and \eqref{eq:samplemean} combine to give
 \begin{multline} \label{eq:cubSummary}
-	\mu : = \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  = \bbE[f(\bsX)] = \int_{\cube} f(\bsx) \, \D \bsx \approx \frac 1n \sum_{i=0}^{n-1} f(\bsX_i) =: \hmu, \\
-	 \bsX \sim \calU\cube, \ \bsX_0, \bsX_1, \ldots \Msim \calU\cube.
+	\mu : = \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  = \bbE[f(\bsX)] = \int_\calX f(\bsx) \, \varrho(\bsx) \, \D \bsx \approx \frac 1n \sum_{i=0}^{n-1} f(\bsX_i) =: \hmu, \\
+	 \bsX_0, \bsX_1, \ldots \sim F.
 \end{multline}
-Moreover, we want to satisfy the error requirement in \eqref{eq:err_req}.
 This requires four components, which we implement as QMCPy classes.
@@
 
-	\item[Discrete Distribution]  produces the sequence $\bsX_0, \bsX_1, \dots$ that mimics $\calU[0,1]^d$;
+	\item[Discrete Distribution] that produces $\bsX_0, \bsX_1, \dots$ mimicking the distribution $F$, which typically is $\calU[0,1]^d$;
 
-	\item[True Measure] $\bst \mapsto \lambda (\bst) \D \bst$  defines the original measure, e.g., Gaussian or Lebesgue;
+	\item[True Measure] $\bst \mapsto \lambda (\bst) \D \bst$ that defines the original integral, e.g., Gaussian or Lebesgue;
 
-	\item[Integrand] $g$  defines the original integrand, and $f$ defines the transformed version to fit the \texttt{DiscreteDistribution}; and
+	\item[Integrand] $g$ that  defines the original integral, plus the transformed version, $f$, to fit the discrete distribution; and
 
-	\item[Stopping Criterion] determines how large $n$ should be to ensure that $\QMCPYabs{\mu - \hmu} \le \varepsilon$ as in (\ref{eq:err_req}).
+	\item[Stopping Criterion] that determines how large $n$ should be to ensure that $\QMCPYabs{\mu - \hmu_n} \le \varepsilon$.
 \end{description}
@@
 The software libraries referenced in Section \ref{sec:available} provide one or more of these components.
-QMCPy combines multiple examples of all these components under an object-oriented framework.
+QMCPy combines multiple examples of them all under and object oriented framework.
 Each example is implemented as a concrete class that realizes the properties and methods required by the abstract class for that component.
-The following sections detail descriptions and specific examples for each component.
+The following sections detail descriptions and use cases for each component and select implementations.
 
-Thorough documentation of all QMCPy classes is available in \cite{QMCPyDocs}.
-Demonstrations of how QMCPy works are given in Google Colab notebooks \cite{QMCPyTutColab2020,QMCPyTutColab2020_paper}.
-The project may be installed from PyPI into a Python 3 environment via the command \texttt{pip install qmcpy}.
-In the code snippets that follow, we assume QMCPy has been imported alongside NumPy \cite{numpy} via the following commands in a Python Console:
+Thorough documentation of all QMCPy classes can be found in \cite{QMCPyDocs}.
+Demonstrations of how QMCPy work are given in Google Colab notebooks \cite{QMCPyTutColab2020,QMCPyTutColab2020_paper} \AGSComment{Update \cite{QMCPyTutColab2020_paper}'s link}.
+The project may be installed from PyPI into your Python 3 environment via the command \texttt{pip install qmcpy}.
+In the codes that follows, we assume QMCPy has been imported alongside Numpy \cite{numpy}
+%and Scipy \cite{scipy} 
+via
 \lstinputlisting[style=Python]{QMCSoftwareArticle/python/import.txt}
@@
 
-LD sequences typically mimic $\calU[0,1]^d$.
-Good sequences mimicking other distributions are obtained by transformations as described in the next section.
-We denote by \texttt{DiscreteDistribution} the abstract class containing LD sequence generators.
-In most cases that have been implemented, the points $\bsX_0, \ldots, \bsX_{n-1}$ have an empirical (discrete) distribution that closely approximates the uniform distribution, say, in the sense of discrepancy.
-We also envision future possible \texttt{DiscreteDistribution} objects that assign unequal weights to the sampling points.
-In any case, the term ``discrete'' refers to the fact that these are sequences of points (and weights), not continuous distributions.
+LD sequences typically mimic $\calU[0,1]^d$, which we assume here.  Good sequences for other distributions are obtained by transformations in the next section.
 
-QMCPy implements \emph{extensible} LD sequences, i.e., those that allow practitioners to obtain and use $\bsX_n, \bsX_{n+1}, \ldots $ without discarding $\bsX_0, \ldots, \bsX_{n-1}$.
-Halton sequences do not have preferred sample sizes $n$, but extensible integration lattices and digital sequences in base $b$ prefer $n$ to be a power of $b$.
-For integration lattices and digital sequences, we have focused on base $b=2$ since this is a popular choice and for convenience in generating extensible sequences.
+QMCPy implements \emph{extensible} LD sequences, i.e., those that allow practitioners to obtain and use $\bsX_n, \bsX_{n+1}, \ldots $ without discarding the existing $\bsX_0, \ldots, \bsX_{n-1}$.
+Halton sequences do not have preferred sample sizes $n$, but integration lattices and digital sequences do.
+These latter two also have an elegant group structure, which we summarize in Table \ref{tab:GroupProp}.
+For simplicity we restrict ourselves to the case where the first $n = 2^m$ points of these LD sequences form a group under the addition operator $\oplus$.
 
-Integration lattices and digital sequences in base $2$ have an elegant group structure, which we summarize in Table \ref{tab:GroupProp}.
-The addition operator  is $\oplus$\footnote{The operator $\oplus$ is commonly used to denote exclusive-or, does correspond to its meaning for digital sequences in base $2$.
-However, we are using it here in a more general sense.}, and its inverse is $\ominus$.
-The unshifted sequence is $\bsZ_0, \bsZ_1, \ldots$ and the randomly shifted sequence is $\bsX_0, \bsX_1, \ldots$
-
-\begin{table}[h]
+\begin{table}
 	\centering
@@
 	\multicolumn{2}{c}{
-	\bsZ_{i} := i_0  \bsZ_1 \oplus i_1 \bsZ_{2} \oplus i_2  \bsZ_{4} \oplus  i_3  \bsZ_{8} \oplus \cdots 
+	\bsZ_{i} := i_0  \bsZ_1 \oplus i_1 \bsZ_{2} \oplus + i_2  \bsZ_{4} \oplus  i_3  \bsZ_{8} + \cdots 
 	\quad
@@
 	\text{Rank-1 Integration Lattices} & \text{Digital Nets} \\
-		\bst \oplus \bsx : = (\bst + \bsx) \bmod \bsone & \bst \oplus \bsx := \text{binary digitwise addition}, \oplus_{\dig} \\ 
-		\text{require } \bsZ_{2^{m}} \oplus \bsZ_{2^{m}} = \bsZ_{\lfloor2^{m-1}\rfloor}
-		\quad \forall m \in \bbN_0 
-%		 \\ \text{require } \begin{array}{l} \bsZ_1 = (1/2, \ldots, 1/2) \\
-%		 	\bsZ_{2^{m}} \oplus \bsZ_{2^{m}} = \bsZ_{2^{m-1}} \quad \forall m \in \bbN 
-%	 	\end{array}
-		\\[.5ex]
+		\bst \oplus \bsx : = (\bst + \bsx) \bmod \bsone & \bst \oplus \bsx := \text{binary digitwise addition} \\ 
+		 \text{require } \begin{array}{l} \bsZ_1 = (1/2, \ldots, 1/2) \\
+		 	\bsZ_{2^{m}} \oplus \bsZ_{2^{m}} = \bsZ_{2^{m-1}} \quad \forall m \in \bbN \end{array}
+		\\
 \toprule
@@
 
-The random shift has been turned off above to illuminate the group structure.
-We normally include the randomization to ensure that there are no points on the boundary of $[0,1]^d$.
+The randomized shift has been turned off above to illuminate the group structure.
+In practice, we normally include the randomization to ensure that there are no points on the boundary of $[0,1]^d$.
 Then, when points are transformed to mimic distributions such as the Gaussian, no LD points will be transformed to infinity.
-Turning off the randomization generates a warning when the \texttt{gen\_samples} method is called.
+Turning off the randomization generates a warning.
 
@@
 Using digitwise addition for digital sequences, we can confirm that according to Table \ref{tab:GroupProp},
-\begin{align*}
+\begin{multline*}
 \bsZ_2 \oplus_{\dig} \bsZ_4 = (0.25,0.75)  \oplus_{\dig} (0.125,0.625) \\
-=  ({}_20.010,{}_20.110)  \oplus_{\dig} ({}_20.001,{}_20.101) &= ({}_20.011,{}_20.011) \\
-& = (0.375,0.375) = \bsZ_6.
-\end{align*}
+=  ({}_20.010,{}_20.110)  \oplus_{\dig} ({}_20.001,{}_20.101) = ({}_20.011,{}_20.011) \\
+= (0.375,0.375) = \bsZ_6.
+\end{multline*}
 
-By contrast, if we construct a digital sequence using the generators for the lattice above with $\bsZ_2 = (0.25, 0.75)$, and $\bsZ_4 = (0.125, 0.375)$, we would obtain
+By contrast, if we construct a digital sequence using the generators for the lattice above with $\bsZ_2 = (0.25, 0.75)$, and $\bsZ_4 = (0.125, 0.875)$, we would obtain
 \begin{multline*}
-\bsZ_6 = \bsZ_2 \oplus_{\dig} \bsZ_4   = ({}_20.010,{}_20.110)  \oplus_{\dig} ({}_20.001,{}_20.011)  \\
-= ({}_20.011,{}_20.101) = (0.375, 0.625),
+\bsZ_6 = \bsZ_2 \oplus_{\dig} \bsZ_4   = ({}_20.010,{}_20.110)  \oplus_{\dig} ({}_20.001,{}_20.111)  \\
+= ({}_20.011,{}_20.001) = (0.375, 0.125),
 \end{multline*}
-which differs from the $\bsZ_6=(0.375, 0.125)$ constructed for lattices.
+which differs from the $\bsZ_6=(0.375, 0.625)$ constructed for lattices.
 To emphasize, lattices and digital sequences are different, even if they share the same generators, $\bsZ_1, \bsZ_2, \bsZ_4, \ldots$.
@@
 A crucial difference between IID generators and LD generators is reflected in the behavior of generating $n$ points.
-For an IID generator, asking for $n$ points repeatedly gives different points each time because they are meant to be random and independent.
+For an IID generator, asking for $n$ points repeatedly gives you different points each time because they are meant to be random and independent.
 \lstinputlisting[style=Python]{QMCSoftwareArticle/python/dd_iidone.txt}
@@
 
-On the other hand, for an LD generator, asking for $n$ points repeatedly gives \emph{the same} points each time because they are meant to be the first $n$ points of a specific LD sequence.  
+On the other hand for an LD generator, asking for $n$ points repeatedly gives you \emph{the same} points each time because they are meant to be the first $n$ points of a given LD sequence.  
 \lstinputlisting[style=Python]{QMCSoftwareArticle/python/dd_latticeone.txt}
-Here we allow the randomization so that the first point in the sequence is not the origin.  To obtain the \emph{next} $n$ points, one may specify the start and ending indices of the sequence.
+Here we allow the randomization so that the first point in the sequence is not the origin.  To obtain the \emph{next} $n$ points one may specify the start and ending indices of the sequence.
 \lstinputlisting[style=Python]{QMCSoftwareArticle/python/dd_latticenextone.txt}
@@
 
-\begin{figure}[t]
-	\includegraphics[width=1\textwidth]{QMCSoftwareArticle/vfigs/dd_lattice_successive.eps}
+\begin{figure}
+	\includegraphics[width=1\textwidth]{QMCSoftwareArticle/figs/dd_lattice_successive.png}
 	\qquad
-	\includegraphics[width=1\textwidth]{QMCSoftwareArticle/vfigs/dd_sobol_successive.eps}
-	\caption{Randomized lattice and Sobol' points mimicking a $\calU[0,1]^2$ measure for $n = 64, 128,$ and 256. Note how increasing the number of points evenly fills in the gaps between the points.}
+	\includegraphics[width=1\textwidth]{QMCSoftwareArticle/figs/dd_sobol_successive.png}
+	\caption{Randomized lattice and Sobol' points mimicking a $\calU[0,1]^2$ measure for $n = 64, 128,$ and 256. Note how increasing the number of points evenly fills in the gaps between points.}
 	\label{fig:increase_n}
@@
 
-The LD sequences implemented as \texttt{DiscreteDistribution} objects mimic the $\calU[0,1]^d$ distribution.
+The LD sequences implemented as \text{DiscreteDistribution} objects typically mimic the $\calU[0,1]^d$ distribution.
 However, we may need sequences to mimic other distributions.
 This is implemented via variable transformations, $\bsPsi$.
-In general, if $\bsX \Msim \calU[0,1]^d$, then
+In general, if $\bsX \sim \calU[0,1]^d$, then
 \begin{subequations} \label{eq:exampleVarTrans}
 \begin{gather}
-\bsT = \bsPsi(\bsX) := \bsa  + (\bsb - \bsa) \odot \bsX \Msim  \calU[\bsa,\bsb], \\
+\bsT = \bsPsi(\bsX) := \bsa  + (\bsb - \bsa) \odot \bsX \sim  \calU[\bsa,\bsb], \\
 \label{eq:exampleVarTransGauss}
-\bsT = \bsPsi(\bsX) := \bsa + \mA \bsPhi^{-1}(\bsX)  \Msim \calN(\bsa, \mSigma), \\
+\bsT = \bsPsi(\bsX) := \bsa + \mA \bsPhi^{-1}(\bsX)  \sim \calN(\bsa, \mSigma), \\
 \nonumber  \text{where }  \bsPhi^{-1}(\bsX) : = \begin{pmatrix} \Phi^{-1}(X_1) \\ \vdots \\ \Phi^{-1}(X_d)\end{pmatrix}, \qquad \mSigma = \mA \mA^T,
@@
 Here, $\bsa$ and $\bsb$ are assumed to be finite, and $\Phi$ is the standard Gaussian distribution function.
-Again we use $\Msim$ to denote mimicry, not necessarily in a probabilistic sense.
+Again we use ``$\sim$'' to denote mimicry, not necessarily in a probabilistic sense.
 
-Figure \ref{fig:tm_ug} displays LD sequences transformed as described above to mimic a uniform and a Gaussian distribution.
-The code to generate these points takes the following form for uniform points based on a Halton sequence:
+Figure \ref{fig:tm_ug} displays LD sequences transformed as described above to mimic a uniform and a Gaussian distribution.  The code to generate these points takes the following form of 
 \lstinputlisting[style=Python]{QMCSoftwareArticle/python/tm_uniform.txt}
-whereas for Gaussian points based on a lattice sequence, we have:
+for uniform points based on a Halton sequence, and 
 \lstinputlisting[style=Python]{QMCSoftwareArticle/python/tm_gaussian.txt}
-Here the covariance decomposition $\mSigma = \mA \mA^T$ is done using principal component analysis. The Cholesky decomposition is also available.
+for Gaussian points based on a lattice sequence.
 
-\begin{figure}[t]
-	\includegraphics[width=.45\textwidth]{QMCSoftwareArticle/vfigs/tm_uniform.eps} 
+\begin{figure}
+	\includegraphics[width=.45\textwidth]{QMCSoftwareArticle/figs/tm_uniform.png} 
 	%	\quad
-	\includegraphics[width=.45\textwidth]{QMCSoftwareArticle/vfigs/tm_gaussian.eps}
-	\caption{Halton samples transformed to mimic a uniform $\calU\left( \begin{bmatrix} -2 \\ 0 \end{bmatrix},\begin{bmatrix} 2 \\ 4 \end{bmatrix} \right)$ 
-	%\frednote{This is not correct, we need the upper and lower endponts of a two dimensional interval.  Also, are these Sobol' or Halton as in the code?} 
-	distribution (left) and lattice samples transformed to mimic a Gaussian $\calN\left(\begin{bmatrix} 3 \\ 2 \end{bmatrix}, \begin{bmatrix} 9 & 5 \\ 5 & 4 \end{bmatrix} \right)$ distribution (right).}
+	\includegraphics[width=.45\textwidth]{QMCSoftwareArticle/figs/tm_gaussian.png}
+	\caption{Sobol' samples transformed to mimic a uniform $\calU\left(\begin{bmatrix} -2 \\ 0 \end{bmatrix}, \begin{bmatrix} 2 \\ 4 \end{bmatrix} \right)$ (left) and  Gaussian $\calN\left(\begin{bmatrix} 3 \\ 2 \end{bmatrix}, \begin{bmatrix} 9 & 5 \\ 5 & 4 \end{bmatrix} \right)$ (right).}
 	\label{fig:tm_ug}
@@
 The Brownian motion distribution arises often in financial risk applications.
-Here the $d$ components of the variable $\bsT$ correspond to the discretized Brownian motion at times $\tau/d, 2\tau/d, \ldots, \tau$, where $\tau$ is the time horizon.
+Here the $d$ components of the variable $\bsT$ correspond to the Brownian motion at times $\tau/d, 2\tau/d, \ldots, \tau$, where $\tau$ is the time horizon.
 The distribution is a special case of the Gaussian with covariance
@@
 \end{equation}
-and mean $\bsa$, which  is proportional to the times $(\tau/d)(1, 2, \ldots, d)^T$. The code for generating a Brownian motion is
+and mean $\bsa$, which  correspond to a drift coefficient times $(\tau/d)(1, 2, \ldots, d)^T$. The code for generating a Brownian motion is
 \lstinputlisting[style=Python]{QMCSoftwareArticle/python/tm_brownian_motion.txt}
@@
 
-\begin{figure}[t]
-	\includegraphics[width=1\textwidth]{QMCSoftwareArticle/vfigs/tm_bm.eps} 
-	\caption{Sobol' samples transformed to mimic a 52-dimensional Brownian Motion without drift (left) and with drift coefficient 2 (right).}
+\begin{figure}
+	\includegraphics[width=1\textwidth]{QMCSoftwareArticle/figs/tm_bm.png} 
+	\caption{Sobol' samples transformed to mimic a 32-dimensional Brownian Motion without drift (left) and with drift coefficient 2 (right).}
 	\label{fig:tm_bm}
@@
 
-Let's return  to the integration problem in \eqref{eq:integral}, which we must rewrite as \eqref{eq:fintegral}.
-We choose a transformation of variables defined as $\bst = \bsPsi(\bsx)$ where $\bsPsi:\cube \to \calT$.
+Let's return now to the integration problem in \eqref{eq:integral}, which we must rewrite as \eqref{eq:fintegral}.
+We choose a transformation of variables defined as $\bst = \bsPsi(\bsx)$ where $\bsPsi:\calX \to \calT$.
 This leads to
@@
 	\nonumber 
- \mu &= \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  = \int_{\cube} g\bigl(\bsPsi(\bsx)\bigr) \, \lambda\bigl(\bsPsi(\bsx)\bigr) \,\QMCPYabs{\bsPsi'(\bsx)}  \, \D \bsx =  \int_{\cube}f(\bsx) \, \D \bsx,  \\
+ \mu &= \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  = \int_\calX g\bigl(\bsPsi(\bsx)\bigr) \, \lambda\bigl(\bsPsi(\bsx)\bigr) \,\QMCPYabs{\frac{\partial \bsPsi}{\partial \bsx}} \, \D \bsx =  \int_\calX f(\bsx) \, \varrho(\bsx) \, \D \bsx  \\
  \label{eq:transVar}
-  & \qquad \qquad \text{where } f(\bsx)  = g\bigl(\bsPsi(\bsx)\bigr)  \, \lambda\bigl(\bsPsi(\bsx)\bigr) \,\QMCPYabs{\bsPsi'(\bsx)} ,
+  & \qquad \qquad \text{where } f(\bsx)  = g\bigl(\bsPsi(\bsx)\bigr)  \, \frac{\lambda(\bsPsi(\bsx))}{\varrho(\bsx)} \,\QMCPYabs{\frac{\partial \bsPsi}{\partial \bsx}},
 \end{align}
-and $\QMCPYabs{\bsPsi'(\bsx)} := \QMCPYabs{\partial \bsPsi/\partial \bsx}$ represents the Jacobian of the variable transformation.
+and $\QMCPYabs{\partial \bsPsi/\partial \bsx}$ represents the Jacobian of the variable transformation.
 The abstract class \texttt{Integrand} provides $f$ based on the user's input of $g$ and the \texttt{TrueMeasure} instance, which defines $\lambda$ and the transformation $\bsPsi$.
@@
 
-We illustrate the \texttt{Integrand} class via an example of Keister \cite{Kei96}:
+We illustrate the use of the \texttt{Integrand} class via an example of Keister \cite{Kei96}:
 \begin{equation} \label{eq:KeisterIntegral}
@@
 \end{equation}
-Since $\lambda$ is the density for $\calN(\bszero,\mI/2)$, it is natural to choose $\bsPsi$ according to \eqref{eq:exampleVarTransGauss} with $\mA = \sqrt{1/2} \, \mI$, in which case $\lambda(\bsPsi(\bsx)) \QMCPYabs{\bsPsi'(\bsx)}  = 1$, and so
+We will use QMC methods to approximate this integral, which means that it will be transformed according to \eqref{eq:transVar} for the case where $\calX = [0,1]^d$ and $\varrho(\bsx) = 1$.
+Since $\lambda$ is the density for $\calN(0,\mI/2)$, it is natural to choose $\bsPsi$ according to \eqref{eq:exampleVarTransGauss} with $\mA = \sqrt{1/2} \, \mI$, in which case $\lambda(\bsPsi(\bsx)) \QMCPYabs{\partial \bsPsi/\partial \bsx} / \varrho(\bsx) = 1$, and so
 \[
@@
 Then we evaluate the sample mean of $n=1000$ $f$ values obtained by sampling at transformed Halton points.
-Notice how a two-dimensional Halton generator is used to construct a Gaussian true measure, which is applied alongside the \texttt{my\_Keister} function to instantiate a customized, QMCPy-compatible integrand for this problem.
 \lstinputlisting[style=Python]{QMCSoftwareArticle/python/i_keister.txt}
 We have no indication yet of how accurate our approximation is.
-That topic is treated in the next section.
+This aspect of QMCPy is described in the next section.
 Figure \ref{fig:ikc} visualizes sampling on the original integrand, $g$, and sampling on the transformed integrand, $f$.
 
-\begin{figure}[t]
-	\includegraphics[height=5cm]{QMCSoftwareArticle/vfigs/i_keister_contours.eps}
-	\caption{Right: Sampling the transformed Keister integrand $f$ at Halton points $\bsX_i \LDsim \calU[0,1]^2$.
-Left: Sampling the original Keister integrand $g$ at $\bsT_i = \bsPsi(\bsX_i) \Msim \calN(\bszero,\mI/2)$ where $\bsPsi$ is defined in \eqref{eq:exampleVarTransGauss}.
+\begin{figure}
+	\includegraphics[height=6cm]{QMCSoftwareArticle/figs/i_keister_contours.png}
+	\caption{Right: Sampling the transformed integrand $f$ at Halton points $\bsx \sim \calU[0,1]^2$.
+Left: Sampling the original integrand $g$ at $\bsPsi(\bsx)$ where $\bsPsi$ is defined in \eqref{eq:exampleVarTransGauss}.
 } \label{fig:ikc}
@@
 	& = \int_{\bbR^d} \underbrace{\cos(\lVert \bst \rVert) \exp(-\bst^T \bst) }_{g(\bst)} \, \underbrace{1}_{\lambda(\bst)} \,\D \bst \\
-	& = \int_{[0,1]^d} \underbrace{\cos(\lVert \bsPsi(\bsx) \rVert) \exp(-\bsPsi^T\!\!(\bsx) \bsPsi(\bsx)) \QMCPYabs{\bsPsi'(\bsx)}  }_{f(\bsx)} \, \D \bsx,
+	& = \int_{[0,1]^d} \underbrace{\cos(\lVert \bsPsi(\bsx) \rVert) \exp(-\bsPsi^T\!\!(\bsx) \bsPsi(\bsx)) \QMCPYabs{\frac{\partial \bsPsi}{\partial \bsx}} }_{f(\bsx)} \, \D \bsx,
 \end{align*}
-where $\bsPsi$ is any transformation from $[0,1]^d$ to $\bbR^d$. Now $\lambda$ is \emph{not} a PDF.  QMCPy can perform the cubature this way as well.
+where $\bsPsi$ is any transformation from $\bbR^d$ to $[0,1]^d$.  QMCPy can perform the cubature this way as well.
 \lstinputlisting[style=Python]{QMCSoftwareArticle/python/i_keisterLebesgue.txt}
-The $\bsPsi$ chosen when transforming uniform sequences on the unit cube to fill $\bbR^d$ is  given by \eqref{eq:exampleVarTransGauss} with $\mA = \mI$.
+The $\bsPsi$ chosen when transforming uniform points on the unit cube to fill the $\bbR^d$ is the one given by \eqref{eq:exampleVarTransGauss} with $\mA = \mI$.
 
-In the examples above, one must input the correct $g$ into \texttt{CustomFun} along with the correct \texttt{TrueMeasure} $\lambda$ to define the integration problem.
-The \texttt{Keister} integrand included in the QMCPy library takes a more flexible approach to defining the integration problem $\mu$ in \eqref{eq:KeisterIntegral}.
-Selecting a different \texttt{sampler} $\bsPsi$ performs  \emph{importance sampling}, which leaves $\mu$ unchanged.
+In the examples above, one needed to input the correct $g$ into \texttt{CustomFun} along with the correct \texttt{TrueMeasure} $\lambda$ to define the integration problem.
+The \texttt{Keister} integrand included in the QMCPy library takes a more flexible approach of defining the integration problem $\mu$ in \eqref{eq:KeisterIntegral}.
+Selecting a different \texttt{sampler} $\bsPsi$ performs  \emph{importance sampling} which leaves $\mu$ unchanged.
 \lstinputlisting[style=Python]{QMCSoftwareArticle/python/i_keisterBuiltIn.txt}
-
-In the first case above, the $\lambda$ in \eqref{eq:transVar} corresponds to the Gaussian density  with mean zero and variance $1/2$ by default, and the corresponding variable transformation, $\bsPsi$, is chosen to make $\lambda\bigl(\bsPsi(\bsx)\bigr) \QMCPYabs{\bsPsi'(\bsx)}  = 1$ and $f(\bsx) = g\left(\bsPsi(\bsx) \right)$.
-In the second case,  we choose an importance sampling density $\lambda_{\IMP}$, corresponding  to standard  Gaussian, and the variable transformation $\bsPsi_{\IMP}$ makes $\lambda_\IMP\bigl(\bsPsi_{\IMP}(\bsx)\bigr) \QMCPYabs{\bsPsi'_{\IMP}(\bsx)}  = 1$.
-Then
-\begin{align}
-	\nonumber 
-	\mu &= \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  = \int_\calT g(\bst) \, \frac{\lambda(\bst)}{\lambda_\IMP(\bst) } \, \lambda_\IMP(\bst) \D \bst  \\ 
-	\nonumber
-	& =  \int_{\cube} g\bigl(\bsPsi_\IMP(\bsx)\bigr) \, \frac{\lambda\bigl(\bsPsi_\IMP(\bsx)\bigr)} {\lambda_\IMP\bigl(\bsPsi_\IMP(\bsx)\bigr) } \lambda_\IMP\bigl(\bsPsi_\IMP(\bsx)\bigr) \,\QMCPYabs{\bsPsi'_{\IMP}(\bsx)}  \, \D \bsx \\
-	\nonumber
-	& =  \int_{\cube} f_\IMP(\bsx) \, \D \bsx  \\
-	\label{eq:transVarImp}
-	& \qquad \qquad \text{where } f_\IMP(\bsx)  = g\bigl(\bsPsi_\IMP(\bsx)\bigr)  \,  \frac{\lambda\bigl(\bsPsi_\IMP(\bsx)\bigr)} {\lambda_\IMP\bigl(\bsPsi_\IMP(\bsx)\bigr) }.
-\end{align}
-Because LD samples mimic $\calU\cube$, choosing a different \texttt{sampler} is equivalent to choosing a different variable transform.
 
@@
 The \texttt{StoppingCriterion} object determines the number of samples $n$ that are required for the sample mean approximation $\hmu$ to be within error tolerance $\varepsilon$ of the true mean $\mu$.
-Several QMC stopping criteria have been implemented in QMCPy, including replications, stopping criteria that track the decay of the Fourier complex exponential or Walsh coefficients of the integrand \cite{HicJim16a,HicEtal17a,JimHic16a}, and stopping criteria based on
-Bayesian credible intervals \cite{RatHic19a,JagHic22a}.
-
-The \texttt{CubQMCSobolG} stopping criterion used in the example below assumes the Walsh-Fourier coefficients of the integrand are absolutely convergent.
-The algorithm iteratively doubles the number of samples used in the integration and estimates the error using the decay of the Walsh coefficients \cite{HicJim16a}.
-When the estimated error is below the user-specified tolerance, it finishes the computation and returns the estimated integral.
+Several QMC stopping criteria have been implemented in QMCPy, including replications, stopping criteria that track the decay of the Fourier complex exponential or Walsh coefficients of the integrand \cite{HicJim16a,HicEtal17a,JimHic16a}, and stopping criteria based on Bayesian credible intervals \cite{RatHic19a,JagHic22a}.
 
 Let us return to the Keister example from the previous section.
-After setting up  a default \texttt{Keister} instance via a Sobol' \texttt{DiscreteDistribution}, we choose a \texttt{StoppingCriterion} object that matches the \texttt{DiscreteDistribution} and input our desired tolerance.
+After setting up  a default \texttt{Keister} instance via a Sobol' \texttt{DiscreteDistribution}, we choose a \texttt{StoppingCriterion} object that matches the \texttt{DiscreteDistribution}, inputting our desired tolerance.
 Calling the  \texttt{integrate} method returns the approximate integral plus some useful information about the computation.
@@
 The second output of the stopping criterion provides helpful diagnostic information.
-This computation requires $n=2^{13}$ Sobol' points and $0.009$ seconds to complete.
-The error bound is $0.000637$, which falls below the absolute tolerance.
+This computation required $n=2^{13}$ Sobol' points and $0.028$ seconds to complete.
+The error bound is $0.000517$, which falls below the absolute tolerance.
 
@@
 By contrast, the  error of QMC using LD sequences is $\calO(n^{-1+\epsilon})$, which implies $\calO(\varepsilon^{-1-\epsilon})$ times and sample sizes.
-We see that QMC methods often require orders of magnitude fewer samples than MC methods to achieve the same error tolerance.
+We see that QMC methods often require orders of magnitude fewer samples that MC methods to achieve the same error tolerance.
 
-\begin{figure}[t]
-	\includegraphics[height=6cm]{QMCSoftwareArticle/vfigs/sc_comp.eps}
+\begin{figure}
+	\includegraphics[height=6cm]{QMCSoftwareArticle/figs/sc_comp.png}
 	\caption{Comparison of run times and sample sizes for computing the $5$-dimensional Keister integral \eqref{eq:KeisterIntegral} using IID and LD lattice sequences for a variety of absolute error tolerances.
-The respective stopping criteria are  \texttt{qp.CubMCG} \cite{HicEtal14a} and  \texttt{qp.CubQMCLatticeG} \cite{JimHic16a}.
+The respective stopping criteria are  \texttt{CubMCG} \AGSComment{CITE} and  \texttt{CubQMCLatticeG} \AGSComment{CITE}.
 The LD sequences provide the desired answer much more efficiently.}
@@
 For another illustration of QMC cubature, we turn to pricing an Asian arithmetic mean call option.
-The (continuous-time) Asian option is defined in terms of the average of the stock price, which is written in terms of an integral.
-The payoff of this option is the positive difference between the strike price, $K$, averaged over the time horizon:
+The payoff of this option is the positive difference between the strike price averaged over the time horizon and the strike price, $K$:
 $$
-\text{payoff}(\bsS) = \max\left(\frac{1}{2d}\sum_{j=1}^d (S_{j-1}+S_j)-K, \; 0\right), \quad \bsS = (S_0, \ldots, S_d).
+\text{payoff}(\bsS) = \max\left(\frac{1}{2d}\sum_{j=1}^d (S_{j-1}+S_j)-K,0\right), \quad \bsS = (S_0, \ldots, S_d).
 $$
-Here $S_j$ denotes the asset price at time $\tau j/d$, and a trapezoidal rule is used for discrete approximation of the integral in time that defines the average. 
-The trapezoidal rule is a more accurate approximation to the integral than a rectangle rule. % Comment from Fred
-A basic model for asset prices is a geometric Brownian motion, 
+where $S_j$ denotes the stock price at time $\tau j/d$.  A basic model for stock prices is a geometric Brownian motion, 
 \[
-S_j(\bsT) = S_0 \exp((r - \sigma^2/2) \tau j/d + \sigma T_j),   \;  j = 1, \ldots, d, \; \bsT = (T_1, \ldots, T_d)\sim \calN(\bszero,\mSigma),
+S_j(\bsT) = S_0 \exp((r - \sigma^2) \tau j/d + \sigma T_j),   \;  j = 1, \ldots, d, \; \bsT = (T_1, \ldots, T_d)\sim \calN(0,\mSigma),
 \]
-where $\mSigma$ is defined in \eqref{eq:BMcov}, $r$ is the interest rate, $\sigma$ is the volatility, and $S_0$ is the initial asset price.
+where $\mSigma$ is defined in \eqref{eq:BMcov}, $r$ is the interest rate, $\sigma$ is the volatility, and $S_0$ is the initial price.
 The fair price of the option is then the expected value of the discounted payoff, namely,
@@
 
-The following code utilizes QMCPy's Asian option \texttt{Integrand} object to approximate the value of an Asian call option for a particular choice of the parameters.
+The following code utilizes QMCPy's Asian option \texttt{Integrand} object to approximate the Asian call option for a particular choice of the parameters.
 \lstinputlisting[style=Python]{QMCSoftwareArticle/python/sc_aco.txt}
-Because this \texttt{Integrand} object has the built-in Brownian motion \texttt{TrueMeasure}, one only need provide the LD sampler.
+Because this \texttt{Integrand} object has the built-in Brownian motion \texttt{TrueMeasure}, one only neeed provide the LD sampler.
 
-Out of the money option price calculations can be sped up by adding an upward drift to the Brownian motion.
-The upward drift produces more in the money paths and also reduces the variation or variance of the final integrand, $f$.
+Out of the money option price calculations can be sped up by adding a drift to the Brownian motion that produces more in the money paths.
 This is a form of importance sampling.
@@
 \lstinputlisting[style=Python]{QMCSoftwareArticle/python/wo_imp_samp_aco.txt}
-Adding the upward drift gives us the answer faster:
+Adding the drift gives us the answer faster:
 \lstinputlisting[style=Python]{QMCSoftwareArticle/python/w_imp_samp_aco.txt}
@@
 
-The improvement in time is less than that in $n$ because the integrand is more expensive to compute when the drift is employed.
-Referring to \eqref{eq:transVarImp}, in the case of no drift, the $\lambda$ corresponds to the density for the discrete Brownian motion, and the variable transformation $\bsPsi$ is chosen so that $f(\bsx) = g\left(\bsPsi(\bsx) \right)$.
-However, in the case of a drift, the integrand becomes  $f_\IMP(\bsx)  = g\bigl(\bsPsi_\IMP(\bsx)\bigr)  \lambda\bigl(\bsPsi_\IMP(\bsx)\bigr)/\lambda_\IMP\bigl(\bsPsi_\IMP(\bsx)\bigr)$, which requires more computation time per integrand value.
+The improvement in time is less than improvement in $n$ because the integrand $f$ is more expensive to compute when the drift is added.
+In the case of no drift, the $\lambda$ in \eqref{eq:transVar} corresponds to the density for the discrete Brownian motion, and the variable transformation $\bsPsi$ is chosen so that $\lambda\bigl(\bsPsi(\cdot)\bigr) \QMCPYabs{\partial \bsPsi/\partial \bsx} = \varrho(\bsx) = 1$ and $f(\bsx) = g\left(\bsPsi(\bsx) \right)$.
+However, if we choose $\lambda_{\IMP}$ as our importance sampling density and $\bsPsi_{\IMP}$ as the variable transformation corresponding to $\lambda_\IMP\bigl(\bsPsi_\IMP(\cdot)\bigr) \QMCPYabs{\partial \bsPsi_\IMP/\partial \bsx} = \varrho(\bsx) = 1$, then
+\begin{align}
+	\nonumber 
+	\mu &= \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  = \int_\calT g(\bst) \, \frac{\lambda(\bst)}{\lambda_\IMP(\bst) } \, \lambda_\IMP(\bst) \D \bst  \\ 
+	\nonumber
+	& =  \int_\calX g\bigl(\bsPsi_\IMP(\bsx)\bigr) \, \frac{\lambda\bigl(\bsPsi_\IMP(\bsx)\bigr)} {\lambda_\IMP\bigl(\bsPsi_\IMP(\bsx)\bigr) } \lambda_\IMP\bigl(\bsPsi_\IMP(\bsx)\bigr) \,\QMCPYabs{\frac{\partial \bsPsi_\IMP}{\partial \bsx}} \, \D \bsx \\
+	\nonumber
+	& =  \int_\calX f_\IMP(\bsx) \, \varrho(\bsx) \, \D \bsx  \\
+	\label{eq:transVarImp}
+	& \qquad \qquad \text{where } f_\IMP(\bsx)  = g\bigl(\bsPsi_\IMP(\bsx)\bigr)  \,  \frac{\lambda\bigl(\bsPsi_\IMP(\bsx)\bigr)} {\lambda_\IMP\bigl(\bsPsi_\IMP(\bsx)\bigr) }
+\end{align}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=1\textwidth]{QMCSoftwareArticle/figs/i_aco.png} 
+	\caption{Left: Brownian Motion paths with drift $a=2$. Right: Corresponding Asian Call option paths with $S_0=30$, $K=45$, $\sigma=0.5$, and $r=0$.}
+	\label{fig:aco}
+\end{figure}
+
+Figure \ref{fig:aco} shows drifted Brownian motion paths and their corresponding option pricing paths. 
+\AGSComment{Bayesian stopping criterion}
 
@@
 
-In this section, we look at the inner workings of QMCPy and point out features we hope will benefit the community of QMC researchers and practitioners.
-We also highlight important nuances of QMC methods and how QMCPy addresses these challenges.
-For details, readers should refer to the QMCPy documentation \cite{QMCPyDocs}.
+In this section, we look at the inner workings of QMCPy and points out available features we hope will benefit the community of QMC researchers and practitioners.
+We also highlight important idiosyncrasies of QMC methods and how QMCPy addresses these challenges.
 
-\subsection{LD Sequences}
+\subsection{Link to LatNetBuilder}
 
-LD sequences are the backbone of QMC methods. QMCPy provides generators that combine research from across the QMC community to enable advanced features and customization options.
+QMCPy provides users access to high quality lattice and digital net generators (defined in Table \ref{tab:GroupProp}).
+A low-discrepancy generator is comprised of two parts: the static generating vectors $\bsZ_1,\bsZ_2,\bsZ_4, \dots \in [0,1)^d$ and the callable generator function.
+QMCPy integrates nicely with latnetbuilder \AGSComment{CITE}, a package centered around search algorithms to find high quality generating vectors.
+QMCPy's lattice and digital net generators can accept a vectors from latnetbuilder and use them to generate custom ordinary lattice, polynomial lattice, polynomial net, or digital net sequence.
+We defer readers to latnetbuilder and QMCPy documentation \AGSComment{CITE BOTH} for details.
 
-Two popular LD sequences are integration lattices and digital nets which we previously outlined in Table \ref{tab:GroupProp}.
-These LD generators are comprised of two parts: the static generating vectors $\bsZ_1,\bsZ_2,\bsZ_4, \ldots \in [0,1)^d$ and the callable generator function.
-By default, QMCPy provides a number of high-quality generating vectors for users to choose from.
-For instance, the default ordinary lattice vector was constructed by Cools, Kuo, and Nuyens \cite{doi:10.1137/06065074X} using component-by-component search, is extensible, has order-2 weights, and supports up to 3600 dimensions and $2^{20}$ samples.
-%https://web.maths.unsw.edu.au/~fkuo/lattice/index.html
-%https://web.maths.unsw.edu.au/~fkuo/sobol/
-However, users who require more samples but fewer dimensions may switch to a generating vector constructed using LatNet Builder \cite{LatNet,LEcEtal22a} to support 750 dimensions and $2^{24}$ samples.
-Moreover, the \texttt{qp.Lattice} and \texttt{qp.DigitalNet} objects allow users to input their own generating vectors to produce highly customized sequences.
-To find such vectors, we recommend using LatNet Builder's construction routines as the results can be easily parsed into a QMCPy-compatible format.
+\subsection{Customization Options for Low Discrepancy Generators}
 
-Along with the selection of a generating vector, QMCPy's low discrepancy sequence routines expose a number of other customization parameters.
-For instance, the lattice generator extends the Magic Point Shop \cite{Nuy17a} to support either linear or natural ordering.
-Digital sequences permit either standard or Gray code ordering and may be randomized via a digital shift optionally combined with a linear scrambling \cite{Mat98}.
-Halton sequences may be randomized via the routines of either Owen \cite{Owe20a} or Hofert and Lemieux \cite{QRNG2020}.
+QMCPy's low discrepancy generator routines also support numerous customizations for users who need specific points sets, orders, or randomizations.
+For example, our in-house developed digital net generator allows randomizations such as digital shifts and linear matrix scrambles, optional graycode orderings, and an optionally non-zero starting dimension.
+Moreover, while most lattice and digital net generators use a fixed generating vector, QMCPy allows users to input their own vector for their specific need, as detailed in the previous subsection.
 
-\subsection{A Word of Caution When Using LD Sequences}
+\subsection{Preferred sample sizes for low-discrepancy generators and}
 
-Although QMCPy's \texttt{DiscreteDistribution}s have many of the same parameters and methods, users should be careful when swapping IID sequences with LD sequences.
-While IID node-sets have no preferred sample size, LD sequences often require special sampling ranges to ensure optimal discrepancy.
-As mentioned  earlier, base-2 digital nets and extensible integration lattices show better evenness for sample sizes that are powers of 2.
-On the other hand, the preferred sample sizes for $d$-dimensional Halton sequences are $n = \prod_{j=1}^d p^{m_j}_j$ where $p_j$ is the $j^{\text{th}}$ prime number and $m_j \in \mathbb{N}_0$ for $j=1,\dots,d$.
-Due to the infrequency of such values, Halton sequences are often regarded as not having a preferred sample size.
+Low-discrepancy sequences differ from IID node sets in that they fill the domain in a more uniform manner.
+As IID nodes are independent, they do not have a preferred sample size.
+That is, if $\bsX \IIDsim \calU[0,1]^d$ then each node is equally likely to fall  anywhere in the unit cube.
+On the other hand, low-discrepancy sequences provide guarantees that the integration domain is more evenly sampled.
+However, these guarantees only occur at specific sample sizes.
+For example, lattice and digital net sequences only fill the domain evenly when $n$ is a power of $2$.
 
-Users may also run into trouble when trying to generate too many points.
-Since QMCPy's generators construct sequences in 32-bit precision, generating greater than $2^{32}$ consecutive samples will cause the sequence to repeat.
-In the future, we plan to expand our generators to support optional 64-bit precision at the cost of greater computational overhead.
+\subsection{\texttt{print}ing  QMCPy Instances}
 
-Another subtlety arises when transforming LD sequences to mimic different distributions.
-As mentioned earlier, unrandomized lattice and digital sequences include the origin, making transformations such as \eqref{eq:exampleVarTransGauss} produce infinite values.
+A helpful features of QMCPy is the ability to print out components instances to see more detailed information.
+This allows users to quickly see the properties that have been set and defaulted for any given component.
+This feature was already used to print the data object returned by the stopping criterion's \texttt{integrate} method in Section \ref{sec:stopping_crit}.
+The following code prints a drifted \texttt{BrownianMotion} instance and displays the connection to an underlying Gaussian measure.
+\lstinputlisting[style=Python]{QMCSoftwareArticle/python/print_bm.txt}
 
-Some popular implementations of LD sequences drop the first point, which is the origin in the absence of randomization.
-The rationale is to avoid the transformation of the origin to infinity when mimicking a Gaussian or other distribution with an infinite sample space.
-Unfortunately, dropping the first point destroys some of the nice properties of the first $n = 2^m$ points of LD sequences, which can degrade the order of convergence for QMC cubature.
-A careful  discussion of this matter is given by~\cite{owen2020dropping}.
-
-\subsection{Transformations}
-
-The transformation $\bsPsi$  connects a \texttt{DiscreteDistribution} and \texttt{TrueMeasure}.
-So far, we have assumed the \texttt{DiscreteDistribution} mimics a $\calU[0,1]^d$ distribution with PDF $\varrho(\bsx)=1$.
-However, it may be advantageous to utilize a \texttt{DiscreteDistribution} that mimics a different distribution.
-%For instance, QMCPy includes a \texttt{DiscreteDistribution} corresponding to an IID standard Gaussian sequence.
-It may be transformed to mimic a general Gaussian distribution \texttt{TrueMeasure} while avoiding the inverse CDF transform required for nodes mimicking $\calU[0,1]^d$.
-% the IID standard Gaussian samples still use inverse CDF, so we don't actually save anything aside from maybe a minor speed up from numpy's faster code
-
-Suppose we have a \texttt{DiscreteDistribution} mimicking  density $\varrho$ supported on $\calX$.  Then using the variable transformation $\bsPsi$, 
-\begin{equation*}
-	\mu = \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  =  \int_{\calX}f(\bsx) \, \varrho(\bsx) \D \bsx \quad \text{for }
-   f(\bsx)  = g\bigl(\bsPsi(\bsx)\bigr)  \, \frac{\lambda\bigl(\bsPsi(\bsx)\bigr)}{\varrho(\bsx)} \, \lvert \bsPsi'(\bsx)\rvert,
-\end{equation*}
-which generalizes \eqref{eq:transVar}.
-QMCPy also includes support for successive changes of measures so users may build complex variable transformations in an intuitive manner.
-Suppose that the variable transformation is a composition of several transformations: $\bsPsi=\widehat{\bsPsi}_L = \bsPsi_L \circ \bsPsi_{L-1} \circ \dots \circ \bsPsi_1$ as in \eqref{eq:transVarImp}.
-Here, $\bsPsi_l:\calX_{l-1} \to \calX_l$,  $\calX_0=\calX$, and $\calX_L=\calT$ so that the transformations are compatible with the \texttt{DiscreteDistribution} and \texttt{TrueMeasure}.
-Let $\widehat{\bsPsi}_l=\bsPsi_l \circ \bsPsi_{l-1} \circ \dots \circ \bsPsi_1$ denote the composition of the first $l$ transforms and assume that $\widehat{\bsPsi}_0(\bsx)=\bsx$, the identity transform.
-Then we may write $\mu =  \int_{\calX}f(\bsx) \, \varrho(\bsx) \D \bsx$ for
-\begin{equation*}
-    f(\bsx) 
- = g\bigl(\widehat{\bsPsi}_L(\bsx)\bigr)\frac{\lambda\bigl(\widehat{\bsPsi}_L(\bsx)\bigr)}{\varrho(\bsx)}\prod_{l=1}^L \bigl\lvert \bsPsi_l'\bigl(\widehat{\bsPsi}_{l-1}(\bsx)\bigr) \bigr\rvert .
-\end{equation*}
-It is often the case that $\bsPsi_l$ is chosen such that $\bsPsi_l(\bsX)$ is stochastically equivalent to a random variable with density $\lambda_l$ on sample space $\calX_l$ when $\bsX$ is a random variable with density $\varrho_l$ on sample space $\calX_{l-1}$.
-This implies  $\varrho_l(\bsx) = \lambda_l(\bsPsi_l(\bsx))\lvert  \bsPsi'_l (\bsx) \rvert$ so that
-\begin{equation*}
-f(\bsx) 
- = g\bigl(\widehat{\bsPsi}_L(\bsx)\bigr)\frac{\lambda\bigl(\widehat{\bsPsi}_L(\bsx)\bigr)}{\varrho(\bsx)}\prod_{l=1}^L \frac{\varrho_l(\widehat{\bsPsi}_{l-1}(\bsx))}{\lambda_l(\widehat{\bsPsi}_l(\bsx))} .
-\end{equation*}
-
-For an example, we return to the Keister integral \eqref{eq:KeisterIntegral}.
-The following code constructs three \texttt{Keister} instances: one without importance sampling, one importance sampled by a Gaussian distribution, and one importance sampled by the composition of a Gaussian distribution with a Kumaraswamy distribution \cite{Kumaraswamy}.
-All \texttt{Integrand}s use a Sobol' \texttt{DiscreteDistribution}, making $\varrho(\bsx)=1$ and $\calX = [0,1]^d$.
-The \texttt{TrueMeasure} is $\calN(\bszero,\mI/2)$ making $\lambda(\bst) = \pi^{-d/2} \exp(-\bst^T \bst)$ and $\calT = \bbR^d$.
-
-The table below displays the variable transformations and the measures for these three cases.
-In all cases $\varrho_1(\bsx) = \cdots = \varrho_L(\bsx) = 1$ because the $\bsPsi_l$ utilize inverse cumulative distributions.
-
-\[\begin{array}{ccccccc}
-	\texttt{Integrand} & L & \lambda_1 & \bsPsi_1 & \lambda_2 & \bsPsi_2 & f \tabularnewline
-	\toprule
-	\texttt{K} & 1 & \calN(\bszero, \mI/2) & \eqref{eq:exampleVarTransGauss}& && g(\bsPsi_1(\cdot)) \tabularnewline [1ex]
-	\texttt{K\_gauss} & 1 & \calN(\bszero,3 \mI/4) & \eqref{eq:exampleVarTransGauss} & && 
-	\displaystyle g(\bsPsi_1(\cdot)) \frac{\lambda(\bsPsi_1(\cdot))}{\lambda_1(\bsPsi_1(\cdot))} \tabularnewline [2.5ex]
-	\texttt{K\_gauss\_kuma} & 2 & \Kum & \bsF^{-1}_\Kum & \calN(\bszero,\mI) &  \eqref{eq:exampleVarTransGauss} &
-	\displaystyle \frac{g(\bsPsi_2(\bsPsi_1(\cdot))) \lambda(\bsPsi_2(\bsPsi_1(\cdot))) }{\lambda_1(\bsPsi_1(\cdot)) \lambda_2(\bsPsi_2(\bsPsi_1(\cdot)))} 
-   \\ \bottomrule
-\end{array}
-\]
-
-Here $\Kum$ denotes the multivariate Kumaraswamy distribution with independent marginals, and $ \boldsymbol{F}^{-1}_\Kum$ denotes the element-wise inverse cumulative distribution function.
-The following code evaluates the Keister integral \eqref{eq:KeisterIntegral} for $d=1$ and error tolerance $\varepsilon = 5 \times 10^{-8}$.
-The timings for each of these different integrands are displayed.
-
-\lstinputlisting[style=Python]{QMCSoftwareArticle/python/composed_IS.txt}
-
-%Successful importance sampling places more points where the original integrand, $g$, varies more.  Equivalently, 
-Successful importance sampling makes the transformed integrand, $f$, more flat.
-The shorter cubature times correspond to flatter integrands, as  illustrated in Figure \ref{fig:mIS}.
-The above example uses $d=1$ to facilitate the plot in Figure \ref{fig:mIS}; however, the same example works for arbitrary dimensions.
-\begin{figure}[t]
-    \centering
-	\includegraphics[width=.8\textwidth]{QMCSoftwareArticle/vfigs/multiple_IS.eps}
-	\caption{Keister functions with and without importance sampling (\texttt{IS}).
-Note that the Keister functions using importance sampling are generally less variable and therefore easier to integrate, as evidenced by the faster integration times.} \label{fig:mIS}
-\end{figure}
-
-Automatic Bayesian stopping criteria choose $n$ needed to satisfy the error tolerance via a Bayesian-credible interval for $\mu$ \cite{RatHic19a,JagHic22a}.  These are also options in QMCPy.
+\AGSComment{
+\begin{itemize}
+    \item we have better generating vector (under latnetbuilder sub-section)
+    \item transform from low discrepancy does not guarantee low discrepancy in new distribution
+    \item Sobol' samples will repeat after $2^{32}$ samples
+    \item multi-level stopping criterion
+    \item PCA and Cholesky decomposition for Gaussian (and Brownian Motion which is a subclass of Gaussian)
+\end{itemize}}
 
@@
 
-QMCPy is ripe for growth and development in several areas.  We hope that the QMC community will join us in making this a reality.
-
-Multi-level (quasi-)Monte Carlo (ML(Q)MC) methods make possible the computation of expectations of functionals of stochastic differential equations and partial differential equations with random coefficients.
-Such use cases appear in quantitative finance and geophysical applications.
-QMCPy's ML(Q)MC's capability is rudimentary, but under active development.
-
-We hope to add a greater variety of use cases and are engaging collaborators to help.
-Sobol' indices, partial differential equations with random coefficients, expected improvement measures for Bayesian optimization, and multivariate probabilities are some of those on our radar.
-
-Recently, several QMC experts have focused on developing LD generators for Python. 
-Well-established packages such as SciPy \cite{SCIPY} and PyTorch \cite{PyTorch} are have developed QMC modules that support numerous LD sequences and related functionalities.
-We plan to integrate the routines as optional backends for QMCPy's LD generators.
-Creating ties to these other packages will allow users to call their preferred generators from within the QMCPy framework.
-Moreover, as features in QMCPy become more common and prove their value, we will try to incorporate them into SciPy and other popular, general-purpose packages.
-
-We also plan to expand our library of digital net generating matrices.
-We wish to incorporate interlaced digital nets, polynomial lattices, and Niederreiter sequences, among others.
-By including high-quality defaults in QMCPy, we hope to make these sequences more readily available to the public.
-
-Our \texttt{DiscreteDistrution} places equal weights on each support point, $\bsX_i$.  In the future, we might generalize this to unequal weights.
-
-QMCPy already includes importance sampling, but the choice of sampling distribution must be chosen a priori.
-We would like to see an automatic, adaptive choice following the developments of \cite{AsmGly2007a,approx_zero_variance_simulation,OweZho00a}.
-
-Control variates can be useful for QMC as well as for IID MC \cite{HicEtal03}.  These should be incorporated into QMCPy in a seamless way.
-
-We close with an invitation.
-Try QMCPy.
-If you find bugs or missing features, please submit an issue to \href{https://github.com/QMCSoftware/QMCSoftware/issues}{\nolinkurl{https://github.com/QMCSoftware/QMCSoftware/issues}}.
-If you wish to add your great algorithm or use case, please submit a pull request to our GitHub repository at \href{https://github.com/QMCSoftware/QMCSoftware/pulls}{\nolinkurl{https://github.com/QMCSoftware/QMCSoftware/pulls}}.
-We hope that the community will embrace QMCPy.
+\AGSComment{
+\begin{itemize}
+    \item Better multi-level stopping criterion (Pieterjan)
+    \item Multi-composed transforms $\bsPsi$ (multiple Jacobian factors)
+    \item Niederreiter sequences (Adrian and Onyekachi)
+    \item Sobol' indicies (Chris Hoyt)
+    \item transformations for discrete distributions that do not mimic $\calU[0,1]^d$
+    \item control variates
+    \item vectorized integrals
+\end{itemize}}
 
@@
 The authors would like to thank the organizers for a wonderful MCQMC 2020. 
-We also thank the referees for their many helpful suggestions.  This work is supported in part by SigOpt and National Science Foundation grant DMS-1522687.
+We also thank the referees for their many helpful suggestions.  This work is supported in part by National Science Foundation grants DMS-1522687 and SigOpt.
 \end{acknowledgement}
@@
 \bibliographystyle{spmpsci.bst}
-\bibliography{FJH23,FJHown23,QMCSoftwareArticle,sc}
+\bibliography{FJH23,FJHown23,QMCSoftwareArticle}
 
-\iffalse
 \AGSComment{
@@
 \subsection{Chaining Transforms}
-With $k$ transforms, let $\widehat{\bsPsi}_j(\bsx) = (\bsPsi_j \circ \bsPsi_{j-1} \circ \dots \circ \bsPsi_1)(\bsx)$ denote the first $j$ compositions and $\bst = \widehat{\bsPsi}_k(\bsx)$ be the complete transformation.
+With $k$ transforms, let $\hat{\bsPsi}_j(\bsx) = (\bsPsi_j \circ \bsPsi_{j-1} \circ \dots \circ \bsPsi_1)(\bsx)$ denote the first $j$ compositions and $\bst = \hat{\bsPsi}_k(\bsx)$ be the complete transformation.
 Then,
-$$f(\bsx) = g(\widehat{\bsPsi}_k(\bsx)) \frac{\lambda(\widehat{\bsPsi}_k(\bsx))}{\varrho(\bsx)} \prod_{i=1}^{k} \left\lvert \frac{\D \, \bsPsi_i}{\D \, \widehat{\bsPsi}_{i-1}(\bsx)} \right\rvert$$
-where $\widehat{\bsPsi}_0(\bsx) = \bsx$.
+$$f(\bsx) = g(\hat{\bsPsi}_k(\bsx)) \frac{\lambda(\hat{\bsPsi}_k(\bsx))}{\varrho(\bsx)} \prod_{i=1}^{k} \left\lvert \frac{\D \, \bsPsi_i}{\D \, \hat{\bsPsi}_{i-1}(\bsx)} \right\rvert$$
+where $\hat{\bsPsi}_0(\bsx) = \bsx$.
 }
@@
 
-\AGSComment{
-\subsection{\texttt{print}ing  QMCPy Instances}
-
-QMCPy provides the a \texttt{print} command to display information about instances of QMCPy classes.
-In the sample code in Section \ref{sec:stopping_crit} this command was used show the output of a \texttt{StoppingCriteroion}'s \texttt{integrate} method.
-
-faulted for any given component.
-This feature was already used to print the data object returned by the stopping criterion's \texttt{integrate} method in Section \ref{sec:stopping_crit}.
-The following code prints a drifted \texttt{BrownianMotion} instance and displays the connection to an underlying Gaussian measure.
-\lstinputlisting[style=Python]{QMCSoftwareArticle/python/print_bm.txt}}
-
-\AGSComment{Do we want the implementation to follow the first or second equation?
-QMCPy current follows first equation (directly computes Jacobians), which may allow for transforms that do not correspond to PDFs.
-However, the second equation would be easier for QMCPy contributors who then do not need to compute Jacobians for their transforms.
-May also be computationally cheaper i.e.
-Gaussian Jacobian requires PDF and inverse CDF computations, where the second equation would save the inverse CDF computation.}
-
-\fi
-
 \end{document}

### 2022-10-31 contributed-books-2/guideline/quickstart.tex
Updates from Overleaf

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a contributed volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[graybox]{svmult}

% choose options for [] as required from the list
% in the Reference Guide

\usepackage{type1cm}        % activate if the above 3 fonts are
\usepackage{nicefrac}
                            % not available on your system
%
\usepackage{makeidx}         % allows index generation
\usepackage{graphicx}        % standard LaTeX graphics tool
                             % when including figure files
\usepackage{multicol}        % used for the two-column index
\usepackage[bottom]{footmisc}% places footnotes at page bottom

\usepackage{newtxtext}       % 
\usepackage[varvw]{newtxmath}       % selects Times Roman as basic font

\usepackage{hyperref}
\usepackage{cprotect}
\def\ttdefault{cmtt}

\pagestyle{plain}

% see the list of further useful packages
% in the Reference Guide

\makeindex             % used for the subject index
                       % please use the style svind.ist with
                       % your makeindex program

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\parindent=0pt%
\parskip=1em%
\raggedbottom%

\def\thechapter{\vspace*{-2pc}}
\def\chaptername{}
\begin{document}

\title{1 Quick Start -- SVMult}

\author{}

\maketitle

\begin{refguide}

\begin{sloppy}

\def\thechapter{\arabic{chapter}}

\vspace*{-13pc}

\section{Initializing the Class}

To format a {\it document for a contributed book} enter
\cprotect\boxtext{\verb|\documentclass{svmult}|}

\vspace*{-5pc}
\hspace*{28pc}\,{\it Tip}: \\
\hspace*{28pc} \hbox{Use the pre-set}\\
\hspace*{28pc} \hbox{templates}

\bigskip at the beginning of your root file. This will set the text area to a \verb|\textwidth| of 117 mm or 27-3/4 pi pi and a \verb|\textheight| of 191 mm or 45-1/6 pi plus a \verb|\headsep| of 12 pt (space between the running head and text).

{\it N.B.} Trim size (physical paper size) is $155 \times 235$ mm or $6\nicefrac{1}{8} \times 9\nicefrac{1}{4}$ in.

For a description of all possible class options provided by {\sc SVMult} see the ``{\sc SVMult} Class Options'' section in the enclosed {\it Reference Guide}.

\section{Required Packages}
The following selection has proved to be essential in preparing a manuscript in the Springer Nature layout.

Invoke the required packages with the command

\cprotect\boxtext{\verb|\usepackage{}|}

\begin{tabular}{p{7.5pc}@{\qquad}p{18.5pc}}
{\tt newtxtext.sty} and {\tt newtxmath.sty} & Supports roman text font provided by a Times clone,  sans serif based on a Helvetica clone,  typewriter faces,  plus math symbol fonts whose math italic letters are from a Times Italic clone\\
{\tt makeidx.sty} &  provides  and interprets the command  \verb|\printindex|  which ``prints'' the index file *.ind (compiled by an index processor) on a chosen page\\
{\tt graphicx.sty} & is a powerful tool for including, rotating, scaling and sizing graphics files (preferably *.eps files)\\
{\tt multicol.sty} & balances out the columns on the last page of, for example, your subject index\\
{\tt footmisc.sty}  & together with style option {\tt [bottom]} places all footnotes at the bottom of the page
\end{tabular}

For a description of other useful packages and {\sc SVMult} class options, special commands and environments tested with the {\sc SVMult} document class see the {\it Reference Guide}.

For a detailed description of how to fine-tune your text, mathematics, and references, of how to process your illustrations, and of how to set up your tables, see the enclosed {\it Author Instructions}.

\section{Important Notes for the Production Process}

Authors can support and speed up the production process by following the below described instructions.

\subsection*{General instructions}

\begin{itemize}
\item Please make sure that any package used in the LaTeX file is compatible with MikTeX2.4/TeXlive2019 or higher.

\item Along with the PDF we are also publishing your LaTeX works in other digital/online versions (in html and ePub). Point to highlight: All LaTeX flexible works to achieve a specific PDF output will not result in the same output in other online/digital versions due to limitations, e.g. dsfonts (double stroke fonts package), eucal (Euler calligraphic font package) --- the result of these packages in PDF will not show the same result in html/ePub. It is recommended to use \verb|\mathbb{...}| and to use \verb|\mathcal|, please do not include the package eucal.

\item Please avoid multiple levels of linked sub-files in chapters: Too many linked files within chapter/book level, too many tree directories within the package will lead to additional work and challenges during the production process.

\item The introduction of multiple math related packages/commands in the same manuscript (e.g. bbm, dsfonts, etc. and \verb|\mathrm|, \verb|\mathsf|, \verb|\mathtt|, \verb|\mathfrak|, \verb|\mathcal|, \verb|\mathbb|, \verb|\mathbbm|, \verb|\mathds| etc.) or the definition of different outputs for the same math command in different chapters (e.g. \verb|\mathcal{X}| as calliography in chapter one and \verb|\mathcal{X}| as fracture or script in chapter two) will lead to difficulties during proof generation. Instead please use default commands (as e.g. \verb|\mathcal| and \verb|\mathbb|) and use these commands consistently throughout all chapters.

\item Usage of nested level macros, improper grouping, introduce \$ within equations (e.g. \verb|$ X + Y = 0 = \hbox{Revenge $X$} $|) will slow down the proofing works and will take additional efforts to produce a quality result and in some cases may lead to author queries and additional corrections.

\item Any color text/shades/backgrounds which are handled in RGB mode will also need appropriate CMYK values in the manuscript. Please include them both one in active and other in comment is also fine.

\noindent
E.g. 
\begin{center}
\verb|	\definecolor{ultramarine}{RGB}{1,1,1}|

\verb|	%%\definecolor{ultramarine}{cmyk}{0,0,0,1}|

\verb|	        \textcolor{ultramarine}{Colored text}|
\end{center}

\noindent
This eliminates incompatible color issues when we switch the color mode during print/online proof generation. Please note that owing to technical restrictions colored texts cannot be included in html and ePub. The colored text will have to be included as images for these online formats.

\end{itemize}

\subsection*{Package}

\subsubsection*{Algorithm}

To typeset algorithms or pseudocode in LaTeX you can use one of the following options:

\begin{itemize}
\item Choose ONE of the (\texttt{algpseudocode} OR \texttt{algcompatible} OR \texttt{algorithmic}) packages to typeset algorithm bodies, and the algorithm package for captioning the algorithm.

\item The \texttt{algorithm2e} package.
\end{itemize}

\noindent
\textbf{Important: Please choose only one of the above groups of packages, and use only the commands and syntax provided by the package you choose}. These packages cannot be loaded simultaneously; otherwise you will get lots of errors.

\subsubsection*{amsmath} 
We recommend not to use ``amsmath'' package at preamble separately when you use ``newtxmath''. It is inbuilt within the package ``newtxmath''.

\end{sloppy}

\end{refguide}

\end{document}

### 2022-10-31 Old Stuff/QMCSoftwareArticle_VersionB.tex
Office Computer Updates

\documentclass[graybox]{svmult}

% choose options for [] as required from the list
% in the Reference Guide

\usepackage{type1cm}        % activate if the above 3 fonts are
                            % not available on your system
%
\usepackage{makeidx}         % allows index generation
\usepackage{graphicx}        % standard LaTeX graphics tool
                             % when including figure files
\usepackage{multicol}        % used for the two-column index
\usepackage[bottom]{footmisc}% places footnotes at page bottom

\usepackage{newtxtext}       % 
\usepackage[varvw]{newtxmath}       % selects Times Roman as basic font

% see the list of further useful packages
% in the Reference Guide

\makeindex             % used for the subject index
                       % please use the style svind.ist with
                       % your makeindex program

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Macros the QMCPy Authors wish to add

%   python styling
\usepackage{listings}
\definecolor{darkgreen}{rgb}{0,0.6,0}
\lstdefinestyle{Python}{
    showstringspaces=false,
    language        = Python,
    basicstyle      = \small\ttfamily,
    morekeywords = {as},
    keywordstyle    = \color{blue},
    stringstyle     = \color{darkgreen},
    commentstyle    = \color{darkgreen}\ttfamily,
	breaklines = true,
	postbreak=\text{$\hookrightarrow$\space},
	% style >>> and ... 
	%   see: https://tex.stackexchange.com/questions/326655/make-a-keyword-in-listings-enviorment
	alsoletter = {>,.} ,
    morekeywords = [2]{>>>,...},
    keywordstyle = [2]\color{cyan}\bfseries}

% work reproducible from https://tinyurl.com/QMCPyArticle2021
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\allowdisplaybreaks

\begin{document}

\title*{Quasi-Monte Carlo Software}
\authorrunning{S.-C.\ T.\ Choi et al.}
\author{Sou-Cheng T. Choi \and Fred J. Hickernell \and R. Jagadeeswaran \and Michael J. McCourt \and Aleksei G. Sorokin}
\institute{Sou-Cheng T. Choi \at Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616; and Kamakura Corporation, 2222 Kalakaua Ave, Suite 1400, Honolulu, HI 96815 \email{schoi32@iit.edu}
\and
Fred J. Hickernell \at Center for Interdisciplinary Scientific Computation and \\
Department of Applied Mathematics, Illinois Institute of Technology \\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{hickernell@iit.edu}
\and
R. Jagadeeswaran \at
Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{jrathin1@iit.edu}; and \\Wi-Tronix LLC, 631 E Boughton Rd, Suite 240, Bolingbrook, IL 60440
\and
Michael J. McCourt \at SigOpt, an Intel company, \\
100 Bush St., Suite 1100, San Francisco, CA 94104
\email{mccourt@sigopt.com}
\and 
Aleksei G. Sorokin \at
Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{asorokin@hawk.iit.edu}}

\maketitle

\abstract{Practitioners wishing to experience the efficiency gains from using low discrepancy sequences need correct, robust, well-written software.  This article, based on our MCQMC 2020 tutorial, describes some of the better quasi-Monte Carlo (QMC) software available.  We highlight the key software components required  by QMC to approximate multivariate integrals or expectations of functions of vector random variables.  We have combined these components in QMCPy, a Python open-source library, which we hope will draw the support of the QMC community.  Here we introduce QMCPy.}

\section{Introduction} \label{sec:intro}

Quasi-Monte Carlo (QMC) methods promise great efficiency gains over independent and identically distributed (IID) Monte Carlo (MC) methods.  In some cases, QMC  achieves one hundredth of the error of IID MC in the same amount of time (see Figure \ref{fig:sc_comp}). Often, these efficiency gains are obtained simply by replacing IID sampling with low discrepancy (LD) sampling, which is the heart of QMC.

Practitioners might wish to test whether QMC would speed up their computation.  Access to the best QMC algorithms available would make that easier.  Theoreticians or algorithm developers might want to demonstrate their ideas on various use cases to show their practical value.

This tutorial points to some of the best QMC software available.  Then we describe  QMCPy \cite{QMCPy2020a}\footnote{QMCPy is in active development. This article is based on version 1.2 on PyPI.}, which is crafted to be a community-owned Python library that combines the best QMC algorithms and interesting use cases from various authors under a common user interface.

The model problem for QMC is approximating a multivariate integral,
\begin{equation} \label{eq:integral}
	\mu := \int_\mathcal{T} g(\boldsymbol{t}) \, \lambda(\boldsymbol{t}) \, \D \boldsymbol{t},
\end{equation}
where $g$ is the integrand, and $\lambda$ is a non-negative weight.  If $\lambda$ is a probability distribution (PDF) for the random variable $\boldsymbol{T}$, then $\mu$ is the mean of $g(\boldsymbol{T})$.  Regardless, we perform a suitable variable transformation to interpret this integral as the  mean of a function of a multivariate, standard uniform random variable:
\begin{equation} \label{eq:fintegral}
	\mu = \mathbb{E}[f(\boldsymbol{X})] =  \int_{[0,1]^d}  f(\boldsymbol{x}) \,  \D \boldsymbol{x} , \qquad \boldsymbol{X} \sim \mathcal{U}[0,1]^d.
\end{equation}

QMC approximates the population mean, $\mu$,  by a sample mean,
\begin{equation} \label{eq:samplemean}
	\widehat{\mu} := \frac 1n \sum_{i=0}^{n-1} f(\boldsymbol{X}_i), \qquad \boldsymbol{X}_0, \boldsymbol{X}_1, \ldots \overset{\textup{M}}{\sim} \mathcal{U}[0,1]^d.
\end{equation}
The choice of the sequence $\{\boldsymbol{X}_i\}_{i=0}^\infty$ and the choice of $n$ to satisfy  the prescribed error requirement,
\begin{equation} \label{eq:err_req}
	\lvert\mu - \widehat{\mu}\rvert \le \varepsilon \qquad \text{absolutely or with high probability},
\end{equation} 
are important decisions, which  QMC software helps the user make.

Here, the notation $\overset{\textup{M}}{\sim}$ means that the sequence mimics the specified, target distribution, but not necessarily in a probabilistic way.  We  use this notation in two forms:  $\overset{\textup{IID}}{\sim}$ and $\overset{\textup{LD}}{\sim}$.

IID sequences must be random. The position of any point is not influenced by any other, so clusters and gaps occur.  A randomly chosen subsequence of an IID sequence is also IID.  When we say that $\boldsymbol{X}_0, \boldsymbol{X}_1, \ldots \overset{\textup{IID}}{\sim} F$ for some distribution $F$, we mean that for any positive integer $n$, the  multivariate probability distribution of $\boldsymbol{X}_0, \ldots, \boldsymbol{X}_{n-1}$ is the product of the marginals, specifically,
\begin{equation*}
	F_{n}(\boldsymbol{x}_0, \ldots, \boldsymbol{x}_{n-1}) = F(\boldsymbol{x}_0) \cdots  F(\boldsymbol{x}_{n-1}).
\end{equation*}
When IID points are used to approximate $\mu$ by the sample mean, the root mean squared error is $\mathcal{O}(n^{-1/2})$.  Fig.~\ref{fig:comparePts} displays IID uniform points, $\boldsymbol{X}^{\textup{IID}}_0, \boldsymbol{X}^{\textup{IID}}_1, \ldots \overset{\textup{IID}}{\sim} \mathcal{U}[0,1]^2$, i.e.,  the target distribution is $F_{\textup{unif}}(\boldsymbol{x}) = x_1 x_2$.

\begin{figure}[t]
	\includegraphics[height=5.8cm]{QMCSoftwareArticle/vfigs/dd_iid_uniform_pts.pdf}
	\quad
	\includegraphics[height=5.8cm]{QMCSoftwareArticle/vfigs/dd_sobol_pts.pdf}
	\caption{IID points (left) contrasted with LD points (right).  The LD points cover the square more evenly.}
	\label{fig:comparePts}
\end{figure}

LD sequences may be deterministic or random, but each point is carefully coordinated with the others so that they fill the domain well.  Subsequences of LD sequences are generally not LD.  When we say that $\boldsymbol{X}_0, \boldsymbol{X}_1, \ldots \overset{\textup{LD}}{\sim} \mathcal{U}[0,1]^d$, we mean that for any positive integer $n$,  the \emph{empirical distribution} of $\boldsymbol{X}_0, \ldots, \boldsymbol{X}_{n-1}$, denoted $F_{\{\boldsymbol{X}_i\}_{i=0}^{n-1}}$,  approximates the uniform distribution, $F_{\textup{unif}}$, well (relative to $n$).  (The empirical distribution of a set assigns equal probability to each point.)

A measure of the difference between the empirical distribution of a set of points and the uniform distribution is called a \emph{discrepancy} and is denoted $D(\{\boldsymbol{X}_i\}_{i=0}^{n-1})$ \cite{DicEtal14a,Hic97a,Hic99a,Nie92}.  This is the origin of the term ``low discrepancy'' points or sequences.  LD points by definition have a smaller discrepancy than IID points.  Fig.~\ref{fig:comparePts} contrasts IID uniform points with LD points, $\boldsymbol{X}^{\textup{LD}}_0, \boldsymbol{X}^{\textup{LD}}_1 \ldots \overset{\textup{LD}}{\sim} \mathcal{U}[0,1]^2$, in this case, linearly scrambled and digitally shifted Sobol' points.

The error in using the sample mean to approximate the integral can be bounded according to the Koksma-Hlawka inequality and its extensions \cite{DicEtal14a,Hic97a,Hic99a,Nie92} as the product of the discrepancy of the sampling sequence and the variation of the integrand, denoted $V(\cdot)$:
\begin{equation}
	\lvert\mu - \widehat{\mu}\rvert = \biggl \lvert \int_{[0,1]^d} f(\boldsymbol{x}) \, \D \bigl(F_{\textup{unif}} - F_{\{\boldsymbol{X}_i\}_{i=0}^{n-1}}\bigr) (\boldsymbol{x}) \biggr \rvert \le D\bigl(\{\boldsymbol{X}_i\}_{i=0}^{n-1}\bigr) V(f).
\end{equation} 
The variation is a (semi-) norm of the integrand in a suitable Banach space.  The discrepancy corresponds to the norm of the error functional for that Banach space.  For typical Banach spaces, the discrepancy of LD points is $\mathcal{O}(n^{-1+\epsilon})$, a higher convergence order than for IID points.  For details,  readers may refer to the references.

Here, we expect the reader to see in Fig.~\ref{fig:comparePts} that the LD points cover the integration domain more evenly than IID points.  LD sampling can be thought of as a more even distribution of the sampling sites than IID.  LD sampling is similar to stratified sampling.  In the examples below, the reader will see the demonstrably smaller cubature errors arising from using LD points.

In the sections that follow, we first overview available QMC software.  We next describe an architecture for good QMC software, i.e., what the key components are and how they should interact.  We then describe how we have implemented this architecture in QMCPy.  Finally, we summarize further directions that we hope QMCPy and other QMC software projects will take.  Those interested in following or contributing to the development of QMCPy are urged to visit the GitHub repository at https://github.com/QMCSoftware/QMCSoftware.

%\href{https://github.com/QMCSoftware/QMCSoftware}{\nolinkurl{https://github.com/QMCSoftware/QMCSoftware}}.

We have endeavored to be as accurate as possible at the time of writing this article.  We hope that progress in QMC software development will make this article happily obsolete in the coming years.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Available Software for QMC} \label{sec:available} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
QMC software spans  LD sequence generators, cubatures, and applications.  Here we review the better-known software, recognizing that some software overlaps multiple categories. Whenever applicable, we state each library's accessibility in QMCPy, or contrast its functionalities with QMCPy's --- where we lag behind in QMCPy, we strive to catch up in the near future.

Software focusing on generating high-quality LD sequences and their generators includes the following, listed in alphabetical order:
\begin{description}%[format=\textup,format=\textbf]
% https://tex.stackexchange.com/questions/74279/how-to-add-bullets-to-description-lists
	\item[\textbf{BRODA}] \sloppypar Commerical and non-commercial software developed jointly with I.M.~Sobol' in C++, MATLAB, and Excel~\cite{BRODA20a}. BRODA can generate Sobol' sequences up to 65,536 dimensions. In comparison, QMCPy supports Sobol' sequences up to 21,201 dimensions.	
	\item[\textbf{Burkhardt}] Various QMC software for generating van der Corput, Faure, Halton, Hammersley,  Niederreiter, or Sobol' sequences in C, C++, Fortran, MATLAB, or Python~\cite{Bur20a}. In QMCPy, we have implemented digital net, lattice, and Halton generators.
	\item[\textbf{LatNet Builder}] The successor to \emph{Lattice Builder}~\cite{LEcMun14a}, this is a C++ library with Python and Java interfaces (in SSJ below) for generating vectors or matrices for lattices and digital nets~\cite{LatNet,LEcEtal22a}. QMCPy contains a module for parsing the resultant vectors or matrices from LatNet Builder for compatibility with our LD point generators. %contains a \href{https://qmcpy.readthedocs.io/en/latest/demo_rst/latnetbuilder.html}{demo for LatNet Builder}. 
	\item[\textbf{MATLAB}] Commercial software for scientific computing \cite{MAT9.10}, which contains Sobol' and Halton sequences in the Statistics and Machine Learning Toolbox. Both generators can be applied jointly with the Parallel Computing Toolbox to accelerate their execution speed. The  dimension of the Sobol' sequences is restricted to 1,111, which is relatively small, yet sufficient for most applications. 
	\item[\textbf{MPS}] Magic Point Shop contains lattices and Sobol' sequences in C++, Python, and MATLAB~\cite{Nuy17a}.   QMCPy started with MPS for developing LD generators.
	\item[\textbf{Owen}] Owen's randomized Halton sequences with dimensions up to 1,000 \cite{Owe20a} and scrambled Sobol' sequences with dimensions up to 21,021  \cite{owen1998scrambling, owen2021r}  in R. QMCPy  supports Owen's Halton randomization method, and we plan to implement Owen's nested uniform scrambling for digital nets in the near future.
	\item[\textbf{PyTorch}] Open-source Python library for deep learning, with unscrambled or scrambled Sobol' sequences \cite{NEURIPS2019_9015,PyTorch}. PyTorch enables seamless utilization of Graphics Processing Units (GPUs) or Field Programmable Gate Arrays (FPGAs).
	\item[\textbf{QMC.jl}] LD Sequences in Julia \cite{Rob20a}. Julia~\cite{bezanson2012julia} is an interpreted language similar to Python and R in terms of ease of use, but is designed to run much faster.
	\item[\textbf{qrng}] Randomized Sobol', Halton, and Korobov sequences in R~\cite{QRNG2020}. The default Halton randomization in QMCPy utilizes the methods from qrng. 
	%translated qrng to Python; see  \url{https://github.com/QMCSoftware/qrngpy}. The Halton randomization from qrng is also optionally available in QMCPy. \AGSComment{We don't really use qrng besides the QRNG Halton randomization.} In addition, the C source files in qrng can be compiled and accessed in QMCPy as one of the backends. In some cases, our version of qrng with its compiled code in QMCPy is more efficient than the original software in R \scnote{Why or how? By taking advantage of MRG63k3a?}; see the demo, \href{https://qmcpy.readthedocs.io/en/latest/demo_rst/quasirandom_generators.html}{\emph{Quasi-Random Sequence Generator Comparison}}, in the QMCPy documentation~\cite{QMCPyDocs}. 
	\item[\textbf{SciPy}] Scientific computing library in Python with Latin hypercube,  Halton, and Sobol' generators \cite{2020SciPy-NMeth}.
	\item[\textbf{TF Quant Finance}] Google's Tensorflow  deep-learning library~\cite{tensorflow2015-whitepaper} specialized for financial modeling \cite{tfqf21}. It contains lattice and Sobol' generators alongside with algorithms sped up with GPUs, FGPAs, or Tensor Processing Units (TPUs).% as opposed to the Central Processing Units (CPUs).		
\end{description}
Software focusing on QMC cubatures and applications includes the following:
\begin{description}%[format=\textup,format=\textbf]
	\item[\textbf{GAIL}] The Guaranteed Automatic Integration Library  contains automatic (Q)MC stopping criteria in MATLAB~\cite{ChoEtal21a,HCJJ14}. These are iterative procedures for one- or high-dimensional integration that take a user's input error tolerance(s) and determine the number of (Q)MC sampling points necessary to achieve user-desired accuracy (almost surely).  Most of GAIL's (Q)MC functions, some with enhancements, are implemented in Python in QMCPy.
	\item[\textbf{ML(Q)MC}] Multi-Level (Q)MC routines in C, C++, MATLAB, Python, and R~\cite{GilesSoft}. We have ported ML(Q)MC functions to QMCPy.
	\item[\textbf{MultilevelEstimators.jl}] ML(Q)MC methods in Julia \cite{Rob21}. The author, Pieterjan Robbe, has contributed cubature algorithms and use cases to QMCPy.
	\item[\textbf{OpenTURNS}] Open source initiative for the Treatment of Uncertainties, Risks'N Statistics \cite{OpenTURNS} written in C++ and Python, leveraging R statistical packages, as well as LAPACK and BLAS for numerical linear algebra.
	\item[\textbf{QMC4PDE}] \sloppypar QMC for elliptic PDEs with random diffusion coefficients in Python~\cite{KuoNuy16a}.
	\item[\textbf{SSJ}] Stochastic Simulation with the \texttt{hups} package in Java \cite{SSJ}.
	\item[\textbf{UQLab}] Framework for Uncertainty Quantification in MATLAB \cite{UQLab2014}.  The core of UQLab is closed source, but a large portion of the library is open source. Recently,  UQ[py]Lab, the beta release of UQLab with Python bindings, is  available as Software as a Service (SaaS) via UQCloud~\cite{lataniotis2021uncertainty}.

\end{description}

The sections that follow describe QMCPy \cite{QMCPy2020a}, which is our attempt to establish a framework for QMC software and to combine the best of the above software under a common user interface written in Python 3.  The choice of language was determined by the desire to make QMC software accessible to a broad audience, especially the technology industry.

\section{Components of QMC Software}
QMC cubature can be summarized as follows.  We want to approximate the expectation, $\mu$, well by the sample mean, $\widehat{\mu}$, where \eqref{eq:integral}, \eqref{eq:fintegral}, and \eqref{eq:samplemean} combine to give
\begin{multline} \label{eq:cubSummary}
	\mu : = \int_\mathcal{T} g(\boldsymbol{t}) \, \lambda(\boldsymbol{t}) \, \D \boldsymbol{t}  = \mathbb{E}[f(\boldsymbol{X})] = \int_{[0,1]^d} f(\boldsymbol{x}) \, \D \boldsymbol{x} \approx \frac 1n \sum_{i=0}^{n-1} f(\boldsymbol{X}_i) =: \widehat{\mu}, \\
	 \boldsymbol{X} \sim \mathcal{U}[0,1]^d, \ \boldsymbol{X}_0, \boldsymbol{X}_1, \ldots \overset{\textup{M}}{\sim} \mathcal{U}[0,1]^d.
\end{multline}
Moreover, we want to satisfy the error requirement in \eqref{eq:err_req}.
This requires four components, which we implement as QMCPy classes.

\begin{description}%[format=\textup,format=\textbf]

	\item[\textbf{Discrete Distribution}]  produces the sequence $\boldsymbol{X}_0, \boldsymbol{X}_1, \dots$ that mimics $\mathcal{U}[0,1]^d$;

	\item[\textbf{True Measure}] $\boldsymbol{t} \mapsto \lambda (\boldsymbol{t}) \D \boldsymbol{t}$  defines the original measure, e.g., Gaussian or Lebesgue;

	\item[\textbf{Integrand}] $g$  defines the original integrand, and $f$ defines the transformed version to fit the \texttt{DiscreteDistribution}; and

	\item[\textbf{Stopping Criterion}] determines how large $n$ should be to ensure that $\lvert \mu - \widehat{\mu}\rvert \le \varepsilon$ as in (\ref{eq:err_req}).
\end{description}

The software libraries referenced in Section \ref{sec:available} provide one or more of these components. QMCPy combines multiple examples of all these components under an object-oriented framework. Each example is implemented as a concrete class that realizes the properties and methods required by the abstract class for that component. The following sections detail descriptions and specific examples for each component.

Thorough documentation of all QMCPy classes is available in \cite{QMCPyDocs}. Demonstrations of how QMCPy works are given in Google Colab notebooks \cite{QMCPyTutColab2020,QMCPyTutColab2020_paper}. The project may be installed from PyPI into a Python 3 environment via the command \texttt{pip install qmcpy}. In the code snippets that follow, we assume QMCPy has been imported alongside NumPy \cite{numpy} via the following commands in a Python Console:
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/import.txt}

\section{Discrete Distributions}

LD sequences typically mimic $\mathcal{U}[0,1]^d$.  Good sequences mimicking other distributions are obtained by transformations as described in the next section.  We denote by \texttt{DiscreteDistribution} the abstract class containing LD sequence generators.  In most cases that have been implemented, the points $\boldsymbol{X}_0, \ldots, \boldsymbol{X}_{n-1}$ have an empirical (discrete) distribution that closely approximates the uniform distribution, say, in the sense of discrepancy.  We also envision future possible \texttt{DiscreteDistribution} objects that assign unequal weights to the sampling points.  In any case, the term ``discrete'' refers to the fact that these are sequences of points (and weights), not continuous distributions.

QMCPy implements \emph{extensible} LD sequences, i.e., those that allow practitioners to obtain and use $\boldsymbol{X}_n, \boldsymbol{X}_{n+1}, \ldots $ without discarding $\boldsymbol{X}_0, \ldots, \boldsymbol{X}_{n-1}$.  Halton sequences do not have preferred sample sizes $n$, but extensible integration lattices and digital sequences in base $b$ prefer $n$ to be a power of $b$.  For integration lattices and digital sequences, we have focused on base $b=2$ since this is a popular choice and for convenience in generating extensible sequences.

Integration lattices and digital sequences in base $2$ have an elegant group structure, which we summarize in Table \ref{tab:GroupProp}.  The addition operator  is $\oplus$\footnote{The operator $\oplus$ is commonly used to denote exclusive-or, does correspond to its meaning for digital sequences in base $2$.  However, we are using it here in a more general sense.}, and its inverse is $\ominus$.  The unshifted sequence is $\boldsymbol{Z}_0, \boldsymbol{Z}_1, \ldots$ and the randomly shifted sequence is $\boldsymbol{X}_0, \boldsymbol{X}_1, \ldots$

\begin{table}[h]
	\centering
	\caption{Properties of lattices and digital net sequences.  Note that they share group properties but also have distinctives.} \label{tab:GroupProp}
\[
	\renewcommand{\arraystretch}{1.3}
\begin{array}{c@{\qquad}c}
	\hline
	\multicolumn{2}{c}{\text{Define \ldots}} \\
	\multicolumn{2}{c}{\boldsymbol{Z}_1, \boldsymbol{Z}_2, \boldsymbol{Z}_4, \ldots \in [0,1)^d \text{ chosen well} } \\
	\multicolumn{2}{c}{
	\boldsymbol{Z}_{i} := i_0  \boldsymbol{Z}_1 \oplus i_1 \boldsymbol{Z}_{2} \oplus i_2  \boldsymbol{Z}_{4} \oplus  i_3  \boldsymbol{Z}_{8} \oplus \cdots 
	\quad
	\text{for }i = i_0 +i_1 2 + i_2 4 + i_3 8 + \cdots, \; i_\ell \in \{0,1\}} \\
    \multicolumn{2}{c}{\boldsymbol{X}_i := \boldsymbol{Z}_i \oplus \boldsymbol{\Delta}, \qquad \text{where }\boldsymbol{\Delta} \overset{\textup{IID}}{\sim} [0,1)^d} \\  \hline
	\text{Rank-1 Integration Lattices} & \text{Digital Nets} \\
		\boldsymbol{t} \oplus \boldsymbol{x} : = (\boldsymbol{t} + \boldsymbol{x}) \bmod \boldsymbol{1} & \boldsymbol{t} \oplus \boldsymbol{x} := \text{binary digitwise addition}, \oplus_{\textup{dig}} \\ 
		\text{require } \boldsymbol{Z}_{2^{m}} \oplus \boldsymbol{Z}_{2^{m}} = \boldsymbol{Z}_{\lfloor2^{m-1}\rfloor}
		\quad \forall m \in \mathbb{N}_0 
%		 \\ \text{require } \begin{array}{l} \boldsymbol{Z}_1 = (1/2, \ldots, 1/2) \\
%		 	\boldsymbol{Z}_{2^{m}} \oplus \boldsymbol{Z}_{2^{m}} = \boldsymbol{Z}_{2^{m-1}} \quad \forall m \in \mathbb{N} 
%	 	\end{array}
		\\[.5ex]
\hline
\multicolumn{2}{c}{\text{Then it follows that \ldots}} \\
	\multicolumn{2}{c}{\left . \begin{array}{r}
			\mathcal{P}_m := \{\boldsymbol{Z}_0, \ldots, \boldsymbol{Z}_{2^m-1}\}, \quad
			\boldsymbol{Z}_i \oplus \boldsymbol{Z}_j \in \mathcal{P}_m \\
			\mathcal{P}_{\boldsymbol{\Delta},m} := \{\boldsymbol{X}_0, \ldots, \boldsymbol{X}_{2^m-1}\}, \quad
			\boldsymbol{X}_i \oplus \boldsymbol{X}_j \ominus \boldsymbol{X}_k \in \mathcal{P}_{\boldsymbol{\Delta},m}
	\end{array} \right \}\quad \begin{array}{l}\forall  i,j,k \in \{0, \ldots, 2^{m} -1\} \\ \forall m \in \mathbb{N}_0\end{array}} \\
\hline
\end{array}
\]
\end{table}

We illustrate lattice and Sobol' sequences using QMCPy. First, we create an instance of a $d=2$ dimensional \texttt{Lattice} object of the  \texttt{DiscreteDistribution} abstract class. Then we generate the first eight (non-randomized) points in this lattice. 
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/dd_lattice.txt}
The first three generators for this lattice are $\boldsymbol{Z}_1 = (0.5, 0.5)$, $\boldsymbol{Z}_2 = (0.25, 0.75)$, and $\boldsymbol{Z}_4 = (0.125, 0.375)$.  One can check that $(\boldsymbol{Z}_2 + \boldsymbol{Z}_4) \bmod \boldsymbol{1} = (0.375, 0.125) = \boldsymbol{Z}_6$, as Table \ref{tab:GroupProp} specifies.

The random shift has been turned off above to illuminate the group structure.  We normally include the randomization to ensure that there are no points on the boundary of $[0,1]^d$.  Then, when points are transformed to mimic distributions such as the Gaussian, no LD points will be transformed to infinity.  Turning off the randomization generates a warning when the \texttt{gen\_samples} method is called.

Now, we generate Sobol' points using a similar process as we did for lattice points.  Sobol' sequences are one of the most popular example of digital sequences.
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/dd_sobol.txt}
Here, $\boldsymbol{Z}_4$ differs from that for lattices, but more importantly, addition for digital sequences differs from that for lattices.  Using digitwise addition for digital sequences, we can confirm that according to Table \ref{tab:GroupProp},
\begin{align*}
\boldsymbol{Z}_2 \oplus_{\textup{dig}} \boldsymbol{Z}_4 = (0.25,0.75)  \oplus_{\textup{dig}} (0.125,0.625) \\
=  ({}_20.010,{}_20.110)  \oplus_{\textup{dig}} ({}_20.001,{}_20.101) &= ({}_20.011,{}_20.011) \\
& = (0.375,0.375) = \boldsymbol{Z}_6.
\end{align*}

By contrast, if we construct a digital sequence using the generators for the lattice above with $\boldsymbol{Z}_2 = (0.25, 0.75)$, and $\boldsymbol{Z}_4 = (0.125, 0.375)$, we would obtain
\begin{multline*}
\boldsymbol{Z}_6 = \boldsymbol{Z}_2 \oplus_{\textup{dig}} \boldsymbol{Z}_4   = ({}_20.010,{}_20.110)  \oplus_{\textup{dig}} ({}_20.001,{}_20.011)  \\
= ({}_20.011,{}_20.101) = (0.375, 0.625),
\end{multline*}
which differs from the $\boldsymbol{Z}_6=(0.375, 0.125)$ constructed for lattices.  To emphasize, lattices and digital sequences are different, even if they share the same generators, $\boldsymbol{Z}_1, \boldsymbol{Z}_2, \boldsymbol{Z}_4, \ldots$.

The examples of \texttt{qp.Lattice} and \texttt{qp.Sobol} illustrate how QMCPy LD generators share a common user interface.  The dimension is specified when the instance is constructed, and the number of points is specified when the \texttt{gen\_samples} method is called.  Following Python practice, parameters can be input without specifying their names if input in the prescribed order.  QMCPy also includes Halton sequences and IID sequences, again deferring details to the QMCPy documentation~\cite{QMCPyDocs}.

A crucial difference between IID generators and LD generators is reflected in the behavior of generating $n$ points.  For an IID generator, asking for $n$ points repeatedly gives different points each time because they are meant to be random and independent.
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/dd_iidone.txt}
Your output may look different depending on the seed used to generate these random numbers.

On the other hand, for an LD generator, asking for $n$ points repeatedly gives \emph{the same} points each time because they are meant to be the first $n$ points of a specific LD sequence.  
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/dd_latticeone.txt}
Here we allow the randomization so that the first point in the sequence is not the origin.  To obtain the \emph{next} $n$ points, one may specify the start and ending indices of the sequence.
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/dd_latticenextone.txt}

Fig.~\ref{fig:increase_n} shows how increasing the number of lattice and Sobol' LD points through powers of two fills in the gaps in an even way.

\begin{figure}[t]
	\includegraphics[width=1\textwidth]{QMCSoftwareArticle/vfigs/dd_lattice_successive.pdf}
	\qquad
	\includegraphics[width=1\textwidth]{QMCSoftwareArticle/vfigs/dd_sobol_successive.pdf}
	\caption{Randomized lattice and Sobol' points mimicking a $\mathcal{U}[0,1]^2$ measure for $n = 64, 128,$ and 256. Note how increasing the number of points evenly fills in the gaps between the points.}
	\label{fig:increase_n}
\end{figure}

\section{True Measures}

The LD sequences implemented as \texttt{DiscreteDistribution} objects mimic the $\mathcal{U}[0,1]^d$ distribution.  However, we may need sequences to mimic other distributions.  This is implemented via variable transformations, $\boldsymbol{\Psi}$.  In general, if $\boldsymbol{X} \overset{\textup{M}}{\sim} \mathcal{U}[0,1]^d$, then
\begin{subequations} \label{eq:exampleVarTrans}
\begin{gather}
\boldsymbol{T} = \boldsymbol{\Psi}(\boldsymbol{X}) := \boldsymbol{a}  + (\boldsymbol{b} - \boldsymbol{a}) \odot \boldsymbol{X} \overset{\textup{M}}{\sim}  \mathcal{U}[\boldsymbol{a},\boldsymbol{b}], \\
\label{eq:exampleVarTransGauss}
\boldsymbol{T} = \boldsymbol{\Psi}(\boldsymbol{X}) := \boldsymbol{a} + \mathsf{A} \boldsymbol{\Phi}^{-1}(\boldsymbol{X})  \overset{\textup{M}}{\sim} \mathcal{N}(\boldsymbol{a}, \Sigma), \\
\nonumber  \text{where }  \boldsymbol{\Phi}^{-1}(\boldsymbol{X}) : = \begin{pmatrix} \Phi^{-1}(X_1) \\ \vdots \\ \Phi^{-1}(X_d)\end{pmatrix}, \qquad \Sigma = \mathsf{A} \mathsf{A}^T,
\end{gather}
\end{subequations}
and $\odot$ denotes term-by-term (Hadamard) multiplication.  Here, $\boldsymbol{a}$ and $\boldsymbol{b}$ are assumed to be finite, and $\Phi$ is the standard Gaussian distribution function.  Again we use $\overset{\textup{M}}{\sim}$ to denote mimicry, not necessarily in a probabilistic sense.

Fig.~\ref{fig:tm_ug} displays LD sequences transformed as described above to mimic a uniform and a Gaussian distribution.  The code to generate these points takes the following form for uniform points based on a Halton sequence: 
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/tm_uniform.txt}
whereas for Gaussian points based on a lattice sequence, we have:
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/tm_gaussian.txt}
Here the covariance decomposition $\Sigma = \mathsf{A} \mathsf{A}^T$ is done using principal component analysis. The Cholesky decomposition is also available.

\begin{figure}[t]
	\includegraphics[width=.45\textwidth]{QMCSoftwareArticle/vfigs/tm_uniform.pdf} 
	%	\quad
	\includegraphics[width=.45\textwidth]{QMCSoftwareArticle/vfigs/tm_gaussian.pdf}
	\caption{Halton samples transformed to mimic a uniform $\mathcal{U}\left( \begin{bmatrix} -2 \\ 0 \end{bmatrix},\begin{bmatrix} 2 \\ 4 \end{bmatrix} \right)$ 
	%\frednote{This is not correct, we need the upper and lower endponts of a two dimensional interval.  Also, are these Sobol' or Halton as in the code?} 
	distribution (left) and lattice samples transformed to mimic a Gaussian $\mathcal{N}\left(\begin{bmatrix} 3 \\ 2 \end{bmatrix}, \begin{bmatrix} 9 & 5 \\ 5 & 4 \end{bmatrix} \right)$ distribution (right).}
	\label{fig:tm_ug}
\end{figure}

The Brownian motion distribution arises often in financial risk applications.  Here the $d$ components of the variable $\boldsymbol{T}$ correspond to the discretized Brownian motion at times $\tau/d, 2\tau/d, \ldots, \tau$, where $\tau$ is the time horizon.  The distribution is a special case of the Gaussian with covariance 
\begin{equation} \label{eq:BMcov}
	\Sigma = (\tau/d) \bigl (\min(j,k) \bigr)_{j,k=1}^d
\end{equation}
and mean $\boldsymbol{a}$, which  is proportional to the times $(\tau/d)(1, 2, \ldots, d)^T$. The code for generating a Brownian motion is
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/tm_brownian_motion.txt}
Fig.~\ref{fig:tm_bm} displays a Brownian motion based on Sobol' sequence with and without a drift.

\begin{figure}[t]
	\includegraphics[width=1\textwidth]{QMCSoftwareArticle/vfigs/tm_bm.pdf} 
	\caption{Sobol' samples transformed to mimic a 52-dimensional Brownian Motion without drift (left) and with drift coefficient 2 (right).}
	\label{fig:tm_bm}
\end{figure}

\section{Integrands}

Let's return  to the integration problem in \eqref{eq:integral}, which we must rewrite as \eqref{eq:fintegral}.  We choose a transformation of variables defined as $\boldsymbol{t} = \boldsymbol{\Psi}(\boldsymbol{x})$ where $\boldsymbol{\Psi}:[0,1]^d \to \mathcal{T}$.  This leads to 
\begin{align}
	\nonumber 
 \mu &= \int_\mathcal{T} g(\boldsymbol{t}) \, \lambda(\boldsymbol{t}) \, \D \boldsymbol{t}  = \int_{[0,1]^d} g\bigl(\boldsymbol{\Psi}(\boldsymbol{x})\bigr) \, \lambda\bigl(\boldsymbol{\Psi}(\boldsymbol{x})\bigr) \,\lvert \boldsymbol{\Psi}'(\boldsymbol{x})\rvert  \, \D \boldsymbol{x} =  \int_{[0,1]^d}f(\boldsymbol{x}) \, \D \boldsymbol{x},  \\
 \label{eq:transVar}
  & \qquad \qquad \text{where } f(\boldsymbol{x})  = g\bigl(\boldsymbol{\Psi}(\boldsymbol{x})\bigr)  \, \lambda\bigl(\boldsymbol{\Psi}(\boldsymbol{x})\bigr) \,\lvert \boldsymbol{\Psi}'(\boldsymbol{x})\rvert ,
\end{align}
and $\lvert \boldsymbol{\Psi}'(\boldsymbol{x})\rvert := \lvert \partial \boldsymbol{\Psi}/\partial \boldsymbol{x}\rvert$ represents the Jacobian of the variable transformation.  The abstract class \texttt{Integrand} provides $f$ based on the user's input of $g$ and the \texttt{TrueMeasure} instance, which defines $\lambda$ and the transformation $\boldsymbol{\Psi}$. Different choices of $\boldsymbol{\Psi}$ lead to different $f$, which may give different rates of convergence of the cubature, $\widehat{\mu}$ to $\mu$.

We illustrate the \texttt{Integrand} class via an example of Keister \cite{Kei96}:
\begin{equation} \label{eq:KeisterIntegral}
	\mu 
	= \int_{\mathbb{R}^d} \cos(\lVert \boldsymbol{t} \rVert) \exp(-\boldsymbol{t}^T \boldsymbol{t}) \, \D \boldsymbol{t} \\ 
	= \int_{\mathbb{R}^d} \underbrace{\pi^{d/2} \cos(\lVert \boldsymbol{t} \rVert)}_{g(\boldsymbol{t})}\, \underbrace{\pi^{-d/2} \exp(-\boldsymbol{t}^T \boldsymbol{t}) }_{\lambda(\boldsymbol{t})} \, \D \boldsymbol{t}.
\end{equation}
Since $\lambda$ is the density for $\mathcal{N}(\boldsymbol{0},\mathsf{I}/2)$, it is natural to choose $\boldsymbol{\Psi}$ according to \eqref{eq:exampleVarTransGauss} with $\mathsf{A} = \sqrt{1/2} \, \mathsf{I}$, in which case $\lambda(\boldsymbol{\Psi}(\boldsymbol{x})) \lvert \boldsymbol{\Psi}'(\boldsymbol{x})\rvert  = 1$, and so 
\[
\mu = \int_{[0,1]^d} \underbrace{\pi^{d/2} \cos(\lVert \boldsymbol{\Psi}(\boldsymbol{x}) \rVert)}_{f(\boldsymbol{x})} \, \D \boldsymbol{x}, \qquad 
\boldsymbol{\Psi}(\boldsymbol{x}) := \sqrt{1/2} \,\boldsymbol{\Phi}^{-1}(\boldsymbol{x}).
\]

The code below sets up an \texttt{Integrand} instance using QMCPy's \texttt{CustomFun} wrapper to tie a user-defined function $g$ into the QMCPy framework.  Then we evaluate the sample mean of $n=1000$ $f$ values obtained by sampling at transformed Halton points. Notice how a two-dimensional Halton generator is used to construct a Gaussian true measure, which is applied alongside the \texttt{my\_Keister} function to instantiate a customized, QMCPy-compatible integrand for this problem.
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/i_keister.txt}
We have no indication yet of how accurate our approximation is.  That topic is treated in the next section.  
Fig.~\ref{fig:ikc} visualizes sampling on the original integrand, $g$, and sampling on the transformed integrand, $f$.

\begin{figure}[t]
	\includegraphics[height=7cm]{QMCSoftwareArticle/vfigs/i_keister_contours.pdf}
	\caption{Right: Sampling the transformed Keister integrand $f$ at Halton points $\boldsymbol{X}_i \overset{\textup{LD}}{\sim} \mathcal{U}[0,1]^2$. Left: Sampling the original Keister integrand $g$ at $\boldsymbol{T}_i = \boldsymbol{\Psi}(\boldsymbol{X}_i) \overset{\textup{M}}{\sim} \mathcal{N}(\boldsymbol{0},\mathsf{I}/2)$ where $\boldsymbol{\Psi}$ is defined in \eqref{eq:exampleVarTransGauss}.  } \label{fig:ikc}
\end{figure}

Another way to approximate the Keister integral in \eqref{eq:KeisterIntegral} is to write it as an integral with respect to the Lebesgue measure:
\begin{align*} 
	\mu 
	& = \int_{\mathbb{R}^d} \underbrace{\cos(\lVert \boldsymbol{t} \rVert) \exp(-\boldsymbol{t}^T \boldsymbol{t}) }_{g(\boldsymbol{t})} \, \underbrace{1}_{\lambda(\boldsymbol{t})} \,\D \boldsymbol{t} \\
	& = \int_{[0,1]^d} \underbrace{\cos(\lVert \boldsymbol{\Psi}(\boldsymbol{x}) \rVert) \exp(-\boldsymbol{\Psi}^T\!\!(\boldsymbol{x}) \boldsymbol{\Psi}(\boldsymbol{x})) \lvert \boldsymbol{\Psi}'(\boldsymbol{x})\rvert  }_{f(\boldsymbol{x})} \, \D \boldsymbol{x},
\end{align*}
where $\boldsymbol{\Psi}$ is any transformation from $[0,1]^d$ to $\mathbb{R}^d$. Now $\lambda$ is \emph{not} a PDF.  QMCPy can perform the cubature this way as well.
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/i_keisterLebesgue.txt}
The $\boldsymbol{\Psi}$ chosen when transforming uniform sequences on the unit cube to fill $\mathbb{R}^d$ is  given by \eqref{eq:exampleVarTransGauss} with $\mathsf{A} = \mathsf{I}$.

In the examples above, one must input the correct $g$ into \texttt{CustomFun} along with the correct \texttt{TrueMeasure} $\lambda$ to define the integration problem. The \texttt{Keister} integrand included in the QMCPy library takes a more flexible approach to defining the integration problem $\mu$ in \eqref{eq:KeisterIntegral}. Selecting a different \texttt{sampler} $\boldsymbol{\Psi}$ performs  \emph{importance sampling}, which leaves $\mu$ unchanged.  
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/i_keisterBuiltIn.txt}

In the first case above, the $\lambda$ in \eqref{eq:transVar} corresponds to the Gaussian density  with mean zero and variance $1/2$ by default, and the corresponding variable transformation, $\boldsymbol{\Psi}$, is chosen to make $\lambda\bigl(\boldsymbol{\Psi}(\boldsymbol{x})\bigr) \lvert \boldsymbol{\Psi}'(\boldsymbol{x})\rvert  = 1$ and $f(\boldsymbol{x}) = g\left(\boldsymbol{\Psi}(\boldsymbol{x}) \right)$.  In the second case,  we choose an importance sampling density $\lambda_{\textup{IS}}$, corresponding  to standard  Gaussian, and the variable transformation $\boldsymbol{\Psi}_{\textup{IS}}$ makes $\lambda_\textup{IS}\bigl(\boldsymbol{\Psi}_{\textup{IS}}(\boldsymbol{x})\bigr) \lvert\boldsymbol{\Psi}'_{\textup{IS}}(\boldsymbol{x})\rvert  = 1$.  Then
\begin{align}
	\nonumber 
	\mu &= \int_\mathcal{T} g(\boldsymbol{t}) \, \lambda(\boldsymbol{t}) \, \D \boldsymbol{t}  = \int_\mathcal{T} g(\boldsymbol{t}) \, \frac{\lambda(\boldsymbol{t})}{\lambda_\textup{IS}(\boldsymbol{t}) } \, \lambda_\textup{IS}(\boldsymbol{t}) \D \boldsymbol{t}  \\ 
	\nonumber
	& =  \int_{[0,1]^d} g\bigl(\boldsymbol{\Psi}_\textup{IS}(\boldsymbol{x})\bigr) \, \frac{\lambda\bigl(\boldsymbol{\Psi}_\textup{IS}(\boldsymbol{x})\bigr)} {\lambda_\textup{IS}\bigl(\boldsymbol{\Psi}_\textup{IS}(\boldsymbol{x})\bigr) } \lambda_\textup{IS}\bigl(\boldsymbol{\Psi}_\textup{IS}(\boldsymbol{x})\bigr) \,\lvert\boldsymbol{\Psi}'_{\textup{IS}}(\boldsymbol{x})\rvert  \, \D \boldsymbol{x} \\
	\nonumber
	& =  \int_{[0,1]^d} f_\textup{IS}(\boldsymbol{x}) \, \D \boldsymbol{x}  \\
	\label{eq:transVarImp}
	& \qquad \qquad \text{where } f_\textup{IS}(\boldsymbol{x})  = g\bigl(\boldsymbol{\Psi}_\textup{IS}(\boldsymbol{x})\bigr)  \,  \frac{\lambda\bigl(\boldsymbol{\Psi}_\textup{IS}(\boldsymbol{x})\bigr)} {\lambda_\textup{IS}\bigl(\boldsymbol{\Psi}_\textup{IS}(\boldsymbol{x})\bigr) }.
\end{align}
Because LD samples mimic $\mathcal{U}[0,1]^d$, choosing a different \texttt{sampler} is equivalent to choosing a different variable transform.

\section{Stopping Criteria} \label{sec:stopping_crit}

The \texttt{StoppingCriterion} object determines the number of samples $n$ that are required for the sample mean approximation $\widehat{\mu}$ to be within error tolerance $\varepsilon$ of the true mean $\mu$.  Several QMC stopping criteria have been implemented in QMCPy, including replications, stopping criteria that track the decay of the Fourier complex exponential or Walsh coefficients of the integrand \cite{HicJim16a,HicEtal17a,JimHic16a}, and stopping criteria based on
Bayesian credible intervals \cite{RatHic19a,JagHic22a}.

The \texttt{CubQMCSobolG} stopping criterion used in the example below assumes the Walsh-Fourier coefficients of the integrand are absolutely convergent. The algorithm iteratively doubles the number of samples used in the integration and estimates the error using the decay of the Walsh coefficients \cite{HicJim16a}. When the estimated error is below the user-specified tolerance, it finishes the computation and returns the estimated integral.

Let us return to the Keister example from the previous section.  After setting up  a default \texttt{Keister} instance via a Sobol' \texttt{DiscreteDistribution}, we choose a \texttt{StoppingCriterion} object that matches the \texttt{DiscreteDistribution} and input our desired tolerance.  Calling the  \texttt{integrate} method returns the approximate integral plus some useful information about the computation.
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/sc_keister_qmc.txt}
The second output of the stopping criterion provides helpful diagnostic information.  This computation requires $n=2^{13}$ Sobol' points and $0.008$ seconds to complete.  The error bound is $0.000606$, which falls below the absolute tolerance.

QMC, which uses LD sequences, is touted as providing substantially greater computational efficiency compared to IID MC.
Fig.~\ref{fig:sc_comp} compares the time and sample sizes needed to compute the $5$-dimensional Keister integral \eqref{eq:KeisterIntegral} using IID sequences and LD lattice sequences. Consistent with what is stated in Section \ref{sec:intro}, the error of IID MC is $\mathcal{O}(n^{-1/2})$, which means that the time and sample size to obtain an absolute error tolerance of $\varepsilon$ is $\mathcal{O}(\varepsilon^{-2})$.  By contrast, the  error of QMC using LD sequences is $\mathcal{O}(n^{-1+\epsilon})$, which implies $\mathcal{O}(\varepsilon^{-1-\epsilon})$ times and sample sizes.  We see that QMC methods often require orders of magnitude fewer samples than MC methods to achieve the same error tolerance.

\begin{figure}[t]
	\includegraphics[height=6cm]{QMCSoftwareArticle/vfigs/sc_comp.pdf}
	\caption{Comparison of run times and sample sizes for computing the $5$-dimensional Keister integral \eqref{eq:KeisterIntegral} using IID and LD lattice sequences for a variety of absolute error tolerances. The respective stopping criteria are  \texttt{qp.CubMCG} \cite{HicEtal14a} and  \texttt{qp.CubQMCLatticeG} \cite{JimHic16a}. The LD sequences provide the desired answer much more efficiently.}
	\label{fig:sc_comp}
\end{figure}

For another illustration of QMC cubature, we turn to pricing an Asian arithmetic mean call option. The (continuous-time) Asian option is defined in terms of the average of the stock price, which is written in terms of an integral. The payoff of this option is the positive difference between the strike price, $K$, averaged over the time horizon: 
$$
\text{payoff}(\boldsymbol{S}) = \max\left(\frac{1}{2d}\sum_{j=1}^d (S_{j-1}+S_j)-K, \; 0\right), \quad \boldsymbol{S} = (S_0, \ldots, S_d).
$$
Here $S_j$ denotes the asset price at time $\tau j/d$, and a trapezoidal rule is used for discrete approximation of the integral in time that defines the average. 
The trapezoidal rule is a more accurate approximation to the integral than a rectangle rule. % Comment from Fred
A basic model for asset prices is a geometric Brownian motion, 
\[
S_j(\boldsymbol{T}) = S_0 \exp((r - \sigma^2/2) \tau j/d + \sigma T_j),   \;  j = 1, \ldots, d, \; \boldsymbol{T} = (T_1, \ldots, T_d)\sim \mathcal{N}(\boldsymbol{0},\Sigma),
\]
where $\Sigma$ is defined in \eqref{eq:BMcov}, $r$ is the interest rate, $\sigma$ is the volatility, and $S_0$ is the initial asset price.  The fair price of the option is then the expected value of the discounted payoff, namely,
\begin{equation*}
	\text{price} = \mu = \mathbb{E}[g(\boldsymbol{T})], \quad \text{where } g(\boldsymbol{t}) = \text{payoff}\bigl(\boldsymbol{S}(\boldsymbol{t}) \bigr) \exp(-r \tau).
\end{equation*}

The following code utilizes QMCPy's Asian option \texttt{Integrand} object to approximate the value of an Asian call option for a particular choice of the parameters.
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/sc_aco.txt}
Because this \texttt{Integrand} object has the built-in Brownian motion \texttt{TrueMeasure}, one only need provide the LD sampler.

Out of the money option price calculations can be sped up by adding an upward drift to the Brownian motion.  The upward drift produces more in the money paths and also reduces the variation or variance of the final integrand, $f$.  This is a form of importance sampling.  Using a Brownian motion without drift we get
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/wo_imp_samp_aco.txt}
Adding the upward drift gives us the answer faster:
\lstinputlisting[style=Python]{QMCSoftwareArticle/python/w_imp_samp_aco.txt}
The choice of a good drift is an art.

The improvement in time is less than that in $n$ because the integrand is more expensive to compute when the drift is employed.  Referring to \eqref{eq:transVarImp}, in the case of no drift, the $\lambda$ corresponds to the density for the discrete Brownian motion, and the variable transformation $\boldsymbol{\Psi}$ is chosen so that $f(\boldsymbol{x}) = g\left(\boldsymbol{\Psi}(\boldsymbol{x}) \right)$.  However, in the case of a drift, the integrand becomes  $f_\textup{IS}(\boldsymbol{x})  = g\bigl(\boldsymbol{\Psi}_\textup{IS}(\boldsymbol{x})\bigr)  \lambda\bigl(\boldsymbol{\Psi}_\textup{IS}(\boldsymbol{x})\bigr)/\lambda_\textup{IS}\bigl(\boldsymbol{\Psi}_\textup{IS}(\boldsymbol{x})\bigr)$, which requires more computation time per integrand value.

\section{Under the Hood}

In this section, we look at the inner workings of QMCPy and point out features we hope will benefit the community of QMC researchers and practitioners. We also highlight important nuances of QMC methods and how QMCPy addresses these challenges. For details, readers should refer to the QMCPy documentation \cite{QMCPyDocs}.

\subsection{LD Sequences}

LD sequences are the backbone of QMC methods. QMCPy provides generators that combine research from across the QMC community to enable advanced features and customization options.

Two popular LD sequences are integration lattices and digital nets which we previously outlined in Table \ref{tab:GroupProp}. These LD generators are comprised of two parts: the static generating vectors $\boldsymbol{Z}_1,\boldsymbol{Z}_2,\boldsymbol{Z}_4, \ldots \in [0,1)^d$ and the callable generator function. By default, QMCPy provides a number of high-quality generating vectors for users to choose from. For instance, the default ordinary lattice vector was constructed by Cools, Kuo, and Nuyens \cite{doi:10.1137/06065074X} using component-by-component search, is extensible, has order-2 weights, and supports up to 3600 dimensions and $2^{20}$ samples. 
%https://web.maths.unsw.edu.au/~fkuo/lattice/index.html
%https://web.maths.unsw.edu.au/~fkuo/sobol/
However, users who require more samples but fewer dimensions may switch to a generating vector constructed using LatNet Builder \cite{LatNet,LEcEtal22a} to support 750 dimensions and $2^{24}$ samples. Moreover, the \texttt{qp.Lattice} and \texttt{qp.DigitalNet} objects allow users to input their own generating vectors to produce highly customized sequences. To find such vectors, we recommend using LatNet Builder's construction routines as the results can be easily parsed into a QMCPy-compatible format.

Along with the selection of a generating vector, QMCPy's low discrepancy sequence routines expose a number of other customization parameters. For instance, the lattice generator extends the Magic Point Shop \cite{Nuy17a} to support either linear or natural ordering. Digital sequences permit either standard or Gray code ordering and may be randomized via a digital shift optionally combined with a linear scrambling~\cite{Mat98}. Halton sequences may be randomized via the routines of either Owen~\cite{Owe20a} or Hofert and Lemieux \cite{QRNG2020}.

\subsection{A Word of Caution When Using LD Sequences}

Although QMCPy's \texttt{DiscreteDistribution}s have many of the same parameters and methods, users should be careful when swapping IID sequences with LD sequences. While IID node-sets have no preferred sample size, LD sequences often require special sampling ranges to ensure optimal discrepancy. As mentioned  earlier, base-2 digital nets and extensible integration lattices show better evenness for sample sizes that are powers of 2. On the other hand, the preferred sample sizes for $d$-dimensional Halton sequences are $n = \prod_{j=1}^d p^{m_j}_j$ where $p_j$ is the $j^{\text{th}}$ prime number and $m_j \in \mathbb{N}_0$ for $j=1,\dots,d$. Due to the infrequency of such values, Halton sequences are often regarded as not having a preferred sample size.

Users may also run into trouble when trying to generate too many points. Since QMCPy's generators construct sequences in 32-bit precision, generating greater than~$2^{32}$ consecutive samples will cause the sequence to repeat. In the future, we plan to expand our generators to support optional 64-bit precision at the cost of greater computational overhead.

Another subtlety arises when transforming LD sequences to mimic different distributions. As mentioned earlier, unrandomized lattice and digital sequences include the origin, making transformations such as \eqref{eq:exampleVarTransGauss} produce infinite values.

Some popular implementations of LD sequences drop the first point, which is the origin in the absence of randomization.  The rationale is to avoid the transformation of the origin to infinity when mimicking a Gaussian or other distribution with an infinite sample space.  Unfortunately, dropping the first point destroys some nice properties of the first $n = 2^m$ points of LD sequences, which can degrade the order of convergence for QMC cubature. A careful  discussion of this matter is given by~\cite{owen2020dropping}.

\subsection{Transformations}

The transformation $\boldsymbol{\Psi}$  connects a \texttt{DiscreteDistribution} and \texttt{TrueMeasure}. So far, we have assumed the \texttt{DiscreteDistribution} mimics a $\mathcal{U}[0,1]^d$ distribution with PDF $\varrho(\boldsymbol{x})=1$. However, it may be advantageous to utilize a \texttt{DiscreteDistribution} that mimics a different distribution. 
%For instance, QMCPy includes a \texttt{DiscreteDistribution} corresponding to an IID standard Gaussian sequence.  It may be transformed to mimic a general Gaussian distribution \texttt{TrueMeasure} while avoiding the inverse CDF transform required for nodes mimicking $\mathcal{U}[0,1]^d$.
% the IID standard Gaussian samples still use inverse CDF, so we don't actually save anything aside from maybe a minor speed up from numpy's faster code

Suppose we have a \texttt{DiscreteDistribution} mimicking  density $\varrho$ supported on $\mathcal{X}$.  Then using the variable transformation $\boldsymbol{\Psi}$, 
\begin{equation*}
	\mu = \int_\mathcal{T} g(\boldsymbol{t}) \, \lambda(\boldsymbol{t}) \, \D \boldsymbol{t}  =  \int_{\mathcal{X}}f(\boldsymbol{x}) \, \varrho(\boldsymbol{x}) \D \boldsymbol{x} \quad \text{for }
   f(\boldsymbol{x})  = g\bigl(\boldsymbol{\Psi}(\boldsymbol{x})\bigr)  \, \frac{\lambda\bigl(\boldsymbol{\Psi}(\boldsymbol{x})\bigr)}{\varrho(\boldsymbol{x})} \, \lvert \boldsymbol{\Psi}'(\boldsymbol{x})\rvert,
\end{equation*}
which generalizes \eqref{eq:transVar}.
QMCPy also includes support for successive changes of measures so users may build complex variable transformations in an intuitive manner. Suppose that the variable transformation is a composition of several transformations: $\boldsymbol{\Psi}=\widehat{\boldsymbol{\Psi}}_L = \boldsymbol{\Psi}_L \circ \boldsymbol{\Psi}_{L-1} \circ \dots \circ \boldsymbol{\Psi}_1$ as in \eqref{eq:transVarImp}. Here, $\boldsymbol{\Psi}_l:\mathcal{X}_{l-1} \to \mathcal{X}_l$,  $\mathcal{X}_0=\mathcal{X}$, and $\mathcal{X}_L=\mathcal{T}$ so that the transformations are compatible with the \texttt{DiscreteDistribution} and \texttt{TrueMeasure}. Let $\widehat{\boldsymbol{\Psi}}_l=\boldsymbol{\Psi}_l \circ \boldsymbol{\Psi}_{l-1} \circ \dots \circ \boldsymbol{\Psi}_1$ denote the composition of the first $l$ transforms and assume that $\widehat{\boldsymbol{\Psi}}_0(\boldsymbol{x})=\boldsymbol{x}$, the identity transform. Then we may write $\mu =  \int_{\mathcal{X}}f(\boldsymbol{x}) \, \varrho(\boldsymbol{x}) \D \boldsymbol{x}$ for 
\begin{equation*}
    f(\boldsymbol{x}) 
 = g\bigl(\widehat{\boldsymbol{\Psi}}_L(\boldsymbol{x})\bigr)\frac{\lambda\bigl(\widehat{\boldsymbol{\Psi}}_L(\boldsymbol{x})\bigr)}{\varrho(\boldsymbol{x})}\prod_{l=1}^L \bigl\lvert \boldsymbol{\Psi}_l'\bigl(\widehat{\boldsymbol{\Psi}}_{l-1}(\boldsymbol{x})\bigr) \bigr\rvert .
\end{equation*}
It is often the case that $\boldsymbol{\Psi}_l$ is chosen such that $\boldsymbol{\Psi}_l(\boldsymbol{X})$ is stochastically equivalent to a random variable with density $\lambda_l$ on sample space $\mathcal{X}_l$ when $\boldsymbol{X}$ is a random variable with density $\varrho_l$ on sample space $\mathcal{X}_{l-1}$.  This implies  $\varrho_l(\boldsymbol{x}) = \lambda_l(\boldsymbol{\Psi}_l(\boldsymbol{x}))\lvert  \boldsymbol{\Psi}'_l (\boldsymbol{x}) \rvert$ so that
\begin{equation*}
f(\boldsymbol{x}) 
 = g\bigl(\widehat{\boldsymbol{\Psi}}_L(\boldsymbol{x})\bigr)\frac{\lambda\bigl(\widehat{\boldsymbol{\Psi}}_L(\boldsymbol{x})\bigr)}{\varrho(\boldsymbol{x})}\prod_{l=1}^L \frac{\varrho_l(\widehat{\boldsymbol{\Psi}}_{l-1}(\boldsymbol{x}))}{\lambda_l(\widehat{\boldsymbol{\Psi}}_l(\boldsymbol{x}))} .
\end{equation*}

For an example, we return to the Keister integral \eqref{eq:KeisterIntegral}. The following code constructs three \texttt{Keister} instances: one without importance sampling, one importance sampled by a Gaussian distribution, and one importance sampled by the composition of a Gaussian distribution with a Kumaraswamy distribution \cite{Kumaraswamy}. All \texttt{Integrand}s use a Sobol' \texttt{DiscreteDistribution}, making $\varrho(\boldsymbol{x})=1$ and $\mathcal{X} = [0,1]^d$. The \texttt{TrueMeasure} is $\mathcal{N}(\boldsymbol{0},\mathsf{I}/2)$ making $\lambda(\boldsymbol{t}) = \pi^{-d/2} \exp(-\boldsymbol{t}^T \boldsymbol{t})$ and $\mathcal{T} = \mathbb{R}^d$.

The table below displays the variable transformations and the measures for these three cases.  In all cases $\varrho_1(\boldsymbol{x}) = \cdots = \varrho_L(\boldsymbol{x}) = 1$ because the $\boldsymbol{\Psi}_l$ utilize inverse cumulative distributions.

\[\begin{array}{ccccccc}
	\texttt{Integrand} & L & \lambda_1 & \boldsymbol{\Psi}_1 & \lambda_2 & \boldsymbol{\Psi}_2 & f \tabularnewline%
	%\setlength{\arrayrulewidth}{1pt}
	\hline
    %\setlength{\arrayrulewidth}{0.4pt}
	\texttt{K} & 1 & \mathcal{N}(\boldsymbol{0}, \mathsf{I}/2) & \eqref{eq:exampleVarTransGauss}& && g(\boldsymbol{\Psi}_1(\cdot)) \tabularnewline [1ex]
	\texttt{K\_gauss} & 1 & \mathcal{N}(\boldsymbol{0},3 \mathsf{I}/4) & \eqref{eq:exampleVarTransGauss} & && 
	\displaystyle g(\boldsymbol{\Psi}_1(\cdot)) \frac{\lambda(\boldsymbol{\Psi}_1(\cdot))}{\lambda_1(\boldsymbol{\Psi}_1(\cdot))} \tabularnewline [2.5ex]
	\texttt{K\_gauss\_kuma} & 2 & \textup{Kum} & \boldsymbol{F}^{-1}_\textup{Kum} & \mathcal{N}(\boldsymbol{0},\mathsf{I}) &  \eqref{eq:exampleVarTransGauss} &
	\displaystyle \frac{g(\boldsymbol{\Psi}_2(\boldsymbol{\Psi}_1(\cdot))) \lambda(\boldsymbol{\Psi}_2(\boldsymbol{\Psi}_1(\cdot))) }{\lambda_1(\boldsymbol{\Psi}_1(\cdot)) \lambda_2(\boldsymbol{\Psi}_2(\boldsymbol{\Psi}_1(\cdot)))} 
   \\ \hline %\bottomrule
\end{array}
\]

Here $\textup{Kum}$ denotes the multivariate Kumaraswamy distribution with independent marginals, and $ \boldsymbol{F}^{-1}_{\textup{Kum}}$ denotes the element-wise inverse cumulative distribution function.  The  code below evaluates the Keister integral \eqref{eq:KeisterIntegral} for $d=1$ and error tolerance $\varepsilon = 5 \times 10^{-8}$.  The timings for each of these different integrands are displayed.

\lstinputlisting[style=Python]{QMCSoftwareArticle/python/composed_IS.txt}

%Successful importance sampling places more points where the original integrand, $g$, varies more.  Equivalently, 
Successful importance sampling makes the transformed integrand, $f$, more flat. The shorter cubature times correspond to flatter integrands, as  illustrated in Fig.~\ref{fig:mIS}. The above example uses $d=1$ to facilitate the plot in 
Fig.~\ref{fig:mIS}; however, the same example works for arbitrary dimensions.
\begin{figure}[t]
    \centering
	\includegraphics[width=.75\textwidth]{QMCSoftwareArticle/vfigs/multiple_IS.pdf}
	\caption{Keister functions with and without importance sampling (\texttt{IS}). Note that the Keister functions using importance sampling are generally less variable and therefore easier to integrate, as evidenced by the faster integration times.} \label{fig:mIS}
\end{figure}

%Automatic Bayesian stopping criteria choose $n$ needed to satisfy the error tolerance via a Bayesian-credible interval for $\mu$ \cite{RatHic19a,JagHic22a}.  These are also options in QMCPy.

\section{Further Work} \label{sec:further}

QMCPy is ripe for growth and development in several areas.  We hope that the QMC community will join us in making this a reality.

Multi-level (quasi-)Monte Carlo (ML(Q)MC) methods make possible the computation of expectations of functionals of stochastic differential equations and partial differential equations with random coefficients.  Such use cases appear in quantitative finance and geophysical applications.  QMCPy's ML(Q)MC's capability is rudimentary, but under active development.

We hope to add a greater variety of use cases and are engaging collaborators to help.  Sobol' indices, partial differential equations with random coefficients, expected improvement measures for Bayesian optimization, and multivariate probabilities are some of those on our radar.

Recently, several QMC experts have focused on developing LD generators for Python. 
Well-established packages such as SciPy \cite{SCIPY} and PyTorch \cite{PyTorch} are have developed QMC modules that support numerous LD sequences and related functionalities. We plan to integrate the routines as optional backends for QMCPy's LD generators. Creating ties to these other packages will allow users to call their preferred generators from within the QMCPy framework.  Moreover, as features in QMCPy become more common and prove their value, we will try to incorporate them into SciPy and other popular, general-purpose packages.

We also plan to expand our library of digital net generating matrices. We wish to incorporate interlaced digital nets, polynomial lattices, and Niederreiter sequences, among others. By including high-quality defaults in QMCPy, we hope to make these sequences more readily available to the public.

Our \texttt{DiscreteDistrution} places equal weights on each support point, $\boldsymbol{X}_i$.  In the future, we might generalize this to unequal weights.

QMCPy already includes importance sampling, but the choice of sampling distribution must be chosen a priori.  We would like to see an automatic, adaptive choice following the developments of \cite{AsmGly2007a,approx_zero_variance_simulation,OweZho00a}.

Control variates can be useful for QMC as well as for IID MC \cite{HicEtal03}.  These should be incorporated into QMCPy in a seamless way.

We close with an invitation.  Try QMCPy.  If you find bugs or missing features, please submit an issue to https://github.com/QMCSoftware/QMCSoftware/issues.  
%\href{https://github.com/QMCSoftware/QMCSoftware/issues}{\nolinkurl{https://github.com/QMCSoftware/QMCSoftware/issues}}.  
If you wish to add your great algorithm or use case, please submit a pull request to our GitHub repository at https://github.com/QMCSoftware/QMCSoftware/pulls. 
%\href{https://github.com/QMCSoftware/QMCSoftware/pulls}{\nolinkurl{https://github.com/QMCSoftware/QMCSoftware/pulls}}.  
We hope that the community will embrace QMCPy.

\begin{acknowledgement}
The authors would like to thank the organizers for a wonderful MCQMC 2020. 
We also thank the referees for their many helpful suggestions.  This work is supported in part by SigOpt and National Science Foundation grant DMS-1522687.
\end{acknowledgement}

%\section*{References}
%\nocite{*}
%\bibliographystyle{spmpsci.bst}
%\bibliography{FJH23,FJHown23,QMCSoftwareArticle,sc}

\begin{thebibliography}{10}
\providecommand{\url}[1]{{#1}}
\providecommand{\urlprefix}{URL }
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{DOI~\discretionary{}{}{}#1}\else
  \providecommand{\doi}{DOI~\discretionary{}{}{}\begingroup
  \urlstyle{rm}\Url}\fi

\bibitem{tensorflow2015-whitepaper}
Abadi, M., Agarwal, A., Barham, P., Brevdo, E., Chen, Z., Citro, C., Corrado,
  G.S., Davis, A., Dean, J., Devin, M., Ghemawat, S., Goodfellow, I., Harp, A.,
  Irving, G., Isard, M., Jia, Y., Jozefowicz, R., Kaiser, L., Kudlur, M.,
  Levenberg, J., Man\'{e}, D., Monga, R., Moore, S., Murray, D., Olah, C.,
  Schuster, M., Shlens, J., Steiner, B., Sutskever, I., Talwar, K., Tucker, P.,
  Vanhoucke, V., Vasudevan, V., Vi\'{e}gas, F., Vinyals, O., Warden, P.,
  Wattenberg, M., Wicke, M., Yu, Y., Zheng, X.: {TensorFlow}: Large-scale
  machine learning on heterogeneous systems (2015).
\newblock \urlprefix\url{https://www.tensorflow.org/}.
\newblock Software available from tensorflow.org

\bibitem{AsmGly2007a}
Asmussen, S., Glynn, P.: Stochastic Simulation: Algorithms and Analysis,
  \emph{Stochastic Modelling and Applied Probability}, vol.~57.
\newblock Springer-Verlag, New York (2007).
\newblock \doi{10.1007/978-0-387-69033-9}

\bibitem{bezanson2012julia}
Bezanson, J., Karpinski, S., Shah, V.B., Edelman, A.: Julia: {A} fast dynamic
  language for technical computing.
\newblock arXiv preprint arXiv:1209.5145  (2012)

\bibitem{Bur20a}
Burkhardt, J.: Various software (2020).
\newblock \urlprefix\url{http://people.sc.fsu.edu/~jburkardt/}

\bibitem{ChoEtal21a}
Choi, S.C.T., Ding, Y., Hickernell, F.J., Jiang, L., {Jim\'enez Rugama},
  {\relax Ll}.A., Li, D., Jagadeeswaran, R., Tong, X., Zhang, K., Zhang, Y.,
  Zhou, X.: {GAIL}: {G}uaranteed {A}utomatic {I}ntegration {L}ibrary (versions
  1.0--2.3.2).
\newblock MATLAB software, \url{http://gailgithub.github.io/GAIL\_Dev/} (2021).
\newblock \doi{10.5281/zenodo.4018189}

\bibitem{QMCPy2020a}
Choi, S.C.T., Hickernell, F.J., Jagadeeswaran, R., McCourt, M., Sorokin, A.:
  {QMCPy}: A quasi-{M}onte {C}arlo {P}ython library (2020).
\newblock \doi{10.5281/zenodo.3964489}.
\newblock \urlprefix\url{https://qmcsoftware.github.io/QMCSoftware/}

\bibitem{QMCPyDocs}
Choi, S.C.T., Hickernell, F.J., Jagadeeswaran, R., McCourt, M.J., Sorokin,
  A.: {QMCPy} documentation (2020).
\newblock \urlprefix\url{https://qmcpy.readthedocs.io/en/latest/}

\bibitem{doi:10.1137/06065074X}
Cools, R., Kuo, F.Y., Nuyens, D.: Constructing embedded lattice rules for
  multivariate integration.
\newblock SIAM Journal on Scientific Computing \textbf{28}(6), 2162--2188
  (2006).
\newblock \doi{10.1137/06065074X}

\bibitem{LatNet}
Darmon, Y., Godin, M., L'Ecuyer, P., Jemel, A., Marion, P., Munger, D.:
  {LatNet} builder (2018).
\newblock \urlprefix\url{https://github.com/umontreal-simul/latnetbuilder}

\bibitem{DicEtal14a}
Dick, J., Kuo, F., Sloan, I.H.: High dimensional integration --- the
  {Q}uasi-{M}onte {C}arlo way.
\newblock Acta Numer. \textbf{22}, 133--288 (2013).
\newblock \doi{10.1017/S0962492913000044}

\bibitem{GilesSoft}
Giles, M.: Multi-level (quasi-){M}onte {C}arlo software (2020).
\newblock \urlprefix\url{https://people.maths.ox.ac.uk/gilesm/mlmc/}

\bibitem{tfqf21}
{Google Inc.}: {TF Quant Finance}: Tensorflow based {Q}uant {F}inance {L}ibrary
  (2021).
\newblock \urlprefix\url{https://github.com/google/tf-quant-finance}

\bibitem{numpy}
Harris, C.R., Millman, K.J., van~der Walt, S.J., Gommers, R., Virtanen, P.,
  Cournapeau, D., Wieser, E., Taylor, J., Berg, S., Smith, N.J., Kern, R.,
  Picus, M., Hoyer, S., van Kerkwijk, M.H., Brett, M., Haldane, A., del
  R{\'{i}}o, J.F., Wiebe, M., Peterson, P., G{\'{e}}rard-Marchant, P.,
  Sheppard, K., Reddy, T., Weckesser, W., Abbasi, H., Gohlke, C., Oliphant,
  T.E.: Array programming with {NumPy}.
\newblock Nature \textbf{585}(7825), 357--362 (2020).
\newblock \doi{10.1038/s41586-020-2649-2}.
\newblock \urlprefix\url{https://doi.org/10.1038/s41586-020-2649-2}

\bibitem{HCJJ14}
Hickernell, F.J., Choi, S.C.T., Jiang, L., Jim{\'e}nez~Rugama, L.A.: {M}onte
  {C}arlo simulation, automatic stopping criteria for.
\newblock Wiley StatsRef: Statistics Reference Online pp. 1--7 (2014)

\bibitem{QMCPyTutColab2020}
Hickernell, F.J., Sorokin, A.: {Q}uasi-{M}onte {C}arlo {(QMC)} software in
  {QMCPy} {G}oogle {C}olaboratory notebook (2020).
\newblock \urlprefix\url{http://tinyurl.com/QMCPyTutorial}

\bibitem{QMCPyTutColab2020_paper}
Hickernell, F.J., Sorokin, A.: {Q}uasi-{M}onte {C}arlo {(QMC)} software in
  {QMCPy} {G}oogle {C}olaboratory notebook for {MCQMC2020} article (2020).
\newblock \urlprefix\url{https://tinyurl.com/QMCPyArticle2021}

\bibitem{Hic97a}
Hickernell, F.J.: A generalized discrepancy and quadrature error bound.
\newblock Math.\ Comp. \textbf{67}, 299--322 (1998).
\newblock \doi{10.1090/S0025-5718-98-00894-1}

\bibitem{Hic99a}
Hickernell, F.J.: Goodness-of-fit statistics, discrepancies and robust
  designs.
\newblock Statist.\ Probab.\ Lett. \textbf{44}, 73--78 (1999).
\newblock \doi{10.1016/S0167-7152(98)00293-4}

\bibitem{HicEtal14a}
Hickernell, F.J., Jiang, L., Liu, Y., Owen, A.B.: Guaranteed conservative
  fixed width confidence intervals via {M}onte {C}arlo sampling.
\newblock In: J.~Dick, F.Y. Kuo, G.W. Peters, I.H. Sloan (eds.) {M}onte {C}arlo
  and Quasi-{M}onte {C}arlo Methods 2012, \emph{Springer Proceedings in
  Mathematics and Statistics}, vol.~65, pp. 105--128. Springer-Verlag, Berlin
  (2013).
\newblock \doi{10.1007/978-3-642-41095-6}

\bibitem{HicJim16a}
Hickernell, F.J., {Jim\'enez Rugama}, {\relax Ll}.A.: Reliable adaptive
  cubature using digital sequences.
\newblock In: R.~Cools, D.~Nuyens (eds.) {M}onte {C}arlo and Quasi-{M}onte
  {C}arlo Methods: {MCQMC}, {L}euven, {B}elgium, {A}pril 2014, \emph{Springer
  Proceedings in Mathematics and Statistics}, vol. 163, pp. 367--383.
  Springer-Verlag, Berlin (2016).
\newblock ArXiv:1410.8615 [math.NA]

\bibitem{HicEtal17a}
Hickernell, F.J., {Jim\'enez Rugama}, {\relax Ll}.A., Li, D.: Adaptive
  quasi-{M}onte {C}arlo methods for cubature.
\newblock In: J.~Dick, F.Y. Kuo, H.~Wo\'zniakowski (eds.) Contemporary
  Computational Mathematics --- a celebration of the 80th birthday of {I}an
  {S}loan, pp. 597--619. Springer-Verlag (2018).
\newblock \doi{10.1007/978-3-319-72456-0}

\bibitem{HicEtal03}
Hickernell, F.J., Lemieux, C., Owen, A.B.: Control variates for
  quasi-{M}onte {C}arlo.
\newblock Statist.\ Sci. \textbf{20}, 1--31 (2005).
\newblock \doi{10.1214/088342304000000468}

\bibitem{QRNG2020}
Hofert, M., Lemieux, C.: qrng {R} package (2017).
\newblock \urlprefix\url{https://cran.r-project.org/web/packages/qrng/qrng.pdf}

\bibitem{RatHic19a}
Jagadeeswaran, R., Hickernell, F.J.: Fast automatic {B}ayesian cubature
  using lattice sampling.
\newblock Stat.\ Comput. \textbf{29}, 1215--1229 (2019).
\newblock \doi{10.1007/s11222-019-09895-9}

\bibitem{JagHic22a}
Jagadeeswaran, R., Hickernell, F.J.: Fast automatic {B}ayesian cubature
  using {S}obol' sampling (2021+).
\newblock In preparation for submission for publication

\bibitem{JimHic16a}
{Jim\'enez Rugama}, {\relax Ll}.A., Hickernell, F.J.: Adaptive
  multidimensional integration based on rank-1 lattices.
\newblock In: R.~Cools, D.~Nuyens (eds.) {M}onte {C}arlo and Quasi-{M}onte
  {C}arlo Methods: {MCQMC}, {L}euven, {B}elgium, {A}pril 2014, \emph{Springer
  Proceedings in Mathematics and Statistics}, vol. 163, pp. 407--422.
  Springer-Verlag, Berlin (2016).
\newblock ArXiv:1411.1966

\bibitem{Kei96}
Keister, B.D.: Multidimensional quadrature algorithms.
\newblock Computers in Physics \textbf{10}, 119--122 (1996).
\newblock \doi{10.1063/1.168565}

\bibitem{BRODA20a}
Kucherenko, S.: {BRODA} (2020).
\newblock \urlprefix\url{https://www.broda.co.uk/index.html}

\bibitem{Kumaraswamy}
{Kumaraswamy}, P.: {A generalized probability density function for
  double-bounded random processes}.
\newblock Journal of Hydrology \textbf{46}(1), 79--88 (1980).
\newblock \doi{10.1016/0022-1694(80)90036-0}

\bibitem{KuoNuy16a}
Kuo, F.Y., Nuyens, D.: Application of quasi-{M}onte {C}arlo methods to elliptic
  {PDE}s with random diffusion coefficients -- a survey of analysis and
  implementation.
\newblock Found.\ Comput.\ Math. \textbf{16}, 1631--1696 (2016).
\newblock \urlprefix\url{https://people.cs.kuleuven.be/~dirk.nuyens/qmc4pde/}

\bibitem{lataniotis2021uncertainty}
Lataniotis, C., Marelli, S., Sudret, B.: Uncertainty quantification in the
  cloud with {UQCloud}.
\newblock In: 4th International Conference on Uncertainty Quantification in
  Computational Sciences and Engineering (UNCECOMP 2021), pp. 209--217 (2021)

\bibitem{SSJ}
L'Ecuyer, P.: {SSJ: Stochastic Simulation in Java} (2020).
\newblock \urlprefix\url{https://github.com/umontreal-simul/ssj}

\bibitem{LEcEtal22a}
L'Ecuyer, P., Marion, P., Godin, M., Puchhammer, F.: A tool for custom
  construction of {QMC} and {RQMC} point sets.
\newblock In: E.~Arnaud, M.~Giles, A.~Keller (eds.) {M}onte {C}arlo and
  Quasi-{M}onte {C}arlo Methods: {MCQMC}, {O}xford, 2020 (2021+)

\bibitem{LEcMun14a}
L'Ecuyer, P., Munger, D.: Algorithm 958: {Lattice Builder}: A general software
  tool for constructing rank-1 latice rules.
\newblock ACM Trans.\ Math.\ Software \textbf{42}, 1--30 (2016)

\bibitem{approx_zero_variance_simulation}
L'Ecuyer, P., Tuffin, B.: Approximate zero-variance simulation.
\newblock In: Proceedings of the 40th Conference on Winter Simulation, WSC '08,
  p. 170–181. Winter Simulation Conference (2008)

\bibitem{UQLab2014}
Marelli, S., Sudret, B.: {UQLab}: {A} framework for uncertainty quantification
  in {MATLAB}.
\newblock In: The 2nd International Conference on Vulnerability and Risk
  Analysis and Management {(ICVRAM 2014)}, pp. 2554--2563. ASCE Library (2014).
\newblock \urlprefix\url{https://www.uqlab.com}

\bibitem{Mat98}
Matou\v{s}ek, J.: On the {$L_2$}-discrepancy for anchored boxes.
\newblock J. Complexity \textbf{14}, 527--556 (1998)

\bibitem{Nie92}
Niederreiter, H.: Random Number Generation and Quasi-{M}onte {C}arlo Methods.
\newblock CBMS-NSF Regional Conference Series in Applied Mathematics. SIAM,
  Philadelphia (1992)

\bibitem{Nuy17a}
Nuyens, D.: Magic point shop (2017).
\newblock
  \urlprefix\url{https://people.cs.kuleuven.be/~dirk.nuyens/qmc-generators/}

\bibitem{OpenTURNS}
{OpenTURNS Developers}: An open source initiative for the {T}reatment of
  {U}ncertainties, {R}isks '{N} {S}tatistics (2020).
\newblock \urlprefix\url{http://www.openturns.org}

\bibitem{owen1998scrambling}
Owen, A.B.: Scrambling {S}obol' and {N}iederreiter--{X}ing points.
\newblock Journal of Complexity \textbf{14}(4), 466--489 (1998)

\bibitem{owen2020dropping}
Owen, A.B.: On dropping the first {S}obol' point (2020).
\newblock ArXiv:2008.08051 [math.NA]

\bibitem{Owe20a}
Owen, A.B.: Randomized {H}alton sequences in {R} (2020).
\newblock \urlprefix\url{http://statweb.stanford.edu/~owen/code/}

\bibitem{owen2021r}
Owen, A.B.: About the {R} function: rsobol  (2021).
\newblock \urlprefix\url{https://statweb.stanford.edu/~owen/reports/seis.pdf}

\bibitem{OweZho00a}
Owen, A.B., Zhou, Y.: Safe and effective importance sampling.
\newblock J. Amer.\ Statist.\ Assoc. \textbf{95}, 135--143 (2000)

\bibitem{NEURIPS2019_9015}
Paszke, A., Gross, S., Massa, F., Lerer, A., Bradbury, J., Chanan, G., Killeen,
  T., Lin, Z., Gimelshein, N., Antiga, L., Desmaison, A., Kopf, A., Yang, E.,
  DeVito, Z., Raison, M., Tejani, A., Chilamkurthy, S., Steiner, B., Fang, L.,
  Bai, J., Chintala, S.: {PyTorch}: An imperative style, high-performance deep
  learning library.
\newblock In: H.~Wallach, H.~Larochelle, A.~Beygelzimer, F.~d\textquotesingle
  Alch\'{e}-Buc, E.~Fox, R.~Garnett (eds.) Advances in Neural Information
  Processing Systems 32, pp. 8024--8035. Curran Associates, Inc. (2019).
\newblock
  \urlprefix\url{http://papers.neurips.cc/paper/9015-pytorch-an-imperative-style-high-performance-deep-learning-library.pdf}

\bibitem{PyTorch}
{PyTorch Developers}: {PyTorch} (2020).
\newblock \urlprefix\url{https://pytorch.org}

\bibitem{Rob20a}
Robbe, P.: Low discrepancy sequences in {J}ulia (2020).
\newblock \urlprefix\url{https://github.com/PieterjanRobbe/QMC.jl}

\bibitem{Rob21}
Robbe, P.: Multilevel {M}onte {C}arlo simulations in {J}ulia (2021).
\newblock
  \urlprefix\url{https://github.com/PieterjanRobbe/MultilevelEstimators.jl}

\bibitem{SCIPY}
{SciPy Developers}: {SciPy Ecosystem} (2018).
\newblock \urlprefix\url{www.scipy.org}

\bibitem{MAT9.10}
{The MathWorks, Inc.}: {MATLAB} R2021a.
\newblock Natick, MA (2020)

\bibitem{2020SciPy-NMeth}
Virtanen, P., Gommers, R., Oliphant, T.E., Haberland, M., Reddy, T.,
  Cournapeau, D., Burovski, E., Peterson, P., Weckesser, W., Bright, J., {van
  der Walt}, S.J., Brett, M., Wilson, J., Millman, K.J., Mayorov, N., Nelson,
  A.R.J., Jones, E., Kern, R., Larson, E., Carey, C.J., Polat, {\.I}., Feng,
  Y., Moore, E.W., {VanderPlas}, J., Laxalde, D., Perktold, J., Cimrman, R.,
  Henriksen, I., Quintero, E.A., Harris, C.R., Archibald, A.M., Ribeiro, A.H.,
  Pedregosa, F., {van Mulbregt}, P., {SciPy 1.0 Contributors}: {{SciPy} 1.0:
  Fundamental Algorithms for Scientific Computing in Python}.
\newblock Nature Methods \textbf{17}, 261--272 (2020).
\newblock \doi{10.1038/s41592-019-0686-2}

\end{thebibliography}

\end{document}

### 2022-10-31 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

--- 2022-10-31 MCQMC2022QMCSoftwareArticle.tex
@@
 
-\title*{Quasi-Monte Carlo Software}
+\title*{Challenges in Developing Great MCQMC Software}
 \authorrunning{S.-C.\ T.\ Choi et al}
-\author{Sou-Cheng Terrya Choi \and Fred J. Hickernell \and R. Jagadeeswaran \and Michael J. McCourt \and Aleksei G. Sorokin}
+\author{Sou-Cheng Terrya Choi \and Fred J. Hickernell \and Aleksei G. Sorokin}
 \institute{Sou-Cheng Terrya Choi \at Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{schoi32@hawk.iit.edu}
@@
 
-\abstract{This is the article based on my MCQMC 2020 Tutorial.
-\AGSComment{Latest code demo at: \href{https://colab.research.google.com/drive/1wijwt0ugsJ9chUHtzSb3lQpc3-N8mCDs?usp=sharing}{https://colab.research.google.com/drive/1wijwt0ugsJ9chUHtzSb3lQpc3-N8mCDs?usp=sharing}}.}
+\abstract{}
+
+\end{document}
+\section{Introduction}
+
+\begin{frame}{A Great MCQMC Software Ecosystem Should Be}
+
+	\vspace{-3ex}
+	\begin{itemize}
+		\item \alert{Correct}, e.g., not omit the zeroth point in a low discrepancy sequence \parencite{Owe22a, scipySobol2020a}
+
+		\item  \alert{Complete}---contain the components or  easily access components in other libraries to solve real, complex problems
+
+		\item  \alert{Accessible}---downloadable; tutorials, demos, discussion forums, etc.\ for (new) users; written in a language that potential users speak; provide a consistent user interface
+
+		\item \alert{Efficient}---in terms of compute time and memory
+
+		\item \alert{Current}---include the latest and best algorithms
+
+		\item  \alert{Sustainable}---have a sufficient user base and developer community for updates and maintenance
+
+		\item \alert{Scalable}---take advantage of advanced computer architectures for speed and to solve large problems
+
+	\end{itemize}
+\end{frame}
+
+\begin{frame}[allowframebreaks]{Selective History of Publicly Available MC\textbf{QMC} Software}
+\vspace{-2ex}
+	\begin{itemize}
+		\item \textcite{LEc2017a} provides a history of \alert{random number generation}
+		\item Notorious \alert{RANDU} from the 1960s \parencite{RANDU}, which fails the spectral test
+		\item \alert{VEGAS} \parencite{Lep78a, Lep21a} importance sampling Monte Carlo algorithm favored by physicists
+		\item \alert{ACM} low discrepancy point generators \parencite{BraFox88,BraFoxNie92,HonHic00a}
+		\item \alert{FinDer} \parencite{PasTra95,FinDer} and \alert{BRODA} \parencite{BRODA20a} targeting quantitative finance
+		\item Korobov cubature and scrambled Sobol' generators in \alert{NAG} \parencite{NAG27} for decades		
+		\item Scrambled Sobol' and Halton generators in \alert{MATLAB} \parencite{MAT9.13} since 2008, and fixed a few years later
+		\item \alert{BUGS} \parencite{BUGSBook, BUGSweb} and \alert{Stan} \parencite{STAN} for Markov Chain Monte Carlo (MCMC)
+		\item \alert{LatMRG} \parencite{LEcCou97}, \alert{RNGStreams} \parencite{LEcEtal02},  \alert{SSJ} \parencite{LEc2002a,SSJ}, and \alert{LatNetBuilder} \parencite{LatNet} \talktime{M 11}
+		\item Low discrepancy generators \parencite{FriKel02,FriKelweb}, \alert{SamplePack} \parencite{SamplePack}, \textcite{GruWeb},  and \alert{MatBuilder} \parencite{paulin2022} \talktime{M 12, Tu 12}
+		\item Sobol' \alert{direction numbers} \parencite{JoeKuo03,JoeKuo08,SobolDirection}
+		\item \alert{Fast CBC}, \alert{Magic Point Shop}, \alert{QMC4PDE} and other code since 2004  at \textcite{NuyWeb} \talktime{Tu 10:30}
+		\item \alert{Multi-level} software \parencite{GilesSoft,GilesQSoft} \talktime{M 10:30}
+		\item Data-driven error bounds and stopping criteria in \alert{GAIL} \parencite{ChoEtal21a} and \alert{QMCPy} \parencite{QMCPy2020a, ChoEtal22a}  \talktime{Tu 11}
+		\item Uncertainty quantification libraries \alert{Dakota} \parencite{DakotaUsersManual}, \alert{UQTk} \parencite{DebEtal04,UQTk}, and \alert{MUQ} \parencite{MUQ} \talktime{Tu 11:30} have some basic level low discrepancy sampling
+	    \item QMC framework in Julia since 2019 \parencite{QMCJulia} by Robbe and others \talktime{M 17}
+		\item Scrambled Sobol' in \alert{SciPy} \parencite{virtanen2020scipy} and \alert{PyTorch} \parencite{paszke2019pytorch} since several years ago
+		\item \alert{TensorFlow} QMC framework \parencite{tfqfQMC2021a} since a year or so ago
+	\end{itemize}
+
+We have stand-alone algorithms, small libraries, and pieces of larger libraries
+\end{frame}
+
+\begin{frame}{Why Do We Need Software?}
+	\vspace{-4ex}
+	\begin{itemize}
+		\item \alert{Theory} 
+		\begin{itemize}
+			\item Explains, justifies, and leads to \alert{algorithms}
+			\item Assumptions$\implies$success can be viewed as failure$\implies$what went wrong
+		\end{itemize}
+		\item \alert{Software} 
+		\begin{itemize}
+			\item Makes \alert{algorithms} practical
+			\item Solves (new) \alert{applications}
+			\item Eliminates \alert{do-it-yourself}
+		\end{itemize}
+		\item \alert{Applications} 
+  		\begin{itemize}
+			\item Make \alert{societal} impact
+			\item Inspire new \alert{theory}
+		\end{itemize}
+
+	\end{itemize}
+
+\uncover<2->{The US Department of Energy is investing in studying how to develop great scientific software, including a recent \emph{Workshop on the Science of Scientific-Software Development and Use} \parencite{Her19a,ASCR-SSSDU}}
+\end{frame}
+
+\section{Structure}
+
+\begin{frame}{Software Architecture}
+	\vspace{-6ex}
+
+\[
+\mu :=  \int_{\ct} g(\vt) \, \lambda(\vt) \, \dif \vt = \cdots = 	
+\underbrace{\Ex[f(\vX)]}_{\text{expectation}} = \underbrace{\int_{[0,1]^d}  f(\vx) \, \dif \vx}_{\text{integration}} \approx  \frac 1{n} \sum_{i=1}^{n} f(\vX_i) =: \hmu_{n}
+\]
+
+\vspace{-3ex}
+
+\begin{description}[alabel]
+	%\setlength{\labelwidth}{5ex}
+
+	\item[Low Discrepancy Generator] producing $\{\vX_1, \vX_2, \dots \}$ that mimics the uniform distribution
+
+	\item[True Measure] that defines the original integral, e.g., Lebesgue; embodies the transformation $\vt = \vPsi(\vx)$
+
+	\item[Integrand] $g$, which defines the original integral, plus the transformed version, $f$, to fit the LD generator
+
+	\item[Stopping Criterion] based on a data-driven error bound, which determines how large $n$ should be to ensure that $\abs{\mu - \hmu_n} \le \varepsilon$
+\end{description}
+
+\vspace{-1ex}
+\uncover<2->{May we agree on a \alert{common software definition} for these these objects like we have for floating point numbers or basic mathematical functions?}
+\end{frame}
+
+\begin{frame}{Choosing a Language, Library, or Environment}
+	\vspace{-5ex}
+	\begin{itemize}
+    \item Language/library choices of the users are driven by \alert{familiarity} and \alert{speed}
+
+  	\item  Favorites aligned with various \alert{communities}, but no one dominates all
+
+	\item Difference between \alert{prototyping} environments and ninja written \alert{production} code
+
+    \item Growth of \alert{multi-processor} environments 
+    \begin{itemize}
+    	\item Provides opportunity
+    	\item Complicates software development; where does parallelism really help?
+    \end{itemize}
+
+    \item<2-> Recommendations if you want to contribute software
+    \begin{itemize}
+    \item Realize your new idea as a \alert{demo} in your favorite library
+    \item Add a significant idea as a \alert{feature} to other libraries
+    \item Move basic features of your library to \alert{larger} library
+    \item Write wrappers or demos to \alert{connect} other libraries to your library
+
+    \end{itemize}
+ \end{itemize}
+\end{frame}
+
+\begin{frame}{Ex.\ QMCPy $+$ FEniCS/Dolfin \parencite{LoggEtal_10_2012}}
+\vspace{-3ex}
+	\begin{itemize}
+		\item Want to demonstrate how our QMC software could work with a popular DE solver for DEs with random coefficients
+		\item Partial success \only<3->{\alert{Why?}}
+	\end{itemize}
+ 	\only<1>{\vspace{-3ex}}\only<2>{\vspace{-2ex}}\only<3>{\vspace{0ex}}
+ \begin{minipage}{0.48\textwidth}
+ 	\only<1-2>{\begin{gather*}
+ 		-\frac{\dif }{\dif x}\biggl(a(x,\vW) \frac{\dif}{\dif x} u(x,\vW) \biggr) = 200( 2x - 1)\\
+ 		 0 \leq x \leq 1, \qquad
+ 		u(0) = u(1) = 0 \\
+ 		a(x,\vW) = 1 + 0.6 \sum_{k=1}^d \frac{W_k T_{k}(2x-1)}{k^2} \\
+ 		 \vW \sim \cu[-1,1]^d \\
+ 		 \Ex[u(0.25,\vW)] = \only<1>{\, ?}\only<2>{ -1.7462}
+ 	\end{gather*}}
+  \only<3->{
+\begin{itemize}
+\item Even with FEniCS/Dolfin's extensive documentation we had \alert{difficulties}
+\begin{itemize}
+    \item Did not understand the latest version, so reverted to an older version
+    \item Had to learn how to change the random instance of the coefficient without tiggering the JIT compiler
+    \item Do not yet know how to express the random coefficient in terms of the covariance kernel
+\end{itemize}
+\end{itemize}
+  }
+ \end{minipage}%
+\hfill
+ \begin{minipage}{0.48\textwidth}
+ 	\only<1-2>{\centering
+ 	\includegraphics<1>[width=\textwidth]{ProgramsImages/axwtrue0.6.eps}
+ 	\includegraphics<2->[width=\textwidth]{ProgramsImages/timing-4.eps}}%
+  \only<3->{
+\begin{itemize}
+\item But
+\begin{itemize}
+    \item It  works
+    \item If we can overcome the challenges, will post a blog and demo at \textcite{QMCBlog}
+\end{itemize}
+\end{itemize}
+
+  }
+\end{minipage}
+
+\end{frame}
+
+\section{Support}
+\begin{frame}{Encouraging Users of Your Library Requires}
+
+	\vspace{-4ex}
+	\begin{itemize}
+		\item A repository where
+		\begin{itemize}
+		\item  Your library can be easily downloaded
+		\item There is documentation
+		\item Issues or queries can be posted
+		\item There are updates
+		\item Tests are run regularly to mitigate against bugs
+		\end{itemize}
+		\item (Elementary) \alert{demos} on how to use the software and highlighting  its advantages
+		\item Demos \alert{connecting}  your library with others
+		\item Recommending your students and postdocs to \alert{use} your library
+		\item \alert{Welcoming} instructions on how to contribute an algorithm or a demo
+	\end{itemize}
+
+\end{frame}
+
+\begin{frame}{Building a Developer Team for Your Library}
+
+	\vspace{-4ex}
+\begin{itemize}
+
+\item Many senior researchers \alert{don't code}, and rely on transient team members (students, postdocs)
+\begin{itemize}
+	\item Encourage a sense of \alert{ownership} that lasts beyond the time your team members are part of your lab
+	\item Encourage those \alert{outside} your own group to contribute
+\end{itemize}
+
+\item Those of us who understand theory and appreciate software need to encourage and steer popular software to ensure that it is implemented well.
+
+\item Recognize good software as \alert{valuable} scholarly output
+
+\item Recognize \alert{research software engineer} as a valued vocation; see \alert{Research Software Engineers International} 	\url{https://researchsoftware.org}, \emph{Research Software Engineers are people who combine professional software expertise with an understanding of research.
+They go by various job titles but the term Research Software Engineer (RSE) is fast gaining international recognition.}
+  \end{itemize}
+\end{frame}
+
+\section{Next Steps}
+
+\begin{frame}{Next Steps}
+
+	\vspace{-2ex}
+	\begin{minipage}{0.48\textwidth}
+		\begin{itemize}
+			\item Which exemplary \alert{theoretical developments} are not yet implemented in readily available software?
+			\item What potential MCQMC user communities are \alert{ill-served} by existing  software libraries?
+			\item Should those of us interested in software gather regularly for \alert{show-and-tell} to foster more collaboration?
+		\end{itemize}
+	\end{minipage}
+	\begin{minipage}{0.48\textwidth}
+		Join us for a \alert{lunch} discussion at Bella Casa tomorrow, July 19, at 12:45 PM
+
+		\vspace{2ex}
+
+		\includegraphics[width=\textwidth]{ProgramsImages/PizeriaBellaCasa.png}
+\end{minipage}
+
+\end{frame}
+
+\begin{frame}[allowframebreaks]{References}
+	\printbibliography
+\end{frame}
 

### 2022-11-01 MCQMC2022ChallengesQMCSoftwareArticle.tex
Office Computer Updates

--- 2022-10-31 MCQMC2022QMCSoftwareArticle.tex
@@
 %%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[graybox,footinfo]{svmult}
+%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% sample root file for your "contribution" to a contributed volume
+%
+% Use this file as a template for your own input.
+%
+%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\smartqed
-\usepackage{mathptmx}       % selects Times Roman as basic font
-\usepackage{helvet}         % selects Helvetica as sans-serif font
-\usepackage{courier}        % selects Courier as typewriter font
+% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass[graybox]{svmult}
+
+% choose options for [] as required from the list
+% in the Reference Guide
+
 \usepackage{type1cm}        % activate if the above 3 fonts are
 % not available on your system
+%
+\usepackage{makeidx}         % allows index generation
 \usepackage{graphicx}        % standard LaTeX graphics tool
 % when including figure files
+\usepackage{multicol}        % used for the two-column index
+\usepackage[bottom]{footmisc}% places footnotes at page bottom
 
-\usepackage{array,colortbl}
-\usepackage{amsmath,amsfonts,amssymb,bm} % no amsthm, Springer defines Theorem, Lemma, etc themselves
-%\usepackage[mathx]{mathabx}
-\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
-\DeclareFontShape{U}{mathx}{m}{n}{
-	<5> <6> <7> <8> <9> <10>
-	<10.95> <12> <14.4> <17.28> <20.74> <24.88>
-	mathx10
-}{}
-\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
-\DeclareFontSubstitution{U}{mathx}{m}{n}
-\DeclareMathAccent{\widecheck}{0}{mathx}{"71}
+\usepackage{newtxtext}       % 
+\usepackage[varvw]{newtxmath}       % selects Times Roman as basic font
 
-% Note that Springer defines the following already:
-%
-% \D upright d for differential d
-% \I upright i for imaginary unit
-% \E upright e for exponential function
-% \tens depicts tensors as sans serif upright
-% \vec depicts vectors as boldface characters instead of the arrow accent
-%
-% Additionally we throw in the following common used macro's:
-\input{macros}
+% see the list of further useful packages
+% in the Reference Guide
 
-% Macros below are now included in macros.tex from MCQMC 2016 web site
-% This spot formerly included macros that are now in macros.tex
+%\makeindex             % used for the subject index
+% please use the style svind.ist with
+% your makeindex program
 
-% indicator boldface 1:
-\DeclareSymbolFont{bbold}{U}{bbold}{m}{n}
-\DeclareSymbolFontAlphabet{\mathbbold}{bbold}
-%\newcommand{\ind}{\mathbbold{1}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\usepackage{microtype} % good font tricks
+%%%%Fred start
+\usepackage{url, xspace}
+\providecommand{\HickernellFJ}{Hickernell\xspace}
 
-\usepackage[colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black]{hyperref}
-\urlstyle{same}
-\usepackage{bookmark}
-\pdfstringdefDisableCommands{\def\and{, }}
-\makeatletter % to avoid hyperref warnings:
-\providecommand*{\toclevel@author}{999}
-\providecommand*{\toclevel@title}{0}
-\makeatother
-
-\usepackage{bbm,mathtools,array,longtable,booktabs,graphicx,color,enumitem,xspace}
-%\input FJHDef.tex
-
-\newcommand{\QMCPYnorm}[2][{}]{\ensuremath{\left \lVert #2 \right \rVert}_{#1}}
-\newcommand{\QMCPYnormnorm}[2][{}]{\ensuremath{\lVert #2 \rVert}_{#1}}
-\newcommand{\QMCPYbignorm}[2][{}]{\ensuremath{\bigl \lVert #2 \bigr \rVert}_{#1}}
-\newcommand{\QMCPYBignorm}[2][{}]{\ensuremath{\Bigl \lVert #2 \Bigr \rVert}_{#1}}
-\newcommand{\QMCPYabs}[1]{\ensuremath{{\left \lvert #1 \right \rvert}}}
-\newcommand{\QMCPYbigabs}[1]{\ensuremath{{\bigl \lvert #1 \bigr \rvert}}}
-
-\providecommand{\HickernellFJ}{Hickernell}
-
-\definecolor{orange}{rgb}{1.0,0.3,0.0}
-\definecolor{violet}{rgb}{0.75,0,1}
-\newcommand{\frednote}[1]{  {\textcolor{red}  {\mbox{**Fred:} #1}}}
-\newcommand{\yuhannote}[1]{ {\textcolor{violet}  {\mbox{**Yuhan:} #1}}}
-\newcommand{\tonynote}[1]{ {\textcolor{orange}  {\mbox{**Tony:} #1}}}
-
-%\journal{Journal of Complexity}
-
-\allowdisplaybreaks[4]
-
-%   python styling
-\usepackage[T1]{fontenc}
-\usepackage{lmodern}
-\usepackage{listings}
-\usepackage{xcolor}
-\definecolor{darkgreen}{rgb}{0,0.6,0}
-\lstdefinestyle{Python}{
-    showstringspaces=false,
-    language        = Python,
-    basicstyle      = \small\ttfamily,
-    morekeywords = {as},
-    keywordstyle    = \color{blue},
-    stringstyle     = \color{darkgreen},
-    commentstyle    = \color{darkgreen}\ttfamily,
-	breaklines = true,
-	postbreak=\text{$\hookrightarrow$\space},
-	% style >>> and ... 
-	%   see: https://tex.stackexchange.com/questions/326655/make-a-keyword-in-listings-enviorment
-	alsoletter = {>,.} ,
-    morekeywords = [2]{>>>,...},
-    keywordstyle = [2]\color{cyan}\bfseries}
-
-\newcommand{\AGSComment}[1]{{\color{cyan} Aleksei: #1}}
-\newcommand{\FJHComment}[1]{{\color{magenta} Fred: #1}}
-
-\newcommand{\hmu}{\widehat{\mu}}
-\newcommand{\IID}{\textup{IID}}
-\newcommand{\LD}{\textup{LD}}
-\newcommand{\IMP}{\textup{imp}}
-\newcommand{\unif}{\textup{unif}}
-\newcommand{\IIDsim}{\overset{\IID}{\sim}}
-\newcommand{\LDsim}{\overset{\LD}{\sim}}
-\newcommand{\dig}{\textup{dig}}
-\newcommand{\mA}{\mathsf{A}}
-\newcommand{\mSigma}{\mathsf{\Sigma}}
-\newcommand{\mLambda}{\mathsf{\Lambda}}
-\newcommand{\mI}{\mathsf{I}}
+%%%%Fred end
 
@@
 
-\title*{Quasi-Monte Carlo Software}
+\title*{Challenges in Developing Great MCQMC Software}
 \authorrunning{S.-C.\ T.\ Choi et al}
-\author{Sou-Cheng Terrya Choi \and Fred J. Hickernell \and R. Jagadeeswaran \and Michael J. McCourt \and Aleksei G. Sorokin}
+\author{Sou-Cheng Terrya Choi \and Yuhan Ding \and Fred J. Hickernell \and Aleksei G. Sorokin}
 \institute{Sou-Cheng Terrya Choi \at Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{schoi32@hawk.iit.edu}
@@
 Department of Applied Mathematics, Illinois Institute of Technology \\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{hickernell@iit.edu}
-\and
-R. Jagadeeswaran \at
-\and
-Michael J. McCourt \at ???
+\and Yuhan Ding \at
 \and 
@@
 
-\abstract{This is the article based on my MCQMC 2020 Tutorial.
-\AGSComment{Latest code demo at: \href{https://colab.research.google.com/drive/1wijwt0ugsJ9chUHtzSb3lQpc3-N8mCDs?usp=sharing}{https://colab.research.google.com/drive/1wijwt0ugsJ9chUHtzSb3lQpc3-N8mCDs?usp=sharing}}.}
+\abstract{}
+
+\section{Introduction}
+
+	\begin{itemize}
+	\item Correct, e.g., not omit the zeroth point in a low discrepancy sequence \cite{Owe22a, scipySobol2020a}
+
+	\item  Complete---contain the components or  easily access components in other libraries to solve real, complex problems
+
+	\item  Accessible---downloadable; tutorials, demos, discussion forums, etc.\ for (new) users; written in a language that potential users speak; provide a consistent user interface
+
+	\item Efficient---in terms of compute time and memory
+
+	\item Current---include the latest and best algorithms
+
+	\item  Sustainable---have a sufficient user base and developer community for updates and maintenance
+
+	\item Scalable---take advantage of advanced computer architectures for speed and to solve large problems
+
+\end{itemize}
+
+	\begin{itemize}
+	\item \cite{LEc2017a} provides a history of random number generation
+	\item Notorious RANDU from the 1960s \cite{RANDU}, which fails the spectral test
+	\item VEGAS \cite{Lep78a, Lep21a} importance sampling Monte Carlo algorithm favored by physicists
+	\item ACM low discrepancy point generators \cite{BraFox88,BraFoxNie92,HonHic00a}
+	\item FinDer \cite{PasTra95,FinDer} and BRODA \cite{BRODA20a} targeting quantitative finance
+	\item Korobov cubature and scrambled Sobol' generators in NAG \cite{NAG27} for decades		
+	\item Scrambled Sobol' and Halton generators in MATLAB \cite{MAT9.13} since 2008, and fixed a few years later
+	\item BUGS \cite{BUGSBook, BUGSweb} and Stan \cite{STAN} for Markov Chain Monte Carlo (MCMC)
+	\item LatMRG \cite{LEcCou97}, RNGStreams \cite{LEcEtal02},  SSJ \cite{LEc2002a,SSJ}, and LatNetBuilder \cite{LatNet} 
+	\item Low discrepancy generators \cite{FriKel02,FriKelweb}, SamplePack \cite{SamplePack}, \cite{GruWeb},  and MatBuilder \cite{paulin2022} 
+	\item Sobol' direction numbers \cite{JoeKuo03,JoeKuo08,SobolDirection}
+	\item Fast CBC, Magic Point Shop, QMC4PDE and other code since 2004  at \cite{NuyWeb} 
+	\item Multi-level software \cite{GilesSoft,GilesQSoft} 
+	\item Data-driven error bounds and stopping criteria in GAIL \cite{ChoEtal21a} and QMCPy \cite{QMCPy2020a, ChoEtal22a}  
+	\item Uncertainty quantification libraries Dakota \cite{DakotaUsersManual}, UQTk \cite{DebEtal04,UQTk}, and MUQ \cite{MUQ}  have some basic level low discrepancy sampling
+	\item QMC framework in Julia since 2019 \cite{QMCJulia} by Robbe and others 
+	\item Scrambled Sobol' in SciPy \cite{virtanen2020scipy} and PyTorch \cite{paszke2019pytorch} since several years ago
+	\item TensorFlow QMC framework \cite{tfqfQMC2021a} since a year or so ago
+\end{itemize}
+
+We have stand-alone algorithms, small libraries, and pieces of larger libraries
+
+\bibliographystyle{spmpsci.bst}
+\bibliography{FJH23,FJHown23,QMCSoftwareArticle,sc}
+
+\end{document}
+
+\begin{frame}[allowframebreaks]{Selective History of Publicly Available MC\textbf{QMC} Software}
+\vspace{-2ex}
+	\begin{itemize}
+		\item \cite{LEc2017a} provides a history of random number generation
+		\item Notorious RANDU from the 1960s \cite{RANDU}, which fails the spectral test
+		\item VEGAS \cite{Lep78a, Lep21a} importance sampling Monte Carlo algorithm favored by physicists
+		\item ACM low discrepancy point generators \cite{BraFox88,BraFoxNie92,HonHic00a}
+		\item FinDer \cite{PasTra95,FinDer} and BRODA \cite{BRODA20a} targeting quantitative finance
+		\item Korobov cubature and scrambled Sobol' generators in NAG \cite{NAG27} for decades		
+		\item Scrambled Sobol' and Halton generators in MATLAB \cite{MAT9.13} since 2008, and fixed a few years later
+		\item BUGS \cite{BUGSBook, BUGSweb} and Stan \cite{STAN} for Markov Chain Monte Carlo (MCMC)
+		\item LatMRG \cite{LEcCou97}, RNGStreams \cite{LEcEtal02},  SSJ \cite{LEc2002a,SSJ}, and LatNetBuilder \cite{LatNet} 
+		\item Low discrepancy generators \cite{FriKel02,FriKelweb}, SamplePack \cite{SamplePack}, \cite{GruWeb},  and MatBuilder \cite{paulin2022} 
+		\item Sobol' direction numbers \cite{JoeKuo03,JoeKuo08,SobolDirection}
+		\item Fast CBC, Magic Point Shop, QMC4PDE and other code since 2004  at \cite{NuyWeb} 
+		\item Multi-level software \cite{GilesSoft,GilesQSoft} 
+		\item Data-driven error bounds and stopping criteria in GAIL \cite{ChoEtal21a} and QMCPy \cite{QMCPy2020a, ChoEtal22a}  
+		\item Uncertainty quantification libraries Dakota \cite{DakotaUsersManual}, UQTk \cite{DebEtal04,UQTk}, and MUQ \cite{MUQ}  have some basic level low discrepancy sampling
+	    \item QMC framework in Julia since 2019 \cite{QMCJulia} by Robbe and others 
+		\item Scrambled Sobol' in SciPy \cite{virtanen2020scipy} and PyTorch \cite{paszke2019pytorch} since several years ago
+		\item TensorFlow QMC framework \cite{tfqfQMC2021a} since a year or so ago
+	\end{itemize}
+
+We have stand-alone algorithms, small libraries, and pieces of larger libraries
+\end{frame}
+
+\begin{frame}{Why Do We Need Software?}
+	\vspace{-4ex}
+	\begin{itemize}
+		\item Theory 
+		\begin{itemize}
+			\item Explains, justifies, and leads to algorithms
+			\item Assumptions$\implies$success can be viewed as failure$\implies$what went wrong
+		\end{itemize}
+		\item Software 
+		\begin{itemize}
+			\item Makes algorithms practical
+			\item Solves (new) applications
+			\item Eliminates do-it-yourself
+		\end{itemize}
+		\item Applications 
+  		\begin{itemize}
+			\item Make societal impact
+			\item Inspire new theory
+		\end{itemize}
+
+	\end{itemize}
+
+\uncover<2->{The US Department of Energy is investing in studying how to develop great scientific software, including a recent \emph{Workshop on the Science of Scientific-Software Development and Use} \cite{Her19a,ASCR-SSSDU}}
+\end{frame}
+
+\section{Structure}
+
+\begin{frame}{Software Architecture}
+	\vspace{-6ex}
+
+\[
+\mu :=  \int_{\ct} g(\vt) \, \lambda(\vt) \, \dif \vt = \cdots = 	
+\underbrace{\Ex[f(\vX)]}_{\text{expectation}} = \underbrace{\int_{[0,1]^d}  f(\vx) \, \dif \vx}_{\text{integration}} \approx  \frac 1{n} \sum_{i=1}^{n} f(\vX_i) =: \hmu_{n}
+\]
+
+\vspace{-3ex}
+
+\begin{description}[alabel]
+	%\setlength{\labelwidth}{5ex}
+
+	\item[Low Discrepancy Generator] producing $\{\vX_1, \vX_2, \dots \}$ that mimics the uniform distribution
+
+	\item[True Measure] that defines the original integral, e.g., Lebesgue; embodies the transformation $\vt = \vPsi(\vx)$
+
+	\item[Integrand] $g$, which defines the original integral, plus the transformed version, $f$, to fit the LD generator
+
+	\item[Stopping Criterion] based on a data-driven error bound, which determines how large $n$ should be to ensure that $\abs{\mu - \hmu_n} \le \varepsilon$
+\end{description}
+
+\vspace{-1ex}
+\uncover<2->{May we agree on a common software definition for these these objects like we have for floating point numbers or basic mathematical functions?}
+\end{frame}
+
+\begin{frame}{Choosing a Language, Library, or Environment}
+	\vspace{-5ex}
+	\begin{itemize}
+    \item Language/library choices of the users are driven by familiarity and speed
+
+  	\item  Favorites aligned with various communities, but no one dominates all
+
+	\item Difference between prototyping environments and ninja written production code
+
+    \item Growth of multi-processor environments 
+    \begin{itemize}
+    	\item Provides opportunity
+    	\item Complicates software development; where does parallelism really help?
+    \end{itemize}
+
+    \item<2-> Recommendations if you want to contribute software
+    \begin{itemize}
+    \item Realize your new idea as a demo in your favorite library
+    \item Add a significant idea as a feature to other libraries
+    \item Move basic features of your library to larger library
+    \item Write wrappers or demos to connect other libraries to your library
+
+    \end{itemize}
+ \end{itemize}
+\end{frame}
+
+\begin{frame}{Ex.\ QMCPy $+$ FEniCS/Dolfin \cite{LoggEtal_10_2012}}
+\vspace{-3ex}
+	\begin{itemize}
+		\item Want to demonstrate how our QMC software could work with a popular DE solver for DEs with random coefficients
+		\item Partial success \only<3->{Why?}
+	\end{itemize}
+ 	\only<1>{\vspace{-3ex}}\only<2>{\vspace{-2ex}}\only<3>{\vspace{0ex}}
+ \begin{minipage}{0.48\textwidth}
+ 	\only<1-2>{\begin{gather*}
+ 		-\frac{\dif }{\dif x}\biggl(a(x,\vW) \frac{\dif}{\dif x} u(x,\vW) \biggr) = 200( 2x - 1)\\
+ 		 0 \leq x \leq 1, \qquad
+ 		u(0) = u(1) = 0 \\
+ 		a(x,\vW) = 1 + 0.6 \sum_{k=1}^d \frac{W_k T_{k}(2x-1)}{k^2} \\
+ 		 \vW \sim \cu[-1,1]^d \\
+ 		 \Ex[u(0.25,\vW)] = \only<1>{\, ?}\only<2>{ -1.7462}
+ 	\end{gather*}}
+  \only<3->{
+\begin{itemize}
+\item Even with FEniCS/Dolfin's extensive documentation we had difficulties
+\begin{itemize}
+    \item Did not understand the latest version, so reverted to an older version
+    \item Had to learn how to change the random instance of the coefficient without tiggering the JIT compiler
+    \item Do not yet know how to express the random coefficient in terms of the covariance kernel
+\end{itemize}
+\end{itemize}
+  }
+ \end{minipage}%
+\hfill
+ \begin{minipage}{0.48\textwidth}
+ 	\only<1-2>{\centering
+ 	\includegraphics<1>[width=\textwidth]{ProgramsImages/axwtrue0.6.eps}
+ 	\includegraphics<2->[width=\textwidth]{ProgramsImages/timing-4.eps}}%
+  \only<3->{
+\begin{itemize}
+\item But
+\begin{itemize}
+    \item It  works
+    \item If we can overcome the challenges, will post a blog and demo at \cite{QMCBlog}
+\end{itemize}
+\end{itemize}
+
+  }
+\end{minipage}
+
+\end{frame}
+
+\section{Support}
+\begin{frame}{Encouraging Users of Your Library Requires}
+
+	\vspace{-4ex}
+	\begin{itemize}
+		\item A repository where
+		\begin{itemize}
+		\item  Your library can be easily downloaded
+		\item There is documentation
+		\item Issues or queries can be posted
+		\item There are updates
+		\item Tests are run regularly to mitigate against bugs
+		\end{itemize}
+		\item (Elementary) demos on how to use the software and highlighting  its advantages
+		\item Demos connecting  your library with others
+		\item Recommending your students and postdocs to use your library
+		\item Welcoming instructions on how to contribute an algorithm or a demo
+	\end{itemize}
+
+\end{frame}
+
+\begin{frame}{Building a Developer Team for Your Library}
+
+	\vspace{-4ex}
+\begin{itemize}
+
+\item Many senior researchers don't code, and rely on transient team members (students, postdocs)
+\begin{itemize}
+	\item Encourage a sense of ownership that lasts beyond the time your team members are part of your lab
+	\item Encourage those outside your own group to contribute
+\end{itemize}
+
+\item Those of us who understand theory and appreciate software need to encourage and steer popular software to ensure that it is implemented well.
+
+\item Recognize good software as valuable scholarly output
+
+\item Recognize research software engineer as a valued vocation; see Research Software Engineers International 	\url{https://researchsoftware.org}, \emph{Research Software Engineers are people who combine professional software expertise with an understanding of research.
+They go by various job titles but the term Research Software Engineer (RSE) is fast gaining international recognition.}
+  \end{itemize}
+\end{frame}
+
+\section{Next Steps}
+
+\begin{frame}{Next Steps}
+
+	\vspace{-2ex}
+	\begin{minipage}{0.48\textwidth}
+		\begin{itemize}
+			\item Which exemplary theoretical developments are not yet implemented in readily available software?
+			\item What potential MCQMC user communities are ill-served by existing  software libraries?
+			\item Should those of us interested in software gather regularly for show-and-tell to foster more collaboration?
+		\end{itemize}
+	\end{minipage}
+	\begin{minipage}{0.48\textwidth}
+		Join us for a lunch discussion at Bella Casa tomorrow, July 19, at 12:45 PM
+
+		\vspace{2ex}
+
+		\includegraphics[width=\textwidth]{ProgramsImages/PizeriaBellaCasa.png}
+\end{minipage}
+
+\end{frame}
+
+\begin{frame}[allowframebreaks]{References}
+	\printbibliography
+\end{frame}
 

### 2022-11-01 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

@@
 
+Monte Carlo and their often more efficient variant, quasi-Monte Carlo (QMC) methods, have developed since the middle of the last century.
+With the explosion in computational science, there is a need for great Monte Carlo and Quasi-Monte Carlo (MCQMC) software to realize the value in MCQMC methods.
+Based on our observations of what software has been developed to date, and what potential is not yet realized, we propose some criteria for developing great MCQMC software and some practical steps forward.
+Our focus is primarily on QMC software.
+
 	\begin{itemize}
@@
 
-\bibliographystyle{spmpsci.bst}
-\bibliography{FJH23,FJHown23,QMCSoftwareArticle,sc}
+\bibliographystyle{spmpsci}
+\bibliography{FJH23,FJHown23}
 
 \end{document}
-
-\begin{frame}[allowframebreaks]{Selective History of Publicly Available MC\textbf{QMC} Software}
-\vspace{-2ex}
-	\begin{itemize}
-		\item \cite{LEc2017a} provides a history of random number generation
-		\item Notorious RANDU from the 1960s \cite{RANDU}, which fails the spectral test
-		\item VEGAS \cite{Lep78a, Lep21a} importance sampling Monte Carlo algorithm favored by physicists
-		\item ACM low discrepancy point generators \cite{BraFox88,BraFoxNie92,HonHic00a}
-		\item FinDer \cite{PasTra95,FinDer} and BRODA \cite{BRODA20a} targeting quantitative finance
-		\item Korobov cubature and scrambled Sobol' generators in NAG \cite{NAG27} for decades		
-		\item Scrambled Sobol' and Halton generators in MATLAB \cite{MAT9.13} since 2008, and fixed a few years later
-		\item BUGS \cite{BUGSBook, BUGSweb} and Stan \cite{STAN} for Markov Chain Monte Carlo (MCMC)
-		\item LatMRG \cite{LEcCou97}, RNGStreams \cite{LEcEtal02},  SSJ \cite{LEc2002a,SSJ}, and LatNetBuilder \cite{LatNet} 
-		\item Low discrepancy generators \cite{FriKel02,FriKelweb}, SamplePack \cite{SamplePack}, \cite{GruWeb},  and MatBuilder \cite{paulin2022} 
-		\item Sobol' direction numbers \cite{JoeKuo03,JoeKuo08,SobolDirection}
-		\item Fast CBC, Magic Point Shop, QMC4PDE and other code since 2004  at \cite{NuyWeb} 
-		\item Multi-level software \cite{GilesSoft,GilesQSoft} 
-		\item Data-driven error bounds and stopping criteria in GAIL \cite{ChoEtal21a} and QMCPy \cite{QMCPy2020a, ChoEtal22a}  
-		\item Uncertainty quantification libraries Dakota \cite{DakotaUsersManual}, UQTk \cite{DebEtal04,UQTk}, and MUQ \cite{MUQ}  have some basic level low discrepancy sampling
-	    \item QMC framework in Julia since 2019 \cite{QMCJulia} by Robbe and others 
-		\item Scrambled Sobol' in SciPy \cite{virtanen2020scipy} and PyTorch \cite{paszke2019pytorch} since several years ago
-		\item TensorFlow QMC framework \cite{tfqfQMC2021a} since a year or so ago
-	\end{itemize}
-
-We have stand-alone algorithms, small libraries, and pieces of larger libraries
-\end{frame}
 

### 2022-12-27 MCQMC2022ChallengesQMCSoftwareArticle.tex
minor updates

@@
 
+%\input FJHDef.tex
+\newcommand{\AGSComment}[1]{{\color{red} #1}}
+
 %%%%Fred end
@@
 
-\title*{Challenges in Developing Great MCQMC Software}
+\title*{Challenges in Developing Great QMC Software}
 \authorrunning{S.-C.\ T.\ Choi et al}
@@
 
-\abstract{}
-
-\section{Introduction}
-
+\abstract{
 Monte Carlo and their often more efficient variant, quasi-Monte Carlo (QMC) methods, have developed since the middle of the last century.
@@
 Our focus is primarily on QMC software.
+}
 
+\newpage
+
+\section{Introduction}
+
+Modern programming languages enable a simple Monte Carlo method approximating $\mu := \mathbb{E}[f(\boldsymbol{X})]$, $\boldsymbol{X} \sim \mathcal{U}[0,1]^d$ to be implemented in three lines of code:
+\begin{enumerate}
+    \item draw $n$ independent identically distributed (IID) standard uniform nodes 
+    $$\boldsymbol{x}_1,\dots,\boldsymbol{x}_n \sim \mathcal{U}[0,1]^d,$$
+    \item evaluate the integrand $f$ at these nodes to obtain $$y_i := f(\boldsymbol{x}_i), \qquad i=1,\dots,n$$
+    \item estimate the true mean $\mu$ by the sample mean
+    \begin{equation}
+        \hat{\mu} := \frac{1}{n} \sum_{i=1}^n y_i.
+        \label{eq:mc_average}
+    \end{equation}  
+\end{enumerate}
+However, practitioners may want to utilize advanced sampling methods such as antithetic sampling, importance sampling, or low discrepancy sampling in the case of Quasi-Monte Carlo (QMC) methods.
+Users may also wish to modify the integrand to accommodate taking the expectation with respect to a non-standard-uniform random variable or to utilize control variates.
+The above enhancements may require taking the weighted average in place of \eqref{eq:mc_average} to maintain an unbiased estimate.
+Practitioners may also be interested in determining $n$ adaptively so the approximation is within some error threshold from the true mean $\mu$.
+
+Developing software to implement such enhanced MC strategies poses a number of challenges.
+First, a development team must determine the value of the software to the users.
+We discuss why great (Q)MC software in necessary in Section \ref{sec:why_we_need_software}.
+Second, a team must determine high level objectives for the software; the tenets of great software are described in Section \ref{sec:tenets_of_great_software}.
+Third, an architecture for the  (Q)MC software should be determined.
+We propose a modularization of the (Q)MC pipeline in Section \ref{sec:qmc_software_architecture}.
+Next, a team must choose which programming languages, hardware environments, and distribution platforms they wish to support.
+We discuss the nuance and challenges in these decisions in Section \ref{sec:language_library_environment}.
+Finally, a project which hopes to stand the test of time must draw attention and support from the broader (Q)MC community.
+To this end, we encourage collaborative community development and describe its' best practices in  Section \ref{sec:encouraging_collaborative_developement}.
+
+\section{Why Develop Software} \label{sec:why_we_need_software}
+
+In 2021, the US Department of Energy held a ``Workshop on the Science of Scientific-Software Development and Use'' \cite{Her19a,ASCR-SSSDU}.
+This workshop not only discussed diagnostics and treatments for the challenges of developing great scientific software, but also emphasized the importance of implementing theoretical advancements into well written, accessible software libraries.
+This section describes both the practical and theoretical importance of developing scientific software.
+
+\subsection{Theory to Software}
+
+The chief purpose of most scientific software developments is to make theoretical advancements available to practitioners.
+Developing theory into practical software leads to algorithms which often require additional considerations.
+For instance, the numerical stability of a method only becomes relevant when considering how to implement a method onto a finite precision computer.
+Taking the time to consider such implications and determine relevant algorithm parameters alleviates these burdens form the user.
+Moreover, the availability of such algorithms enables practitioners with a variety of applications to quickly solve problems.
+
+\subsection{Software to Theory}
+
+An often overlooked benefit of scientific software  is the feedback garnered for future theoretical development. A few motivating examples are
 \begin{itemize}
-	\item Correct, e.g., not omit the zeroth point in a low discrepancy sequence \cite{Owe22a, scipySobol2020a}
+    \item the early application of QMC to high dimensional option pricing and 
+    \item the early application of neural networks to problems in computer vision.
+\end{itemize}
+The unexpected performance in these examples motivated theoretical developments to justify their success.
+Moreover, the availability of high quality software will often attract practitioners to a certain methodology over competitors, and in turn may inspire further theoretical development.
 
-	\item  Complete---contain the components or  easily access components in other libraries to solve real, complex problems
+\section{Tenets of Great Software} \label{sec:tenets_of_great_software}
 
-	\item  Accessible---downloadable; tutorials, demos, discussion forums, etc.\ for (new) users; written in a language that potential users speak; provide a consistent user interface
+In this section we lay out common aspirations of scientific software.
+While some objectives are expected by practitioners, e.g.
+correctly written software, others may better be characterized as features that are pursued as a project matures, e.g.
+supporting scalable computation.
+The following paragraphs highlight aspirations with more important, expected behaviours coming earlier in the discussion.
 
-	\item Efficient---in terms of compute time and memory
+Software is expected to be \emph{correct}.
+A practitioner generating a digital sequence expects the routine to produce theoretically accurate points.
+For example, the QMC community has helped developers ensure implementations of low discrepancy sequence generators include the zeroth point \cite{Owe22a, scipySobol2020a} and  have correct randomization routines.
+%, and disallow thinning / leaping: taking every $k^\text{th}$ value in the sequence.
 
-	\item Current---include the latest and best algorithms
+Algorithms should be implemented in an \emph{efficient} manner.
+Both the compute time and memory requirements should match theoretical developments which in turn should be as close to the state of the art as possible.
+As a classical example, it is expected that an implementation of the discrete Fourier transform for $n$ points has complexity $\mathcal{O}(n\log n)$.
 
-	\item  Sustainable---have a sufficient user base and developer community for updates and maintenance
+Software should be \emph{accessible} in terms of installation, presentation, navigation, exemplification, and communication.
+A library should be straightforward to install from a common distribution platform, consistent in user interface, intuitive to navigate to modules and documentation, comprehensible in its' examples, and supportive in user engagement.
+Software accessibility considerations are further discussed in Section \ref{sec:language_library_environment}.
 
-	\item Scalable---take advantage of advanced computer architectures for speed and to solve large problems
+Libraries should be \emph{sustainable} for future development.
+The goal should be to have a sufficient user base and developer community for updates and maintenance.
+Similar to accessibility, sustainability is often overlooked for aspirations discussed later in this section.
+However, we contend that a sub-optimal but user-friendly library with community engagement may be more impactful than an optimal library which lacks support.
+We promote collaborative, community driven development further in Section \ref{sec:encouraging_collaborative_developement}.
 
-\end{itemize}
+Software should be \emph{integrated} into the environment's scientific software ecosystem.
+More simply, libraries should play nicely together and utilize the mature developments of other teams.
+A simple example in \AGSComment{CITE} shows a straightforward integration between the QMC package QMCPy \cite{QMCPy2020a} and the PDE package FEniCS/Dolfin \cite{LoggEtal_10_2012}, both in Python.
 
+Routines should be \emph{scalable} to advanced computer architectures.
+The widespread availability of on demand computing with multi-core and multi-GPU machines encourages software that can take advantage of parallel computation.
+Scalable algorithms lend themselves to a larger class or problems and provide practitioners the opportunity to more quickly test ideas.
+
+\section{(Quasi-)Monte Carlo Software Architecture} \label{sec:qmc_software_architecture}
+
+An important step in the development of any software is to determine its' architecture.
+How should components interact with each other?
+Which aspects should be modular?
+Object oriented?
+Procedural?
+Answers to these questions form the blueprint of software design and guide implementation.
+
+This section proposes an object oriented design for (Q)MC software.
+The authors have implemented this architecture into the QMCPy \cite{QMCPy2020a} Python package which is further detailed in the MCQMC 2020 proceedings article \AGSComment{CITE}.
+While this is not the only possible architecture for (Q)MC problems, we hope the high level object definitions facilitate easy collaboration and integration for research across the (Q)MC community.
+
+The standard (Q)MC problem is to approximate the true mean $\mu := \mathbb{E}[g(\boldsymbol{T})]$ where $\boldsymbol{T}$ is some $d$ dimensional random variable.
+For nice $\boldsymbol{T}$, one may perform a change of variables to write $\mu = \mathbb{E}[f(\boldsymbol{X})]$ where $\boldsymbol{X} \sim \mathcal{U}[0,1]^d$.
+For example, if $\boldsymbol{T} \sim \mathcal{N}(\boldsymbol{0},\boldsymbol{I})$, then one may set $f(\boldsymbol{x}) := g(\boldsymbol{\Phi}^{-1}(\boldsymbol{x}))$ for $\boldsymbol{x} \in [0,1]^d$ where $\boldsymbol{\Phi}^{-1}$ is the inverse CDF of a standard normal distribution taken element wise.
+One may then approximate $\mu$ by the $n$ sample average $\hat{\mu} = 1/n \sum_{i=1}^n f(\boldsymbol{x}_i)$ where $\{\boldsymbol{x}_i\}_{i=1}^n$ are chosen so their empirical distribution is ``close'' to the uniform distribution.
+The approximation error $\varepsilon = \lvert \mu - \hat{\mu} \rvert$ is $\mathcal{O}(n^{-1/2})$ when $\{\boldsymbol{x}_i\}_{i=1}^n$ are chosen to be IID standard uniform as in the case of simple Monte Carlo (MC) and almost $\mathcal{O}(n^{-1})$ when $\{\boldsymbol{x}_i\}_{i=1}^n$ are chosen from a low discrepancy sequence as done for Quasi-Monte Carlo (QMC).
+Instead of selecting the number of samples $n$ and then determining the error $\varepsilon$, it is often desirable to select $\varepsilon$ and then adaptively determine $n$.
+From this view, $n$ is $\mathcal{O}(\varepsilon^{-2})$ for MC and almost $\mathcal{O}(\varepsilon^{-1})$ for QMC.
+A more detailed account of MC and QMC can be found in \AGSComment{CITE}.
+
+The (Q)MC pipeline above can be decomposed into $4$ main components:  
+\begin{description}
+    \item[\textbf{Generator:}] producing $\{\boldsymbol{x}_i\}_{i=1}^n$ whose empirical distribution is ``close'' to the standard uniform distribution in $d$ dimensions for any $n \geq 1$.
+May produce IID or low discrepancy sequences depending on the problem.
+    \item[\textbf{Measure:}] that defines the distribution of $\boldsymbol{T}$ e.g., uniform, Gaussian, Lebesgue. This definition facilitates choosing a change of variables to determine $f$. 
+    \item[\textbf{Integrand:}] $g$, which defines the original integral. When provided knowledge of the true measure, the integrand can also evaluate $f$ via the change of variables.  
+    \item[\textbf{Stopping Criterion:}] based on a data-driven error bound, which determines how large $n$ should be to ensure that $\varepsilon$ is less than some user-provided error threshold.
+\end{description}
+A variety of QMC materials and the their categorizations into the above framework are described in the following list.
 \begin{itemize}
-	\item \cite{LEc2017a} provides a history of random number generation
-	\item Notorious RANDU from the 1960s \cite{RANDU}, which fails the spectral test
-	\item VEGAS \cite{Lep78a, Lep21a} importance sampling Monte Carlo algorithm favored by physicists
+    %\item \cite{LEc2017a} provides a history of random number generation
+    %\item Notorious RANDU from the 1960s \cite{RANDU}, which fails the spectral test
+    %\item VEGAS \cite{Lep78a, Lep21a} importance sampling Monte Carlo algorithm favored by physicists
     \item ACM low discrepancy point generators \cite{BraFox88,BraFoxNie92,HonHic00a}
@@
     \item Scrambled Sobol' and Halton generators in MATLAB \cite{MAT9.13} since 2008, and fixed a few years later
-	\item BUGS \cite{BUGSBook, BUGSweb} and Stan \cite{STAN} for Markov Chain Monte Carlo (MCMC)
+    %\item BUGS \cite{BUGSBook, BUGSweb} and Stan \cite{STAN} for Markov Chain Monte Carlo (MCMC)
     \item LatMRG \cite{LEcCou97}, RNGStreams \cite{LEcEtal02},  SSJ \cite{LEc2002a,SSJ}, and LatNetBuilder \cite{LatNet} 
@@
     \item TensorFlow QMC framework \cite{tfqfQMC2021a} since a year or so ago
+    \item CUBA \cite{CUBA}
 \end{itemize}
 
-We have stand-alone algorithms, small libraries, and pieces of larger libraries
+\section{Choosing a Environment} \label{sec:language_library_environment}
+
+This section discusses the challenges in choosing environments for development and distribution.
+The development environment includes the choice of which languages, architectures, testing methods, and documentation methods to support.
+The deployment environment will often be influenced by the choice of programming languages.
+
+The choice of programming language should be aligned with the communities the team wishes to reach with their software.
+The choice should also account for developer familiarity and ability to quickly prototype ideas.
+A team should align their choice of language to balance the desire to have user-friendly prototyping code versus the desire to have high performance computing (HPC) ``ninja'' code written for production setting.
+To the second point, it is also important to consider which language will best facilitate the parallel and GPU computations the team wishes to support.
+
+Many programming languages come bundled with their own preferred distribution systems.
+Python has PyPI, R has CRAN, Julia has Pkg.
+However, many projects support multiple languages and have complex dependency requirements.
+In such cases, it can be helpful to containerize your application using a service like Docker.
+
+\section{Encouraging Collaborative Development} \label{sec:encouraging_collaborative_developement}
+
+We now discuss ways a team may encourage collaborative development while touching on a few best practices.
+As discussed earlier, collaborative development promotes greater community engagement, ecosystem integration, and software longevity.
+
+The authors have found the Git version control system to be an invaluable for software development.
+Git hosts such as GitHub or Bitbucket may offer teams additional features such as tracking issues, managing discussion threads, and setting up automated workflows for jobs such as testing and documentation compilation.
+
+Accessibility is key to community engagment.
+It is paramount that your package be easy to install, well documented, and tested regularly.
+Aside from static communications, a team should also be engaged in addressing users bugs and be attentive to collaborating with proposed code changes.
+
+It is also important for a development team to make a clear path for collaborators to contributors to the software.
+Templates for issues and pull requests are a good place to start.
+Contributors should also be encouraged to write robust tests for their updates and highlight their new features in a demo for greater accessibility.
 
@@
 \bibliography{FJH23,FJHown23}
+
+\AGSComment{
+\tableofcontents
+\section*{TODO}
+\begin{itemize}
+    \item Flesh out Section \ref{sec:encouraging_collaborative_developement}
+    \item QMC Software Architecture section: better way then description of components and list of softwares? 
+    \item For Section \ref{sec:tenets_of_great_software}, the slides also include 
+    \begin{itemize}
+        \item  Complete---contain the components or  easily access components in other libraries to solve real, complex problems
+        \item Current---include the latest and best algorithms
+    \end{itemize}
+    which I think fall under the umbrella of software being \emph{integrated} with the community.
+    \item update QMCPy citation \cite{QMCPy2020a}
+    \item figures?
+    \item I have left out or been breif about material from the following slides: QMCPy + FEniCS/Dolfin,  Building a Developer Team for Your Library
+\end{itemize}
+}
 

### 2023-01-20 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf
start UMBridge MUQ example

@@
 %%%%Fred start
-\usepackage{url, xspace}
+\usepackage{xspace}
+\usepackage[hyphens]{url}
 \providecommand{\HickernellFJ}{Hickernell\xspace}
@@
 Contributors should also be encouraged to write robust tests for their updates and highlight their new features in a demo for greater accessibility.
+
+\section{Example Collaboration with UMBridge and QMCPy} \label{sec:example_collaboration_umbridge}
+
+This section discusses a collaboration between the UM-Bridge (the UQ and Model Bridge) \AGSComment{CITE} and QMCPy (the Quasi-Monte Carlo Python library) \cite{QMCPy2020a} softwares.
+The UM-Bridge documentation describes its' mission is to, ``provide a unified interface for numerical models that is accessible from virtually any programming language or framework.
+It is primarily intended for coupling advanced models (e.g.
+simulations of complex physical processes) to advanced statistics or optimization methods.'' QMCPy is a high level framework for combining QMC components in a compatible and consistent manner.
+The collaboration between these two packages focused on making the UQ methods of QMCPy compatible with the abstract model representations from UM-Bridge.
+Specifically, UM-Bridge models are made compatible with the QMCPy framework through an integrand wrapper for seamless compatibility with a variety of existing point generators, transformations, and stopping criterion already available in QMCPy.
+
+To exemplify the integration, we focus on quantifying uncertainty propagation of material properties of a cantilevered beam.
+It is assumed the beam is split into $3$ regions of random stiffness dependent on uncertain material properties.
+Given a sample of $3$ stiffness values, the UM-Bridge model provides the displacement at $31$ equidistant nodes along the beam.
+Our task is to quantify the uncertainty in displacement at each of these $31$ locations along the beam.
+
+A fully coded example of this problem is available at \url{https://github.com/QMCSoftware/QMCSoftware/blob/master/demos/umbridge.ipynb}.
+At a high level, one begins by running the UM-Bridge model in a Docker container.
+A model is then initialized in the UM-Bridge Python client which is then provided to QMCPy alongside a low discrepancy point generator and distribution governing the $3$ beam stiffness values.
+QMCPy then performs adaptive QMC cubature to simultaneously approximate the expected displacement at the $31$ locations of interest.
+When the cubature method queries the model, the QMCPy wrapper around UM-Bridge model sends requests to the Docker UM-Bridge model in parallel.
+The load balancing provided by UM-Bridge may be utilized to evaluate the model in parallel, a potentially advantageous feature for models which are costly to evaluate.
+Figure \ref{fig:umbridge_muq} depicts the $31$ approximate expected displacement values subject to uniform stiffness.
+
+\begin{figure}
+    \centering
+    \includegraphics[height=4cm]{umbridge_muq.png}
+    \caption{QMCPy approximate expected displacement at $31$ locations subject to random uniform stiffness. }
+    \label{fig:umbridge_muq}
+\end{figure}
 

### 2023-01-29 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

@@
 %%%%Fred start
+\usepackage{cite}
 \usepackage{xspace}
@@
 
-\title*{Challenges in Developing Great QMC Software}
+\title*{Challenges in Developing Great Quasi-Monte Carlo Software}
 \authorrunning{S.-C.\ T.\ Choi et al}
@@
 \abstract{
-Monte Carlo and their often more efficient variant, quasi-Monte Carlo (QMC) methods, have developed since the middle of the last century.
-With the explosion in computational science, there is a need for great Monte Carlo and Quasi-Monte Carlo (MCQMC) software to realize the value in MCQMC methods.
-Based on our observations of what software has been developed to date, and what potential is not yet realized, we propose some criteria for developing great MCQMC software and some practical steps forward.
-Our focus is primarily on QMC software.
+Quasi-Monte Carlo (QMC) methods have developed over several decades.
+With the explosion in computational science, there is a need for great software to realize that implements QMC algorithms.
+We review what QMC software has been developed to date, propose some criteria for developing great QMC software, and suggest some steps toward achieving great software.
 }
@@
 
-Modern programming languages enable a simple Monte Carlo method approximating $\mu := \mathbb{E}[f(\boldsymbol{X})]$, $\boldsymbol{X} \sim \mathcal{U}[0,1]^d$ to be implemented in three lines of code:
+A QMC approximation of $\mu := \mathbb{E}[f(\boldsymbol{X})]$, $\boldsymbol{X} \sim \mathcal{U}[0,1]^d$ can be implemented in a few steps:
 \begin{enumerate}
-    \item draw $n$ independent identically distributed (IID) standard uniform nodes 
-    $$\boldsymbol{x}_1,\dots,\boldsymbol{x}_n \sim \mathcal{U}[0,1]^d,$$
-    \item evaluate the integrand $f$ at these nodes to obtain $$y_i := f(\boldsymbol{x}_i), \qquad i=1,\dots,n$$
-    \item estimate the true mean $\mu$ by the sample mean
+    \item Draw a sequence of $n$ a low discrepancy (LD) \cite{Nie92, SloJoe94, DicEtal14a, DicEtal22a} nodes,
+    $\boldsymbol{x}_1,\dots,\boldsymbol{x}_n$ that mimic $\mathcal{U}[0,1]^d$.
+    \item Evaluate the integrand $f$ at these nodes to obtain $f(\boldsymbol{x}_i)$, $i=1,\dots,n$.
+    \item Estimate the true mean, $\mu$, by the sample mean
     \begin{equation}
-        \hat{\mu} := \frac{1}{n} \sum_{i=1}^n y_i.
+        \hat{\mu} := \frac{1}{n} \sum_{i=1}^n f(\boldsymbol{x}_i).
         \label{eq:mc_average}
@@
 \end{enumerate}
-However, practitioners may want to utilize advanced sampling methods such as antithetic sampling, importance sampling, or low discrepancy sampling in the case of Quasi-Monte Carlo (QMC) methods.
-Users may also wish to modify the integrand to accommodate taking the expectation with respect to a non-standard-uniform random variable or to utilize control variates.
-The above enhancements may require taking the weighted average in place of \eqref{eq:mc_average} to maintain an unbiased estimate.
-Practitioners may also be interested in determining $n$ adaptively so the approximation is within some error threshold from the true mean $\mu$.
+However, the practice of QMC is often more complicated.
+The original problem may need to be rewritten fit above form and/or to facilitate a good approximation with small $n$.
+Practitioners may wish to determine $n$ adaptively to satisfy a desired error tolerance.
+
+In the next section we describe why QMC software is important.  We then discuss characteristics of great QMC software:
+
+\begin{itemize}
+\item Correct,
+
+\item Efficient in computational time and memory,
+
+\item Accessible to practitioners and theorists alike,
+
+\item Sustainable by a community that owns it,
+
+\item Integrated with related libraries, and
+
+\item Scalable to advanced hardware architectures.
+
+\end{itemize}
 
@@
 
-In 2021, the US Department of Energy held a ``Workshop on the Science of Scientific-Software Development and Use'' \cite{Her19a,ASCR-SSSDU}.
+Recent interest in quality scientific software is typified by the 2021 US Department of Energy \emph{Workshop on the Science of Scientific-Software Development and Use} \cite{Her19a,ASCR-SSSDU,osti_1846008}.
 This workshop not only discussed diagnostics and treatments for the challenges of developing great scientific software, but also emphasized the importance of implementing theoretical advancements into well written, accessible software libraries.

### 2023-02-04 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

@@
 
-Recent interest in quality scientific software is typified by the 2021 US Department of Energy \emph{Workshop on the Science of Scientific-Software Development and Use} \cite{Her19a,ASCR-SSSDU,osti_1846008}.
+Recent interest in quality scientific software is exemplified by the 2021 US Department of Energy \emph{Workshop on the Science of Scientific-Software Development and Use} \cite{Her19a,ASCR-SSSDU,osti_1846008}.
 This workshop not only discussed diagnostics and treatments for the challenges of developing great scientific software, but also emphasized the importance of implementing theoretical advancements into well written, accessible software libraries.

### 2023-02-04 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

--- 2023-01-29 MCQMC2022ChallengesQMCSoftwareArticle.tex
@@
 
-\section{Why Develop Software} \label{sec:why_we_need_software}
-
-Recent interest in quality scientific software is typified by the 2021 US Department of Energy \emph{Workshop on the Science of Scientific-Software Development and Use} \cite{Her19a,ASCR-SSSDU,osti_1846008}.
+\section{Why Develop QMC Software} \label{sec:why_we_need_software}
+
+Recent interest in quality scientific software is exemplified by the 2021 US Department of Energy \emph{Workshop on the Science of Scientific-Software Development and Use} \cite{ASCR-SSSDU,osti_1846008}.
 This workshop not only discussed diagnostics and treatments for the challenges of developing great scientific software, but also emphasized the importance of implementing theoretical advancements into well written, accessible software libraries.
-This section describes both the practical and theoretical importance of developing scientific software.
-
-\subsection{Theory to Software}
-
-The chief purpose of most scientific software developments is to make theoretical advancements available to practitioners.
-Developing theory into practical software leads to algorithms which often require additional considerations.
-For instance, the numerical stability of a method only becomes relevant when considering how to implement a method onto a finite precision computer.
-Taking the time to consider such implications and determine relevant algorithm parameters alleviates these burdens form the user.
-Moreover, the availability of such algorithms enables practitioners with a variety of applications to quickly solve problems.
-
-\subsection{Software to Theory}
-
-An often overlooked benefit of scientific software  is the feedback garnered for future theoretical development. A few motivating examples are
-\begin{itemize}
-    \item the early application of QMC to high dimensional option pricing and 
-    \item the early application of neural networks to problems in computer vision.
-\end{itemize}
-The unexpected performance in these examples motivated theoretical developments to justify their success.
-Moreover, the availability of high quality software will often attract practitioners to a certain methodology over competitors, and in turn may inspire further theoretical development.
+Three cross-cutting themes that arose in this workshop were
+\begin{itemize}
+    \item  We need to consider both human and technical elements to better understand how to improve the development and use of scientific software.
+
+    \item We need to address urgent challenges in workforce recruitment and retention in the computing sciences with growth through expanded diversity, stable career paths, and the creation of a community and culture that attract and retain new generations of scientists.
+
+    \item Scientific software has become essential to all areas of science and technology, creating opportunities for expanded partnerships, collaboration, and impact.
+\end{itemize}
+These themes apply to QMC software in particular, as well as scientific software in general.
+
+\subsection{QMC Theory to Software}
+
+QMC software makes theoretical advances in QMC available to practitioners.
+However, translating pseudo-code into good executable code is nuanced.
+Software developers must write code that is computationally efficient, numerical stable, and provides reasonable default choices of tuning parameters.
+Good QMC software relieves users of these concerns.
+
+\subsection{QMC Software to Theory}
+
+Good QMC software opens up new application areas for QMC methods by allowing practitioners to compare new methods to existing ones.
+QMC software has been successful in quantitative finance, uncertainty quantification, and image rendering.
+Unexpectedly good or bad computational results leads to open theoretical questions.
+For example, the early application of QMC to high dimensional integrals arising in financial risk \cite{PasTra95} led to a wave of theoretical results on the tractability of integration in weighted spaces \cite{Woz99a,DicEtal14a,NovWoz10a}.
 

### 2023-02-04 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

@@
 %%%%Fred start
+\include{common-macros.tex}
 \usepackage{cite}
@@
 
-A QMC approximation of $\mu := \mathbb{E}[f(\boldsymbol{X})]$, $\boldsymbol{X} \sim \mathcal{U}[0,1]^d$ can be implemented in a few steps:
+A QMC approximation of $\mu := \EE[f(\bsX)]$, $\bsX \sim \calU[0,1]^d$ can be implemented in a few steps:
 \begin{enumerate}
     \item Draw a sequence of $n$ a low discrepancy (LD) \cite{Nie92, SloJoe94, DicEtal14a, DicEtal22a} nodes,
-    $\boldsymbol{x}_1,\dots,\boldsymbol{x}_n$ that mimic $\mathcal{U}[0,1]^d$.
-    \item Evaluate the integrand $f$ at these nodes to obtain $f(\boldsymbol{x}_i)$, $i=1,\dots,n$.
+    $\bsx_1,\dots,\bsx_n$ that mimic $\calU[0,1]^d$.
+    \item Evaluate the integrand $f$ at these nodes to obtain $f(\bsx_i)$, $i=1,\dots,n$.
     \item Estimate the true mean, $\mu$, by the sample mean
     \begin{equation}
-        \hat{\mu} := \frac{1}{n} \sum_{i=1}^n f(\boldsymbol{x}_i).
+        \hat{\mu} := \frac{1}{n} \sum_{i=1}^n f(\bsx_i).
         \label{eq:mc_average}
@@
 \begin{itemize}
+\item Integrated with related libraries (Sect.\ \ref{CDHS_sec:integrated}),
+
 \item Correct,
@@
 
-\item Sustainable by a community that owns it,
-
-\item Integrated with related libraries, and
+\item Sustainable by a community that owns it, and
 
@@
 First, a development team must determine the value of the software to the users.
-We discuss why great (Q)MC software in necessary in Section \ref{sec:why_we_need_software}.
-Second, a team must determine high level objectives for the software; the tenets of great software are described in Section \ref{sec:tenets_of_great_software}.
+We discuss why great (Q)MC software in necessary in Section \ref{CDHS_sec:why_we_need_software}.
+Second, a team must determine high level objectives for the software; the tenets of great software are described in Section \ref{CDHS_sec:tenets_of_great_software}.
 Third, an architecture for the  (Q)MC software should be determined.
-We propose a modularization of the (Q)MC pipeline in Section \ref{sec:qmc_software_architecture}.
+We propose a modularization of the (Q)MC pipeline in Section \ref{CDHS_sec:qmc_software_architecture}.
 Next, a team must choose which programming languages, hardware environments, and distribution platforms they wish to support.
-We discuss the nuance and challenges in these decisions in Section \ref{sec:language_library_environment}.
+We discuss the nuance and challenges in these decisions in Section \ref{CDHS_sec:language_library_environment}.
 Finally, a project which hopes to stand the test of time must draw attention and support from the broader (Q)MC community.
-To this end, we encourage collaborative community development and describe its' best practices in  Section \ref{sec:encouraging_collaborative_developement}.
-
-\section{Why Develop QMC Software} \label{sec:why_we_need_software}
+To this end, we encourage collaborative community development and describe its' best practices in  Section \ref{CDHS_sec:encouraging_collaborative_developement}.
+
+\section{Why Develop QMC Software} \label{CDHS_sec:why_we_need_software}
 
@@
 
-\section{Tenets of Great Software} \label{sec:tenets_of_great_software}
+\section{Integrated} \label{CDHS_sec:integrated}
+We now discuss characteristics of great QMC software outlined in the introduction.
+
+Expanding on the problem formulation in the introduction, we want to approximate the integral or expectation, $\mu$, well by the sample mean, $\hat{\mu}$:
+\begin{gather} 
+\label{eq:cubSummary}
+	\mu : = \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  = \bbE[f(\bsX)] = \int_{[0,1]^d} f(\bsx)  \, \D \bsx \approx \frac 1n \sum_{i=0}^{n-1} f(\bsx_i) =: \hat{\mu}_n, \\
+ \label{CDHS_eq:error_tol}
+ \abs{\mu - \hat{\mu}_n} \le \varepsilon.
+\end{gather}
+Progressing from the left side of this equation to the right requires several steps.
+
+The \emph{integrand}, $g$, and the non-negative weight, $\lambda$, define the original integral arising from an application.
+The function $g$ may represent the discounted payoff of a financial derivative, or the velocity of flow at a point in rock with stochastic porosity.
+If $\lambda$ is a probability density function for a random variable $\bsT$, then $\mu = \EE[g(\bsT)]$.
+
+To put $\mu$ into the form that is directly accessible to QMC methods requires a \emph{variable transformation}, $\bsPsi$, such that 
+\begin{equation*}
+    \Psi((0,1)^d) = \calT, \qquad f(x) = g(\bsPsi(\bsx)) \lambda(\bsPsi(\bsx)) \abs{\frac{\partial \bsPsi}{\partial{\bsx}}}.
+\end{equation*}
+Such a transformation is typically non-unique, and the choice of a good one is equivalent to importance sampling.
+
+An \emph{LD generator}, such as a lattice, digital net, or Halton point generator supplies the sequence $\bsx_1, \bsx_2, \ldots$.
+These sequences may be deterministic or random and are intended to have an empirical distribution that approximates well the uniform distribution.
+
+Finally, a \emph{stopping criterion} is determines the sample size, $n$, required to satisfy the error criterion \eqref{CDHS_eq:error_tol} given in terms of the absolute error tolerance, $\varepsilon$.
+One can also imagine error criteria involving a relative error tolerance or a combination of absolute and relative error tolerances.
+
+No one software library contains all of these features, which is why great QMC software must be well integrated into other software libraries.
+This may mean large libraries including QMC routines and/or building connections between QMC libraries and other libraries.
+
+The most prevalent generally available QMC software are LD generators.
+They appear in ACM publications \cite{BraFox88,BraFoxNie92,HonHic00a}, FinDer \cite{PasTra95,FinDer}, BRODA \cite{BRODA20a}, NAG \cite{NAG27}, and MATLAB \cite{MAT9.13}.
+
+LatNetBuilder \cite{LatNet} 
+    \item Low discrepancy generators \cite{FriKel02,FriKelweb}, SamplePack \cite{SamplePack}, \cite{GruWeb},  and MatBuilder \cite{paulin2022} 
+    \item Sobol' direction numbers \cite{JoeKuo03,JoeKuo08,SobolDirection}
+    \item Fast CBC, Magic Point Shop, QMC4PDE and other code since 2004  at \cite{NuyWeb} 
+    \item Multi-level software \cite{GilesSoft,GilesQSoft} 
+    \item Data-driven error bounds and stopping criteria in GAIL \cite{ChoEtal21a} and QMCPy \cite{QMCPy2020a, ChoEtal22a}  
+    \item Uncertainty quantification libraries Dakota \cite{DakotaUsersManual}, UQTk \cite{DebEtal04,UQTk}, and MUQ \cite{MUQ}  have some basic level low discrepancy sampling
+    \item QMC framework in Julia since 2019 \cite{QMCJulia} by Robbe and others 
+    \item Scrambled Sobol' in SciPy \cite{virtanen2020scipy} and PyTorch \cite{paszke2019pytorch} since several years ago
+    \item TensorFlow QMC framework \cite{tfqfQMC2021a} since a year or so ago
+    \item CUBA \cite{CUBA}
+
+This requires four components, which we implement as QMCPy classes.
+
+Software should be \emph{integrated} into the environment's scientific software ecosystem.
+More simply, libraries should play nicely together and utilize the mature developments of other teams.
+A simple example in \AGSComment{CITE} shows a straightforward integration between the QMC package QMCPy \cite{QMCPy2020a} and the PDE package FEniCS/Dolfin \cite{LoggEtal_10_2012}, both in Python.
+
+\section{Tenets of Great Software} \label{CDHS_sec:tenets_of_great_software}
 
@@
 A library should be straightforward to install from a common distribution platform, consistent in user interface, intuitive to navigate to modules and documentation, comprehensible in its' examples, and supportive in user engagement.
-Software accessibility considerations are further discussed in Section \ref{sec:language_library_environment}.
+Software accessibility considerations are further discussed in Section \ref{CDHS_sec:language_library_environment}.
 
@@
 However, we contend that a sub-optimal but user-friendly library with community engagement may be more impactful than an optimal library which lacks support.
-We promote collaborative, community driven development further in Section \ref{sec:encouraging_collaborative_developement}.
+We promote collaborative, community driven development further in Section \ref{CDHS_sec:encouraging_collaborative_developement}.
 
@@
 
-\section{(Quasi-)Monte Carlo Software Architecture} \label{sec:qmc_software_architecture}
+\section{(Quasi-)Monte Carlo Software Architecture} \label{CDHS_sec:qmc_software_architecture}
 
@@
 
-\section{Choosing a Environment} \label{sec:language_library_environment}
+\section{Choosing a Environment} \label{CDHS_sec:language_library_environment}
 
@@
 
-\section{Encouraging Collaborative Development} \label{sec:encouraging_collaborative_developement}
+\section{Encouraging Collaborative Development} \label{CDHS_sec:encouraging_collaborative_developement}
 
@@
 
-\section{Example Collaboration with UMBridge and QMCPy} \label{sec:example_collaboration_umbridge}
+\section{Example Collaboration with UMBridge and QMCPy} \label{CDHS_sec:example_collaboration_umbridge}
 
@@
 \begin{itemize}
-    \item Flesh out Section \ref{sec:encouraging_collaborative_developement}
+    \item Flesh out Section \ref{CDHS_sec:encouraging_collaborative_developement}
     \item QMC Software Architecture section: better way then description of components and list of softwares? 
-    \item For Section \ref{sec:tenets_of_great_software}, the slides also include 
+    \item For Section \ref{CDHS_sec:tenets_of_great_software}, the slides also include 
     \begin{itemize}
@@
 
-\section{Introduction} \label{sec:intro}
+\section{Introduction} \label{CDHS_sec:intro}
 Quasi-Monte Carlo (QMC) methods promise great efficiency gains over independent and identically distributed (IID) Monte Carlo (MC) methods.
@@
 
-\section{Available Software for QMC} \label{sec:available} 
+\section{Available Software for QMC} \label{CDHS_sec:available} 
 QMC software spans  LD sequence generators, algorithms, and applications.  We review the better known software collections, recognizing that some software overlaps multiple categories.
@@
 
-The software libraries referenced in Section \ref{sec:available} provide one or more of these components.
+The software libraries referenced in Section \ref{CDHS_sec:available} provide one or more of these components.
 QMCPy combines multiple examples of them all under and object oriented framework.
@@
 
-\section{Stopping Criteria} \label{sec:stopping_crit}
+\section{Stopping Criteria} \label{CDHS_sec:stopping_crit}
 
@@
 Figure \ref{fig:sc_comp} compares the time and sample sizes needed to compute the $5$-dimensional Keister integral \eqref{eq:KeisterIntegral} using IID sequences and LD lattice sequences.
-Consistent with what is stated in Section \ref{sec:intro}, the error of IID MC is $\calO(n^{-1/2})$, which means that the time and sample size to obtain an absolute error tolerance of $\varepsilon$ is $\calO(\varepsilon^{-2})$.
+Consistent with what is stated in Section \ref{CDHS_sec:intro}, the error of IID MC is $\calO(n^{-1/2})$, which means that the time and sample size to obtain an absolute error tolerance of $\varepsilon$ is $\calO(\varepsilon^{-2})$.
 By contrast, the  error of QMC using LD sequences is $\calO(n^{-1+\epsilon})$, which implies $\calO(\varepsilon^{-1-\epsilon})$ times and sample sizes.
@@
 This allows users to quickly see the properties that have been set and defaulted for any given component.
-This feature was already used to print the data object returned by the stopping criterion's \texttt{integrate} method in Section \ref{sec:stopping_crit}.
+This feature was already used to print the data object returned by the stopping criterion's \texttt{integrate} method in Section \ref{CDHS_sec:stopping_crit}.
 The following code prints a drifted \texttt{BrownianMotion} instance and displays the connection to an underlying Gaussian measure.
@@
 
-\section{Further Work} \label{sec:further}
+\section{Further Work} \label{CDHS_sec:further}
 

### 2023-02-10 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

@@
 These sequences may be deterministic or random and are intended to have an empirical distribution that approximates well the uniform distribution.
+LD generators are the most prevalent generally available components of QMC software.
+These generators appear in Association of Computing Machinery  publications \cite{BraFox88,BraFoxNie92,HonHic00a}, FinDer \cite{PasTra95,FinDer}, BRODA \cite{BRODA20a}, NAG \cite{NAG27}, and MATLAB \cite{MAT9.13}.
 
@@
 This may mean large libraries including QMC routines and/or building connections between QMC libraries and other libraries.
-
-The most prevalent generally available QMC software are LD generators.
-They appear in ACM publications \cite{BraFox88,BraFoxNie92,HonHic00a}, FinDer \cite{PasTra95,FinDer}, BRODA \cite{BRODA20a}, NAG \cite{NAG27}, and MATLAB \cite{MAT9.13}.
 

### 2023-02-11 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

@@
 
-\item Correct,
-
-\item Efficient in computational time and memory,
-
-\item Accessible to practitioners and theorists alike,
-
-\item Sustainable by a community that owns it, and
-
-\item Scalable to advanced hardware architectures.
+\item Correct (Sect.\ \ref{CDHS_sec:correct}),
+
+\item Efficient in computational time and memory (Sect.\ \ref{CDHS_sec:efficient}),
+
+\item Accessible to practitioners and theorists alike (Sect.\ \ref{CDHS_sec:accessible}),
+
+\item Sustainable by a community that owns it (Sect.\ \ref{CDHS_sec:sustainable}), and
+
+\item Scalable to advanced hardware architectures (Sect.\  \ref{CDHS_sec:scalable}).
 
@@
 First, a development team must determine the value of the software to the users.
-We discuss why great (Q)MC software in necessary in Section \ref{CDHS_sec:why_we_need_software}.
-Second, a team must determine high level objectives for the software; the tenets of great software are described in Section \ref{CDHS_sec:tenets_of_great_software}.
+We discuss why great (Q)MC software in necessary in Sect.\ \ref{CDHS_sec:why_we_need_software}.
+Second, a team must determine high level objectives for the software; the tenets of great software are described in Sect.\ \ref{CDHS_sec:tenets_of_great_software}.
 Third, an architecture for the  (Q)MC software should be determined.
-We propose a modularization of the (Q)MC pipeline in Section \ref{CDHS_sec:qmc_software_architecture}.
+We propose a modularization of the (Q)MC pipeline in Sect.\ \ref{CDHS_sec:qmc_software_architecture}.
 Next, a team must choose which programming languages, hardware environments, and distribution platforms they wish to support.
-We discuss the nuance and challenges in these decisions in Section \ref{CDHS_sec:language_library_environment}.
+We discuss the nuance and challenges in these decisions in Sect.\ \ref{CDHS_sec:language_library_environment}.
 Finally, a project which hopes to stand the test of time must draw attention and support from the broader (Q)MC community.
-To this end, we encourage collaborative community development and describe its' best practices in  Section \ref{CDHS_sec:encouraging_collaborative_developement}.
+To this end, we encourage collaborative community development and describe its' best practices in  Sect.\ \ref{CDHS_sec:encouraging_collaborative_developement}.
 
@@
 
+We next discuss characteristics of great QMC software outlined in the introduction.  For each characteristic we identify what is lacking and what might be done to remedy the lack.
+
 \section{Integrated} \label{CDHS_sec:integrated}
-We now discuss characteristics of great QMC software outlined in the introduction.
-
-Expanding on the problem formulation in the introduction, we want to approximate the integral or expectation, $\mu$, well by the sample mean, $\hat{\mu}$:
+
+Expanding on the problem formulation in the introduction, we want to approximate well the integral or expectation, $\mu$, by the sample mean, $\hat{\mu}$:
 \begin{gather} 
@@
 \end{gather}
-Progressing from the left side of this equation to the right requires several steps.
+Progressing from the original problem to a satisfactory solution requires several software components.
+Great QMC software environment allows different implementations of these components to be interchanged.
+
+\subsection{LD Generators} A lattice, digital net, or Halton point generator typically supplies the sequence $\bsx_1, \bsx_2, \ldots$.
+These sequences may be deterministic or random and are intended to have an empirical distribution that approximates well the uniform distribution.
+
+LD generators are the most prevalent generally available components of QMC software.
+These generators appear in Association of Computing Machinery  publications \cite{BraFox88,BraFoxNie92,HonHic00a}, FinDer \cite{PasTra95,FinDer}, libseq \cite{FriKel02,FriKelweb}, BRODA \cite{BRODA20a}, NAG \cite{NAG27}, MATLAB \cite{MAT9.13}, SamplePack \cite{SamplePack}, Gr\"nschlo{\ss}'s website \cite{GruWeb}, the Magic Point Shop \cite{NuyWeb}, Julia \cite{QMCJulia}, SciPy \cite{virtanen2020scipy}, PyTorch \cite{paszke2019pytorch}, and TensorFlow \cite{tfqfQMC2021a}, and MatBuilder \cite{paulin2022}.
+Joe and Kuo
 
@@
 
-An \emph{LD generator}, such as a lattice, digital net, or Halton point generator supplies the sequence $\bsx_1, \bsx_2, \ldots$.
-These sequences may be deterministic or random and are intended to have an empirical distribution that approximates well the uniform distribution.
-LD generators are the most prevalent generally available components of QMC software.
-These generators appear in Association of Computing Machinery  publications \cite{BraFox88,BraFoxNie92,HonHic00a}, FinDer \cite{PasTra95,FinDer}, BRODA \cite{BRODA20a}, NAG \cite{NAG27}, and MATLAB \cite{MAT9.13}.
+ Joe and Kuo
 
@@
 
+R by Lemieux
+
 LatNetBuilder \cite{LatNet} 
-    \item Low discrepancy generators \cite{FriKel02,FriKelweb}, SamplePack \cite{SamplePack}, \cite{GruWeb},  and MatBuilder \cite{paulin2022} 
+   and MatBuilder \cite{paulin2022} 
     \item Sobol' direction numbers \cite{JoeKuo03,JoeKuo08,SobolDirection}
@@
     \item Uncertainty quantification libraries Dakota \cite{DakotaUsersManual}, UQTk \cite{DebEtal04,UQTk}, and MUQ \cite{MUQ}  have some basic level low discrepancy sampling
-    \item QMC framework in Julia since 2019 \cite{QMCJulia} by Robbe and others 
-    \item Scrambled Sobol' in SciPy \cite{virtanen2020scipy} and PyTorch \cite{paszke2019pytorch} since several years ago
-    \item TensorFlow QMC framework \cite{tfqfQMC2021a} since a year or so ago
+ago
     \item CUBA \cite{CUBA}
@@
 A simple example in \AGSComment{CITE} shows a straightforward integration between the QMC package QMCPy \cite{QMCPy2020a} and the PDE package FEniCS/Dolfin \cite{LoggEtal_10_2012}, both in Python.
+
+* Standard definitions, e.g., for LD generator
+* UM-Bridge, Dolfin
+* Move to other libraries
+* Demos, bridges
+
+\section{Correct} \label{CDHS_sec:correct}
+* Matlab randomization
+* Don't omit first point
+*
+
+\section{Efficient} \label{CDHS_sec:efficient}
+* New architectures
+* Memory and FLOPS
+
+\section{Accessible} \label{CDHS_sec:accessible}
+* Documentation
+* Demos
+* Tutorials
+* Blogs
+* Supported, issues, features
+* Up to date
+*
+
+\section{Sustainable} \label{CDHS_sec:sustainable} Community Owned
+* Developers, contributors
+* Use to illustrate research code
+* Use cases, new applications
+* Referenced
+* Build STEM pipeline
+
+\section{Scalable} \label{CDHS_sec:scalable}
 
@@
 A library should be straightforward to install from a common distribution platform, consistent in user interface, intuitive to navigate to modules and documentation, comprehensible in its' examples, and supportive in user engagement.
-Software accessibility considerations are further discussed in Section \ref{CDHS_sec:language_library_environment}.
+Software accessibility considerations are further discussed in Sect.\ \ref{CDHS_sec:language_library_environment}.
 
@@
 However, we contend that a sub-optimal but user-friendly library with community engagement may be more impactful than an optimal library which lacks support.
-We promote collaborative, community driven development further in Section \ref{CDHS_sec:encouraging_collaborative_developement}.
+We promote collaborative, community driven development further in Sect.\ \ref{CDHS_sec:encouraging_collaborative_developement}.
 

### 2023-02-11 MCQMC2022ChallengesQMCSoftwareArticle.tex
2023 Mac 14 Updates

@@
 %%%%Fred start
-\include{common-macros.tex}
+\input{common-macros.tex}
 \usepackage{cite}
@@
 LD generators are the most prevalent generally available components of QMC software.
-These generators appear in Association of Computing Machinery  publications \cite{BraFox88,BraFoxNie92,HonHic00a}, FinDer \cite{PasTra95,FinDer}, libseq \cite{FriKel02,FriKelweb}, BRODA \cite{BRODA20a}, NAG \cite{NAG27}, MATLAB \cite{MAT9.13}, SamplePack \cite{SamplePack}, Gr\"nschlo{\ss}'s website \cite{GruWeb}, the Magic Point Shop \cite{NuyWeb}, Julia \cite{QMCJulia}, SciPy \cite{virtanen2020scipy}, PyTorch \cite{paszke2019pytorch}, and TensorFlow \cite{tfqfQMC2021a}, and MatBuilder \cite{paulin2022}.
+These generators appear in Association of Computing Machinery  publications \cite{BraFox88,BraFoxNie92,HonHic00a}, FinDer \cite{PasTra95,FinDer}, libseq \cite{FriKel02,FriKelweb}, BRODA \cite{BRODA20a}, NAG \cite{NAG27}, MATLAB \cite{MAT9.13}, SamplePack \cite{SamplePack}, Gr\"nschlo{\ss}'s website \cite{GruWeb}, the Magic Point Shop \cite{NuyWeb}, R \cite{QRNG2020}, Julia \cite{QMCJulia}, SciPy \cite{virtanen2020scipy}, PyTorch \cite{paszke2019pytorch}, and TensorFlow \cite{tfqfQMC2021a}, and MatBuilder \cite{paulin2022}.
 Joe and Kuo
@@
 This may mean large libraries including QMC routines and/or building connections between QMC libraries and other libraries.
-
-R by Lemieux
 

### 2023-02-11 MCQMC2022ChallengesQMCSoftwareArticle.tex
2023 Mac 14 Updates

@@
 
-\subsection{LD Generators} A lattice, digital net, or Halton point generator typically supplies the sequence $\bsx_1, \bsx_2, \ldots$.
+\subsection{LD Sequence Generators} A lattice, digital sequence, or Halton sequence generator typically supplies the sequence $\bsx_1, \bsx_2, \ldots$.
 These sequences may be deterministic or random and are intended to have an empirical distribution that approximates well the uniform distribution.
@@
 These generators appear in Association of Computing Machinery  publications \cite{BraFox88,BraFoxNie92,HonHic00a}, FinDer \cite{PasTra95,FinDer}, libseq \cite{FriKel02,FriKelweb}, BRODA \cite{BRODA20a}, NAG \cite{NAG27}, MATLAB \cite{MAT9.13}, SamplePack \cite{SamplePack}, Gr\"nschlo{\ss}'s website \cite{GruWeb}, the Magic Point Shop \cite{NuyWeb}, R \cite{QRNG2020}, Julia \cite{QMCJulia}, SciPy \cite{virtanen2020scipy}, PyTorch \cite{paszke2019pytorch}, and TensorFlow \cite{tfqfQMC2021a}, and MatBuilder \cite{paulin2022}.
-Joe and Kuo
-
-The \emph{integrand}, $g$, and the non-negative weight, $\lambda$, define the original integral arising from an application.
-The function $g$ may represent the discounted payoff of a financial derivative, or the velocity of flow at a point in rock with stochastic porosity.
+
+Lattice and digital sequences are not unique, and improved or alternative versions continue to be constructed.
+Joe and Kuo \cite{JoeKuo03,JoeKuo08,SobolDirection} have proposed  Sobol' generator direction numbers, which are now widely used, but were not part of the early LD sequence software.
+L'Ecuyer and colleagues have created LatNetBuilder \cite{LatNet} to construct good lattice and polynomial lattice sequences based on user defined criteria.
+As was raised during a lunch at MCQMC 2022, we should agree on a consistent format for storing the parameters that define good LD sequences so that they can by shared across software libraries.
+
+\subsection{Integrands and Variable Transformations} The original integral or expectation arising from the application is defined in terms of the integrand, $g$, and the non-negative weight, $\lambda$.
+For example,
+\begin{itemize}
+	\item $g$ may represent the discounted payoff of a financial derivative and $\lambda$ be the probability density function (PDF) for a discretized Brownian motion \cite{Gla03},
+
+	\item  $g$ may represent the velocity of flow at a point in rock and $\lambda$ is the PDF of the porosity field , or
+
+	\item $g\lambda$ is the unnormalized Bayesian posterior density multiplied by parameter of interest.
+\end{itemize},
 If $\lambda$ is a probability density function for a random variable $\bsT$, then $\mu = \EE[g(\bsT)]$.
@@
 
+\iffalse
 LatNetBuilder \cite{LatNet}
@@
     \item CUBA \cite{CUBA}
+\fi
 

### 2023-02-13 MCQMC2022ChallengesQMCSoftwareArticle.tex
2023 Mac 14 Updates

@@
 \begin{itemize}
-	\item $g$ may represent the discounted payoff of a financial derivative and $\lambda$ be the probability density function (PDF) for a discretized Brownian motion \cite{Gla03},
-
-	\item  $g$ may represent the velocity of flow at a point in rock and $\lambda$ is the PDF of the porosity field , or
+	\item $g$ is the discounted payoff of a financial derivative and $\lambda$ be the probability density function (PDF) for a discretized Brownian motion \cite{Gla03},
+
+	\item  $g$ is the velocity of a fluid at a point in rock with a random porosity field whose discretized PDF is $\lambda$, or
 

### 2023-02-13 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

@@
 \end{gather}
-Progressing from the original problem to a satisfactory solution requires several software components.
-Great QMC software environment allows different implementations of these components to be interchanged.
+Progressing from the original problem to a satisfactory solution requires several software components, which we describe in the subsections below.
+Great QMC software environments allows different implementations of these components to be interchanged.
 
@@
 
+For the comprehensive software environments, such as NAG, MATLAB, R, SciPy, and TensorFlow, including the LD sequence generators allow users to a ????
+
 \subsection{Integrands and Variable Transformations} The original integral or expectation arising from the application is defined in terms of the integrand, $g$, and the non-negative weight, $\lambda$.
@@
 \begin{itemize}
-	\item $g$ is the discounted payoff of a financial derivative and $\lambda$ be the probability density function (PDF) for a discretized Brownian motion \cite{Gla03},
-
-	\item  $g$ is the velocity of a fluid at a point in rock with a random porosity field whose discretized PDF is $\lambda$, or
+	\item $g$ is the discounted payoff of a financial derivative and $\lambda$ is the probability density function (PDF) for a discretized Brownian motion \cite{Gla03},
+
+	\item  $g$ is the velocity of a fluid at a point in rock with a random porosity field whose discretized PDF is $\lambda$ \cite{}, or
 
 	\item $g\lambda$ is the unnormalized Bayesian posterior density multiplied by parameter of interest.
-\end{itemize},
-If $\lambda$ is a probability density function for a random variable $\bsT$, then $\mu = \EE[g(\bsT)]$.
+\end{itemize}
 
@@
 \begin{equation*}
-    \Psi((0,1)^d) = \calT, \qquad f(x) = g(\bsPsi(\bsx)) \lambda(\bsPsi(\bsx)) \abs{\frac{\partial \bsPsi}{\partial{\bsx}}}.
+    \Psi((0,1)^d) = \calT, \qquad f(\bsx) = g(\bsPsi(\bsx)) \lambda(\bsPsi(\bsx)) \abs{\frac{\partial \bsPsi}{\partial\bsx}}.
 \end{equation*}
 Such a transformation is typically non-unique, and the choice of a good one is equivalent to importance sampling.
+If $\lambda$ is a PDF for a random variable $\bsT$, then $\mu = \EE[g(\bsT)]$.
+Moreover, one might choose $\bsPsi$ such that $\bsT = \bsPsi(\bsX)$, in which case $\lambda(\bsPsi(\bsx)) \abs{\partial \bsPsi/\partial\bsx} = 1$ and $f(\bsx) = g(\bsPsi(\bsx))$.
 

### 2023-03-04 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

@@
 LD generators are the most prevalent generally available components of QMC software.
-These generators appear in Association of Computing Machinery  publications \cite{BraFox88,BraFoxNie92,HonHic00a}, FinDer \cite{PasTra95,FinDer}, libseq \cite{FriKel02,FriKelweb}, BRODA \cite{BRODA20a}, NAG \cite{NAG27}, MATLAB \cite{MAT9.13}, SamplePack \cite{SamplePack}, Gr\"nschlo{\ss}'s website \cite{GruWeb}, the Magic Point Shop \cite{NuyWeb}, R \cite{QRNG2020}, Julia \cite{QMCJulia}, SciPy \cite{virtanen2020scipy}, PyTorch \cite{paszke2019pytorch}, and TensorFlow \cite{tfqfQMC2021a}, and MatBuilder \cite{paulin2022}.
+These generators appear in Association of Computing Machinery  publications \cite{BraFox88,BraFoxNie92,HonHic00a}, FinDer \cite{PasTra95,FinDer}, libseq \cite{FriKel02,FriKelweb}, BRODA \cite{BRODA20a}, NAG \cite{NAG27}, MATLAB \cite{MAT9.13}, SamplePack \cite{SamplePack}, Gr\"unschlo{\ss}'s website \cite{GruWeb}, the Magic Point Shop \cite{NuyWeb}, R \cite{QRNG2020}, Julia \cite{QMCJulia}, SciPy \cite{virtanen2020scipy}, PyTorch \cite{paszke2019pytorch}, and TensorFlow \cite{tfqfQMC2021a}, MatBuilder \cite{paulin2022}, and  QMCPy \cite{QMCPy2020a}.
 
@@
 
-For the comprehensive software environments, such as NAG, MATLAB, R, SciPy, and TensorFlow, including the LD sequence generators allow users to a ????
+For the comprehensive software environments, such as NAG, MATLAB, R, SciPy, and TensorFlow, the LD sequence generators included already allow users to easily access them while drawing from other packages in these environments.
+However, the developers of smaller libraries, like BRODA and QMCPy should assume the responsibility of showing how their libraries can connect well with others.
 
@@
 
- Joe and Kuo
-
+Complex integrands might be best evaluated by libraries that have no direct connection to QMC software libraries.
+In the uncertainty quantification (UQ) space, the recently developed UM-Bridge \cite{umbridge} (the UQ and Model Bridge), which connects libraries that generate sequences of model parameters, $\bsx_1, \bsx_2, \ldots$ with libraries that evaluate models, $f(\bsx_1), f(\bsx_2), \ldots$.
+We illustrate how UM-Bridge connects QMCPy with a differential equation solver for a cantilevered beam with random Young's modulus \cite{ParSee22a}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.3\textwidth]{cantileveredbeam.eps}
+    \caption{Caption}
+    \label{fig:my_label}
+\end{figure}
+
+\subsection{Stopping Criteria}
 Finally, a \emph{stopping criterion} is determines the sample size, $n$, required to satisfy the error criterion \eqref{CDHS_eq:error_tol} given in terms of the absolute error tolerance, $\varepsilon$.
@@
 
+\subsection{??}
 No one software library contains all of these features, which is why great QMC software must be well integrated into other software libraries.

### 2023-03-04 MCQMC2022ChallengesQMCSoftwareArticle.tex
Home Computer Updates
Updates from Overleaf

@@
 \providecommand{\HickernellFJ}{Hickernell\xspace}
+\graphicspath{{ProgramsImages/}}
 
@@
 LD generators are the most prevalent generally available components of QMC software.
-These generators appear in Association of Computing Machinery  publications \cite{BraFox88,BraFoxNie92,HonHic00a}, FinDer \cite{PasTra95,FinDer}, libseq \cite{FriKel02,FriKelweb}, BRODA \cite{BRODA20a}, NAG \cite{NAG27}, MATLAB \cite{MAT9.13}, SamplePack \cite{SamplePack}, Gr\"unschlo{\ss}'s website \cite{GruWeb}, the Magic Point Shop \cite{NuyWeb}, R \cite{QRNG2020}, Julia \cite{QMCJulia}, SciPy \cite{virtanen2020scipy}, PyTorch \cite{paszke2019pytorch}, and TensorFlow \cite{tfqfQMC2021a}, MatBuilder \cite{paulin2022}, and  QMCPy \cite{QMCPy2020a}.
+These generators appear in Association of Computing Machinery  publications \cite{BraFox88,BraFoxNie92,HonHic00a}, FinDer \cite{PasTra95,FinDer}, libseq \cite{FriKel02,FriKelweb}, BRODA \cite{BRODA20a}, NAG \cite{NAG27}, MATLAB \cite{MAT9.13}, SamplePack \cite{SamplePack}, Gr\"unschlo{\ss}'s website \cite{GruWeb}, the Magic Point Shop \cite{NuyWeb}, R \cite{QRNG2020}, Julia \cite{QMCJulia}, PyTorch \cite{paszke2019pytorch}, SAS \cite{SAS_LD}, SciPy \cite{virtanen2020scipy}, and TensorFlow \cite{tfqfQMC2021a}, MatBuilder \cite{paulin2022}, and  QMCPy \cite{QMCPy2020a}.
 
@@
 For the comprehensive software environments, such as NAG, MATLAB, R, SciPy, and TensorFlow, the LD sequence generators included already allow users to easily access them while drawing from other packages in these environments.
-However, the developers of smaller libraries, like BRODA and QMCPy should assume the responsibility of showing how their libraries can connect well with others.
+Some large environments, like Dakota \cite{DakotaUsersManual} and SAS have limited LD generators and should expand their offerings.
+Developers of smaller libraries, like BRODA and QMCPy, should take initiative to show how their libraries can connect well with others.
 
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{cantileveredbeam.eps}
+    \includegraphics[width=0.7\textwidth]{ldparallelbeam.eps}
+    \caption{The expected deflection of a beam with random Young's modulus.
+QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.
+For small error tolerances, the time and sample size is $\calO(n^{-1})$, and parallel processing gives a speed-up of factor four to five.}
+    \label{fig:beam}
+\end{figure}
+
 Complex integrands might be best evaluated by libraries that have no direct connection to QMC software libraries.
 In the uncertainty quantification (UQ) space, the recently developed UM-Bridge \cite{umbridge} (the UQ and Model Bridge), which connects libraries that generate sequences of model parameters, $\bsx_1, \bsx_2, \ldots$ with libraries that evaluate models, $f(\bsx_1), f(\bsx_2), \ldots$.
-We illustrate how UM-Bridge connects QMCPy with a differential equation solver for a cantilevered beam with random Young's modulus \cite{ParSee22a}.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.3\textwidth]{cantileveredbeam.eps}
-    \caption{Caption}
-    \label{fig:my_label}
-\end{figure}
+We illustrate in Figure \ref{fig:beam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam with random Young's modulus \cite{ParSee22a} where the randomness is given by $\bsX \sim \calU[1,1.2]^3$.
+The beam position is discretized at $30$ points.
+QMCPy computes the expected deflection of the beam a these points for a sequence of error tolerances using lattice rules.
+By taking advantage of virtual machines and docker containers, the evaluation of the beam deflection can be computed in parallel in multiple cores of the laptop computer to decrease execution time.
+
+We need more connections of QMC software to other libraries, like what has been done with  QMCPy with and UM-Bridge.
+This will facilitate the newer and better QMC algorithms tackling new applications.
 
 \subsection{Stopping Criteria}
-Finally, a \emph{stopping criterion} is determines the sample size, $n$, required to satisfy the error criterion \eqref{CDHS_eq:error_tol} given in terms of the absolute error tolerance, $\varepsilon$.
-One can also imagine error criteria involving a relative error tolerance or a combination of absolute and relative error tolerances.
-
-\subsection{??}
-No one software library contains all of these features, which is why great QMC software must be well integrated into other software libraries.
-This may mean large libraries including QMC routines and/or building connections between QMC libraries and other libraries.
-
-\iffalse
-LatNetBuilder \cite{LatNet}
-   and MatBuilder \cite{paulin2022}
-    \item Sobol' direction numbers \cite{JoeKuo03,JoeKuo08,SobolDirection}
-    \item Fast CBC, Magic Point Shop, QMC4PDE and other code since 2004  at \cite{NuyWeb}
-    \item Multi-level software \cite{GilesSoft,GilesQSoft}
-    \item Data-driven error bounds and stopping criteria in GAIL \cite{ChoEtal21a} and QMCPy \cite{QMCPy2020a, ChoEtal22a}
-    \item Uncertainty quantification libraries Dakota \cite{DakotaUsersManual}, UQTk \cite{DebEtal04,UQTk}, and MUQ \cite{MUQ}  have some basic level low discrepancy sampling
-ago
-    \item CUBA \cite{CUBA}
-\fi
-
-This requires four components, which we implement as QMCPy classes.
-
-Software should be \emph{integrated} into the environment's scientific software ecosystem.
-More simply, libraries should play nicely together and utilize the mature developments of other teams.
-A simple example in \AGSComment{CITE} shows a straightforward integration between the QMC package QMCPy \cite{QMCPy2020a} and the PDE package FEniCS/Dolfin \cite{LoggEtal_10_2012}, both in Python.
-
-* Standard definitions, e.g., for LD generator
-* UM-Bridge, Dolfin
-* Move to other libraries
-* Demos, bridges
+The cantilevered beam example illustrated in Figure \ref{fig:beam} utilizes QMCPy's stopping criterion to choose the sample size, $n$, based on the input error tolerance, $\varepsilon$, and the observed discrete Fourier series coefficients of the integrand, $f$.
+See \cite{HicJim16a} for details.
+Theoretically justified stopping criteria for choosing sample sizes are important and a relatively recent development \cite{HicEtal14a,HicJim16a,JimHic16a,RatHic19a,JagHic22a}.
+Some of these stopping criteria for absolute error criteria, \eqref{CDHS_eq:error_tol},  have been extended to relative error criteria or a combination of absolute and relative error criteria, as well as stopping criteria for approximating functions of several integrals \cite{Jia16a,GilJim16b,HicEtal17a,JagSor23a}.
+
+We emphasize again that no one software library contains all of these features, which is why great QMC software must be well integrated into other software libraries.
+This may mean large libraries incorporating more QMC routines and/or building connections between QMC libraries and other libraries.
 
 \section{Correct} \label{CDHS_sec:correct}
+Although correctness is an obvious feature of great QMC software, it cannot always be assumed in popular packages.
+
 * Matlab randomization

### 2023-03-04 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

@@
 \section{Correct} \label{CDHS_sec:correct}
-Although correctness is an obvious feature of great QMC software, it cannot always be assumed in popular packages.
-
-* Matlab randomization
-* Don't omit first point
-*
+Although correctness is an obvious feature of great QMC software, it cannot always be assumed in popular packages.  Here are two examples.
+
+MATLAB's original randomization of Sobol' points was incorrectly implemented.
+This was discovered by Llu/'is Antoni Jim\'enez Rugama who informed the developer.
+The error was corrected in MATLAB R2018a.
+
+SciPy and PyTorch's original implementations of Sobol' sequences omitted the first point.
+The rationale is that the unscrambled first point of the Sobol' sequence is $\bsx_1 = \bszero$, and this point becomes the undesirable $\boldsymbol{-\infty}$ when transformed to mimic a Gaussian distribution.
+A vigorous discussion ensued \cite{scipySobol2020a}, which prompted an article by Art Owen \cite{Owe22a} on why one should not drop the first point and destroy the next structure.
+Instead one should randomly scramble the Sobol' sequence (and other LD sequences) so that $\bsx_1 \ne \bszero$.
+Although SciPy and PyTorch were corrected, other libraries with LD sequences still omit the first point.
+
+QMC software is sometimes written by those not intimately involved in the QMC community.
+Those of us who are most familiar with good QMC practice need to vigilantly be aware of QMC software in popular practices and inform the developers when bugs are spotted.
 
 \section{Efficient} \label{CDHS_sec:efficient}
+
 * New architectures

### 2023-03-04 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

@@
 Instead one should randomly scramble the Sobol' sequence (and other LD sequences) so that $\bsx_1 \ne \bszero$.
-Although SciPy and PyTorch were corrected, other libraries with LD sequences still omit the first point.
-
-QMC software is sometimes written by those not intimately involved in the QMC community.
+Although SciPy and PyTorch were corrected, other software libraries with LD sequences still omit the first point.
+
+QMC software is sometimes written by those not intimately a part of the QMC community.
 Those of us who are most familiar with good QMC practice need to vigilantly be aware of QMC software in popular practices and inform the developers when bugs are spotted.
@@
 \section{Efficient} \label{CDHS_sec:efficient}
-
-* New architectures
-* Memory and FLOPS
+Another obvious feature of great QMC software is numerical efficiency.
+This requires developers to identify bottlenecks in performance.
+Intensive computation may need to be rewritten in lower level languages, like C.
+Code may need to be refactored to avoid excessive calls to memory.
+As new hardware architectures arise, QMC software should adapt.
 
 \section{Accessible} \label{CDHS_sec:accessible}
-* Documentation
-* Demos
-* Tutorials
-* Blogs
-* Supported, issues, features
-* Up to date
-*
-
-\section{Sustainable} \label{CDHS_sec:sustainable} Community Owned
+One attribute of great QMC software that often receives insufficient attention is its accessibility.
+Developers may write code for their own purposes and move on.
+Great software should be stored on a repository, where the latest version can be downloaded.
+There should be documentation that guides the user through installation and explains the library's features.
+Demos are crucial for showing how the software work and forming a basis for a new user to write code to solve their own examples.
+There should be a place for users to report bugs and make feature requests.
+Tutorials help newcomers to QMC understand its advantages.
+Blogs educate a wider audience and enlarge our QMC community.
+
+\section{Sustainable} \label{CDHS_sec:sustainable} 
+Great QMC software should be alive.
+It should be under active development or merged into a libraary that is actively maintained.
+Commercial software has a infrastructure to sustain the algorithms in their libraries, but without attention, those algorithms can become stale.
+Open source software requires a community of active users and developers to keep the library up-to-date and bug-free.
+
+Community Owned
 * Developers, contributors
@@
 * Build STEM pipeline
-
-\section{Scalable} \label{CDHS_sec:scalable}
 

### 2023-03-06 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

@@
 \authorrunning{S.-C.\ T.\ Choi et al}
-\author{Sou-Cheng Terrya Choi \and Yuhan Ding \and Fred J. Hickernell \and Aleksei G. Sorokin}
-\institute{Sou-Cheng Terrya Choi \at Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{schoi32@hawk.iit.edu}
+\author{Sou-Cheng Terrya Choi \and Yuhan Ding \and Fred J. Hickernell \and R. Jagadeeswaran Aleksei G. Sorokin}
+\institute{
+Sou-Cheng Terrya Choi \at Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{schoi32@hawk.iit.edu}
+\and 
+Yuhan Ding \at
+ Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{yding2@hawk.iit.edu}
 \and
@@
 Department of Applied Mathematics, Illinois Institute of Technology \\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{hickernell@iit.edu}
-\and Yuhan Ding \at
+\and
+R. Jagadeeswaran \at ???
 \and
@@
 }
-
-\newpage
 
@@
 However, the practice of QMC is often more complicated.
-The original problem may need to be rewritten fit above form and/or to facilitate a good approximation with small $n$.
-Practitioners may wish to determine $n$ adaptively to satisfy a desired error tolerance.
+The original problem may need to be rewritten to fit above form and/or to facilitate a good approximation with small $n$.
+Practitioners may wish to determine $n$ adaptively to satisfy a prescribed error tolerance.
 
@@
 \begin{itemize}
-\item Integrated with related libraries (Sect.\ \ref{CDHS_sec:integrated}),
-
-\item Correct (Sect.\ \ref{CDHS_sec:correct}),
-
-\item Efficient in computational time and memory (Sect.\ \ref{CDHS_sec:efficient}),
-
-\item Accessible to practitioners and theorists alike (Sect.\ \ref{CDHS_sec:accessible}),
-
-\item Sustainable by a community that owns it (Sect.\ \ref{CDHS_sec:sustainable}), and
-
-\item Scalable to advanced hardware architectures (Sect.\  \ref{CDHS_sec:scalable}).
-
-\end{itemize}
-
-Developing software to implement such enhanced MC strategies poses a number of challenges.
-First, a development team must determine the value of the software to the users.
-We discuss why great (Q)MC software in necessary in Sect.\ \ref{CDHS_sec:why_we_need_software}.
-Second, a team must determine high level objectives for the software; the tenets of great software are described in Sect.\ \ref{CDHS_sec:tenets_of_great_software}.
-Third, an architecture for the  (Q)MC software should be determined.
-We propose a modularization of the (Q)MC pipeline in Sect.\ \ref{CDHS_sec:qmc_software_architecture}.
-Next, a team must choose which programming languages, hardware environments, and distribution platforms they wish to support.
-We discuss the nuance and challenges in these decisions in Sect.\ \ref{CDHS_sec:language_library_environment}.
-Finally, a project which hopes to stand the test of time must draw attention and support from the broader (Q)MC community.
-To this end, we encourage collaborative community development and describe its' best practices in  Sect.\ \ref{CDHS_sec:encouraging_collaborative_developement}.
-
-\section{Why Develop QMC Software} \label{CDHS_sec:why_we_need_software}
+\item Integrated with related libraries (Sect.\ \ref{CDHJS_sec:integrated}),
+
+\item Correct (Sect.\ \ref{CDHJS_sec:correct}),
+
+\item Efficient in computational time and memory (Sect.\ \ref{CDHJS_sec:efficient}),
+
+\item Accessible to practitioners and theorists alike (Sect.\ \ref{CDHJS_sec:accessible}), and
+
+\item Sustainable by a community that owns it (Sect.\ \ref{CDHJS_sec:sustainable}).
+
+\end{itemize}
+
+In each section we describe the challenges faced and how they might be overcome.
+
+\section{Why Develop QMC Software} \label{CDHJS_sec:why_we_need_software}
 
@@
 Software developers must write code that is computationally efficient, numerical stable, and provides reasonable default choices of tuning parameters.
-Good QMC software relieves users of these concerns.
+Great QMC software relieves users of these concerns.
 
@@
 
-Good QMC software opens up new application areas for QMC methods by allowing practitioners to compare new methods to existing ones.
+Great QMC software opens up new application areas for QMC methods by allowing practitioners to compare new methods to existing ones.
 QMC software has been successful in quantitative finance, uncertainty quantification, and image rendering.
@@
 
-\section{Integrated} \label{CDHS_sec:integrated}
+\section{Integrated} \label{CDHJS_sec:integrated}
 
@@
 	\mu : = \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  = \bbE[f(\bsX)] = \int_{[0,1]^d} f(\bsx)  \, \D \bsx \approx \frac 1n \sum_{i=0}^{n-1} f(\bsx_i) =: \hat{\mu}_n, \\
- \label{CDHS_eq:error_tol}
+ \label{CDHJS_eq:error_tol}
  \abs{\mu - \hat{\mu}_n} \le \varepsilon.
@@
 Progressing from the original problem to a satisfactory solution requires several software components, which we describe in the subsections below.
-Great QMC software environments allows different implementations of these components to be interchanged.
+Great QMC software environments allows different implementations of these components to be freely interchanged.
 
@@
 
-For the comprehensive software environments, such as NAG, MATLAB, R, SciPy, and TensorFlow, the LD sequence generators included already allow users to easily access them while drawing from other packages in these environments.
-Some large environments, like Dakota \cite{DakotaUsersManual} and SAS have limited LD generators and should expand their offerings.
-Developers of smaller libraries, like BRODA and QMCPy, should take initiative to show how their libraries can connect well with others.
-
-\subsection{Integrands and Variable Transformations} The original integral or expectation arising from the application is defined in terms of the integrand, $g$, and the non-negative weight, $\lambda$.
-For example,
+Users computing solutions in comprehensive software environments, such as NAG, MATLAB, R, SciPy, and TensorFlow, have convenient access to LD sequence generators.
+However, most of these could provide more flexibility in their LD offerings.
+Some large environments, such as  Dakota \cite{DakotaUsersManual} and SAS, have quite limited LD generators and should expand their offerings.
+Developers of smaller libraries with more QMC features, like BRODA and QMCPy, should demonstrate how their libraries can connect well with other software environments.
+An example involving QMCPy is given in the next subsection.
+
+\subsection{Integrands and Variable Transformations} \label{CDHJS:sec:integrands}
+The original integral or expectation arising from the application is defined in terms of the integrand, $g$, and the non-negative weight, $\lambda$.  For example,
 \begin{itemize}
@@
 
-	\item  $g$ is the velocity of a fluid at a point in rock with a random porosity field whose discretized PDF is $\lambda$ \cite{}, or
-
-	\item $g\lambda$ is the unnormalized Bayesian posterior density multiplied by parameter of interest.
-\end{itemize}
-
-To put $\mu$ into the form that is directly accessible to QMC methods requires a \emph{variable transformation}, $\bsPsi$, such that
+	\item  $g$ is the velocity of a fluid at a point in rock with a random porosity field whose discretized PDF is $\lambda$ \cite{KuoNuy16a}, or
+
+	\item $g\lambda$ is the unnormalized Bayesian posterior density multiplied by parameter of interest \cite{GelEtal13}.
+\end{itemize}
+
+To put $\mu$ into the form that is directly accessible to QMC methods requires a variable transformation, $\bsPsi$, such that
 \begin{equation*}
@@
 Complex integrands might be best evaluated by libraries that have no direct connection to QMC software libraries.
-In the uncertainty quantification (UQ) space, the recently developed UM-Bridge \cite{umbridge} (the UQ and Model Bridge), which connects libraries that generate sequences of model parameters, $\bsx_1, \bsx_2, \ldots$ with libraries that evaluate models, $f(\bsx_1), f(\bsx_2), \ldots$.
-We illustrate in Figure \ref{fig:beam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam with random Young's modulus \cite{ParSee22a} where the randomness is given by $\bsX \sim \calU[1,1.2]^3$.
+In the uncertainty quantification (UQ) space, the recently developed UQ and Model Bridge (UM-Bridge) \cite{umbridge}, connects libraries that generate sequences of model parameters, $\bsx_1, \bsx_2, \ldots$ with libraries that evaluate models, $f(\bsx_1), f(\bsx_2), \ldots$.
+We illustrate in Figure \ref{fig:beam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam with random Young's modulus \cite{ParSee22a}, where the randomness is given by $\bsX \sim \calU[1,1.2]^3$.
 The beam position is discretized at $30$ points.
@@
 Theoretically justified stopping criteria for choosing sample sizes are important and a relatively recent development \cite{HicEtal14a,HicJim16a,JimHic16a,RatHic19a,JagHic22a}.
-Some of these stopping criteria for absolute error criteria, \eqref{CDHS_eq:error_tol},  have been extended to relative error criteria or a combination of absolute and relative error criteria, as well as stopping criteria for approximating functions of several integrals \cite{Jia16a,GilJim16b,HicEtal17a,JagSor23a}.
+Some of these stopping criteria for absolute error criteria, \eqref{CDHJS_eq:error_tol},  have been extended to relative error criteria or a combination of absolute and relative error criteria, as well as stopping criteria for approximating functions of several integrals \cite{Jia16a,GilJim16b,HicEtal17a,JagSor23a}.
 
@@
 
-\section{Correct} \label{CDHS_sec:correct}
+\section{Correct} \label{CDHJS_sec:correct}
 Although correctness is an obvious feature of great QMC software, it cannot always be assumed in popular packages.  Here are two examples.
@@
 
-\section{Efficient} \label{CDHS_sec:efficient}
+\section{Efficient} \label{CDHJS_sec:efficient}
 Another obvious feature of great QMC software is numerical efficiency.
@@
 
-\section{Accessible} \label{CDHS_sec:accessible}
+\section{Accessible} \label{CDHJS_sec:accessible}
 One attribute of great QMC software that often receives insufficient attention is its accessibility.
@@
 
-\section{Sustainable} \label{CDHS_sec:sustainable} 
+\section{Sustainable} \label{CDHJS_sec:sustainable} 
 Great QMC software should be alive.
@@
 
-\section{Tenets of Great Software} \label{CDHS_sec:tenets_of_great_software}
-
-In this section we lay out common aspirations of scientific software.
-While some objectives are expected by practitioners, e.g.
-correctly written software, others may better be characterized as features that are pursued as a project matures, e.g.
-supporting scalable computation.
-The following paragraphs highlight aspirations with more important, expected behaviours coming earlier in the discussion.
-
-Software is expected to be \emph{correct}.
-A practitioner generating a digital sequence expects the routine to produce theoretically accurate points.
-For example, the QMC community has helped developers ensure implementations of low discrepancy sequence generators include the zeroth point \cite{Owe22a, scipySobol2020a} and  have correct randomization routines.
-%, and disallow thinning / leaping: taking every $k^\text{th}$ value in the sequence.
-
-Algorithms should be implemented in an \emph{efficient} manner.
-Both the compute time and memory requirements should match theoretical developments which in turn should be as close to the state of the art as possible.
-As a classical example, it is expected that an implementation of the discrete Fourier transform for $n$ points has complexity $\mathcal{O}(n\log n)$.
-
-Software should be \emph{accessible} in terms of installation, presentation, navigation, exemplification, and communication.
-A library should be straightforward to install from a common distribution platform, consistent in user interface, intuitive to navigate to modules and documentation, comprehensible in its' examples, and supportive in user engagement.
-Software accessibility considerations are further discussed in Sect.\ \ref{CDHS_sec:language_library_environment}.
-
-Libraries should be \emph{sustainable} for future development.
-The goal should be to have a sufficient user base and developer community for updates and maintenance.
-Similar to accessibility, sustainability is often overlooked for aspirations discussed later in this section.
-However, we contend that a sub-optimal but user-friendly library with community engagement may be more impactful than an optimal library which lacks support.
-We promote collaborative, community driven development further in Sect.\ \ref{CDHS_sec:encouraging_collaborative_developement}.
-
-Software should be \emph{integrated} into the environment's scientific software ecosystem.
-More simply, libraries should play nicely together and utilize the mature developments of other teams.
-A simple example in \AGSComment{CITE} shows a straightforward integration between the QMC package QMCPy \cite{QMCPy2020a} and the PDE package FEniCS/Dolfin \cite{LoggEtal_10_2012}, both in Python.
-
-Routines should be \emph{scalable} to advanced computer architectures.
-The widespread availability of on demand computing with multi-core and multi-GPU machines encourages software that can take advantage of parallel computation.
-Scalable algorithms lend themselves to a larger class or problems and provide practitioners the opportunity to more quickly test ideas.
-
-\section{(Quasi-)Monte Carlo Software Architecture} \label{CDHS_sec:qmc_software_architecture}
-
-An important step in the development of any software is to determine its' architecture.
-How should components interact with each other?
-Which aspects should be modular?
-Object oriented?
-Procedural?
-Answers to these questions form the blueprint of software design and guide implementation.
-
-This section proposes an object oriented design for (Q)MC software.
-The authors have implemented this architecture into the QMCPy \cite{QMCPy2020a} Python package which is further detailed in the MCQMC 2020 proceedings article \AGSComment{CITE}.
-While this is not the only possible architecture for (Q)MC problems, we hope the high level object definitions facilitate easy collaboration and integration for research across the (Q)MC community.
-
-The standard (Q)MC problem is to approximate the true mean $\mu := \mathbb{E}[g(\boldsymbol{T})]$ where $\boldsymbol{T}$ is some $d$ dimensional random variable.
-For nice $\boldsymbol{T}$, one may perform a change of variables to write $\mu = \mathbb{E}[f(\boldsymbol{X})]$ where $\boldsymbol{X} \sim \mathcal{U}[0,1]^d$.
-For example, if $\boldsymbol{T} \sim \mathcal{N}(\boldsymbol{0},\boldsymbol{I})$, then one may set $f(\boldsymbol{x}) := g(\boldsymbol{\Phi}^{-1}(\boldsymbol{x}))$ for $\boldsymbol{x} \in [0,1]^d$ where $\boldsymbol{\Phi}^{-1}$ is the inverse CDF of a standard normal distribution taken element wise.
-One may then approximate $\mu$ by the $n$ sample average $\hat{\mu} = 1/n \sum_{i=1}^n f(\boldsymbol{x}_i)$ where $\{\boldsymbol{x}_i\}_{i=1}^n$ are chosen so their empirical distribution is ``close'' to the uniform distribution.
-The approximation error $\varepsilon = \lvert \mu - \hat{\mu} \rvert$ is $\mathcal{O}(n^{-1/2})$ when $\{\boldsymbol{x}_i\}_{i=1}^n$ are chosen to be IID standard uniform as in the case of simple Monte Carlo (MC) and almost $\mathcal{O}(n^{-1})$ when $\{\boldsymbol{x}_i\}_{i=1}^n$ are chosen from a low discrepancy sequence as done for Quasi-Monte Carlo (QMC).
-Instead of selecting the number of samples $n$ and then determining the error $\varepsilon$, it is often desirable to select $\varepsilon$ and then adaptively determine $n$.
-From this view, $n$ is $\mathcal{O}(\varepsilon^{-2})$ for MC and almost $\mathcal{O}(\varepsilon^{-1})$ for QMC.
-A more detailed account of MC and QMC can be found in \AGSComment{CITE}.
-
-The (Q)MC pipeline above can be decomposed into $4$ main components:
-\begin{description}
-    \item[\textbf{Generator:}] producing $\{\boldsymbol{x}_i\}_{i=1}^n$ whose empirical distribution is ``close'' to the standard uniform distribution in $d$ dimensions for any $n \geq 1$.
-May produce IID or low discrepancy sequences depending on the problem.
-    \item[\textbf{Measure:}] that defines the distribution of $\boldsymbol{T}$ e.g., uniform, Gaussian, Lebesgue. This definition facilitates choosing a change of variables to determine $f$.
-    \item[\textbf{Integrand:}] $g$, which defines the original integral. When provided knowledge of the true measure, the integrand can also evaluate $f$ via the change of variables.
-    \item[\textbf{Stopping Criterion:}] based on a data-driven error bound, which determines how large $n$ should be to ensure that $\varepsilon$ is less than some user-provided error threshold.
-\end{description}
-A variety of QMC materials and the their categorizations into the above framework are described in the following list.
-\begin{itemize}
-    %\item \cite{LEc2017a} provides a history of random number generation
-    %\item Notorious RANDU from the 1960s \cite{RANDU}, which fails the spectral test
-    %\item VEGAS \cite{Lep78a, Lep21a} importance sampling Monte Carlo algorithm favored by physicists
-    \item ACM low discrepancy point generators \cite{BraFox88,BraFoxNie92,HonHic00a}
-    \item FinDer \cite{PasTra95,FinDer} and BRODA \cite{BRODA20a} targeting quantitative finance
-    \item Korobov cubature and scrambled Sobol' generators in NAG \cite{NAG27} for decades
-    \item Scrambled Sobol' and Halton generators in MATLAB \cite{MAT9.13} since 2008, and fixed a few years later
-    %\item BUGS \cite{BUGSBook, BUGSweb} and Stan \cite{STAN} for Markov Chain Monte Carlo (MCMC)
-    \item LatMRG \cite{LEcCou97}, RNGStreams \cite{LEcEtal02},  SSJ \cite{LEc2002a,SSJ}, and LatNetBuilder \cite{LatNet}
-    \item Low discrepancy generators \cite{FriKel02,FriKelweb}, SamplePack \cite{SamplePack}, \cite{GruWeb},  and MatBuilder \cite{paulin2022}
-    \item Sobol' direction numbers \cite{JoeKuo03,JoeKuo08,SobolDirection}
-    \item Fast CBC, Magic Point Shop, QMC4PDE and other code since 2004  at \cite{NuyWeb}
-    \item Multi-level software \cite{GilesSoft,GilesQSoft}
-    \item Data-driven error bounds and stopping criteria in GAIL \cite{ChoEtal21a} and QMCPy \cite{QMCPy2020a, ChoEtal22a}
-    \item Uncertainty quantification libraries Dakota \cite{DakotaUsersManual}, UQTk \cite{DebEtal04,UQTk}, and MUQ \cite{MUQ}  have some basic level low discrepancy sampling
-    \item QMC framework in Julia since 2019 \cite{QMCJulia} by Robbe and others
-    \item Scrambled Sobol' in SciPy \cite{virtanen2020scipy} and PyTorch \cite{paszke2019pytorch} since several years ago
-    \item TensorFlow QMC framework \cite{tfqfQMC2021a} since a year or so ago
-    \item CUBA \cite{CUBA}
-\end{itemize}
-
-\section{Choosing a Environment} \label{CDHS_sec:language_library_environment}
-
-This section discusses the challenges in choosing environments for development and distribution.
-The development environment includes the choice of which languages, architectures, testing methods, and documentation methods to support.
-The deployment environment will often be influenced by the choice of programming languages.
-
-The choice of programming language should be aligned with the communities the team wishes to reach with their software.
-The choice should also account for developer familiarity and ability to quickly prototype ideas.
-A team should align their choice of language to balance the desire to have user-friendly prototyping code versus the desire to have high performance computing (HPC) ``ninja'' code written for production setting.
-To the second point, it is also important to consider which language will best facilitate the parallel and GPU computations the team wishes to support.
-
-Many programming languages come bundled with their own preferred distribution systems.
-Python has PyPI, R has CRAN, Julia has Pkg.
-However, many projects support multiple languages and have complex dependency requirements.
-In such cases, it can be helpful to containerize your application using a service like Docker.
-
-\section{Encouraging Collaborative Development} \label{CDHS_sec:encouraging_collaborative_developement}
-
-We now discuss ways a team may encourage collaborative development while touching on a few best practices.
-As discussed earlier, collaborative development promotes greater community engagement, ecosystem integration, and software longevity.
-
-The authors have found the Git version control system to be an invaluable for software development.
-Git hosts such as GitHub or Bitbucket may offer teams additional features such as tracking issues, managing discussion threads, and setting up automated workflows for jobs such as testing and documentation compilation.
-
-Accessibility is key to community engagment.
-It is paramount that your package be easy to install, well documented, and tested regularly.
-Aside from static communications, a team should also be engaged in addressing users bugs and be attentive to collaborating with proposed code changes.
-
-It is also important for a development team to make a clear path for collaborators to contributors to the software.
-Templates for issues and pull requests are a good place to start.
-Contributors should also be encouraged to write robust tests for their updates and highlight their new features in a demo for greater accessibility.
-
-\section{Example Collaboration with UMBridge and QMCPy} \label{CDHS_sec:example_collaboration_umbridge}
-
-This section discusses a collaboration between the UM-Bridge (the UQ and Model Bridge) \AGSComment{CITE} and QMCPy (the Quasi-Monte Carlo Python library) \cite{QMCPy2020a} softwares.
-The UM-Bridge documentation describes its' mission is to, ``provide a unified interface for numerical models that is accessible from virtually any programming language or framework.
-It is primarily intended for coupling advanced models (e.g.
-simulations of complex physical processes) to advanced statistics or optimization methods.'' QMCPy is a high level framework for combining QMC components in a compatible and consistent manner.
-The collaboration between these two packages focused on making the UQ methods of QMCPy compatible with the abstract model representations from UM-Bridge.
-Specifically, UM-Bridge models are made compatible with the QMCPy framework through an integrand wrapper for seamless compatibility with a variety of existing point generators, transformations, and stopping criterion already available in QMCPy.
-
-To exemplify the integration, we focus on quantifying uncertainty propagation of material properties of a cantilevered beam.
-It is assumed the beam is split into $3$ regions of random stiffness dependent on uncertain material properties.
-Given a sample of $3$ stiffness values, the UM-Bridge model provides the displacement at $31$ equidistant nodes along the beam.
-Our task is to quantify the uncertainty in displacement at each of these $31$ locations along the beam.
-
-A fully coded example of this problem is available at \url{https://github.com/QMCSoftware/QMCSoftware/blob/master/demos/umbridge.ipynb}.
-At a high level, one begins by running the UM-Bridge model in a Docker container.
-A model is then initialized in the UM-Bridge Python client which is then provided to QMCPy alongside a low discrepancy point generator and distribution governing the $3$ beam stiffness values.
-QMCPy then performs adaptive QMC cubature to simultaneously approximate the expected displacement at the $31$ locations of interest.
-When the cubature method queries the model, the QMCPy wrapper around UM-Bridge model sends requests to the Docker UM-Bridge model in parallel.
-The load balancing provided by UM-Bridge may be utilized to evaluate the model in parallel, a potentially advantageous feature for models which are costly to evaluate.
-Figure \ref{fig:umbridge_muq} depicts the $31$ approximate expected displacement values subject to uniform stiffness.
-
-\begin{figure}
-    \centering
-    \includegraphics[height=4cm]{umbridge_muq.png}
-    \caption{QMCPy approximate expected displacement at $31$ locations subject to random uniform stiffness. }
-    \label{fig:umbridge_muq}
-\end{figure}
-
 \bibliographystyle{spmpsci}
@@
 \AGSComment{
-\tableofcontents
-\section*{TODO}
-\begin{itemize}
-    \item Flesh out Section \ref{CDHS_sec:encouraging_collaborative_developement}
-    \item QMC Software Architecture section: better way then description of components and list of softwares?
-    \item For Section \ref{CDHS_sec:tenets_of_great_software}, the slides also include
-    \begin{itemize}
-        \item  Complete---contain the components or  easily access components in other libraries to solve real, complex problems
-        \item Current---include the latest and best algorithms
-    \end{itemize}
-    which I think fall under the umbrella of software being \emph{integrated} with the community.
-    \item update QMCPy citation \cite{QMCPy2020a}
-    \item figures?
-    \item I have left out or been breif about material from the following slides: QMCPy + FEniCS/Dolfin,  Building a Developer Team for Your Library
-\end{itemize}
+
 }
@@
 
-\section{Introduction} \label{CDHS_sec:intro}
+\section{Introduction} \label{CDHJS_sec:intro}
 Quasi-Monte Carlo (QMC) methods promise great efficiency gains over independent and identically distributed (IID) Monte Carlo (MC) methods.
@@
 
-\section{Available Software for QMC} \label{CDHS_sec:available}
+\section{Available Software for QMC} \label{CDHJS_sec:available}
 QMC software spans  LD sequence generators, algorithms, and applications.  We review the better known software collections, recognizing that some software overlaps multiple categories.
@@
 
-The software libraries referenced in Section \ref{CDHS_sec:available} provide one or more of these components.
+The software libraries referenced in Section \ref{CDHJS_sec:available} provide one or more of these components.
 QMCPy combines multiple examples of them all under and object oriented framework.
@@
 
-\section{Stopping Criteria} \label{CDHS_sec:stopping_crit}
+\section{Stopping Criteria} \label{CDHJS_sec:stopping_crit}
 
@@
 Figure \ref{fig:sc_comp} compares the time and sample sizes needed to compute the $5$-dimensional Keister integral \eqref{eq:KeisterIntegral} using IID sequences and LD lattice sequences.
-Consistent with what is stated in Section \ref{CDHS_sec:intro}, the error of IID MC is $\calO(n^{-1/2})$, which means that the time and sample size to obtain an absolute error tolerance of $\varepsilon$ is $\calO(\varepsilon^{-2})$.
+Consistent with what is stated in Section \ref{CDHJS_sec:intro}, the error of IID MC is $\calO(n^{-1/2})$, which means that the time and sample size to obtain an absolute error tolerance of $\varepsilon$ is $\calO(\varepsilon^{-2})$.
 By contrast, the  error of QMC using LD sequences is $\calO(n^{-1+\epsilon})$, which implies $\calO(\varepsilon^{-1-\epsilon})$ times and sample sizes.
@@
 This allows users to quickly see the properties that have been set and defaulted for any given component.
-This feature was already used to print the data object returned by the stopping criterion's \texttt{integrate} method in Section \ref{CDHS_sec:stopping_crit}.
+This feature was already used to print the data object returned by the stopping criterion's \texttt{integrate} method in Section \ref{CDHJS_sec:stopping_crit}.
 The following code prints a drifted \texttt{BrownianMotion} instance and displays the connection to an underlying Gaussian measure.
@@
 
-\section{Further Work} \label{CDHS_sec:further}
+\section{Further Work} \label{CDHJS_sec:further}
 
@@
 \end{document}
+
+\section{Tenets of Great Software} \label{CDHJS_sec:tenets_of_great_software}
+
+In this section we lay out common aspirations of scientific software.
+While some objectives are expected by practitioners, e.g.
+correctly written software, others may better be characterized as features that are pursued as a project matures, e.g.
+supporting scalable computation.
+The following paragraphs highlight aspirations with more important, expected behaviours coming earlier in the discussion.
+
+Software is expected to be \emph{correct}.
+A practitioner generating a digital sequence expects the routine to produce theoretically accurate points.
+For example, the QMC community has helped developers ensure implementations of low discrepancy sequence generators include the zeroth point \cite{Owe22a, scipySobol2020a} and  have correct randomization routines.
+%, and disallow thinning / leaping: taking every $k^\text{th}$ value in the sequence.
+
+Algorithms should be implemented in an \emph{efficient} manner.
+Both the compute time and memory requirements should match theoretical developments which in turn should be as close to the state of the art as possible.
+As a classical example, it is expected that an implementation of the discrete Fourier transform for $n$ points has complexity $\mathcal{O}(n\log n)$.
+
+Software should be \emph{accessible} in terms of installation, presentation, navigation, exemplification, and communication.
+A library should be straightforward to install from a common distribution platform, consistent in user interface, intuitive to navigate to modules and documentation, comprehensible in its' examples, and supportive in user engagement.
+Software accessibility considerations are further discussed in Sect.\ \ref{CDHJS_sec:language_library_environment}.
+
+Libraries should be \emph{sustainable} for future development.
+The goal should be to have a sufficient user base and developer community for updates and maintenance.
+Similar to accessibility, sustainability is often overlooked for aspirations discussed later in this section.
+However, we contend that a sub-optimal but user-friendly library with community engagement may be more impactful than an optimal library which lacks support.
+We promote collaborative, community driven development further in Sect.\ \ref{CDHJS_sec:encouraging_collaborative_developement}.
+
+Software should be \emph{integrated} into the environment's scientific software ecosystem.
+More simply, libraries should play nicely together and utilize the mature developments of other teams.
+A simple example in \AGSComment{CITE} shows a straightforward integration between the QMC package QMCPy \cite{QMCPy2020a} and the PDE package FEniCS/Dolfin \cite{LoggEtal_10_2012}, both in Python.
+
+Routines should be \emph{scalable} to advanced computer architectures.
+The widespread availability of on demand computing with multi-core and multi-GPU machines encourages software that can take advantage of parallel computation.
+Scalable algorithms lend themselves to a larger class or problems and provide practitioners the opportunity to more quickly test ideas.
+
+\section{(Quasi-)Monte Carlo Software Architecture} \label{CDHJS_sec:qmc_software_architecture}
+
+An important step in the development of any software is to determine its' architecture.
+How should components interact with each other?
+Which aspects should be modular?
+Object oriented?
+Procedural?
+Answers to these questions form the blueprint of software design and guide implementation.
+
+This section proposes an object oriented design for (Q)MC software.
+The authors have implemented this architecture into the QMCPy \cite{QMCPy2020a} Python package which is further detailed in the MCQMC 2020 proceedings article \AGSComment{CITE}.
+While this is not the only possible architecture for (Q)MC problems, we hope the high level object definitions facilitate easy collaboration and integration for research across the (Q)MC community.
+
+The standard (Q)MC problem is to approximate the true mean $\mu := \mathbb{E}[g(\boldsymbol{T})]$ where $\boldsymbol{T}$ is some $d$ dimensional random variable.
+For nice $\boldsymbol{T}$, one may perform a change of variables to write $\mu = \mathbb{E}[f(\boldsymbol{X})]$ where $\boldsymbol{X} \sim \mathcal{U}[0,1]^d$.
+For example, if $\boldsymbol{T} \sim \mathcal{N}(\boldsymbol{0},\boldsymbol{I})$, then one may set $f(\boldsymbol{x}) := g(\boldsymbol{\Phi}^{-1}(\boldsymbol{x}))$ for $\boldsymbol{x} \in [0,1]^d$ where $\boldsymbol{\Phi}^{-1}$ is the inverse CDF of a standard normal distribution taken element wise.
+One may then approximate $\mu$ by the $n$ sample average $\hat{\mu} = 1/n \sum_{i=1}^n f(\boldsymbol{x}_i)$ where $\{\boldsymbol{x}_i\}_{i=1}^n$ are chosen so their empirical distribution is ``close'' to the uniform distribution.
+The approximation error $\varepsilon = \lvert \mu - \hat{\mu} \rvert$ is $\mathcal{O}(n^{-1/2})$ when $\{\boldsymbol{x}_i\}_{i=1}^n$ are chosen to be IID standard uniform as in the case of simple Monte Carlo (MC) and almost $\mathcal{O}(n^{-1})$ when $\{\boldsymbol{x}_i\}_{i=1}^n$ are chosen from a low discrepancy sequence as done for Quasi-Monte Carlo (QMC).
+Instead of selecting the number of samples $n$ and then determining the error $\varepsilon$, it is often desirable to select $\varepsilon$ and then adaptively determine $n$.
+From this view, $n$ is $\mathcal{O}(\varepsilon^{-2})$ for MC and almost $\mathcal{O}(\varepsilon^{-1})$ for QMC.
+A more detailed account of MC and QMC can be found in \AGSComment{CITE}.
+
+The (Q)MC pipeline above can be decomposed into $4$ main components:
+\begin{description}
+    \item[\textbf{Generator:}] producing $\{\boldsymbol{x}_i\}_{i=1}^n$ whose empirical distribution is ``close'' to the standard uniform distribution in $d$ dimensions for any $n \geq 1$.
+May produce IID or low discrepancy sequences depending on the problem.
+    \item[\textbf{Measure:}] that defines the distribution of $\boldsymbol{T}$ e.g., uniform, Gaussian, Lebesgue. This definition facilitates choosing a change of variables to determine $f$.
+    \item[\textbf{Integrand:}] $g$, which defines the original integral. When provided knowledge of the true measure, the integrand can also evaluate $f$ via the change of variables.
+    \item[\textbf{Stopping Criterion:}] based on a data-driven error bound, which determines how large $n$ should be to ensure that $\varepsilon$ is less than some user-provided error threshold.
+\end{description}
+A variety of QMC materials and the their categorizations into the above framework are described in the following list.
+\begin{itemize}
+    %\item \cite{LEc2017a} provides a history of random number generation
+    %\item Notorious RANDU from the 1960s \cite{RANDU}, which fails the spectral test
+    %\item VEGAS \cite{Lep78a, Lep21a} importance sampling Monte Carlo algorithm favored by physicists
+    \item ACM low discrepancy point generators \cite{BraFox88,BraFoxNie92,HonHic00a}
+    \item FinDer \cite{PasTra95,FinDer} and BRODA \cite{BRODA20a} targeting quantitative finance
+    \item Korobov cubature and scrambled Sobol' generators in NAG \cite{NAG27} for decades
+    \item Scrambled Sobol' and Halton generators in MATLAB \cite{MAT9.13} since 2008, and fixed a few years later
+    %\item BUGS \cite{BUGSBook, BUGSweb} and Stan \cite{STAN} for Markov Chain Monte Carlo (MCMC)
+    \item LatMRG \cite{LEcCou97}, RNGStreams \cite{LEcEtal02},  SSJ \cite{LEc2002a,SSJ}, and LatNetBuilder \cite{LatNet}
+    \item Low discrepancy generators \cite{FriKel02,FriKelweb}, SamplePack \cite{SamplePack}, \cite{GruWeb},  and MatBuilder \cite{paulin2022}
+    \item Sobol' direction numbers \cite{JoeKuo03,JoeKuo08,SobolDirection}
+    \item Fast CBC, Magic Point Shop, QMC4PDE and other code since 2004  at \cite{NuyWeb}
+    \item Multi-level software \cite{GilesSoft,GilesQSoft}
+    \item Data-driven error bounds and stopping criteria in GAIL \cite{ChoEtal21a} and QMCPy \cite{QMCPy2020a, ChoEtal22a}
+    \item Uncertainty quantification libraries Dakota \cite{DakotaUsersManual}, UQTk \cite{DebEtal04,UQTk}, and MUQ \cite{MUQ}  have some basic level low discrepancy sampling
+    \item QMC framework in Julia since 2019 \cite{QMCJulia} by Robbe and others
+    \item Scrambled Sobol' in SciPy \cite{virtanen2020scipy} and PyTorch \cite{paszke2019pytorch} since several years ago
+    \item TensorFlow QMC framework \cite{tfqfQMC2021a} since a year or so ago
+    \item CUBA \cite{CUBA}
+\end{itemize}
+
+\section{Choosing a Environment} \label{CDHJS_sec:language_library_environment}
+
+This section discusses the challenges in choosing environments for development and distribution.
+The development environment includes the choice of which languages, architectures, testing methods, and documentation methods to support.
+The deployment environment will often be influenced by the choice of programming languages.
+
+The choice of programming language should be aligned with the communities the team wishes to reach with their software.
+The choice should also account for developer familiarity and ability to quickly prototype ideas.
+A team should align their choice of language to balance the desire to have user-friendly prototyping code versus the desire to have high performance computing (HPC) ``ninja'' code written for production setting.
+To the second point, it is also important to consider which language will best facilitate the parallel and GPU computations the team wishes to support.
+
+Many programming languages come bundled with their own preferred distribution systems.
+Python has PyPI, R has CRAN, Julia has Pkg.
+However, many projects support multiple languages and have complex dependency requirements.
+In such cases, it can be helpful to containerize your application using a service like Docker.
+
+\section{Encouraging Collaborative Development} \label{CDHJS_sec:encouraging_collaborative_developement}
+
+We now discuss ways a team may encourage collaborative development while touching on a few best practices.
+As discussed earlier, collaborative development promotes greater community engagement, ecosystem integration, and software longevity.
+
+The authors have found the Git version control system to be an invaluable for software development.
+Git hosts such as GitHub or Bitbucket may offer teams additional features such as tracking issues, managing discussion threads, and setting up automated workflows for jobs such as testing and documentation compilation.
+
+Accessibility is key to community engagment.
+It is paramount that your package be easy to install, well documented, and tested regularly.
+Aside from static communications, a team should also be engaged in addressing users bugs and be attentive to collaborating with proposed code changes.
+
+It is also important for a development team to make a clear path for collaborators to contributors to the software.
+Templates for issues and pull requests are a good place to start.
+Contributors should also be encouraged to write robust tests for their updates and highlight their new features in a demo for greater accessibility.
+
+\section{Example Collaboration with UMBridge and QMCPy} \label{CDHJS_sec:example_collaboration_umbridge}
+
+This section discusses a collaboration between the UM-Bridge (the UQ and Model Bridge) \AGSComment{CITE} and QMCPy (the Quasi-Monte Carlo Python library) \cite{QMCPy2020a} softwares.
+The UM-Bridge documentation describes its' mission is to, ``provide a unified interface for numerical models that is accessible from virtually any programming language or framework.
+It is primarily intended for coupling advanced models (e.g.
+simulations of complex physical processes) to advanced statistics or optimization methods.'' QMCPy is a high level framework for combining QMC components in a compatible and consistent manner.
+The collaboration between these two packages focused on making the UQ methods of QMCPy compatible with the abstract model representations from UM-Bridge.
+Specifically, UM-Bridge models are made compatible with the QMCPy framework through an integrand wrapper for seamless compatibility with a variety of existing point generators, transformations, and stopping criterion already available in QMCPy.
+
+To exemplify the integration, we focus on quantifying uncertainty propagation of material properties of a cantilevered beam.
+It is assumed the beam is split into $3$ regions of random stiffness dependent on uncertain material properties.
+Given a sample of $3$ stiffness values, the UM-Bridge model provides the displacement at $31$ equidistant nodes along the beam.
+Our task is to quantify the uncertainty in displacement at each of these $31$ locations along the beam.
+
+A fully coded example of this problem is available at \url{https://github.com/QMCSoftware/QMCSoftware/blob/master/demos/umbridge.ipynb}.
+At a high level, one begins by running the UM-Bridge model in a Docker container.
+A model is then initialized in the UM-Bridge Python client which is then provided to QMCPy alongside a low discrepancy point generator and distribution governing the $3$ beam stiffness values.
+QMCPy then performs adaptive QMC cubature to simultaneously approximate the expected displacement at the $31$ locations of interest.
+When the cubature method queries the model, the QMCPy wrapper around UM-Bridge model sends requests to the Docker UM-Bridge model in parallel.
+The load balancing provided by UM-Bridge may be utilized to evaluate the model in parallel, a potentially advantageous feature for models which are costly to evaluate.
+Figure \ref{fig:umbridge_muq} depicts the $31$ approximate expected displacement values subject to uniform stiffness.
+
+\begin{figure}
+    \centering
+    \includegraphics[height=4cm]{umbridge_muq.png}
+    \caption{QMCPy approximate expected displacement at $31$ locations subject to random uniform stiffness. }
+    \label{fig:umbridge_muq}
+\end{figure}
+
+\section*{TODO}
+\begin{itemize}
+    \item Flesh out Section \ref{CDHJS_sec:encouraging_collaborative_developement}
+    \item QMC Software Architecture section: better way then description of components and list of softwares?
+    \item For Section \ref{CDHJS_sec:tenets_of_great_software}, the slides also include
+    \begin{itemize}
+        \item  Complete---contain the components or  easily access components in other libraries to solve real, complex problems
+        \item Current---include the latest and best algorithms
+    \end{itemize}
+    which I think fall under the umbrella of software being \emph{integrated} with the community.
+    \item update QMCPy citation \cite{QMCPy2020a}
+    \item figures?
+    \item I have left out or been breif about material from the following slides: QMCPy + FEniCS/Dolfin,  Building a Developer Team for Your Library
+\end{itemize}
+
+Developing software to implement such enhanced MC strategies poses a number of challenges.
+First, a development team must determine the value of the software to the users.
+We discuss why great (Q)MC software in necessary in Sect.\ \ref{CDHJS_sec:why_we_need_software}.
+Second, a team must determine high level objectives for the software; the tenets of great software are described in Sect.\ \ref{CDHJS_sec:tenets_of_great_software}.
+Third, an architecture for the  (Q)MC software should be determined.
+We propose a modularization of the (Q)MC pipeline in Sect.\ \ref{CDHJS_sec:qmc_software_architecture}.
+Next, a team must choose which programming languages, hardware environments, and distribution platforms they wish to support.
+We discuss the nuance and challenges in these decisions in Sect.\ \ref{CDHJS_sec:language_library_environment}.
+Finally, a project which hopes to stand the test of time must draw attention and support from the broader (Q)MC community.
+To this end, we encourage collaborative community development and describe its' best practices in  Sect.\ \ref{CDHJS_sec:encouraging_collaborative_developement}.

### 2023-03-06 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

@@
 We illustrate in Figure \ref{fig:beam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam with random Young's modulus \cite{ParSee22a}, where the randomness is given by $\bsX \sim \calU[1,1.2]^3$.
-The beam position is discretized at $30$ points.
+The beam position is discretized at $31$ points.
 QMCPy computes the expected deflection of the beam a these points for a sequence of error tolerances using lattice rules.
-By taking advantage of virtual machines and docker containers, the evaluation of the beam deflection can be computed in parallel in multiple cores of the laptop computer to decrease execution time.
+By taking advantage of virtual machines and docker containers, the evaluation of the beam deflection can be computed in parallel in multiple cores of a single CPU to decrease execution time.
 
 We need more connections of QMC software to other libraries, like what has been done with  QMCPy with and UM-Bridge.
-This will facilitate the newer and better QMC algorithms tackling new applications.
+This will facilitate allow newer and better QMC algorithms to tackle new applications.
 
@@
 The cantilevered beam example illustrated in Figure \ref{fig:beam} utilizes QMCPy's stopping criterion to choose the sample size, $n$, based on the input error tolerance, $\varepsilon$, and the observed discrete Fourier series coefficients of the integrand, $f$.
-See \cite{HicJim16a} for details.
+See \cite{HicJim16a} for details of this stopping criterion for lattice rules.
 Theoretically justified stopping criteria for choosing sample sizes are important and a relatively recent development \cite{HicEtal14a,HicJim16a,JimHic16a,RatHic19a,JagHic22a}.
@@
 
-We emphasize again that no one software library contains all of these features, which is why great QMC software must be well integrated into other software libraries.
-This may mean large libraries incorporating more QMC routines and/or building connections between QMC libraries and other libraries.
+We emphasize again that no one software library contains all of the best LD generators, integrands, variable transformations, and stopping criteria, which is why great QMC software must be well integrated into other software libraries.
+This may take the form of large libraries incorporating more QMC routines and/or building connections among QMC libraries and with other non-QMC libraries.
 
@@
 MATLAB's original randomization of Sobol' points was incorrectly implemented.
-This was discovered by Llu/'is Antoni Jim\'enez Rugama who informed the developer.
+This was discovered by Llu\'is Antoni Jim\'enez Rugama who informed the developer.
 The error was corrected in MATLAB R2018a.
@@
 SciPy and PyTorch's original implementations of Sobol' sequences omitted the first point.
-The rationale is that the unscrambled first point of the Sobol' sequence is $\bsx_1 = \bszero$, and this point becomes the undesirable $\boldsymbol{-\infty}$ when transformed to mimic a Gaussian distribution.
-A vigorous discussion ensued \cite{scipySobol2020a}, which prompted an article by Art Owen \cite{Owe22a} on why one should not drop the first point and destroy the next structure.
+The rationale was that the unscrambled first point of the Sobol' sequence is $\bsx_1 = \bszero$, which becomes the undesirable $\boldsymbol{-\infty}$ when transformed to mimic a Gaussian distribution.
+A vigorous discussion ensued \cite{scipySobol2020a}, which prompted an article by Art Owen \cite{Owe22a} on why one should not drop the first point and destroy the digital net structure.
 Instead one should randomly scramble the Sobol' sequence (and other LD sequences) so that $\bsx_1 \ne \bszero$.
-Although SciPy and PyTorch were corrected, other software libraries with LD sequences still omit the first point.
-
-QMC software is sometimes written by those not intimately a part of the QMC community.
-Those of us who are most familiar with good QMC practice need to vigilantly be aware of QMC software in popular practices and inform the developers when bugs are spotted.
+Although SciPy and PyTorch were corrected, other software libraries with LD sequence generators still omit the first point.
+
+QMC software is sometimes written by those not intimately familiar with QMC.
+Those of us most familiar with good QMC practice need to vigilantly be aware of QMC algorithms in popular software and inform the developers when their software contains bugs or deviates from good practice.
 
@@
 
+Efficiency gains in QMC software can also be the result clever algorithm choices.
+The Bayesian stopping rules for lattice and Sobol sequences in \cite{RatHic19a,JagHic22a} choose covariance kernels that match the corresponding LD sequence.
+This reduces the computations from what would be $\calO(n^3)$ for arbitrary covariance kernels to only $\calO(n \log n)$ for these well-chosen kernels.
+
 \section{Accessible} \label{CDHJS_sec:accessible}
@@
 Developers may write code for their own purposes and move on.
-Great software should be stored on a repository, where the latest version can be downloaded.
+Great software should be stored in a repository, where the latest version can be downloaded.
 There should be documentation that guides the user through installation and explains the library's features.
-Demos are crucial for showing how the software work and forming a basis for a new user to write code to solve their own examples.
+Demos are crucial for showing how the software works and for providing a template for new users to write their own code.
 There should be a place for users to report bugs and make feature requests.
@@
 \section{Sustainable} \label{CDHJS_sec:sustainable} 
-Great QMC software should be alive.
-It should be under active development or merged into a libraary that is actively maintained.
-Commercial software has a infrastructure to sustain the algorithms in their libraries, but without attention, those algorithms can become stale.
+Great QMC software libraries should be alive.
+They should be under active development or merged into a another library that is actively maintained.
+Commercial software has an infrastructure to sustain their algorithms, but without attention, those algorithms may become stale.
 Open source software requires a community of active users and developers to keep the library up-to-date and bug-free.
 
-Community Owned
-* Developers, contributors
-* Use to illustrate research code
-* Use cases, new applications
-* Referenced
-* Build STEM pipeline
+The QMC community can help sustain great QMC software in several ways.
+Let's use existing QMC software libraries to build our research code and showcase our new algorithms.
+Doing this allows the next generation to more easily reproduce our results and build upon them.
+We should post demos of our new ideas on our own web pages, and if possible in the software libraries that we are using.
+QMCPy welcomes blogs, short articles introducing a broader audience to important ideas, illustrated by software.
+When we see bugs or missing features in QMC software, let's persistently request them.
+The more robust our QMC software, the easier it is to open new application areas to QMC and grow our QMC community.
+
+Many of us QMC researchers don't code.
+We rely on transient team members (students and postdocs) to write the code that showcases our new ideas.
+Building our own QMC libraries or contributing to existing ones helps ensure that the code your team writes will last beyond the time of the membership in your group.
+
+Encouraging students and postdocs to invest time in developing QMC software will require us as their mentors valuing software as a research output.
+Great software should be published with a digital object identifier (DOI).
+The term research software engineer (RSE) is becoming a more valued vocation.
+According to the website for  Research Software Engineers International 	\url{https://researchsoftware.org}, \emph{Research Software Engineers are people who combine professional software expertise with an understanding of research.}
+
+\section{Summary}
+QMC has experienced decades of exciting growth in theory since the late 1950s.
+The earliest publicly available QMC software dates to the early 1990s and its growth has not matched that of theory.
+As a QMC community, we have a responsibility to implement our great theory into great QMC software.
 
@@
 \bibliography{FJH23,FJHown23}
-
-\AGSComment{
-
-}
 

### 2023-03-06 MCQMC2022ChallengesQMCSoftwareArticle.tex
2023 Mac 14 Updates
Updates from Overleaf

@@
 \and
-R. Jagadeeswaran \at ???
+R. Jagadeeswaran \at
+Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{jrathin1@hawk.iit.edu}
 \and
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{cantileveredbeam.eps}
+    \includegraphics[width=0.5\textwidth]{cantileveredbeamwords.eps}
     \includegraphics[width=0.7\textwidth]{ldparallelbeam.eps}
-    \caption{The expected deflection of a beam with random Young's modulus.
+    \caption{The expected deflection of a cantilevered beam with random Young's modulus.
 QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.

### 2023-03-06 MCQMC2022ChallengesQMCSoftwareArticle.tex
aleksei: small wording changes

@@
 Quasi-Monte Carlo (QMC) methods have developed over several decades.
-With the explosion in computational science, there is a need for great software to realize that implements QMC algorithms.
+With the explosion in computational science, there is a need for great software that implements QMC algorithms.
 We review what QMC software has been developed to date, propose some criteria for developing great QMC software, and suggest some steps toward achieving great software.
@@
 However, the practice of QMC is often more complicated.
-The original problem may need to be rewritten to fit above form and/or to facilitate a good approximation with small $n$.
+The original problem may need to be rewritten to fit the above form and/or to facilitate a good approximation with small $n$.
 Practitioners may wish to determine $n$ adaptively to satisfy a prescribed error tolerance.
@@
 \label{eq:cubSummary}
-	\mu : = \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  = \bbE[f(\bsX)] = \int_{[0,1]^d} f(\bsx)  \, \D \bsx \approx \frac 1n \sum_{i=0}^{n-1} f(\bsx_i) =: \hat{\mu}_n, \\
+	\mu : = \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  = \bbE[f(\bsX)] = \int_{[0,1]^d} f(\bsx)  \, \D \bsx \approx \frac 1n \sum_{i=1}^{n} f(\bsx_i) =: \hat{\mu}_n, \\
  \label{CDHJS_eq:error_tol}
@@
 QMCPy computes the expected deflection of the beam a these points for a sequence of error tolerances using lattice rules.
-By taking advantage of virtual machines and docker containers, the evaluation of the beam deflection can be computed in parallel in multiple cores of a single CPU to decrease execution time.
-
-We need more connections of QMC software to other libraries, like what has been done with  QMCPy with and UM-Bridge.
-This will facilitate allow newer and better QMC algorithms to tackle new applications.
+By taking advantage of virtual machines and docker containers, UM-Bridge provides automatic load balancing to facilitate seamless parallel model evaluation.% the evaluation of the beam deflection can be computed in parallel in multiple cores of a single CPU to decrease execution time.
+
+We need more connections of QMC software to other libraries, like what has been done with  QMCPy and UM-Bridge.  This will allow newer and better QMC algorithms to tackle new applications.
 
@@
 QMC software is sometimes written by those not intimately familiar with QMC.
-Those of us most familiar with good QMC practice need to vigilantly be aware of QMC algorithms in popular software and inform the developers when their software contains bugs or deviates from good practice.
+Those of us most familiar with good QMC practice need to vigilantly be aware of QMC algorithms in popular software and inform the developers when their software contains bugs or deviates from best practice.
 

### 2023-03-06 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

@@
 The rationale was that the unscrambled first point of the Sobol' sequence is $\bsx_1 = \bszero$, which becomes the undesirable $\boldsymbol{-\infty}$ when transformed to mimic a Gaussian distribution.
-A vigorous discussion ensued \cite{scipySobol2020a}, which prompted an article by Art Owen \cite{Owe22a} on why one should not drop the first point and destroy the digital net structure.
+A vigorous discussion ensued \cite{scipySobol2020a,pytorchSoboldiscussion2020a}, which prompted an article by Art Owen \cite{Owe22a} on why one should not drop the first point and destroy the digital net structure.
 Instead one should randomly scramble the Sobol' sequence (and other LD sequences) so that $\bsx_1 \ne \bszero$.

### 2023-03-06 MCQMC2022ChallengesQMCSoftwareArticle.tex
SC: Fixed a few latex errors. Added a suggestion.

@@
 \newcommand{\AGSComment}[1]{{\color{red} #1}}
+\newcommand{\FredComment}[1]{{\color{blue} Fred: #1}}
+\newcommand{\SCComment}[1]{{\color{green} Sou-Cheng: #1}}
+\newcommand{\YDComment}[1]{{\color{orange} Yuhan: #1}}
+\newcommand{\JagsComment}[1]{{\color{violet} Jags: #1}}
+\DeclareMathOperator{\disc}{disc}
 
@@
 In each section we describe the challenges faced and how they might be overcome.
+
+\SCComment{Somewhere in this section or Section 8, perhaps we could say that we will use QMCPy for illustration. Something like: We draw on our collective experiences as 
+members of the academic community in the (Q)MC area; developers of open-source and commercial scientific software; users of (Q)MC and a wide range of scientific software as research scientists or data scientists.
+We highlight the insights we have gained in developing the (Q)MC scientific software packages, QMCPy~\cite{} and GAIL~\cite{}.
+In particular, QMCPy is not only a QMC software, but also a scientific middleware framework, which provides access to  more than X (Q)MC scientific software such as A~\cite{}, B~\cite{}, C~\cite{}.}
 

### 2023-03-07 MCQMC2022ChallengesQMCSoftwareArticle.tex
Intermediate changes

@@
 
-In the next section we describe why QMC software is important.  We then discuss characteristics of great QMC software:
+In the next section, we describe why QMC software is important.  We then discuss the characteristics of great QMC software:
 
@@
 However, translating pseudo-code into good executable code is nuanced.
-Software developers must write code that is computationally efficient, numerical stable, and provides reasonable default choices of tuning parameters.
+Software developers must write code that is computationally efficient, numerically stable, and provides reasonable default choices of tuning parameters.
 Great QMC software relieves users of these concerns.
@@
 QMC software has been successful in quantitative finance, uncertainty quantification, and image rendering.
-Unexpectedly good or bad computational results leads to open theoretical questions.
+Unexpectedly good or bad computational results lead to open theoretical questions.
 For example, the early application of QMC to high dimensional integrals arising in financial risk \cite{PasTra95} led to a wave of theoretical results on the tractability of integration in weighted spaces \cite{Woz99a,DicEtal14a,NovWoz10a}.
 
-We next discuss characteristics of great QMC software outlined in the introduction.  For each characteristic we identify what is lacking and what might be done to remedy the lack.
+We next discuss the characteristics of great QMC software outlined in the introduction.  For each characteristic, we identify what is lacking and what might be done to remedy the lack.
 
@@
 Progressing from the original problem to a satisfactory solution requires several software components, which we describe in the subsections below.
-Great QMC software environments allows different implementations of these components to be freely interchanged.
+Great QMC software environments allow different implementations of these components to be freely interchanged.
 
@@
 Joe and Kuo \cite{JoeKuo03,JoeKuo08,SobolDirection} have proposed  Sobol' generator direction numbers, which are now widely used, but were not part of the early LD sequence software.
-L'Ecuyer and colleagues have created LatNetBuilder \cite{LatNet} to construct good lattice and polynomial lattice sequences based on user defined criteria.
-As was raised during a lunch at MCQMC 2022, we should agree on a consistent format for storing the parameters that define good LD sequences so that they can by shared across software libraries.
+L'Ecuyer and colleagues have created LatNetBuilder \cite{LatNet} to construct good lattice and polynomial lattice sequences based on user-defined criteria.
+As was raised during a lunch at MCQMC 2022, we should agree on a consistent format for storing the parameters that define good LD sequences so that they can be shared across software libraries.
 
@@
 The beam position is discretized at $31$ points.
-QMCPy computes the expected deflection of the beam a these points for a sequence of error tolerances using lattice rules.
+QMCPy computes the expected deflection of the beam at these points for a sequence of error tolerances using lattice rules.
 By taking advantage of virtual machines and docker containers, UM-Bridge provides automatic load balancing to facilitate seamless parallel model evaluation.% the evaluation of the beam deflection can be computed in parallel in multiple cores of a single CPU to decrease execution time.
@@
 QMC software is sometimes written by those not intimately familiar with QMC.
-Those of us most familiar with good QMC practice need to vigilantly be aware of QMC algorithms in popular software and inform the developers when their software contains bugs or deviates from best practice.
+Those of us most familiar with good QMC practices need to vigilantly be aware of QMC algorithms in popular software and inform the developers when their software contains bugs or deviates from best practices.
 
@@
 This requires developers to identify bottlenecks in performance.
-Intensive computation may need to be rewritten in lower level languages, like C.
+The intensive computations may need to be rewritten in lower-level languages, like C.
 Code may need to be refactored to avoid excessive calls to memory.
@@
 
-Efficiency gains in QMC software can also be the result clever algorithm choices.
+Efficiency gains in QMC software can also be the result of clever algorithm choices.
 The Bayesian stopping rules for lattice and Sobol sequences in \cite{RatHic19a,JagHic22a} choose covariance kernels that match the corresponding LD sequence.

### 2023-03-07 MCQMC2022ChallengesQMCSoftwareArticle.tex
2023 Mac 14 Updates

@@
 \DeclareMathOperator{\disc}{disc}
+\newcommand{\FJHNote}[1]{{\color{blue}Fred: #1}}
 
@@
 With the explosion in computational science, there is a need for great software that implements QMC algorithms.
-We review what QMC software has been developed to date, propose some criteria for developing great QMC software, and suggest some steps toward achieving great software.
+We summarize the  QMC software that has been developed to date, propose some criteria for developing great QMC software, and suggest some steps toward achieving great software.
 }
@@
     \item Evaluate the integrand $f$ at these nodes to obtain $f(\bsx_i)$, $i=1,\dots,n$.
-    \item Estimate the true mean, $\mu$, by the sample mean
+    \item Estimate the true mean, $\mu$, by the sample mean,
     \begin{equation}
@@
 \end{itemize}
-
 In each section we describe the challenges faced and how they might be overcome.
 
-\SCComment{Somewhere in this section or Section 8, perhaps we could say that we will use QMCPy for illustration. Something like: We draw on our collective experiences as 
-members of the academic community in the (Q)MC area; developers of open-source and commercial scientific software; users of (Q)MC and a wide range of scientific software as research scientists or data scientists.
-We highlight the insights we have gained in developing the (Q)MC scientific software packages, QMCPy~\cite{} and GAIL~\cite{}.
+We draw on our collective experiences as
+members of the academic QMC community in the, developers of open-source and commercial scientific software, and users of a wide range of scientific software as research scientists or data scientists.
+Many of the insights that we have gained have come through our development of the Guaranteeed Automatic Integration Library (GAIL)~\cite{ChoEtal22a} in MATLAB and the Quasi-Monte Carlo Python library (QMCPy)~ \cite{QMCPy2020a}.
+\FJHNote{Sou-Cheng, not sure what this means.
 In particular, QMCPy is not only a QMC software, but also a scientific middleware framework, which provides access to  more than X (Q)MC scientific software such as A~\cite{}, B~\cite{}, C~\cite{}.}
@@
 This workshop not only discussed diagnostics and treatments for the challenges of developing great scientific software, but also emphasized the importance of implementing theoretical advancements into well written, accessible software libraries.
-Three cross-cutting themes that arose in this workshop were
+Three cross-cutting themes arose in this workshop:
 \begin{itemize}
@@
 For example, the early application of QMC to high dimensional integrals arising in financial risk \cite{PasTra95} led to a wave of theoretical results on the tractability of integration in weighted spaces \cite{Woz99a,DicEtal14a,NovWoz10a}.
+
+Great QMC software in eliminates the need researchers to resort to  ``do-it-yourself'' for established algorithms.  The less code we have to write, the fewer errors.
 
@@
 
-Lattice and digital sequences are not unique, and improved or alternative versions continue to be constructed.
-Joe and Kuo \cite{JoeKuo03,JoeKuo08,SobolDirection} have proposed  Sobol' generator direction numbers, which are now widely used, but were not part of the early LD sequence software.
-L'Ecuyer and colleagues have created LatNetBuilder \cite{LatNet} to construct good lattice and polynomial lattice sequences based on user-defined criteria.
-As was raised during a lunch at MCQMC 2022, we should agree on a consistent format for storing the parameters that define good LD sequences so that they can be shared across software libraries.
+Lattice and digital sequences are not unique, and improved or alternate versions continue to be constructed.
+Stephen Joe and Frances Kuo \cite{JoeKuo03,JoeKuo08,SobolDirection} have proposed  Sobol' generator direction numbers, which are now widely used, but were not part of early LD sequence software.
+Pierre L'Ecuyer and collaborators have created LatNetBuilder \cite{LatNet} to construct good lattice and polynomial lattice sequences based on user-defined criteria.
+During a lunch at MCQMC 2022, it was proposed that we  agree on a consistent format for storing the parameters that define good LD sequences so that they can be shared across software libraries.
 
 Users computing solutions in comprehensive software environments, such as NAG, MATLAB, R, SciPy, and TensorFlow, have convenient access to LD sequence generators.
-However, most of these could provide more flexibility in their LD offerings.
+However, most of these implementations in large libraries could provide more flexible LD offerings.
 Some large environments, such as  Dakota \cite{DakotaUsersManual} and SAS, have quite limited LD generators and should expand their offerings.
 Developers of smaller libraries with more QMC features, like BRODA and QMCPy, should demonstrate how their libraries can connect well with other software environments.
-An example involving QMCPy is given in the next subsection.
+An example of QMCPy connecting with other software is given in the next subsection.
 
@@
 If $\lambda$ is a PDF for a random variable $\bsT$, then $\mu = \EE[g(\bsT)]$.
-Moreover, one might choose $\bsPsi$ such that $\bsT = \bsPsi(\bsX)$, in which case $\lambda(\bsPsi(\bsx)) \abs{\partial \bsPsi/\partial\bsx} = 1$ and $f(\bsx) = g(\bsPsi(\bsx))$.
+In this case, one might choose $\bsPsi$ such that $\bsT \sim \bsPsi(\bsX)$, in which case $\lambda(\bsPsi(\bsx)) \abs{\partial \bsPsi/\partial\bsx} = 1$ and $f(\bsx) = g(\bsPsi(\bsx))$.
 
@@
 QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.
-For small error tolerances, the time and sample size is $\calO(n^{-1})$, and parallel processing gives a speed-up of factor four to five.}
+For small error tolerances, the time and sample size is $\calO(n^{-1})$, and parallel processing gives a speed-up of factor four to five.
+These plots may be reproduced using the Jupyter notebook \url{???}.}
     \label{fig:beam}
@@
 \section{Correct} \label{CDHJS_sec:correct}
-Although correctness is an obvious feature of great QMC software, it cannot always be assumed in popular packages.  Here are two examples.
+Although correctness is an obvious feature of great QMC software, it should not always be assumed in popular packages.  Here are two examples.
 
 MATLAB's original randomization of Sobol' points was incorrectly implemented.
-This was discovered by Llu\'is Antoni Jim\'enez Rugama who informed the developer.
+This was discovered by Llu\'is Antoni Jim\'enez Rugama, who informed the developer.
 The error was corrected in MATLAB R2018a.
@@
 QMC software is sometimes written by those not intimately familiar with QMC.
-Those of us most familiar with good QMC practices need to vigilantly be aware of QMC algorithms in popular software and inform the developers when their software contains bugs or deviates from best practices.
+Those of us most familiar with good QMC practices need to pay attention to QMC algorithms in popular software libraries and inform the developers when their software contains bugs or deviates from best practices.
 
@@
 This requires developers to identify bottlenecks in performance.
-The intensive computations may need to be rewritten in lower-level languages, like C.
+Intensive computations may need to be rewritten in lower-level languages, like C.
 Code may need to be refactored to avoid excessive calls to memory.
-As new hardware architectures arise, QMC software should adapt.
-
-Efficiency gains in QMC software can also be the result of clever algorithm choices.
+As new hardware architectures arise, QMC software should adapt to take advantage of them.
+
+Efficiency gains in QMC software can also result from clever algorithm choices.
 The Bayesian stopping rules for lattice and Sobol sequences in \cite{RatHic19a,JagHic22a} choose covariance kernels that match the corresponding LD sequence.
 This reduces the computations from what would be $\calO(n^3)$ for arbitrary covariance kernels to only $\calO(n \log n)$ for these well-chosen kernels.
+
+In spite of the desire for efficiency, great QMC software will be written in a variety of computer languages since different communities tend to favor different languages.
+Trade-offs will be made between efficiency and convenience.
 
@@
 Demos are crucial for showing how the software works and for providing a template for new users to write their own code.
-There should be a place for users to report bugs and make feature requests.
 Tutorials help newcomers to QMC understand its advantages.
 Blogs educate a wider audience and enlarge our QMC community.
+A suite of unit tests should be run regularly to ensure that new features do not introduce bugs.
+There should be a place for users to report bugs and make feature requests.
 
@@
 Great QMC software libraries should be alive.
-They should be under active development or merged into a another library that is actively maintained.
-Commercial software has an infrastructure to sustain their algorithms, but without attention, those algorithms may become stale.
-Open source software requires a community of active users and developers to keep the library up-to-date and bug-free.
-
-The QMC community can help sustain great QMC software in several ways.
-Let's use existing QMC software libraries to build our research code and showcase our new algorithms.
-Doing this allows the next generation to more easily reproduce our results and build upon them.
-We should post demos of our new ideas on our own web pages, and if possible in the software libraries that we are using.
+They should be under active development or merged into other libraries that are actively maintained.
+Commercial software has an infrastructure to maintain their algorithms, but these algorithms need periodic refreshing, Open source software requires a community of active users and developers to keep the library up-to-date and bug-free.
+
+The QMC community can help sustain great QMC software in several ways.  \begin{itemize}
+	\item Let's build our research code and showcase our new algorithms using existing QMC software libraries or build and maintain new libraries, if needed.
+This allows the next generation to more easily reproduce our results and build upon them.
+
+	\item Many of us QMC researchers rely on transient team members (students and postdocs) to write the code that illustrates our new ideas.
+Building our own QMC libraries or contributing to existing ones helps ensure that the code our team members write lasts beyond their involvement in our teams.
+
+	\item It is relatively easy to post demos of our new ideas on our own web pages, and if possible in the software libraries that we are using.
 QMCPy welcomes blogs, short articles introducing a broader audience to important ideas, illustrated by software.
-When we see bugs or missing features in QMC software, let's persistently request them.
+
+	\item When we see bugs or missing features in QMC software, let's persistently request fixes and improvements.
 The more robust our QMC software, the easier it is to open new application areas to QMC and grow our QMC community.
-
-Many of us QMC researchers don't code.
-We rely on transient team members (students and postdocs) to write the code that showcases our new ideas.
-Building our own QMC libraries or contributing to existing ones helps ensure that the code your team writes will last beyond the time of the membership in your group.
+\end{itemize}
 
@@
 \end{document}
-
-\begin{frame}{Why Do We Need Software?}
-	\vspace{-4ex}
-	\begin{itemize}
-		\item Theory
-		\begin{itemize}
-			\item Explains, justifies, and leads to algorithms
-			\item Assumptions$\implies$success can be viewed as failure$\implies$what went wrong
-		\end{itemize}
-		\item Software
-		\begin{itemize}
-			\item Makes algorithms practical
-			\item Solves (new) applications
-			\item Eliminates do-it-yourself
-		\end{itemize}
-		\item Applications
-  		\begin{itemize}
-			\item Make societal impact
-			\item Inspire new theory
-		\end{itemize}
-
-	\end{itemize}
-
-\uncover<2->{The US Department of Energy is investing in studying how to develop great scientific software, including a recent \emph{Workshop on the Science of Scientific-Software Development and Use} \cite{Her19a,ASCR-SSSDU}}
-\end{frame}
-
-\section{Structure}
-
-\begin{frame}{Software Architecture}
-	\vspace{-6ex}
-
-\[
-\mu :=  \int_{\ct} g(\vt) \, \lambda(\vt) \, \dif \vt = \cdots =
-\underbrace{\Ex[f(\vX)]}_{\text{expectation}} = \underbrace{\int_{[0,1]^d}  f(\vx) \, \dif \vx}_{\text{integration}} \approx  \frac 1{n} \sum_{i=1}^{n} f(\vX_i) =: \hmu_{n}
-\]
-
-\vspace{-3ex}
-
-\begin{description}[alabel]
-	%\setlength{\labelwidth}{5ex}
-
-	\item[Low Discrepancy Generator] producing $\{\vX_1, \vX_2, \dots \}$ that mimics the uniform distribution
-
-	\item[True Measure] that defines the original integral, e.g., Lebesgue; embodies the transformation $\vt = \vPsi(\vx)$
-
-	\item[Integrand] $g$, which defines the original integral, plus the transformed version, $f$, to fit the LD generator
-
-	\item[Stopping Criterion] based on a data-driven error bound, which determines how large $n$ should be to ensure that $\abs{\mu - \hmu_n} \le \varepsilon$
-\end{description}
-
-\vspace{-1ex}
-\uncover<2->{May we agree on a common software definition for these these objects like we have for floating point numbers or basic mathematical functions?}
-\end{frame}
-
-\begin{frame}{Choosing a Language, Library, or Environment}
-	\vspace{-5ex}
-	\begin{itemize}
-    \item Language/library choices of the users are driven by familiarity and speed
-
-  	\item  Favorites aligned with various communities, but no one dominates all
-
-	\item Difference between prototyping environments and ninja written production code
-
-    \item Growth of multi-processor environments
-    \begin{itemize}
-    	\item Provides opportunity
-    	\item Complicates software development; where does parallelism really help?
-    \end{itemize}
-
-    \item<2-> Recommendations if you want to contribute software
-    \begin{itemize}
-    \item Realize your new idea as a demo in your favorite library
-    \item Add a significant idea as a feature to other libraries
-    \item Move basic features of your library to larger library
-    \item Write wrappers or demos to connect other libraries to your library
-
-    \end{itemize}
- \end{itemize}
-\end{frame}
-
-\begin{frame}{Ex.\ QMCPy $+$ FEniCS/Dolfin \cite{LoggEtal_10_2012}}
-\vspace{-3ex}
-	\begin{itemize}
-		\item Want to demonstrate how our QMC software could work with a popular DE solver for DEs with random coefficients
-		\item Partial success \only<3->{Why?}
-	\end{itemize}
- 	\only<1>{\vspace{-3ex}}\only<2>{\vspace{-2ex}}\only<3>{\vspace{0ex}}
- \begin{minipage}{0.48\textwidth}
- 	\only<1-2>{\begin{gather*}
- 		-\frac{\dif }{\dif x}\biggl(a(x,\vW) \frac{\dif}{\dif x} u(x,\vW) \biggr) = 200( 2x - 1)\\
- 		 0 \leq x \leq 1, \qquad
- 		u(0) = u(1) = 0 \\
- 		a(x,\vW) = 1 + 0.6 \sum_{k=1}^d \frac{W_k T_{k}(2x-1)}{k^2} \\
- 		 \vW \sim \cu[-1,1]^d \\
- 		 \Ex[u(0.25,\vW)] = \only<1>{\, ?}\only<2>{ -1.7462}
- 	\end{gather*}}
-  \only<3->{
-\begin{itemize}
-\item Even with FEniCS/Dolfin's extensive documentation we had difficulties
-\begin{itemize}
-    \item Did not understand the latest version, so reverted to an older version
-    \item Had to learn how to change the random instance of the coefficient without tiggering the JIT compiler
-    \item Do not yet know how to express the random coefficient in terms of the covariance kernel
-\end{itemize}
-\end{itemize}
-  }
- \end{minipage}%
-\hfill
- \begin{minipage}{0.48\textwidth}
- 	\only<1-2>{\centering
- 	\includegraphics<1>[width=\textwidth]{ProgramsImages/axwtrue0.6.eps}
- 	\includegraphics<2->[width=\textwidth]{ProgramsImages/timing-4.eps}}%
-  \only<3->{
-\begin{itemize}
-\item But
-\begin{itemize}
-    \item It  works
-    \item If we can overcome the challenges, will post a blog and demo at \cite{QMCBlog}
-\end{itemize}
-\end{itemize}
-
-  }
-\end{minipage}
-
-\end{frame}
-
-\section{Support}
-\begin{frame}{Encouraging Users of Your Library Requires}
-
-	\vspace{-4ex}
-	\begin{itemize}
-		\item A repository where
-		\begin{itemize}
-		\item  Your library can be easily downloaded
-		\item There is documentation
-		\item Issues or queries can be posted
-		\item There are updates
-		\item Tests are run regularly to mitigate against bugs
-		\end{itemize}
-		\item (Elementary) demos on how to use the software and highlighting  its advantages
-		\item Demos connecting  your library with others
-		\item Recommending your students and postdocs to use your library
-		\item Welcoming instructions on how to contribute an algorithm or a demo
-	\end{itemize}
-
-\end{frame}
-
-\begin{frame}{Building a Developer Team for Your Library}
-
-	\vspace{-4ex}
-\begin{itemize}
-
-\item Many senior researchers don't code, and rely on transient team members (students, postdocs)
-\begin{itemize}
-	\item Encourage a sense of ownership that lasts beyond the time your team members are part of your lab
-	\item Encourage those outside your own group to contribute
-\end{itemize}
-
-\item Those of us who understand theory and appreciate software need to encourage and steer popular software to ensure that it is implemented well.
-
-\item Recognize good software as valuable scholarly output
-
-\item Recognize research software engineer as a valued vocation; see Research Software Engineers International 	\url{https://researchsoftware.org}, \emph{Research Software Engineers are people who combine professional software expertise with an understanding of research.
-They go by various job titles but the term Research Software Engineer (RSE) is fast gaining international recognition.}
-  \end{itemize}
-\end{frame}
-
-\section{Next Steps}
-
-\begin{frame}{Next Steps}
-
-	\vspace{-2ex}
-	\begin{minipage}{0.48\textwidth}
-		\begin{itemize}
-			\item Which exemplary theoretical developments are not yet implemented in readily available software?
-			\item What potential MCQMC user communities are ill-served by existing  software libraries?
-			\item Should those of us interested in software gather regularly for show-and-tell to foster more collaboration?
-		\end{itemize}
-	\end{minipage}
-	\begin{minipage}{0.48\textwidth}
-		Join us for a lunch discussion at Bella Casa tomorrow, July 19, at 12:45 PM
-
-		\vspace{2ex}
-
-		\includegraphics[width=\textwidth]{ProgramsImages/PizeriaBellaCasa.png}
-\end{minipage}
-
-\end{frame}
-
-\begin{frame}[allowframebreaks]{References}
-	\printbibliography
-\end{frame}
 

### 2023-03-08 MCQMC2022ChallengesQMCSoftwareArticle.tex
2023 Mac 14 Updates

@@
 This reduces the computations from what would be $\calO(n^3)$ for arbitrary covariance kernels to only $\calO(n \log n)$ for these well-chosen kernels.
+Consequently, computing the stopping criterion is roughly the same order of magnitude as computing the solution.
 
@@
 \section{Accessible} \label{CDHJS_sec:accessible}
-One attribute of great QMC software that often receives insufficient attention is its accessibility.
-Developers may write code for their own purposes and move on.
+One attribute of great QMC software that often receives insufficient attention is its accessibility.  Developers may write code for their own purposes and move on.
+
 Great software should be stored in a repository, where the latest version can be downloaded.
@@
 There should be a place for users to report bugs and make feature requests.
+
+Accessible QMC software has greater impact.  Accessibility also depends on the next characteristic of great QMC software.
 
@@
 \bibliography{FJH23,FJHown23}
-
-\end{document}
-
-\section{Introduction} \label{CDHJS_sec:intro}
-Quasi-Monte Carlo (QMC) methods promise great efficiency gains over independent and identically distributed (IID) Monte Carlo (MC) methods.
-In some cases QMC  achieves one hundredth of the error of IID MC in the same amount of time.
-Often, these efficiency gains are obtained simply by  replacing IID sampling by low discrepancy (LD) sampling, which is the heart of QMC.
-
-As a practitioner, you might wish to test whether QMC would speed your computation.
-You would like easy access to the best QMC algorithms available.
-As a theoretician or an algorithm developer, you would want to demonstrate your best ideas on  various  use cases to show their practical value.
-
-This tutorial points to some of the best QMC software available.
-Moreover, we describe QMCPy \cite{QMCPy2020a}, which is designed to be a community owned Python library that combines the best QMC algorithms from various authors under a common user interface.
-
-The model problem for QMC is approximating an integral,
-\begin{equation} \label{eq:integral}
-	\mu := \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst,
-\end{equation}
-where $g$ is the integrand, and $\lambda$ is a non-negative weight.
-We use $\mu$ to denote the value of this integral because we interpret it as the population mean of a random variable after a suitable variable transformation:
-\begin{equation} \label{eq:fintegral}
-	\mu = \bbE[f(\bsX)] = \int_\calX f(\bsx) \, \varrho(\bsx) \, \D \bsx =  \int_\calX f(\bsx) \,  \D F(\bsx) ,
-\end{equation}
-where $\varrho$ is a probability density with corresponding probability distribution $F$.
-In practice, $\calX$ often corresponds to the unit cube, $[0,1]^d$, and $F$ corresponds to the uniform distribution.
-
-QMC approximates this population mean by a sample mean,
-\begin{equation} \label{eq:samplemean}
-	\hmu := \frac 1n \sum_{i=0}^{n-1} f(\bsX_i), \qquad \bsX_0, \bsX_1, \ldots \sim F.
-\end{equation}
-The choice of this sequence and the choice of $n$ to satisfy  the prescribed error tolerance are important decisions, which  QMC software helps the user make.
-
-Here, the notation ``$\sim$'' means that the sequence mimics the specified, target distribution, but not necessarily in a probabilistic way.
-We  use this notation in two forms:  $\IIDsim$ and $\LDsim$.
-
-IID sequences must be random.
-The position of any point is not influenced by any other, so clusters and gaps occur.
-A subsequence of IID points chosen randomly is also IID.
-When we say that $\bsX_0, \bsX_1, \ldots \IIDsim F$, we mean that for any positive integer $n$, the  multivariate probability distribution of $\bsX_0, \ldots, \bsX_{n-1}$ is the product of the marginals, specifically,
-\begin{equation*}
-	F_{n}(\bsx_0, \ldots, \bsx_{n-1}) = F(\bsx_0) \cdots  F(\bsx_{n-1}).
-\end{equation*}
-When IID points are used to approximate $\mu$ by the sample mean, the error is $\calO(n^{-1/2})$.
-Figure \ref{fig:comparePts} displays IID uniform points, $\bsX^{\IID}_0, \bsX^{\IID}_1, \ldots \IIDsim \calU[0,1]^2$, where the target distribution is $F(\bsx) = x_1 x_2$.
-
-\begin{figure}
-	\includegraphics[height=5cm]{QMCSoftwareArticle/figs/dd_iid_uniform_pts.png}
-	\qquad
-	\includegraphics[height=5cm]{QMCSoftwareArticle/figs/dd_sobol_pts.png}
-	\caption{IID points contrasted with LD points.  The LD points cover the square more evenly.}
-	\label{fig:comparePts}
-\end{figure}
-
-LD sequences may be deterministic or random, but each point is carefully coordinated with the other so that they fill the domain well.
-When we say that $\bsX_0, \bsX_1, \ldots \LDsim F$, we mean that for any positive integer $n$,  the \emph{empirical distribution} of $\bsX_0, \ldots, \bsX_{n-1}$, denoted $F_{\{\bsX_i\}_{i=0}^{n-1}}$,  approximates the target distribution, $F$, well (relative to $n$).
-(The empirical distribution of a set assigns equal probability to each point.)  The measure of the difference between the empirical distribution of a set of points and the target distribution is called a \emph{discrepancy} and is denoted $D(\{\bsX_i\}_{i=0}^{n-1}, F)$.
-This is the origin of the term ``low discrepancy'' points or sequences.
-LD points by definition have a smaller discrepancy than IID points.
-Figure \ref{fig:comparePts} contrasts IID uniform points with LD points, $\bsX^{\LD}_0, \bsX^{\LD}_1 \ldots \LDsim \calU[0,1]^2$, in this case linearly scrambled Sobol' points.
-For nearly all LD sequences, the target distribution is $\calU[0,1]^d$.
-
-The error in using the sample mean to approximate the integral can be bounded according to the Koksma-Hlawka inequality and its extensions as the product of the discrepancy of the sampling sequence and the variation of the integrand, denoted $V(\cdot)$:
-\begin{equation}
-	\QMCPYabs{\mu - \hmu} = \QMCPYabs{\int_{\calX} f(\bsx) \, \D (F - F_{\{\bsX_i\}_{i=0}^{n-1}}) (\bsx)} \le D(\{\bsX_i\}_{i=0}^{n-1}, F) V(f),
-\end{equation}
-The variation is a (semi-) norm of the integrand in a suitable Banach space.
-The discrepancy corresponds to the norm of the error functional for that Banach space.
-For typical Banach spaces, the discrepancy of LD points is $\calO(n^{-1+\epsilon})$, which is a higher convergence order than for IID points.
-For details, the reader is referred to the references.
-For our purposes, we expect the reader to see in Figure \ref{fig:comparePts} that the LD points cover the integration domain more evenly than IID points.
-In the examples below the reader will see the demonstrably smaller cubature errors arising from using LD points.
-
-In the sections that follow we first overview available QMC software.
-We next describe an architecture for good QMC software, i.e., what are the key components and how should they interact.
-We then describe how we have implemented this architecture in QMCPy.
-Finally, we summarize further directions that we hope QMCPy and related software projects will take.
-Those interested in following the development of QMCPy or even contributing to that development are urged to visit the GitHub repository at \href{https://qmcsoftware.github.io/QMCSoftware/}{\nolinkurl{https://qmcsoftware.github.io/QMCSoftware/}}.
-
-\section{Available Software for QMC} \label{CDHJS_sec:available}
-QMC software spans  LD sequence generators, algorithms, and applications.  We review the better known software collections, recognizing that some software overlaps multiple categories.
-Software focusing on generating high quality LD sequences  or their generators includes
-\begin{description}[format=\textup,format=\textbf]
-% https://tex.stackexchange.com/questions/74279/how-to-add-bullets-to-description-lists
-	\item[BRODA] Sobol' sequences in C, MATLAB, and Excel \cite{BRODA20a},
-	\item[Burkhardt] various QMC software in C++, Fortran, MATLAB, \& Python \cite{Bur20a},
-	\item[LatNet Builder] Generating vectors/matrices for lattices and digital nets \cite{LEcEtal22a,LatNet},
-	\item[MATLAB] Sobol' and Halton sequences, commercial \cite{MAT9.9},
-	\item[MPS] Magic Point Shop, lattices and Sobol' sequences \cite{Nuy17a},
-	\item[Owen] Randomized Halton sequences in R \cite{Owe20a},
-	\item[PyTorch] Scrambled Sobol' sequences \cite{PyTorch},
-	\item[QMC.jl] LD Sequences in Julia \cite{Rob20a}, and
-	\item [qrng]  Sobol', Halton, and Korobov sequences in R \cite{QRNG2020}.
-\end{description}
-Software focusing on QMC algorithms and applications includes
-\begin{description}[format=\textup,format=\textbf]
-	\item[GAIL] Automatic (Q)MC stopping criteria in MATLAB \cite{ChoEtal20a},
-	\item[ML(Q)MC] Multi-Level (Quasi-)Monte Carlo routines in C++, MATLAB, Python, and R \cite{GilesSoft},
-	\item[OpenTURNS] Open source initiative for the Treatment of Uncertainties, Risks 'N Statistics in Python \cite{OpenTURNS},
-	\item[QMC4PDE] QMC for elliptic PDEs with random diffusion coefficients \cite{KuoNuy16a},
-	\item[SSJ] Stochastic Simulation in Java \cite{SSJ}, and
-	\item[UQLab] Framework for Uncertainty Quantification in MATLAB \cite{UQLab2014}.
-\end{description}
-
-The sections that follow describe QMCPy \cite{QMCPy2020a}, which is our attempt to combine the best of the above software under a common user interface written in Python 3.
-The choice of language was determined by the desire to make QMC software accessible to a broad audience, especially the tech industry.
-
-\section{Components of QMC Software}
-QMC cubature can be summarized as follows.
-We want to approximate the integral oe expectation, $\mu$, well by the sample mean, $\hmu$, where \eqref{eq:integral}, \eqref{eq:fintegral}, and \eqref{eq:samplemean} combine to give
-\begin{multline} \label{eq:cubSummary}
-	\mu : = \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  = \bbE[f(\bsX)] = \int_\calX f(\bsx) \, \varrho(\bsx) \, \D \bsx \approx \frac 1n \sum_{i=0}^{n-1} f(\bsX_i) =: \hmu, \\
-	 \bsX_0, \bsX_1, \ldots \sim F.
-\end{multline}
-This requires four components, which we implement as QMCPy classes.
-
-\begin{description}[format=\textup,format=\textbf]
-
-	\item[Discrete Distribution] that produces $\bsX_0, \bsX_1, \dots$ mimicking the distribution $F$, which typically is $\calU[0,1]^d$;
-
-	\item[True Measure] $\bst \mapsto \lambda (\bst) \D \bst$ that defines the original integral, e.g., Gaussian or Lebesgue;
-
-	\item[Integrand] $g$ that  defines the original integral, plus the transformed version, $f$, to fit the discrete distribution; and
-
-	\item[Stopping Criterion] that determines how large $n$ should be to ensure that $\QMCPYabs{\mu - \hmu_n} \le \varepsilon$.
-\end{description}
-
-The software libraries referenced in Section \ref{CDHJS_sec:available} provide one or more of these components.
-QMCPy combines multiple examples of them all under and object oriented framework.
-Each example is implemented as a concrete class that realizes the properties and methods required by the abstract class for that component.
-The following sections detail descriptions and use cases for each component and select implementations.
-
-Thorough documentation of all QMCPy classes can be found in \cite{QMCPyDocs}.
-Demonstrations of how QMCPy work are given in Google Colab notebooks \cite{QMCPyTutColab2020,QMCPyTutColab2020_paper} \AGSComment{Update \cite{QMCPyTutColab2020_paper}'s link}.
-The project may be installed from PyPI into your Python 3 environment via the command \texttt{pip install qmcpy}.
-In the codes that follows, we assume QMCPy has been imported alongside Numpy \cite{numpy}
-%and Scipy \cite{scipy}
-via
-\lstinputlisting[style=Python]{QMCSoftwareArticle/python/import.txt}
-
-\section{Discrete Distributions}
-
-LD sequences typically mimic $\calU[0,1]^d$, which we assume here.  Good sequences for other distributions are obtained by transformations in the next section.
-
-QMCPy implements \emph{extensible} LD sequences, i.e., those that allow practitioners to obtain and use $\bsX_n, \bsX_{n+1}, \ldots $ without discarding the existing $\bsX_0, \ldots, \bsX_{n-1}$.
-Halton sequences do not have preferred sample sizes $n$, but integration lattices and digital sequences do.
-These latter two also have an elegant group structure, which we summarize in Table \ref{tab:GroupProp}.
-For simplicity we restrict ourselves to the case where the first $n = 2^m$ points of these LD sequences form a group under the addition operator $\oplus$.
-
-\begin{table}
-	\centering
-	\caption{Properties of lattices and digital net sequences.  Note that they share group properties but also have distinctives.} \label{tab:GroupProp}
-\[
-	\renewcommand{\arraystretch}{1.3}
-\begin{array}{c@{\qquad}c}
-	\toprule
-	\multicolumn{2}{c}{\text{Define \ldots}} \\
-	\multicolumn{2}{c}{\bsZ_1, \bsZ_2, \bsZ_4, \ldots \in [0,1)^d \text{ chosen well} } \\
-	\multicolumn{2}{c}{
-	\bsZ_{i} := i_0  \bsZ_1 \oplus i_1 \bsZ_{2} \oplus + i_2  \bsZ_{4} \oplus  i_3  \bsZ_{8} + \cdots
-	\quad
-	\text{for }i = i_0 +i_1 2 + i_2 4 + i_3 8 + \cdots, \; i_\ell \in \{0,1\}} \\
-    \multicolumn{2}{c}{\bsX_i := \bsZ_i \oplus \bsDelta, \qquad \text{where }\bsDelta \IIDsim [0,1)^d} \\  \hline
-	\text{Rank-1 Integration Lattices} & \text{Digital Nets} \\
-		\bst \oplus \bsx : = (\bst + \bsx) \bmod \bsone & \bst \oplus \bsx := \text{binary digitwise addition} \\
-		 \text{require } \begin{array}{l} \bsZ_1 = (1/2, \ldots, 1/2) \\
-		 	\bsZ_{2^{m}} \oplus \bsZ_{2^{m}} = \bsZ_{2^{m-1}} \quad \forall m \in \bbN \end{array}
-		\\
-\toprule
-\multicolumn{2}{c}{\text{Then it follows that \ldots}} \\
-	\multicolumn{2}{c}{\left . \begin{array}{r}
-			\calP_m := \{\bsZ_0, \ldots, \bsZ_{2^m-1}\}, \quad
-			\bsZ_i \oplus \bsZ_j \in \calP_m \\
-			\calP_{\bsDelta,m} := \{\bsX_0, \ldots, \bsX_{2^m-1}\}, \quad
-			\bsX_i \oplus \bsX_j \ominus \bsX_k \in \calP_{\bsDelta,m}
-	\end{array} \right \}\quad \begin{array}{l}\forall  i,j,k \in \{0, \ldots, 2^{m} -1\} \\ \forall m \in \bbN_0\end{array}} \\
-\bottomrule
-\end{array}
-\]
-\end{table}
-
-We illustrate lattice and Sobol' sequences using QMCPy.
-First, we create an instance of a $d=2$ dimensional \texttt{Lattice} object of the  \texttt{DiscreteDistribution} abstract class.
-Then we generate the first eight (non-randomized) points in this lattice.
-\lstinputlisting[style=Python]{QMCSoftwareArticle/python/dd_lattice.txt}
-The first three generators for this lattice are $\bsZ_1 = (0.5, 0.5)$, $\bsZ_2 = (0.25, 0.75)$, and $\bsZ_4 = (0.125, 0.375)$.
-One can check that $(\bsZ_2 + \bsZ_4) \bmod \bsone = (0.375, 0.125) = \bsZ_6$, as Table \ref{tab:GroupProp} specifies.
-
-The randomized shift has been turned off above to illuminate the group structure.
-In practice, we normally include the randomization to ensure that there are no points on the boundary of $[0,1]^d$.
-Then, when points are transformed to mimic distributions such as the Gaussian, no LD points will be transformed to infinity.
-Turning off the randomization generates a warning.
-
-Now, we generate Sobol' points using a similar process as we did for lattice points.  Sobol' sequences are one of the most popular example of digital sequences.
-\lstinputlisting[style=Python]{QMCSoftwareArticle/python/dd_sobol.txt}
-Here, $\bsZ_4$ differs from that for lattices, but more importantly, addition for digital sequences differs from that for lattices.
-Using digitwise addition for digital sequences, we can confirm that according to Table \ref{tab:GroupProp},
-\begin{multline*}
-\bsZ_2 \oplus_{\dig} \bsZ_4 = (0.25,0.75)  \oplus_{\dig} (0.125,0.625) \\
-=  ({}_20.010,{}_20.110)  \oplus_{\dig} ({}_20.001,{}_20.101) = ({}_20.011,{}_20.011) \\
-= (0.375,0.375) = \bsZ_6.
-\end{multline*}
-
-By contrast, if we construct a digital sequence using the generators for the lattice above with $\bsZ_2 = (0.25, 0.75)$, and $\bsZ_4 = (0.125, 0.875)$, we would obtain
-\begin{multline*}
-\bsZ_6 = \bsZ_2 \oplus_{\dig} \bsZ_4   = ({}_20.010,{}_20.110)  \oplus_{\dig} ({}_20.001,{}_20.111)  \\
-= ({}_20.011,{}_20.001) = (0.375, 0.125),
-\end{multline*}
-which differs from the $\bsZ_6=(0.375, 0.625)$ constructed for lattices.
-To emphasize, lattices and digital sequences are different, even if they share the same generators, $\bsZ_1, \bsZ_2, \bsZ_4, \ldots$.
-
-The examples of \texttt{qp.Lattice} and \texttt{qp.Sobol} illustrate how QMCPy LD generators share a common user interface.
-The dimension is specified when the instance is constructed, and the number of points is specified when the \texttt{gen\_samples} method is called.
-Following Python practice, parameters can be input without specifying their names if they are input in the prescribed order.
-QMCPy also includes Halton sequences and IID sequences, again deferring details to the QMCPy documentation \cite{QMCPyDocs}.
-
-A crucial difference between IID generators and LD generators is reflected in the behavior of generating $n$ points.
-For an IID generator, asking for $n$ points repeatedly gives you different points each time because they are meant to be random and independent.
-\lstinputlisting[style=Python]{QMCSoftwareArticle/python/dd_iidone.txt}
-Your output may look different depending on the seed used to generate these random numbers.
-
-On the other hand for an LD generator, asking for $n$ points repeatedly gives you \emph{the same} points each time because they are meant to be the first $n$ points of a given LD sequence.
-\lstinputlisting[style=Python]{QMCSoftwareArticle/python/dd_latticeone.txt}
-Here we allow the randomization so that the first point in the sequence is not the origin.  To obtain the \emph{next} $n$ points one may specify the start and ending indices of the sequence.
-\lstinputlisting[style=Python]{QMCSoftwareArticle/python/dd_latticenextone.txt}
-
-Figure \ref{fig:increase_n} shows how increasing the number of lattice and Sobol' LD points through powers of two fills in the gaps in an even way.
-
-\begin{figure}
-	\includegraphics[width=1\textwidth]{QMCSoftwareArticle/figs/dd_lattice_successive.png}
-	\qquad
-	\includegraphics[width=1\textwidth]{QMCSoftwareArticle/figs/dd_sobol_successive.png}
-	\caption{Randomized lattice and Sobol' points mimicking a $\calU[0,1]^2$ measure for $n = 64, 128,$ and 256. Note how increasing the number of points evenly fills in the gaps between points.}
-	\label{fig:increase_n}
-\end{figure}
-
-\section{True Measures}
-
-The LD sequences implemented as \text{DiscreteDistribution} objects typically mimic the $\calU[0,1]^d$ distribution.
-However, we may need sequences to mimic other distributions.
-This is implemented via variable transformations, $\bsPsi$.
-In general, if $\bsX \sim \calU[0,1]^d$, then
-\begin{subequations} \label{eq:exampleVarTrans}
-\begin{gather}
-\bsT = \bsPsi(\bsX) := \bsa  + (\bsb - \bsa) \odot \bsX \sim  \calU[\bsa,\bsb], \\
-\label{eq:exampleVarTransGauss}
-\bsT = \bsPsi(\bsX) := \bsa + \mA \bsPhi^{-1}(\bsX)  \sim \calN(\bsa, \mSigma), \\
-\nonumber  \text{where }  \bsPhi^{-1}(\bsX) : = \begin{pmatrix} \Phi^{-1}(X_1) \\ \vdots \\ \Phi^{-1}(X_d)\end{pmatrix}, \qquad \mSigma = \mA \mA^T,
-\end{gather}
-\end{subequations}
-and $\odot$ denotes term-by-term (Hadamard) multiplication.
-Here, $\bsa$ and $\bsb$ are assumed to be finite, and $\Phi$ is the standard Gaussian distribution function.
-Again we use ``$\sim$'' to denote mimicry, not necessarily in a probabilistic sense.
-
-Figure \ref{fig:tm_ug} displays LD sequences transformed as described above to mimic a uniform and a Gaussian distribution.  The code to generate these points takes the following form of
-\lstinputlisting[style=Python]{QMCSoftwareArticle/python/tm_uniform.txt}
-for uniform points based on a Halton sequence, and
-\lstinputlisting[style=Python]{QMCSoftwareArticle/python/tm_gaussian.txt}
-for Gaussian points based on a lattice sequence.
-
-\begin{figure}
-	\includegraphics[width=.45\textwidth]{QMCSoftwareArticle/figs/tm_uniform.png}
-	%	\quad
-	\includegraphics[width=.45\textwidth]{QMCSoftwareArticle/figs/tm_gaussian.png}
-	\caption{Sobol' samples transformed to mimic a uniform $\calU\left(\begin{bmatrix} -2 \\ 0 \end{bmatrix}, \begin{bmatrix} 2 \\ 4 \end{bmatrix} \right)$ (left) and  Gaussian $\calN\left(\begin{bmatrix} 3 \\ 2 \end{bmatrix}, \begin{bmatrix} 9 & 5 \\ 5 & 4 \end{bmatrix} \right)$ (right).}
-	\label{fig:tm_ug}
-\end{figure}
-
-The Brownian motion distribution arises often in financial risk applications.
-Here the $d$ components of the variable $\bsT$ correspond to the Brownian motion at times $\tau/d, 2\tau/d, \ldots, \tau$, where $\tau$ is the time horizon.
-The distribution is a special case of the Gaussian with covariance
-\begin{equation} \label{eq:BMcov}
-	\mSigma = (\tau/d) \bigl (\min(j,k) \bigr)_{j,k=1}^d
-\end{equation}
-and mean $\bsa$, which  correspond to a drift coefficient times $(\tau/d)(1, 2, \ldots, d)^T$. The code for generating a Brownian motion is
-\lstinputlisting[style=Python]{QMCSoftwareArticle/python/tm_brownian_motion.txt}
-Figure \ref{fig:tm_bm} displays a Brownian motion based on Sobol' sequence with and without a drift.
-
-\begin{figure}
-	\includegraphics[width=1\textwidth]{QMCSoftwareArticle/figs/tm_bm.png}
-	\caption{Sobol' samples transformed to mimic a 32-dimensional Brownian Motion without drift (left) and with drift coefficient 2 (right).}
-	\label{fig:tm_bm}
-\end{figure}
-
-\section{Integrands}
-
-Let's return now to the integration problem in \eqref{eq:integral}, which we must rewrite as \eqref{eq:fintegral}.
-We choose a transformation of variables defined as $\bst = \bsPsi(\bsx)$ where $\bsPsi:\calX \to \calT$.
-This leads to
-\begin{align}
-	\nonumber
- \mu &= \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  = \int_\calX g\bigl(\bsPsi(\bsx)\bigr) \, \lambda\bigl(\bsPsi(\bsx)\bigr) \,\QMCPYabs{\frac{\partial \bsPsi}{\partial \bsx}} \, \D \bsx =  \int_\calX f(\bsx) \, \varrho(\bsx) \, \D \bsx  \\
- \label{eq:transVar}
-  & \qquad \qquad \text{where } f(\bsx)  = g\bigl(\bsPsi(\bsx)\bigr)  \, \frac{\lambda(\bsPsi(\bsx))}{\varrho(\bsx)} \,\QMCPYabs{\frac{\partial \bsPsi}{\partial \bsx}},
-\end{align}
-and $\QMCPYabs{\partial \bsPsi/\partial \bsx}$ represents the Jacobian of the variable transformation.
-The abstract class \texttt{Integrand} provides $f$ based on the user's input of $g$ and the \texttt{TrueMeasure} instance, which defines $\lambda$ and the transformation $\bsPsi$.
-Different choices of $\bsPsi$ lead to different $f$, which may give different rates of convergence of the cubature, $\hmu$ to $\mu$.
-
-We illustrate the use of the \texttt{Integrand} class via an example of Keister \cite{Kei96}:
-\begin{equation} \label{eq:KeisterIntegral}
-	\mu
-	= \int_{\bbR^d} \cos(\lVert \bst \rVert) \exp(-\bst^T \bst) \, \D \bst \\
-	= \int_{\bbR^d} \underbrace{\pi^{d/2} \cos(\lVert \bst \rVert)}_{g(\bst)}\, \underbrace{\pi^{-d/2} \exp(-\bst^T \bst) }_{\lambda(\bst)} \, \D \bst.
-\end{equation}
-We will use QMC methods to approximate this integral, which means that it will be transformed according to \eqref{eq:transVar} for the case where $\calX = [0,1]^d$ and $\varrho(\bsx) = 1$.
-Since $\lambda$ is the density for $\calN(0,\mI/2)$, it is natural to choose $\bsPsi$ according to \eqref{eq:exampleVarTransGauss} with $\mA = \sqrt{1/2} \, \mI$, in which case $\lambda(\bsPsi(\bsx)) \QMCPYabs{\partial \bsPsi/\partial \bsx} / \varrho(\bsx) = 1$, and so
-\[
-\mu = \int_{[0,1]^d} \underbrace{\pi^{d/2} \cos(\lVert \bsPsi(\bsx) \rVert)}_{f(\bsx)} \, \D \bsx, \qquad
-\bsPsi(\bsx) := \sqrt{1/2} \,\bsPhi^{-1}(\bsx).
-\]
-
-The code below sets up an \texttt{Integrand} instance using QMCPy's \texttt{CustomFun} wrapper to tie a user-defined function $g$ into the QMCPy framework.
-Then we evaluate the sample mean of $n=1000$ $f$ values obtained by sampling at transformed Halton points.
-\lstinputlisting[style=Python]{QMCSoftwareArticle/python/i_keister.txt}
-We have no indication yet of how accurate our approximation is.
-This aspect of QMCPy is described in the next section.
-Figure \ref{fig:ikc} visualizes sampling on the original integrand, $g$, and sampling on the transformed integrand, $f$.
-
-\begin{figure}
-	\includegraphics[height=6cm]{QMCSoftwareArticle/figs/i_keister_contours.png}
-	\caption{Right: Sampling the transformed integrand $f$ at Halton points $\bsx \sim \calU[0,1]^2$.
-Left: Sampling the original integrand $g$ at $\bsPsi(\bsx)$ where $\bsPsi$ is defined in \eqref{eq:exampleVarTransGauss}.
-} \label{fig:ikc}
-\end{figure}
-
-Another way to approximate the Keister integral in \eqref{eq:KeisterIntegral} is to write it as an integral with respect to the Lebesgue measure:
-\begin{align*}
-	\mu
-	& = \int_{\bbR^d} \underbrace{\cos(\lVert \bst \rVert) \exp(-\bst^T \bst) }_{g(\bst)} \, \underbrace{1}_{\lambda(\bst)} \,\D \bst \\
-	& = \int_{[0,1]^d} \underbrace{\cos(\lVert \bsPsi(\bsx) \rVert) \exp(-\bsPsi^T\!\!(\bsx) \bsPsi(\bsx)) \QMCPYabs{\frac{\partial \bsPsi}{\partial \bsx}} }_{f(\bsx)} \, \D \bsx,
-\end{align*}
-where $\bsPsi$ is any transformation from $\bbR^d$ to $[0,1]^d$.  QMCPy can perform the cubature this way as well.
-\lstinputlisting[style=Python]{QMCSoftwareArticle/python/i_keisterLebesgue.txt}
-The $\bsPsi$ chosen when transforming uniform points on the unit cube to fill the $\bbR^d$ is the one given by \eqref{eq:exampleVarTransGauss} with $\mA = \mI$.
-
-In the examples above, one needed to input the correct $g$ into \texttt{CustomFun} along with the correct \texttt{TrueMeasure} $\lambda$ to define the integration problem.
-The \texttt{Keister} integrand included in the QMCPy library takes a more flexible approach of defining the integration problem $\mu$ in \eqref{eq:KeisterIntegral}.
-Selecting a different \texttt{sampler} $\bsPsi$ performs  \emph{importance sampling} which leaves $\mu$ unchanged.
-\lstinputlisting[style=Python]{QMCSoftwareArticle/python/i_keisterBuiltIn.txt}
-
-\section{Stopping Criteria} \label{CDHJS_sec:stopping_crit}
-
-The \texttt{StoppingCriterion} object determines the number of samples $n$ that are required for the sample mean approximation $\hmu$ to be within error tolerance $\varepsilon$ of the true mean $\mu$.
-Several QMC stopping criteria have been implemented in QMCPy, including replications, stopping criteria that track the decay of the Fourier complex exponential or Walsh coefficients of the integrand \cite{HicJim16a,HicEtal17a,JimHic16a}, and stopping criteria based on Bayesian credible intervals \cite{RatHic19a,JagHic22a}.
-
-Let us return to the Keister example from the previous section.
-After setting up  a default \texttt{Keister} instance via a Sobol' \texttt{DiscreteDistribution}, we choose a \texttt{StoppingCriterion} object that matches the \texttt{DiscreteDistribution}, inputting our desired tolerance.
-Calling the  \texttt{integrate} method returns the approximate integral plus some useful information about the computation.
-\lstinputlisting[style=Python]{QMCSoftwareArticle/python/sc_keister_qmc.txt}
-The second output of the stopping criterion provides helpful diagnostic information.
-This computation required $n=2^{13}$ Sobol' points and $0.028$ seconds to complete.
-The error bound is $0.000517$, which falls below the absolute tolerance.
-
-QMC, which uses LD sequences, is touted as providing substantially greater computational efficiency compared to IID MC.
-Figure \ref{fig:sc_comp} compares the time and sample sizes needed to compute the $5$-dimensional Keister integral \eqref{eq:KeisterIntegral} using IID sequences and LD lattice sequences.
-Consistent with what is stated in Section \ref{CDHJS_sec:intro}, the error of IID MC is $\calO(n^{-1/2})$, which means that the time and sample size to obtain an absolute error tolerance of $\varepsilon$ is $\calO(\varepsilon^{-2})$.
-By contrast, the  error of QMC using LD sequences is $\calO(n^{-1+\epsilon})$, which implies $\calO(\varepsilon^{-1-\epsilon})$ times and sample sizes.
-We see that QMC methods often require orders of magnitude fewer samples that MC methods to achieve the same error tolerance.
-
-\begin{figure}
-	\includegraphics[height=6cm]{QMCSoftwareArticle/figs/sc_comp.png}
-	\caption{Comparison of run times and sample sizes for computing the $5$-dimensional Keister integral \eqref{eq:KeisterIntegral} using IID and LD lattice sequences for a variety of absolute error tolerances.
-The respective stopping criteria are  \texttt{CubMCG} \AGSComment{CITE} and  \texttt{CubQMCLatticeG} \AGSComment{CITE}.
-The LD sequences provide the desired answer much more efficiently.}
-	\label{fig:sc_comp}
-\end{figure}
-
-For another illustration of QMC cubature, we turn to pricing an Asian arithmetic mean call option.
-The payoff of this option is the positive difference between the strike price averaged over the time horizon and the strike price, $K$:
-$$
-\text{payoff}(\bsS) = \max\left(\frac{1}{2d}\sum_{j=1}^d (S_{j-1}+S_j)-K,0\right), \quad \bsS = (S_0, \ldots, S_d).
-$$
-where $S_j$ denotes the stock price at time $\tau j/d$.  A basic model for stock prices is a geometric Brownian motion,
-\[
-S_j(\bsT) = S_0 \exp((r - \sigma^2) \tau j/d + \sigma T_j),   \;  j = 1, \ldots, d, \; \bsT = (T_1, \ldots, T_d)\sim \calN(0,\mSigma),
-\]
-where $\mSigma$ is defined in \eqref{eq:BMcov}, $r$ is the interest rate, $\sigma$ is the volatility, and $S_0$ is the initial price.
-The fair price of the option is then the expected value of the discounted payoff, namely,
-\begin{equation*}
-	\text{price} = \mu = \bbE[g(\bsT)], \quad \text{where } g(\bst) = \text{payoff}\bigl(\bsS(\bst) \bigr) \exp(-r \tau).
-\end{equation*}
-
-The following code utilizes QMCPy's Asian option \texttt{Integrand} object to approximate the Asian call option for a particular choice of the parameters.
-\lstinputlisting[style=Python]{QMCSoftwareArticle/python/sc_aco.txt}
-Because this \texttt{Integrand} object has the built-in Brownian motion \texttt{TrueMeasure}, one only neeed provide the LD sampler.
-
-Out of the money option price calculations can be sped up by adding a drift to the Brownian motion that produces more in the money paths.
-This is a form of importance sampling.
-Using a Brownian motion without drift we get
-\lstinputlisting[style=Python]{QMCSoftwareArticle/python/wo_imp_samp_aco.txt}
-Adding the drift gives us the answer faster:
-\lstinputlisting[style=Python]{QMCSoftwareArticle/python/w_imp_samp_aco.txt}
-The choice of a good drift is an art.
-
-The improvement in time is less than improvement in $n$ because the integrand $f$ is more expensive to compute when the drift is added.
-In the case of no drift, the $\lambda$ in \eqref{eq:transVar} corresponds to the density for the discrete Brownian motion, and the variable transformation $\bsPsi$ is chosen so that $\lambda\bigl(\bsPsi(\cdot)\bigr) \QMCPYabs{\partial \bsPsi/\partial \bsx} = \varrho(\bsx) = 1$ and $f(\bsx) = g\left(\bsPsi(\bsx) \right)$.
-However, if we choose $\lambda_{\IMP}$ as our importance sampling density and $\bsPsi_{\IMP}$ as the variable transformation corresponding to $\lambda_\IMP\bigl(\bsPsi_\IMP(\cdot)\bigr) \QMCPYabs{\partial \bsPsi_\IMP/\partial \bsx} = \varrho(\bsx) = 1$, then
-\begin{align}
-	\nonumber
-	\mu &= \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  = \int_\calT g(\bst) \, \frac{\lambda(\bst)}{\lambda_\IMP(\bst) } \, \lambda_\IMP(\bst) \D \bst  \\
-	\nonumber
-	& =  \int_\calX g\bigl(\bsPsi_\IMP(\bsx)\bigr) \, \frac{\lambda\bigl(\bsPsi_\IMP(\bsx)\bigr)} {\lambda_\IMP\bigl(\bsPsi_\IMP(\bsx)\bigr) } \lambda_\IMP\bigl(\bsPsi_\IMP(\bsx)\bigr) \,\QMCPYabs{\frac{\partial \bsPsi_\IMP}{\partial \bsx}} \, \D \bsx \\
-	\nonumber
-	& =  \int_\calX f_\IMP(\bsx) \, \varrho(\bsx) \, \D \bsx  \\
-	\label{eq:transVarImp}
-	& \qquad \qquad \text{where } f_\IMP(\bsx)  = g\bigl(\bsPsi_\IMP(\bsx)\bigr)  \,  \frac{\lambda\bigl(\bsPsi_\IMP(\bsx)\bigr)} {\lambda_\IMP\bigl(\bsPsi_\IMP(\bsx)\bigr) }
-\end{align}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=1\textwidth]{QMCSoftwareArticle/figs/i_aco.png}
-	\caption{Left: Brownian Motion paths with drift $a=2$. Right: Corresponding Asian Call option paths with $S_0=30$, $K=45$, $\sigma=0.5$, and $r=0$.}
-	\label{fig:aco}
-\end{figure}
-
-Figure \ref{fig:aco} shows drifted Brownian motion paths and their corresponding option pricing paths.
-\AGSComment{Bayesian stopping criterion}
-
-\section{Under the Hood}
-
-In this section, we look at the inner workings of QMCPy and points out available features we hope will benefit the community of QMC researchers and practitioners.
-We also highlight important idiosyncrasies of QMC methods and how QMCPy addresses these challenges.
-
-\subsection{Link to LatNetBuilder}
-
-QMCPy provides users access to high quality lattice and digital net generators (defined in Table \ref{tab:GroupProp}).
-A low-discrepancy generator is comprised of two parts: the static generating vectors $\bsZ_1,\bsZ_2,\bsZ_4, \dots \in [0,1)^d$ and the callable generator function.
-QMCPy integrates nicely with latnetbuilder \AGSComment{CITE}, a package centered around search algorithms to find high quality generating vectors.
-QMCPy's lattice and digital net generators can accept a vectors from latnetbuilder and use them to generate custom ordinary lattice, polynomial lattice, polynomial net, or digital net sequence.
-We defer readers to latnetbuilder and QMCPy documentation \AGSComment{CITE BOTH} for details.
-
-\subsection{Customization Options for Low Discrepancy Generators}
-
-QMCPy's low discrepancy generator routines also support numerous customizations for users who need specific points sets, orders, or randomizations.
-For example, our in-house developed digital net generator allows randomizations such as digital shifts and linear matrix scrambles, optional graycode orderings, and an optionally non-zero starting dimension.
-Moreover, while most lattice and digital net generators use a fixed generating vector, QMCPy allows users to input their own vector for their specific need, as detailed in the previous subsection.
-
-\subsection{Preferred sample sizes for low-discrepancy generators and}
-
-Low-discrepancy sequences differ from IID node sets in that they fill the domain in a more uniform manner.
-As IID nodes are independent, they do not have a preferred sample size.
-That is, if $\bsX \IIDsim \calU[0,1]^d$ then each node is equally likely to fall  anywhere in the unit cube.
-On the other hand, low-discrepancy sequences provide guarantees that the integration domain is more evenly sampled.
-However, these guarantees only occur at specific sample sizes.
-For example, lattice and digital net sequences only fill the domain evenly when $n$ is a power of $2$.
-
-\subsection{\texttt{print}ing  QMCPy Instances}
-
-A helpful features of QMCPy is the ability to print out components instances to see more detailed information.
-This allows users to quickly see the properties that have been set and defaulted for any given component.
-This feature was already used to print the data object returned by the stopping criterion's \texttt{integrate} method in Section \ref{CDHJS_sec:stopping_crit}.
-The following code prints a drifted \texttt{BrownianMotion} instance and displays the connection to an underlying Gaussian measure.
-\lstinputlisting[style=Python]{QMCSoftwareArticle/python/print_bm.txt}
-
-\AGSComment{
-\begin{itemize}
-    \item we have better generating vector (under latnetbuilder sub-section)
-    \item transform from low discrepancy does not guarantee low discrepancy in new distribution
-    \item Sobol' samples will repeat after $2^{32}$ samples
-    \item multi-level stopping criterion
-    \item PCA and Cholesky decomposition for Gaussian (and Brownian Motion which is a subclass of Gaussian)
-\end{itemize}}
-
-\section{Further Work} \label{CDHJS_sec:further}
-
-\AGSComment{
-\begin{itemize}
-    \item Better multi-level stopping criterion (Pieterjan)
-    \item Multi-composed transforms $\bsPsi$ (multiple Jacobian factors)
-    \item Niederreiter sequences (Adrian and Onyekachi)
-    \item Sobol' indicies (Chris Hoyt)
-    \item transformations for discrete distributions that do not mimic $\calU[0,1]^d$
-    \item control variates
-    \item vectorized integrals
-\end{itemize}}
-
-\begin{acknowledgement}
-The authors would like to thank the organizers for a wonderful MCQMC 2020.
-We also thank the referees for their many helpful suggestions.  This work is supported in part by National Science Foundation grants DMS-1522687 and SigOpt.
-\end{acknowledgement}
-
-%\section*{References}
-%\nocite{*}
-\bibliographystyle{spmpsci.bst}
-\bibliography{FJH23,FJHown23,QMCSoftwareArticle}
-
-\AGSComment{
-
-\section{Extra Derivations}
-
-\subsection{Importance sampling with Gaussian}
-
-$$ \bsPsi(\bsx) = A\Phi^{-1}(\bsx) + \mu \quad \text{where} \quad \Sigma = AA^T$$
-$$ \left\lvert \frac{\D \bsPsi}{\D \bsx} \right\rvert = \det(J) \quad \text{for} \quad J_{ij} = \frac{\D \Psi_i}{\D x_j} = \frac{A_{ij}}{\phi(\Phi^{-1}(x_j))}$$
-$$ \det(J) = \det\left(A . \text{diag}\left(\frac{1}{\phi(\Phi^{-1}(\bsx))}\right)\right) = \frac{\det(A)}{\prod_{j=1}^d \phi(\Phi^{-1}(x_j))}$$}
-
-\FJHComment{Given $g$ and $\lambda$ to define your integral, suppose that $\lambda(\bst)$ is some multiple of some multivariate normal distribution:
-	\begin{equation*}
-		\lambda(\bst)  = \frac{c \exp\bigl(-(\bst -\bsb)^T \mLambda^{-1}(\bst -\bsb)/2\bigr)}{\sqrt{ (2\pi)^d \det(\mLambda)}}
-	\end{equation*}
-Suppose that we use the variable transformation $\bsPsi$ defined in \eqref{eq:exampleVarTransGauss}.  By what you have derived above:
-\begin{align*}
-	\QMCPYabs{\frac{\partial \bsPsi}{\partial \bsx}}
-	& = \frac{\det(\mA^T)}{\prod_{j=1}^d \phi(\Phi^{-1}(x_j))} = \sqrt{(2\pi)^d \det(\mSigma)} \exp\bigl( \QMCPYnormnorm[2]{\bsPhi^{-1}(\bsx)}^2/2\bigr) \\
-	\MoveEqLeft[2]{\lambda\bigl(\bsPsi(\bsx)\bigr) \QMCPYabs{\frac{\partial \bsPsi}{\partial \bsx}}} \\
-	& =  c \sqrt{\frac{\det(\mSigma)}{\det(\mLambda)}} \exp \bigl( [ \bsPhi^{-1}(\bsx)^T \bsPhi^{-1}(\bsx) \\
-	& \qquad \qquad -   (\mA\bsPhi^{-1}(\bsx) +\bsa -\bsb)^T\mLambda^{-1}(\mA\bsPhi^{-1}(\bsx) +\bsa -\bsb)        ] /2   \bigr) \\
-	& =  c \sqrt{\frac{\det(\mSigma)}{\det(\mLambda)}} \exp \bigl( [ \bsPhi^{-1}(\bsx)^T \bsPhi^{-1}(\bsx)
-	-   \bsPhi^{-1}(\bsx)^T\mA^T\mLambda^{-1}\mA \bsPhi^{-1}(\bsx)\\
-	& \qquad \qquad  - 2(\bsa -\bsb)^T \mLambda^{-1}\mA\bsPhi^{-1}(\bsx) - (\bsa -\bsb)^T \mLambda^{-1} (\bsa -\bsb)] /2   \bigr)
-\end{align*}
-If $\mLambda = \mSigma$, then this simplifies to
-\begin{equation*}
-		\lambda\bigl(\bsPsi(\bsx)\bigr) \QMCPYabs{\frac{\partial \bsPsi}{\partial \bsx}}
-	 =  c \exp \bigl( - (\bsa -\bsb)^T \mA^{-T} [
-	2 \bsPhi^{-1}(\bsx) + \mA^{-1}(\bsa -\bsb)] /2   \bigr) .
-\end{equation*}
-If $\mLambda = \lambda^2 \mI$ and  $\mA = \sigma \mI$, then
-\begin{align*}
-	\lambda\bigl(\bsPsi(\bsx)\bigr) \QMCPYabs{\frac{\partial \bsPsi}{\partial \bsx}}
-	& = \frac{c\sigma}{\lambda}
-	\exp \bigl( [ (\lambda^2 -\sigma^2)\bsPhi^{-1}(\bsx)^T\bsPhi^{-1}(\bsx)
-	\\
-	& -  (\bsa -\bsb)^T [
-	2\sigma \bsPhi^{-1}(\bsx) + (\bsa -\bsb)] /(2\lambda^2)   \bigr) .
-\end{align*}
-}
-
-\AGSComment{
-
-\subsection{Chaining Transforms}
-With $k$ transforms, let $\hat{\bsPsi}_j(\bsx) = (\bsPsi_j \circ \bsPsi_{j-1} \circ \dots \circ \bsPsi_1)(\bsx)$ denote the first $j$ compositions and $\bst = \hat{\bsPsi}_k(\bsx)$ be the complete transformation.
-Then,
-$$f(\bsx) = g(\hat{\bsPsi}_k(\bsx)) \frac{\lambda(\hat{\bsPsi}_k(\bsx))}{\varrho(\bsx)} \prod_{i=1}^{k} \left\lvert \frac{\D \, \bsPsi_i}{\D \, \hat{\bsPsi}_{i-1}(\bsx)} \right\rvert$$
-where $\hat{\bsPsi}_0(\bsx) = \bsx$.
-}
-
-\FJHComment{
-
-\subsection{Software Framework}
-
-\newcommand{\dd}{\texttt{DiscreteDistribution}\xspace}
-\newcommand{\tm}{\texttt{TrueMeasure}\xspace}
-
-\begin{itemize}
-\item \dd $\rightarrow$ \tm gives an instance that can generate a sequence that mimics the measure specified by \tm if \tm is a probability measure; it also has a \texttt{gen\_samples} method.
-
-\item if \tm is not a probability measure, then \dd $\rightarrow$ \tm sets up a transformation from $\calX$ to $\calT$.
-
-\item \texttt{CustomFun} takes a user-defined function $g$ and a \tm instance and uses the $\lambda$ inferred from the \tm to form $f$.
-
-\item Built-in \texttt{Integrand} classes know what $g$ and $f$ are.
-
-\begin{itemize}
-    \item One way is to have these \texttt{Integrand} classes assume a \tm and only need a \dd input. If you do not like the default \tm, then you need to input your own.
-
-    \item Another way is to have these \texttt{Integrand} classes take a \tm input and use that, plus their knowledge of $g$ and $\lambda$ to find $f$, perhaps by what could be considered importance sampling.
-
-\end{itemize}
-\end{itemize}
-}
 

### 2023-03-08 MCQMC2022ChallengesQMCSoftwareArticle.tex
SC: Fix typo and compilation errors

@@
 \usepackage[bottom]{footmisc}% places footnotes at page bottom
+\usepackage{epstopdf,psfrag}
 
@@
 \begin{enumerate}
-    \item Draw a sequence of $n$ a low discrepancy (LD) \cite{Nie92, SloJoe94, DicEtal14a, DicEtal22a} nodes,
+    \item Draw a sequence of $n$ low-discrepancy (LD) \cite{Nie92, SloJoe94, DicEtal14a, DicEtal22a} nodes,
     $\bsx_1,\dots,\bsx_n$ that mimic $\calU[0,1]^d$.
@@
 For small error tolerances, the time and sample size is $\calO(n^{-1})$, and parallel processing gives a speed-up of factor four to five.
-These plots may be reproduced using the Jupyter notebook \url{???}.}
+These plots may be reproduced using the Jupyter notebook \SCComment{Add url}.}
     \label{fig:beam}

### 2023-03-08 MCQMC2022ChallengesQMCSoftwareArticle.tex
SC: Explain more...

@@
 Many of the insights that we have gained have come through our development of the Guaranteeed Automatic Integration Library (GAIL)~\cite{ChoEtal22a} in MATLAB and the Quasi-Monte Carlo Python library (QMCPy)~ \cite{QMCPy2020a}.
-\FJHNote{Sou-Cheng, not sure what this means.
-In particular, QMCPy is not only a QMC software, but also a scientific middleware framework, which provides access to  more than X (Q)MC scientific software such as A~\cite{}, B~\cite{}, C~\cite{}.}
+\FJHNote{Sou-Cheng, not sure what this means.} \SCComment{Fred, more elaboration here: In particular, QMCPy is not just a QMC software containing our own algorithms; from the outset, it is designed to serve as a middleware framework, allowing users to tap into a collection of QMC libraries originally written in various languages by different researchers at different times.
+Using object-oriented design, QMCPy specifies a few  extensible high-level mathematical objects essential in multiple integration.
+Developers of other (Q)MC software could easily extend these objects to incorporate their existing or working algorithms by a combination of mechanism: object inheritance; Python package inclusion; direct code translation; interfaces to compiled C or Java libraries.
+Hence QMCPy has been able to provide access to an increasing number of well established (Q)MC algorithmic or application software such as QRNG~\cite{}, MLMC~\cite{}, GAIL~\cite{}, UM-Bridge~\cite{}.}
 

### 2023-03-08 MCQMC2022ChallengesQMCSoftwareArticle.tex
SC: refine

@@
 Many of the insights that we have gained have come through our development of the Guaranteeed Automatic Integration Library (GAIL)~\cite{ChoEtal22a} in MATLAB and the Quasi-Monte Carlo Python library (QMCPy)~ \cite{QMCPy2020a}.
-\FJHNote{Sou-Cheng, not sure what this means.} \SCComment{Fred, more elaboration here: In particular, QMCPy is not just a QMC software containing our own algorithms; from the outset, it is designed to serve as a middleware framework, allowing users to tap into a collection of QMC libraries originally written in various languages by different researchers at different times.
-Using object-oriented design, QMCPy specifies a few  extensible high-level mathematical objects essential in multiple integration.
+\FJHNote{Sou-Cheng, not sure what this means.} \SCComment{Fred, more elaboration here, but maybe it fits better in Section 3: In particular, QMCPy is not just a QMC software containing our own algorithms; from the outset, it is designed to serve as a middleware framework, allowing users to tap into a collection of QMC libraries originally written in various languages by different researchers at different times.
+Using object-oriented design, QMCPy specifies a few  extensible high-level mathematical objects essential in multiple integration (see Sect.\ \ref{CDHJS_sec:integrated}).
 Developers of other (Q)MC software could easily extend these objects to incorporate their existing or working algorithms by a combination of mechanism: object inheritance; Python package inclusion; direct code translation; interfaces to compiled C or Java libraries.

### 2023-03-09 MCQMC2022ChallengesQMCSoftwareArticle.tex
aleksei: wording updates and a few suggested changes

@@
 We draw on our collective experiences as
-members of the academic QMC community in the, developers of open-source and commercial scientific software, and users of a wide range of scientific software as research scientists or data scientists.
+members of the academic QMC community, developers of open-source and commercial scientific software, and users of a wide range of scientific software as research scientists or data scientists.
 Many of the insights that we have gained have come through our development of the Guaranteeed Automatic Integration Library (GAIL)~\cite{ChoEtal22a} in MATLAB and the Quasi-Monte Carlo Python library (QMCPy)~ \cite{QMCPy2020a}.
@@
 
-Great QMC software in eliminates the need researchers to resort to  ``do-it-yourself'' for established algorithms.  The less code we have to write, the fewer errors.
+Great QMC software eliminates the need researchers to resort to  ``do-it-yourself'' for established algorithms.
+The less code we have to write, the fewer errors.
+\AGSComment{Aleksei: Put in previous subsection.}
 
@@
 \begin{equation*}
-    \Psi((0,1)^d) = \calT, \qquad f(\bsx) = g(\bsPsi(\bsx)) \lambda(\bsPsi(\bsx)) \abs{\frac{\partial \bsPsi}{\partial\bsx}}.
+    \bsPsi((0,1)^d) = \calT, \qquad f(\bsx) = g(\bsPsi(\bsx)) \lambda(\bsPsi(\bsx)) \abs{\frac{\partial \bsPsi}{\partial\bsx}}.
 \end{equation*}
@@
 In this case, one might choose $\bsPsi$ such that $\bsT \sim \bsPsi(\bsX)$, in which case $\lambda(\bsPsi(\bsx)) \abs{\partial \bsPsi/\partial\bsx} = 1$ and $f(\bsx) = g(\bsPsi(\bsx))$.
+\AGSComment{Aleksei: Software such as SciPy and QMCPy facilitate such variable transformations by providing default $\bsPsi$ for common $\bsT$.
+This alleviates the burden on the user to determine a transformation and implement the Jacobian.
+}
 

### 2023-03-09 MCQMC2022ChallengesQMCSoftwareArticle.tex
SC: Added some comments and suggested changes.

@@
 \title*{Challenges in Developing Great Quasi-Monte Carlo Software}
-\authorrunning{S.-C.\ T.\ Choi et al}
-\author{Sou-Cheng Terrya Choi \and Yuhan Ding \and Fred J. Hickernell \and R. Jagadeeswaran Aleksei G. Sorokin}
+\authorrunning{S.-C.\ T.\ Choi et al.}
+\author{Sou-Cheng T.~Choi \and Yuhan Ding \and Fred J. Hickernell \and Jagadeeswaran  Rathinavel \and Aleksei G. Sorokin}
 \institute{
-Sou-Cheng Terrya Choi \at Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{schoi32@hawk.iit.edu}
+Sou-Cheng T.~Choi \at Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{schoi32@iit.edu}
 \and
@@
 \and
-R. Jagadeeswaran \at
+Jagadeeswaran  Rathinavel \at
 Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{jrathin1@hawk.iit.edu}
@@
 We summarize the  QMC software that has been developed to date, propose some criteria for developing great QMC software, and suggest some steps toward achieving great software.
+\SCComment{We demonstrate these criteria and steps in practice with Quasi-Monte Carlo Python library (QMCPy), an open-source community software framework, extensible by design with common programming interfaces to an increasing number of existing or emerging QMC libraries developed by the greater community of QMC researchers.}
 }
@@
 Users computing solutions in comprehensive software environments, such as NAG, MATLAB, R, SciPy, and TensorFlow, have convenient access to LD sequence generators.
-However, most of these implementations in large libraries could provide more flexible LD offerings.
-Some large environments, such as  Dakota \cite{DakotaUsersManual} and SAS, have quite limited LD generators and should expand their offerings.
-Developers of smaller libraries with more QMC features, like BRODA and QMCPy, should demonstrate how their libraries can connect well with other software environments.
+However, most of these implementations in large libraries could provide more flexible \SCComment{what does more flexible means?} LD offerings.
+Some large environments such as  Dakota \cite{DakotaUsersManual} have quite limited LD generators and should expand their offerings.
+\SCComment{Other established software such as SAS expand offerings of trustworthy and efficient analytics through continuous internal implementation as well as flexible integration of third-party open-source software; see \url{https://www.sas.com/en_us/software/viya/open.html}.
+Such latter strategy represents opportunities for}  developers of smaller libraries with more QMC features, like BRODA and QMCPy, to demonstrate how their libraries can connect well with other software environments.
 An example of QMCPy connecting with other software is given in the next subsection.
@@
 QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.
-For small error tolerances, the time and sample size is $\calO(n^{-1})$, and parallel processing gives a speed-up of factor four to five.
-These plots may be reproduced using the Jupyter notebook \SCComment{Add url}.}
+For small error tolerances, the time and sample size is $\calO(n^{-1})$ \SCComment{typo:$\calO(\varepsilon^{-1})$}, and parallel processing gives a speed-up of factor four to five.
+These plots may be reproduced using the Jupyter notebook \SCComment{(1) Add url.
+(2) What is the meaning of min and sec on the y-axis of lower left graph (The label says (s), which means seconds right?)?
+Are there numerical values on (left and right) y-axis?
+(3) The font size of the axis labels for the lower graphs could be made larger more like the top graph.
+(4) I guess the blue colors are covered by the orange colors in the bottom right graph.
+Maybe we can use blue crosses and orange broken lines instead.}}
     \label{fig:beam}

### 2023-03-09 MCQMC2022ChallengesQMCSoftwareArticle.tex
SC: A few more minor changes

@@
 Practitioners may wish to determine $n$ adaptively to satisfy a prescribed error tolerance.
+\SCComment{Use $\hat{\mu}_n$ in (1) so that it is consistent with (3).}
 
@@
 
-Recent interest in quality scientific software is exemplified by the 2021 US Department of Energy \emph{Workshop on the Science of Scientific-Software Development and Use} \cite{ASCR-SSSDU,osti_1846008}.
+Recent interest in quality scientific software is exemplified by the 2021 US Department of Energy \SCComment{report,} \emph{Workshop on the Science of Scientific-Software Development and Use} \cite{ASCR-SSSDU, osti_1846008}.
 This workshop not only discussed diagnostics and treatments for the challenges of developing great scientific software, but also emphasized the importance of implementing theoretical advancements into well written, accessible software libraries.
@@
 
-LD generators are the most prevalent generally available components of QMC software.
+LD generators are the most prevalent generally available \SCComment{do we need `generally available'?} components of QMC software.
 These generators appear in Association of Computing Machinery  publications \cite{BraFox88,BraFoxNie92,HonHic00a}, FinDer \cite{PasTra95,FinDer}, libseq \cite{FriKel02,FriKelweb}, BRODA \cite{BRODA20a}, NAG \cite{NAG27}, MATLAB \cite{MAT9.13}, SamplePack \cite{SamplePack}, Gr\"unschlo{\ss}'s website \cite{GruWeb}, the Magic Point Shop \cite{NuyWeb}, R \cite{QRNG2020}, Julia \cite{QMCJulia}, PyTorch \cite{paszke2019pytorch}, SAS \cite{SAS_LD}, SciPy \cite{virtanen2020scipy}, and TensorFlow \cite{tfqfQMC2021a}, MatBuilder \cite{paulin2022}, and  QMCPy \cite{QMCPy2020a}.
@@
 \subsection{Integrands and Variable Transformations} \label{CDHJS:sec:integrands}
-The original integral or expectation arising from the application is defined in terms of the integrand, $g$, and the non-negative weight, $\lambda$.  For example,
+The original integral or expectation arising from the \SCComment{an?} application is defined in terms of the integrand, $g$, and the non-negative weight, $\lambda$.  For example,
 \begin{itemize}
@@
 QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.
-For small error tolerances, the time and sample size is $\calO(n^{-1})$ \SCComment{typo:$\calO(\varepsilon^{-1})$}, and parallel processing gives a speed-up of factor four to five.
+For small error tolerances, the time and sample size is $\calO(n^{-1})$ \SCComment{typo:$\calO(\varepsilon^{-1})$}, and parallel processing gives a speed-up factor of four to five.
 These plots may be reproduced using the Jupyter notebook \SCComment{(1) Add url.

### 2023-03-12 MCQMC2022ChallengesQMCSoftwareArticle.tex
sc:  fix typos

--- 2023-03-09 MCQMC2022ChallengesQMCSoftwareArticle.tex
@@
     \begin{equation}
-        \hat{\mu} := \frac{1}{n} \sum_{i=1}^n f(\bsx_i).
+        \hat{\mu_n} := \frac{1}{n} \sum_{i=1}^n f(\bsx_i).
         \label{eq:mc_average}
@@
 However, the practice of QMC is often more complicated.
-The original problem may need to be rewritten to fit the above form and/or to facilitate a good approximation with small $n$.
+The original problem may need to be rewritten to fit the above form and/or to facilitate a good approximation with a small number of sampling points, $n$.
 Practitioners may wish to determine $n$ adaptively to satisfy a prescribed error tolerance.
@@
 \end{itemize}
-In each section we describe the challenges faced and how they might be overcome.
+In each section, we describe the challenges faced and how they might be overcome.
 
@@
 members of the academic QMC community, developers of open-source and commercial scientific software, and users of a wide range of scientific software as research scientists or data scientists.
-Many of the insights that we have gained have come through our development of the Guaranteeed Automatic Integration Library (GAIL)~\cite{ChoEtal22a} in MATLAB and the Quasi-Monte Carlo Python library (QMCPy)~ \cite{QMCPy2020a}.
+Many of the insights that we have gained have come through our development of the Guaranteed Automatic Integration Library (GAIL)~\cite{ChoEtal22a} in MATLAB and the Quasi-Monte Carlo Python library (QMCPy)~ \cite{QMCPy2020a}.
 \FJHNote{Sou-Cheng, not sure what this means.} \SCComment{Fred, more elaboration here, but maybe it fits better in Section 3: In particular, QMCPy is not just a QMC software containing our own algorithms; from the outset, it is designed to serve as a middleware framework, allowing users to tap into a collection of QMC libraries originally written in various languages by different researchers at different times.
@@
 Developers of other (Q)MC software could easily extend these objects to incorporate their existing or working algorithms by a combination of mechanism: object inheritance; Python package inclusion; direct code translation; interfaces to compiled C or Java libraries.
-Hence QMCPy has been able to provide access to an increasing number of well established (Q)MC algorithmic or application software such as QRNG~\cite{}, MLMC~\cite{}, GAIL~\cite{}, UM-Bridge~\cite{}.}
+Hence QMCPy has been able to provide access to an increasing number of well-established (Q)MC algorithmic or application software.}
 
@@
 
-Recent interest in quality scientific software is exemplified by the 2021 US Department of Energy \emph{Workshop on the Science of Scientific-Software Development and Use} \cite{ASCR-SSSDU,osti_1846008}.
-This workshop not only discussed diagnostics and treatments for the challenges of developing great scientific software, but also emphasized the importance of implementing theoretical advancements into well written, accessible software libraries.
+Recent interest in quality scientific software is exemplified by the 2021 US Department of Energy report, \emph{Workshop on the Science of Scientific-Software Development and Use} \cite{ASCR-SSSDU, osti_1846008}.
+This workshop not only discussed diagnostics and treatments for the challenges of developing great scientific software, but also emphasized the importance of implementing theoretical advancements into well-written, accessible software libraries.
 Three cross-cutting themes arose in this workshop:
@@
 
-Great QMC software eliminates the need researchers to resort to  ``do-it-yourself'' for established algorithms.
+Great QMC software eliminates the need for researchers to resort to  ``do-it-yourself'' for established algorithms.
 The less code we have to write, the fewer errors.
@@
 
-Expanding on the problem formulation in the introduction, we want to approximate well the integral or expectation, $\mu$, by the sample mean, $\hat{\mu}$:
+Expanding on the problem formulation in the introduction, we want to approximate well the integral or expectation, $\mu$, by the sample mean, $\hat{\mu_n}$:
 \begin{gather}
@@
 
-LD generators are the most prevalent generally available components of QMC software.
-These generators appear in Association of Computing Machinery  publications \cite{BraFox88,BraFoxNie92,HonHic00a}, FinDer \cite{PasTra95,FinDer}, libseq \cite{FriKel02,FriKelweb}, BRODA \cite{BRODA20a}, NAG \cite{NAG27}, MATLAB \cite{MAT9.13}, SamplePack \cite{SamplePack}, Gr\"unschlo{\ss}'s website \cite{GruWeb}, the Magic Point Shop \cite{NuyWeb}, R \cite{QRNG2020}, Julia \cite{QMCJulia}, PyTorch \cite{paszke2019pytorch}, SAS \cite{SAS_LD}, SciPy \cite{virtanen2020scipy}, and TensorFlow \cite{tfqfQMC2021a}, MatBuilder \cite{paulin2022}, and  QMCPy \cite{QMCPy2020a}.
+LD generators are the most prevalent, generally available components of QMC software.
+These generators appear in Association of Computing Machinery  publications \cite{BraFox88,BraFoxNie92,HonHic00a}, FinDer \cite{PasTra95,FinDer}, libseq \cite{FriKel02,FriKelweb}, BRODA \cite{BRODA20a}, NAG \cite{NAG27}, MATLAB \cite{MAT9.13}, SamplePack \cite{SamplePack}, Gr\"unschlo{\ss}'s website \cite{GruWeb}, the Magic Point Shop \cite{NuyWeb}, R \cite{QRNG2020}, Julia \cite{QMCJulia}, PyTorch \cite{paszke2019pytorch}, SAS \cite{SAS_LD}, SciPy \cite{virtanen2020scipy},  TensorFlow \cite{tfqfQMC2021a}, MatBuilder \cite{paulin2022}, and  QMCPy \cite{QMCPy2020a}.
 
@@
 Users computing solutions in comprehensive software environments, such as NAG, MATLAB, R, SciPy, and TensorFlow, have convenient access to LD sequence generators.
-However, most of these implementations in large libraries could provide more flexible \SCComment{what does more flexible means?} LD offerings.
+However, most of these implementations in large libraries could provide more flexible \SCComment{what does more flexible mean?} LD offerings.
 Some large environments such as  Dakota \cite{DakotaUsersManual} have quite limited LD generators and should expand their offerings.
 \SCComment{Other established software such as SAS expand offerings of trustworthy and efficient analytics through continuous internal implementation as well as flexible integration of third-party open-source software; see \url{https://www.sas.com/en_us/software/viya/open.html}.
-Such latter strategy represents opportunities for}  developers of smaller libraries with more QMC features, like BRODA and QMCPy, to demonstrate how their libraries can connect well with other software environments.
+Such a latter software development strategy presents opportunities for}  developers of smaller libraries with more QMC features, like BRODA and QMCPy, to demonstrate how their libraries can connect well with other software environments.
 An example of QMCPy connecting with other software is given in the next subsection.
@@
 \subsection{Integrands and Variable Transformations} \label{CDHJS:sec:integrands}
-The original integral or expectation arising from the application is defined in terms of the integrand, $g$, and the non-negative weight, $\lambda$.  For example,
+The original integral or expectation arising from an application is defined in terms of the integrand, $g$, and the non-negative weight, $\lambda$.  For example,
 \begin{itemize}
@@
 QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.
-For small error tolerances, the time and sample size is $\calO(n^{-1})$ \SCComment{typo:$\calO(\varepsilon^{-1})$}, and parallel processing gives a speed-up of factor four to five.
-These plots may be reproduced using the Jupyter notebook \SCComment{(1) Add url.
+For small error tolerances, the time and sample size is $\calO(\varepsilon^{-1})$, and parallel processing gives a speed-up factor of four to five.
+These plots may be reproduced using the Jupyter notebook \SCComment{(1) Add url or citation.
 (2) What is the meaning of min and sec on the y-axis of lower left graph (The label says (s), which means seconds right?)?
@@
 The cantilevered beam example illustrated in Figure \ref{fig:beam} utilizes QMCPy's stopping criterion to choose the sample size, $n$, based on the input error tolerance, $\varepsilon$, and the observed discrete Fourier series coefficients of the integrand, $f$.
-See \cite{HicJim16a} for details of this stopping criterion for lattice rules.
+See \cite{HicJim16a} for details of this stopping criterion for lattice rules and \cite{HicEtal18a} for a review on automatic QMC stopping criteria.
 Theoretically justified stopping criteria for choosing sample sizes are important and a relatively recent development \cite{HicEtal14a,HicJim16a,JimHic16a,RatHic19a,JagHic22a}.
@@
 They should be under active development or merged into other libraries that are actively maintained.
-Commercial software has an infrastructure to maintain their algorithms, but these algorithms need periodic refreshing, Open source software requires a community of active users and developers to keep the library up-to-date and bug-free.
+Commercial software has an infrastructure to maintain their algorithms, but these algorithms need periodic refreshing.
+Open source software requires a community of active users and developers to keep the library up-to-date and bug-free.
 
@@
 The more robust our QMC software, the easier it is to open new application areas to QMC and grow our QMC community.
+
+    \item \SCComment{Let's publish our QMC  software with journals such as ACM TOMS, JORS, etc. to obtain code review and
+    feedback.}
+
+    \item \SCComment{When we find QMC scientific software useful in our research or applications, let's encourage and recognize such software by giving them proper citations~\cite{smith2016software}.
+}
+
+    \item \SCComment{Encourage QMC community members to learn about software development through interdisciplinary education, internship, or hackathon. }
+
 \end{itemize}
@@
 Great software should be published with a digital object identifier (DOI).
-The term research software engineer (RSE) is becoming a more valued vocation.
-According to the website for  Research Software Engineers International \url{https://researchsoftware.org}, \emph{Research Software Engineers are people who combine professional software expertise with an understanding of research.}
+\SCComment{The previous sentence is about publishing software.
+The next point is about a career path.}  The term research software engineer (RSE) is becoming a more valued vocation.
+According to the website for  Research Software Engineers International, \url{https://researchsoftware.org}, \emph{Research Software Engineers are people who combine professional software expertise with an understanding of research.}
 
@@
 
+\SCComment{Change `sobol' to `Sobol' in reference items 29 and 30. Change `scipy' to `SciPy' in 44 and 45. }
 \bibliographystyle{spmpsci}
-\bibliography{FJH23,FJHown23}
+\bibliography{FJH23,FJHown23,sc}
 

### 2023-03-14 MCQMC2022ChallengesQMCSoftwareArticle.tex
sc: Added examples in Section 5.

@@
 \section{Efficient} \label{CDHJS_sec:efficient}
-Another obvious feature of great QMC software is numerical efficiency.
+Another obvious feature of great QMC software is computational efficiency.
 This requires developers to identify bottlenecks in performance.
@@
 Efficiency gains in QMC software can also result from clever algorithm choices.
-The Bayesian stopping rules for lattice and Sobol sequences in \cite{RatHic19a,JagHic22a} choose covariance kernels that match the corresponding LD sequence.
+\SCComment{Adaptive QMC algorithms as in \cite{JimHic16a, HicJim16a, HicEtal17a} can be used to dynamically increase the number of sampling points based on estimated error bounds.
+They can lead to significant efficiency gains by reducing the number of unnecessary function evaluations.} The Bayesian stopping rules for lattice and Sobol sequences in \cite{RatHic19a,JagHic22a} choose covariance kernels that match the corresponding LD sequence.
 This reduces the computations from what would be $\calO(n^3)$ for arbitrary covariance kernels to only $\calO(n \log n)$ for these well-chosen kernels.
@@
 Blogs educate a wider audience and enlarge our QMC community.
-A suite of unit tests should be run regularly to ensure that new features do not introduce bugs.
-There should be a place for users to report bugs and make feature requests.
+A suite of unit tests should be run regularly to ensure that new features do not introduce bugs.\SCComment{Move the last sentence to section on Correct} There should be a place for users to report bugs and make feature requests.
 

### 2023-03-14 MCQMC2022ChallengesQMCSoftwareArticle.tex
sc: move unit test discussion from Section 6 to Section 4

@@
 
+\SCComment{The above examples and discussion also underscore the importance of continuous testing in QMC software development.
+This involves creating a comprehensive test suite that covers a wide range of scenarios and input parameters.
+Additionally, QMC software should be benchmarked against known results to ensure that it is both accurate and efficient.}
+
 \section{Efficient} \label{CDHJS_sec:efficient}
@@
 Blogs educate a wider audience and enlarge our QMC community.
-A suite of unit tests should be run regularly to ensure that new features do not introduce bugs.\SCComment{Move the last sentence to section on Correct} There should be a place for users to report bugs and make feature requests.
+There should be a place for users to report bugs and make feature requests.
 

### 2023-03-17 MCQMC2022ChallengesQMCSoftwareArticle.tex
2023 Mac 14 Updates
Updates from Overleaf

@@
 %%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%% Author sign off on this first submission, last revision
+% Sou-Cheng Choi signed off on
+% Yuhan Ding signed off on
+% Fred J. Hickernell signed off on
+% Jagadeeswaran Rathinavel signed off on
+% Aleksei Sorokin signed off on
 
@@
 We summarize the  QMC software that has been developed to date, propose some criteria for developing great QMC software, and suggest some steps toward achieving great software.
-\SCComment{We demonstrate these criteria and steps in practice with Quasi-Monte Carlo Python library (QMCPy), an open-source community software framework, extensible by design with common programming interfaces to an increasing number of existing or emerging QMC libraries developed by the greater community of QMC researchers.}
+We illustrate these criteria and steps in practice with Quasi-Monte Carlo Python library (QMCPy), an open-source community software framework, extensible by design with common programming interfaces to an increasing number of existing or emerging QMC libraries developed by the greater community of QMC researchers.
 }

### 2023-03-17 MCQMC2022ChallengesQMCSoftwareArticle.tex
Accepting Sou-Cheng's and Aleksei's comments.

@@
 Many of the insights that we have gained have come through our development of the Guaranteed Automatic Integration Library (GAIL)~\cite{ChoEtal22a} in MATLAB and the Quasi-Monte Carlo Python library (QMCPy)~ \cite{QMCPy2020a}.
-\FJHNote{Sou-Cheng, not sure what this means.} \SCComment{Fred, more elaboration here, but maybe it fits better in Section 3: In particular, QMCPy is not just a QMC software containing our own algorithms; from the outset, it is designed to serve as a middleware framework, allowing users to tap into a collection of QMC libraries originally written in various languages by different researchers at different times.
-Using object-oriented design, QMCPy specifies a few  extensible high-level mathematical objects essential in multiple integration (see Sect.\ \ref{CDHJS_sec:integrated}).
-Developers of other (Q)MC software could easily extend these objects to incorporate their existing or working algorithms by a combination of mechanism: object inheritance; Python package inclusion; direct code translation; interfaces to compiled C or Java libraries.
-Hence QMCPy has been able to provide access to an increasing number of well-established (Q)MC algorithmic or application software.}
 
@@
 
+Great QMC software eliminates the need for researchers to resort to  ``do-it-yourself'' for established algorithms.  The less code we have to write, the fewer errors.
+
 \subsection{QMC Software to Theory}
@@
 For example, the early application of QMC to high dimensional integrals arising in financial risk \cite{PasTra95} led to a wave of theoretical results on the tractability of integration in weighted spaces \cite{Woz99a,DicEtal14a,NovWoz10a}.
-
-Great QMC software eliminates the need for researchers to resort to  ``do-it-yourself'' for established algorithms.
-The less code we have to write, the fewer errors.
-\AGSComment{Aleksei: Put in previous subsection.}
 
@@
 Lattice and digital sequences are not unique, and improved or alternate versions continue to be constructed.
-Stephen Joe and Frances Kuo \cite{JoeKuo03,JoeKuo08,SobolDirection} have proposed  Sobol' generator direction numbers, which are now widely used, but were not part of early LD sequence software.
+Stephen Joe and Frances Kuo \cite{JoeKuo03,JoeKuo08a,SobolDirection} have proposed  Sobol' generator direction numbers, which are now widely used, but were not part of early LD sequence software.
 Pierre L'Ecuyer and collaborators have created LatNetBuilder \cite{LatNet} to construct good lattice and polynomial lattice sequences based on user-defined criteria.
@@
 Users computing solutions in comprehensive software environments, such as NAG, MATLAB, R, SciPy, and TensorFlow, have convenient access to LD sequence generators.
-However, most of these implementations in large libraries could provide more flexible \SCComment{what does more flexible mean?} LD offerings.
+However, most of these implementations in large libraries would benefit a wider menu of LD offerings, including custom generators from LatNetBuilder or other sources.
 Some large environments such as  Dakota \cite{DakotaUsersManual} have quite limited LD generators and should expand their offerings.
-\SCComment{Other established software such as SAS expand offerings of trustworthy and efficient analytics through continuous internal implementation as well as flexible integration of third-party open-source software; see \url{https://www.sas.com/en_us/software/viya/open.html}.
-Such a latter software development strategy presents opportunities for}  developers of smaller libraries with more QMC features, like BRODA and QMCPy, to demonstrate how their libraries can connect well with other software environments.
+SAS facilitates flexible integration of third-party open-source software; see \url{https://www.sas.com/en_us/software/viya/open.html}.
+This presents opportunities for  developers of smaller libraries with more QMC features, like BRODA and QMCPy, to connect their libraries with larger software environments.
 An example of QMCPy connecting with other software is given in the next subsection.
+
+In fact, QMCPy is designed not simply to contain our own algorithms.
+It is designed to serve as a middleware framework, allowing users to tap into a collection of QMC libraries originally written in various languages by different researchers at different times.
+Using object-oriented design, QMCPy specifies a few  extensible high-level mathematical objects as described in this section and in \cite{ChoEtal22a}.
+Developers of other (Q)MC software can  extend these objects to incorporate their existing algorithms by a combination of mechanisms: object inheritance, Python package inclusion, direct code translation, and/or interfaces to compiled C or Java libraries.
 
@@
 In this case, one might choose $\bsPsi$ such that $\bsT \sim \bsPsi(\bsX)$, in which case $\lambda(\bsPsi(\bsx)) \abs{\partial \bsPsi/\partial\bsx} = 1$ and $f(\bsx) = g(\bsPsi(\bsx))$.
-\AGSComment{Aleksei: Software such as SciPy and QMCPy facilitate such variable transformations by providing default $\bsPsi$ for common $\bsT$.
+Software such as SciPy and QMCPy facilitate such variable transformations by providing default $\bsPsi$ for common $\bsT$.
 This alleviates the burden on the user to determine a transformation and implement the Jacobian.
-}
 
@@
 
-\SCComment{The above examples and discussion also underscore the importance of continuous testing in QMC software development.
+The above examples and discussion also underscore the importance of continuous testing in QMC software development.
 This involves creating a comprehensive test suite that covers a wide range of scenarios and input parameters.
-Additionally, QMC software should be benchmarked against known results to ensure that it is both accurate and efficient.}
+Additionally, QMC software should be benchmarked against known results to ensure that it is both accurate and efficient.
 
@@
 Efficiency gains in QMC software can also result from clever algorithm choices.
-\SCComment{Adaptive QMC algorithms as in \cite{JimHic16a, HicJim16a, HicEtal17a} can be used to dynamically increase the number of sampling points based on estimated error bounds.
-They can lead to significant efficiency gains by reducing the number of unnecessary function evaluations.} The Bayesian stopping rules for lattice and Sobol sequences in \cite{RatHic19a,JagHic22a} choose covariance kernels that match the corresponding LD sequence.
+Adaptive QMC algorithms as in \cite{JimHic16a, HicJim16a, HicEtal17a, RatHic19a,JagHic22a} can be used to dynamically increase the number of sampling points based on estimated error bounds.
+They can lead to significant efficiency gains by reducing the number of unnecessary function evaluations.
+The Bayesian stopping rules for lattice and Sobol sequences in \cite{RatHic19a,JagHic22a} choose covariance kernels that match the corresponding LD sequence.
 This reduces the computations from what would be $\calO(n^3)$ for arbitrary covariance kernels to only $\calO(n \log n)$ for these well-chosen kernels.
@@
 
-    \item \SCComment{Let's publish our QMC  software with journals such as ACM TOMS, JORS, etc. to obtain code review and
-    feedback.}
+    \item Let's publish our QMC  software with journals such as ACM TOMS, JORS, etc. to obtain code review and
+    feedback.
 
-    \item \SCComment{When we find QMC scientific software useful in our research or applications, let's encourage and recognize such software by giving them proper citations~\cite{smith2016software}.
-}
+    \item When we find QMC scientific software useful in our research or applications, let's encourage and recognize such software by properly citing it~\cite{smith2016software}.
 
-    \item \SCComment{Encourage QMC community members to learn about software development through interdisciplinary education, internship, or hackathon. }
+    \item Encourage QMC community members to learn about software development through interdisciplinary education, internships, or hackathons.
 
@@
 Great software should be published with a digital object identifier (DOI).
-\SCComment{The previous sentence is about publishing software.
-The next point is about a career path.}  The term research software engineer (RSE) is becoming a more valued vocation.
+
+The term research software engineer (RSE) is becoming a more valued vocation.
 According to the website for  Research Software Engineers International, \url{https://researchsoftware.org}, \emph{Research Software Engineers are people who combine professional software expertise with an understanding of research.}
@@
 
-\SCComment{Change `sobol' to `Sobol' in reference items 29 and 30. Change `scipy' to `SciPy' in 44 and 45. }
 \bibliographystyle{spmpsci}

### 2023-03-17 MCQMC2022ChallengesQMCSoftwareArticle.tex
Fred 2023 Mac 14 Updates

@@
     \includegraphics[width=0.5\textwidth]{cantileveredbeamwords.eps}
-    \includegraphics[width=0.7\textwidth]{ldparallelbeam.eps}
+    \includegraphics[width=1\textwidth]{ldparallelbeam.eps}
     \caption{The expected deflection of a cantilevered beam with random Young's modulus.
@@
 For small error tolerances, the time and sample size is $\calO(\varepsilon^{-1})$, and parallel processing gives a speed-up factor of four to five.
-These plots may be reproduced using the Jupyter notebook \SCComment{(1) Add url or citation.
-(2) What is the meaning of min and sec on the y-axis of lower left graph (The label says (s), which means seconds right?)?
-Are there numerical values on (left and right) y-axis?
-(3) The font size of the axis labels for the lower graphs could be made larger more like the top graph.
-(4) I guess the blue colors are covered by the orange colors in the bottom right graph.
-Maybe we can use blue crosses and orange broken lines instead.}}
+These plots may be reproduced using the Jupyter notebook \url{https://github.com/QMCSoftware/QMCSoftware/blob/develop/demos/talk_paper_demos/MCQMC2022_Article_Figures.ipynb}  }
     \label{fig:beam}

### 2023-03-17 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf
Fred 2023 Mac 14 Updates
aleksei's small changes and sign off

@@
 % Jagadeeswaran Rathinavel signed off on
-% Aleksei Sorokin signed off on
+% Aleksei Sorokin signed off on 3/17/2023
 
@@
 We summarize the  QMC software that has been developed to date, propose some criteria for developing great QMC software, and suggest some steps toward achieving great software.
-We illustrate these criteria and steps in practice with Quasi-Monte Carlo Python library (QMCPy), an open-source community software framework, extensible by design with common programming interfaces to an increasing number of existing or emerging QMC libraries developed by the greater community of QMC researchers.
+We illustrate these criteria and steps in practice with the Quasi-Monte Carlo Python library (QMCPy), an open-source community software framework, extensible by design with common programming interfaces to an increasing number of existing or emerging QMC libraries developed by the greater community of QMC researchers.
 }
@@
     \begin{equation}
-        \hat{\mu_n} := \frac{1}{n} \sum_{i=1}^n f(\bsx_i).
+        \hat{\mu}_n := \frac{1}{n} \sum_{i=1}^n f(\bsx_i).
         \label{eq:mc_average}
@@
 
-Expanding on the problem formulation in the introduction, we want to approximate well the integral or expectation, $\mu$, by the sample mean, $\hat{\mu_n}$:
+Expanding on the problem formulation in the introduction, we want to approximate well the integral or expectation, $\mu$, by the sample mean, $\hat{\mu}_n$:
 \begin{gather}
@@
 For small error tolerances, the time and sample size is $\calO(\varepsilon^{-1})$, and parallel processing gives a speed-up factor of four to five.
-These plots may be reproduced using the Jupyter notebook \url{https://github.com/QMCSoftware/QMCSoftware/blob/develop/demos/talk_paper_demos/MCQMC2022_Article_Figures.ipynb}  }
+These plots may be reproduced using the Jupyter notebook \cite{MCQMC2022Figs}.
+}
     \label{fig:beam}

### 2023-03-17 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

@@
 % Yuhan Ding signed off on
-% Fred J. Hickernell signed off on
+% Fred J. Hickernell signed off on 3/17
 % Jagadeeswaran Rathinavel signed off on
@@
 Lattice and digital sequences are not unique, and improved or alternate versions continue to be constructed.
-Stephen Joe and Frances Kuo \cite{JoeKuo03,JoeKuo08a,SobolDirection} have proposed  Sobol' generator direction numbers, which are now widely used, but were not part of early LD sequence software.
+Stephen Joe and Frances Kuo \cite{JoeKuo03,KuoJoe08a,SobolDirection} have proposed  Sobol' generator direction numbers, which are now widely used, but were not part of early LD sequence software.
 Pierre L'Ecuyer and collaborators have created LatNetBuilder \cite{LatNet} to construct good lattice and polynomial lattice sequences based on user-defined criteria.
@@
 \bibliographystyle{spmpsci}
-\bibliography{FJH23,FJHown23,sc}
+\bibliography{FJH23,FJHown23}
 
@@
     \item Low discrepancy generators \cite{FriKel02,FriKelweb}, SamplePack \cite{SamplePack}, \cite{GruWeb},  and MatBuilder \cite{paulin2022}
-    \item Sobol' direction numbers \cite{JoeKuo03,JoeKuo08,SobolDirection}
+    \item Sobol' direction numbers \cite{JoeKuo03,KuoJoe08a,SobolDirection}
     \item Fast CBC, Magic Point Shop, QMC4PDE and other code since 2004  at \cite{NuyWeb}

### 2023-03-17 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf
sc: add an affiliation

@@
 
-%%%%%%% Author sign off on this first submission, last revision
-% Sou-Cheng Choi signed off on
+%%%%%%% Author sign off on this first submission, last revision 3/17/2023
+% Sou-Cheng Choi signed off on 3/17/20223
 % Yuhan Ding signed off on
-% Fred J. Hickernell signed off on 3/17
+% Fred J. Hickernell signed off on 3/17/2023
 % Jagadeeswaran Rathinavel signed off on
@@
 \institute{
-Sou-Cheng T.~Choi \at Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{schoi32@iit.edu}
+Sou-Cheng T.~Choi \at Department of Applied Mathematics, Illinois Institute of Technology; SAS Institute Inc. \\RE 220, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{schoi32@iit.edu}
 \and
@@
 These sequences may be deterministic or random and are intended to have an empirical distribution that approximates well the uniform distribution.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{latticepts.eps}
+    \caption{A two-dimensional projection of}
+    \label{fig:latticepts}
+\end{figure}
 

### 2023-03-17 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

@@
 These sequences may be deterministic or random and are intended to have an empirical distribution that approximates well the uniform distribution.
+Figure \ref{CDHJSfiglatticepts} illustrates an example of a lattice sequence.
 
@@
     \includegraphics[width=\textwidth]{latticepts.eps}
-    \caption{A two-dimensional projection of}
-    \label{fig:latticepts}
+    \caption{A two-dimensional projection of lattice points for increasing sample size.
+The later points fill in gaps because they are highly correlated with the earlier points.
+These plots may be reproduced using the Jupyter notebook \cite{MCQMC2022Figs}.
+}
+    \label{CDHJSfiglatticepts}
 \end{figure}
@@
 }
-    \label{fig:beam}
+    \label{CDHJSfigbeam}
 \end{figure}
@@
 In the uncertainty quantification (UQ) space, the recently developed UQ and Model Bridge (UM-Bridge) \cite{umbridge}, connects libraries that generate sequences of model parameters, $\bsx_1, \bsx_2, \ldots$ with libraries that evaluate models, $f(\bsx_1), f(\bsx_2), \ldots$.
-We illustrate in Figure \ref{fig:beam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam with random Young's modulus \cite{ParSee22a}, where the randomness is given by $\bsX \sim \calU[1,1.2]^3$.
+We illustrate in Figure \ref{CDHJSfigbeam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam with random Young's modulus \cite{ParSee22a}, where the randomness is given by $\bsX \sim \calU[1,1.2]^3$.
 The beam position is discretized at $31$ points.
@@
 \subsection{Stopping Criteria}
-The cantilevered beam example illustrated in Figure \ref{fig:beam} utilizes QMCPy's stopping criterion to choose the sample size, $n$, based on the input error tolerance, $\varepsilon$, and the observed discrete Fourier series coefficients of the integrand, $f$.
+The cantilevered beam example illustrated in Figure \ref{CDHJSfigbeam} utilizes QMCPy's stopping criterion to choose the sample size, $n$, based on the input error tolerance, $\varepsilon$, and the observed discrete Fourier series coefficients of the integrand, $f$.
 See \cite{HicJim16a} for details of this stopping criterion for lattice rules and \cite{HicEtal18a} for a review on automatic QMC stopping criteria.
@@
 The load balancing provided by UM-Bridge may be utilized to evaluate the model in parallel, a potentially advantageous feature for models which are costly to evaluate.
-Figure \ref{fig:umbridge_muq} depicts the $31$ approximate expected displacement values subject to uniform stiffness.
+Figure \ref{CDHJSfigumbridge_muq} depicts the $31$ approximate expected displacement values subject to uniform stiffness.
 
@@
     \caption{QMCPy approximate expected displacement at $31$ locations subject to random uniform stiffness. }
-    \label{fig:umbridge_muq}
+    \label{CDHJSfigumbridge_muq}
 \end{figure}

### 2023-03-17 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

@@
 % Fred J. Hickernell signed off on 3/17/2023
-% Jagadeeswaran Rathinavel signed off on
+% Jagadeeswaran Rathinavel signed off on 3/17/2023
 % Aleksei Sorokin signed off on 3/17/2023
@@
 These sequences may be deterministic or random and are intended to have an empirical distribution that approximates well the uniform distribution.
-Figure \ref{CDHJSfiglatticepts} illustrates an example of a lattice sequence.
+Figure \ref{CDHJSfiglatticepts} illustrates an example of a (randomly shifted) lattice sequence.
 
@@
 It is designed to serve as a middleware framework, allowing users to tap into a collection of QMC libraries originally written in various languages by different researchers at different times.
-Using object-oriented design, QMCPy specifies a few  extensible high-level mathematical objects as described in this section and in \cite{ChoEtal22a}.
+Using the object-oriented design, QMCPy specifies a few  extensible high-level mathematical objects as described in this section and in \cite{ChoEtal22a}.
 Developers of other (Q)MC software can  extend these objects to incorporate their existing algorithms by a combination of mechanisms: object inheritance, Python package inclusion, direct code translation, and/or interfaces to compiled C or Java libraries.
@@
 
-	\item $g\lambda$ is the unnormalized Bayesian posterior density multiplied by parameter of interest \cite{GelEtal13}.
+	\item $g\lambda$ is the unnormalized Bayesian posterior density multiplied by the parameter of interest \cite{GelEtal13}.
 \end{itemize}

### 2023-03-17 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

@@
 Pierre L'Ecuyer and collaborators have created LatNetBuilder \cite{LatNet} to construct good lattice and polynomial lattice sequences based on user-defined criteria.
-During a lunch at MCQMC 2022, it was proposed that we  agree on a consistent format for storing the parameters that define good LD sequences so that they can be shared across software libraries.
+During lunch at MCQMC 2022, it was proposed that we  agree on a consistent format for storing the parameters that define good LD sequences so that they can be shared across software libraries.
 
@@
 Some large environments such as  Dakota \cite{DakotaUsersManual} have quite limited LD generators and should expand their offerings.
-SAS facilitates flexible integration of third-party open-source software; see \url{https://www.sas.com/en_us/software/viya/open.html}.
+SAS facilitates flexible integration of third-party open-source software; see \url{https://www.sas.com/en_us/software/viya/open.html}
  This presents opportunities for  developers of smaller libraries with more QMC features, like BRODA and QMCPy, to connect their libraries with larger software environments.

### 2023-03-18 MCQMC2022ChallengesQMCSoftwareArticle.tex
sc: fix broken url with hyperref
sc: use hyperref with url

@@
 % RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\PassOptionsToPackage{hyphens}{url}
 \documentclass[graybox]{svmult}
@@
 \usepackage{xspace}
+\usepackage[bookmarks=false]{hyperref}
 \usepackage[hyphens]{url}
@@
 Some large environments such as  Dakota \cite{DakotaUsersManual} have quite limited LD generators and should expand their offerings.
-SAS facilitates flexible integration of third-party open-source software; see \url{https://www.sas.com/en_us/software/viya/open.html}.
+SAS facilitates flexible integration of third-party open-source software; see \href{https://www.sas.com/en_us/software/viya/open.html}{\url{https://www.sas.com/en_us/software/viya/open.html}}.
 This presents opportunities for developers of smaller libraries with more QMC features, like BRODA and QMCPy, to connect their libraries with larger software environments.
@@
 
-The term research software engineer (RSE) is becoming a more valued vocation.
-According to the website for  Research Software Engineers International, \url{https://researchsoftware.org}, \emph{Research Software Engineers are people who combine professional software expertise with an understanding of research.}
+The term research software engineer (RSE) is becoming a more valued vocation.  According to the website for  Research Software Engineers International, 
+\href{https://researchsoftware.org}{\url{https://researchsoftware.org}}, \emph{Research Software Engineers are people who combine professional software expertise with an understanding of research.}
 

### 2023-03-19 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

@@
 
-%%%%%%% Author sign off on this first submission, last revision 3/17/2023
+%%%%%%% Author sign off on this first submission, last revision 3/19/2023
 % Sou-Cheng Choi signed off on 3/17/20223
-% Yuhan Ding signed off on
+% Yuhan Ding signed off on 3/19/2023
 % Fred J. Hickernell signed off on 3/17/2023
@@
 \end{document}
-
-\section{Tenets of Great Software} \label{CDHJS_sec:tenets_of_great_software}
-
-In this section we lay out common aspirations of scientific software.
-While some objectives are expected by practitioners, e.g.
-correctly written software, others may better be characterized as features that are pursued as a project matures, e.g.
-supporting scalable computation.
-The following paragraphs highlight aspirations with more important, expected behaviours coming earlier in the discussion.
-
-Software is expected to be \emph{correct}.
-A practitioner generating a digital sequence expects the routine to produce theoretically accurate points.
-For example, the QMC community has helped developers ensure implementations of low discrepancy sequence generators include the zeroth point \cite{Owe22a, scipySobol2020a} and  have correct randomization routines.
-%, and disallow thinning / leaping: taking every $k^\text{th}$ value in the sequence.
-
-Algorithms should be implemented in an \emph{efficient} manner.
-Both the compute time and memory requirements should match theoretical developments which in turn should be as close to the state of the art as possible.
-As a classical example, it is expected that an implementation of the discrete Fourier transform for $n$ points has complexity $\mathcal{O}(n\log n)$.
-
-Software should be \emph{accessible} in terms of installation, presentation, navigation, exemplification, and communication.
-A library should be straightforward to install from a common distribution platform, consistent in user interface, intuitive to navigate to modules and documentation, comprehensible in its' examples, and supportive in user engagement.
-Software accessibility considerations are further discussed in Sect.\ \ref{CDHJS_sec:language_library_environment}.
-
-Libraries should be \emph{sustainable} for future development.
-The goal should be to have a sufficient user base and developer community for updates and maintenance.
-Similar to accessibility, sustainability is often overlooked for aspirations discussed later in this section.
-However, we contend that a sub-optimal but user-friendly library with community engagement may be more impactful than an optimal library which lacks support.
-We promote collaborative, community driven development further in Sect.\ \ref{CDHJS_sec:encouraging_collaborative_developement}.
-
-Software should be \emph{integrated} into the environment's scientific software ecosystem.
-More simply, libraries should play nicely together and utilize the mature developments of other teams.
-A simple example in \AGSComment{CITE} shows a straightforward integration between the QMC package QMCPy \cite{QMCPy2020a} and the PDE package FEniCS/Dolfin \cite{LoggEtal_10_2012}, both in Python.
-
-Routines should be \emph{scalable} to advanced computer architectures.
-The widespread availability of on demand computing with multi-core and multi-GPU machines encourages software that can take advantage of parallel computation.
-Scalable algorithms lend themselves to a larger class or problems and provide practitioners the opportunity to more quickly test ideas.
-
-\section{(Quasi-)Monte Carlo Software Architecture} \label{CDHJS_sec:qmc_software_architecture}
-
-An important step in the development of any software is to determine its' architecture.
-How should components interact with each other?
-Which aspects should be modular?
-Object oriented?
-Procedural?
-Answers to these questions form the blueprint of software design and guide implementation.
-
-This section proposes an object oriented design for (Q)MC software.
-The authors have implemented this architecture into the QMCPy \cite{QMCPy2020a} Python package which is further detailed in the MCQMC 2020 proceedings article \AGSComment{CITE}.
-While this is not the only possible architecture for (Q)MC problems, we hope the high level object definitions facilitate easy collaboration and integration for research across the (Q)MC community.
-
-The standard (Q)MC problem is to approximate the true mean $\mu := \mathbb{E}[g(\boldsymbol{T})]$ where $\boldsymbol{T}$ is some $d$ dimensional random variable.
-For nice $\boldsymbol{T}$, one may perform a change of variables to write $\mu = \mathbb{E}[f(\boldsymbol{X})]$ where $\boldsymbol{X} \sim \mathcal{U}[0,1]^d$.
-For example, if $\boldsymbol{T} \sim \mathcal{N}(\boldsymbol{0},\boldsymbol{I})$, then one may set $f(\boldsymbol{x}) := g(\boldsymbol{\Phi}^{-1}(\boldsymbol{x}))$ for $\boldsymbol{x} \in [0,1]^d$ where $\boldsymbol{\Phi}^{-1}$ is the inverse CDF of a standard normal distribution taken element wise.
-One may then approximate $\mu$ by the $n$ sample average $\hat{\mu} = 1/n \sum_{i=1}^n f(\boldsymbol{x}_i)$ where $\{\boldsymbol{x}_i\}_{i=1}^n$ are chosen so their empirical distribution is ``close'' to the uniform distribution.
-The approximation error $\varepsilon = \lvert \mu - \hat{\mu} \rvert$ is $\mathcal{O}(n^{-1/2})$ when $\{\boldsymbol{x}_i\}_{i=1}^n$ are chosen to be IID standard uniform as in the case of simple Monte Carlo (MC) and almost $\mathcal{O}(n^{-1})$ when $\{\boldsymbol{x}_i\}_{i=1}^n$ are chosen from a low discrepancy sequence as done for Quasi-Monte Carlo (QMC).
-Instead of selecting the number of samples $n$ and then determining the error $\varepsilon$, it is often desirable to select $\varepsilon$ and then adaptively determine $n$.
-From this view, $n$ is $\mathcal{O}(\varepsilon^{-2})$ for MC and almost $\mathcal{O}(\varepsilon^{-1})$ for QMC.
-A more detailed account of MC and QMC can be found in \AGSComment{CITE}.
-
-The (Q)MC pipeline above can be decomposed into $4$ main components:
-\begin{description}
-    \item[\textbf{Generator:}] producing $\{\boldsymbol{x}_i\}_{i=1}^n$ whose empirical distribution is ``close'' to the standard uniform distribution in $d$ dimensions for any $n \geq 1$.
-May produce IID or low discrepancy sequences depending on the problem.
-    \item[\textbf{Measure:}] that defines the distribution of $\boldsymbol{T}$ e.g., uniform, Gaussian, Lebesgue. This definition facilitates choosing a change of variables to determine $f$.
-    \item[\textbf{Integrand:}] $g$, which defines the original integral. When provided knowledge of the true measure, the integrand can also evaluate $f$ via the change of variables.
-    \item[\textbf{Stopping Criterion:}] based on a data-driven error bound, which determines how large $n$ should be to ensure that $\varepsilon$ is less than some user-provided error threshold.
-\end{description}
-A variety of QMC materials and the their categorizations into the above framework are described in the following list.
-\begin{itemize}
-    %\item \cite{LEc2017a} provides a history of random number generation
-    %\item Notorious RANDU from the 1960s \cite{RANDU}, which fails the spectral test
-    %\item VEGAS \cite{Lep78a, Lep21a} importance sampling Monte Carlo algorithm favored by physicists
-    \item ACM low discrepancy point generators \cite{BraFox88,BraFoxNie92,HonHic00a}
-    \item FinDer \cite{PasTra95,FinDer} and BRODA \cite{BRODA20a} targeting quantitative finance
-    \item Korobov cubature and scrambled Sobol' generators in NAG \cite{NAG27} for decades
-    \item Scrambled Sobol' and Halton generators in MATLAB \cite{MAT9.13} since 2008, and fixed a few years later
-    %\item BUGS \cite{BUGSBook, BUGSweb} and Stan \cite{STAN} for Markov Chain Monte Carlo (MCMC)
-    \item LatMRG \cite{LEcCou97}, RNGStreams \cite{LEcEtal02},  SSJ \cite{LEc2002a,SSJ}, and LatNetBuilder \cite{LatNet}
-    \item Low discrepancy generators \cite{FriKel02,FriKelweb}, SamplePack \cite{SamplePack}, \cite{GruWeb},  and MatBuilder \cite{paulin2022}
-    \item Sobol' direction numbers \cite{JoeKuo03,KuoJoe08a,SobolDirection}
-    \item Fast CBC, Magic Point Shop, QMC4PDE and other code since 2004  at \cite{NuyWeb}
-    \item Multi-level software \cite{GilesSoft,GilesQSoft}
-    \item Data-driven error bounds and stopping criteria in GAIL \cite{ChoEtal21a} and QMCPy \cite{QMCPy2020a, ChoEtal22a}
-    \item Uncertainty quantification libraries Dakota \cite{DakotaUsersManual}, UQTk \cite{DebEtal04,UQTk}, and MUQ \cite{MUQ}  have some basic level low discrepancy sampling
-    \item QMC framework in Julia since 2019 \cite{QMCJulia} by Robbe and others
-    \item Scrambled Sobol' in SciPy \cite{virtanen2020scipy} and PyTorch \cite{paszke2019pytorch} since several years ago
-    \item TensorFlow QMC framework \cite{tfqfQMC2021a} since a year or so ago
-    \item CUBA \cite{CUBA}
-\end{itemize}
-
-\section{Choosing a Environment} \label{CDHJS_sec:language_library_environment}
-
-This section discusses the challenges in choosing environments for development and distribution.
-The development environment includes the choice of which languages, architectures, testing methods, and documentation methods to support.
-The deployment environment will often be influenced by the choice of programming languages.
-
-The choice of programming language should be aligned with the communities the team wishes to reach with their software.
-The choice should also account for developer familiarity and ability to quickly prototype ideas.
-A team should align their choice of language to balance the desire to have user-friendly prototyping code versus the desire to have high performance computing (HPC) ``ninja'' code written for production setting.
-To the second point, it is also important to consider which language will best facilitate the parallel and GPU computations the team wishes to support.
-
-Many programming languages come bundled with their own preferred distribution systems.
-Python has PyPI, R has CRAN, Julia has Pkg.
-However, many projects support multiple languages and have complex dependency requirements.
-In such cases, it can be helpful to containerize your application using a service like Docker.
-
-\section{Encouraging Collaborative Development} \label{CDHJS_sec:encouraging_collaborative_developement}
-
-We now discuss ways a team may encourage collaborative development while touching on a few best practices.
-As discussed earlier, collaborative development promotes greater community engagement, ecosystem integration, and software longevity.
-
-The authors have found the Git version control system to be an invaluable for software development.
-Git hosts such as GitHub or Bitbucket may offer teams additional features such as tracking issues, managing discussion threads, and setting up automated workflows for jobs such as testing and documentation compilation.
-
-Accessibility is key to community engagment.
-It is paramount that your package be easy to install, well documented, and tested regularly.
-Aside from static communications, a team should also be engaged in addressing users bugs and be attentive to collaborating with proposed code changes.
-
-It is also important for a development team to make a clear path for collaborators to contributors to the software.
-Templates for issues and pull requests are a good place to start.
-Contributors should also be encouraged to write robust tests for their updates and highlight their new features in a demo for greater accessibility.
-
-\section{Example Collaboration with UMBridge and QMCPy} \label{CDHJS_sec:example_collaboration_umbridge}
-
-This section discusses a collaboration between the UM-Bridge (the UQ and Model Bridge) \AGSComment{CITE} and QMCPy (the Quasi-Monte Carlo Python library) \cite{QMCPy2020a} softwares.
-The UM-Bridge documentation describes its' mission is to, ``provide a unified interface for numerical models that is accessible from virtually any programming language or framework.
-It is primarily intended for coupling advanced models (e.g.
-simulations of complex physical processes) to advanced statistics or optimization methods.'' QMCPy is a high level framework for combining QMC components in a compatible and consistent manner.
-The collaboration between these two packages focused on making the UQ methods of QMCPy compatible with the abstract model representations from UM-Bridge.
-Specifically, UM-Bridge models are made compatible with the QMCPy framework through an integrand wrapper for seamless compatibility with a variety of existing point generators, transformations, and stopping criterion already available in QMCPy.
-
-To exemplify the integration, we focus on quantifying uncertainty propagation of material properties of a cantilevered beam.
-It is assumed the beam is split into $3$ regions of random stiffness dependent on uncertain material properties.
-Given a sample of $3$ stiffness values, the UM-Bridge model provides the displacement at $31$ equidistant nodes along the beam.
-Our task is to quantify the uncertainty in displacement at each of these $31$ locations along the beam.
-
-A fully coded example of this problem is available at \url{https://github.com/QMCSoftware/QMCSoftware/blob/master/demos/umbridge.ipynb}.
-At a high level, one begins by running the UM-Bridge model in a Docker container.
-A model is then initialized in the UM-Bridge Python client which is then provided to QMCPy alongside a low discrepancy point generator and distribution governing the $3$ beam stiffness values.
-QMCPy then performs adaptive QMC cubature to simultaneously approximate the expected displacement at the $31$ locations of interest.
-When the cubature method queries the model, the QMCPy wrapper around UM-Bridge model sends requests to the Docker UM-Bridge model in parallel.
-The load balancing provided by UM-Bridge may be utilized to evaluate the model in parallel, a potentially advantageous feature for models which are costly to evaluate.
-Figure \ref{CDHJSfigumbridge_muq} depicts the $31$ approximate expected displacement values subject to uniform stiffness.
-
-\begin{figure}
-    \centering
-    \includegraphics[height=4cm]{umbridge_muq.png}
-    \caption{QMCPy approximate expected displacement at $31$ locations subject to random uniform stiffness. }
-    \label{CDHJSfigumbridge_muq}
-\end{figure}
-
-\section*{TODO}
-\begin{itemize}
-    \item Flesh out Section \ref{CDHJS_sec:encouraging_collaborative_developement}
-    \item QMC Software Architecture section: better way then description of components and list of softwares?
-    \item For Section \ref{CDHJS_sec:tenets_of_great_software}, the slides also include
-    \begin{itemize}
-        \item  Complete---contain the components or  easily access components in other libraries to solve real, complex problems
-        \item Current---include the latest and best algorithms
-    \end{itemize}
-    which I think fall under the umbrella of software being \emph{integrated} with the community.
-    \item update QMCPy citation \cite{QMCPy2020a}
-    \item figures?
-    \item I have left out or been breif about material from the following slides: QMCPy + FEniCS/Dolfin,  Building a Developer Team for Your Library
-\end{itemize}
-
-Developing software to implement such enhanced MC strategies poses a number of challenges.
-First, a development team must determine the value of the software to the users.
-We discuss why great (Q)MC software in necessary in Sect.\ \ref{CDHJS_sec:why_we_need_software}.
-Second, a team must determine high level objectives for the software; the tenets of great software are described in Sect.\ \ref{CDHJS_sec:tenets_of_great_software}.
-Third, an architecture for the  (Q)MC software should be determined.
-We propose a modularization of the (Q)MC pipeline in Sect.\ \ref{CDHJS_sec:qmc_software_architecture}.
-Next, a team must choose which programming languages, hardware environments, and distribution platforms they wish to support.
-We discuss the nuance and challenges in these decisions in Sect.\ \ref{CDHJS_sec:language_library_environment}.
-Finally, a project which hopes to stand the test of time must draw attention and support from the broader (Q)MC community.
-To this end, we encourage collaborative community development and describe its' best practices in  Sect.\ \ref{CDHJS_sec:encouraging_collaborative_developement}.

### 2023-06-18 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
SC: responded to comments #1 and #3 in referee report.

--- 2023-03-19 MCQMC2022ChallengesQMCSoftwareArticle.tex
@@
 
-\begin{figure}
+\begin{figure}[t]
     \centering
@@
 Stephen Joe and Frances Kuo \cite{JoeKuo03,KuoJoe08a,SobolDirection} have proposed  Sobol' generator direction numbers, which are now widely used, but were not part of early LD sequence software.
-Pierre L'Ecuyer and collaborators have created LatNetBuilder \cite{LatNet} to construct good lattice and polynomial lattice sequences based on user-defined criteria.
+Pierre L'Ecuyer and collaborators have created LatNet Builder \cite{LEcEtal22a} to construct good lattice and polynomial lattice sequences based on user-defined criteria.
 During lunch at MCQMC 2022, it was proposed that we  agree on a consistent format for storing the parameters that define good LD sequences so that they can be shared across software libraries.
+\SCComment{Changed LatNet Builder's reference in response to Referee's \#3, but the citation relies on another reference ArnGilKel22a which is not found}
 
 Users computing solutions in comprehensive software environments, such as NAG, MATLAB, R, SciPy, and TensorFlow, have convenient access to LD sequence generators.
-However, most of these implementations in large libraries would benefit a wider menu of LD offerings, including custom generators from LatNetBuilder or other sources.
+However, most of these implementations in large libraries would benefit a wider menu of LD offerings, including custom generators from LatNet Builder or other sources.
 Some large environments such as  Dakota \cite{DakotaUsersManual} have quite limited LD generators and should expand their offerings.
@@
 
-\begin{figure}
+\begin{figure}[t]
     \centering 
+    \raisebox{0.5cm}{\includegraphics[width=0.4\textwidth]{BeamDrawing.png}}
     \includegraphics[width=0.5\textwidth]{cantileveredbeamwords.eps}
     \includegraphics[width=1\textwidth]{ldparallelbeam.eps}
-    \caption{The expected deflection of a cantilevered beam with random Young's modulus.
+    \caption{The expected deflection (upper left, image taken from~\cite{ParSee22a} with permission \SCComment{to be obtained}) of a cantilevered beam  (upper right, \SCComment{what is $n$ for this figure?}) with random Young's modulus.
 QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.
-For small error tolerances, the time and sample size is $\calO(\varepsilon^{-1})$, and parallel processing gives a speed-up factor of four to five.
+For small error tolerances, the time and sample size is $\calO(\varepsilon^{-1})$ (lower right), and parallel processing gives a speed-up factor of four to five (lower left).
 These plots may be reproduced using the Jupyter notebook \cite{MCQMC2022Figs}.
@@
 In the uncertainty quantification (UQ) space, the recently developed UQ and Model Bridge (UM-Bridge) \cite{umbridge}, connects libraries that generate sequences of model parameters, $\bsx_1, \bsx_2, \ldots$ with libraries that evaluate models, $f(\bsx_1), f(\bsx_2), \ldots$.
-We illustrate in Figure \ref{CDHJSfigbeam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam with random Young's modulus \cite{ParSee22a}, where the randomness is given by $\bsX \sim \calU[1,1.2]^3$.
-The beam position is discretized at $31$ points.
-QMCPy computes the expected deflection of the beam at these points for a sequence of error tolerances using lattice rules.
+
+We illustrate in Figure~\ref{CDHJSfigbeam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam, with a random Young's modulus \SCComment{$m(x)$ of length $L$, assumed to be piecewise constant over three regions, each of which with a uniform distribution of uncertainty~\cite{ParSee22a}, i.e., $m(x) = \left[\begin{matrix} m_1(x), m_2(x), m_3(x) \end{matrix} \right]$ with $m_i(x) \sim \calU[1,1.2]^3$.
+The beam position $x$ is discretized at $N=31$ equidistant points, and
+the vertical deflection $u(x)$ when the beam is subject to an external load $f(x)$ satisfies the Euler-Bernoulli partial differential equation (PDE):
+$$ \frac{\partial^2}{\partial x^2}\left[ \exp(m(x)) \frac{\partial^2  u(x)}{\partial x^2}\right] = f(x),$$
+%where $m(x) = \log E(x)$ is the log of an effective stiffness $E(x)$.  The  
+with the following boundary conditions:
+$$ \left.u(x)\right|_{x=0} = 0, \qquad 
+\left.\frac{\partial u(x)}{\partial x}\right|_{x=0} = 0, \qquad
+\left.\frac{\partial^2 u(x)}{\partial x^2}\right|_{x=L} = 0, \qquad  \left.\frac{\partial^3 u(x)}{\partial x^3}\right|_{x=L} = 0. $$
+This PDE can be discretized with finite differences yielding a linear system $K(m) \hat{u} = \hat{f}$, where $\hat{u}$ and $\hat{f} \in \mathbb{R}^N$. }
+%approximates $u(x)$ and $m(x)$ at the nodes, respectively. 
+%The forward model $F:\mathbb{R}^3 \rightarrow \mathbb{R}^N $ maps $m = \left[\begin{matrix}m_1, m_2, m_3 \end{matrix}\right]^T$ to $\hat{u}$, while $Q : \mathbb{R}^N \rightarrow \mathbb{R}^2 $ selects the estimated solution at the 10th and 25th nodes, i.e., $ Q(\hat{u}) := \left[\begin{matrix} \hat{u}_{10}, \hat{u}_{25} \end{matrix} \right]^T$.
+The objective is to determine the distributions of $Q(F(m))$.
+QMCPy computes the expected deflection of the beam at the node points for a sequence of error tolerances using lattice rules.
 By taking advantage of virtual machines and docker containers, UM-Bridge provides automatic load balancing to facilitate seamless parallel model evaluation.% the evaluation of the beam deflection can be computed in parallel in multiple cores of a single CPU to decrease execution time.
@@
 
+\SCComment{Perhaps we could respond to referee's comment \#2 and \#4 here by framing it as future work.}
+
+\section*{Acknowledgements}\SCComment{Added this section}
+The authors are grateful to the referee's constructive comments.
+
 \bibliographystyle{spmpsci}

### 2023-06-18 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
SC: change f to g to make it less confusing.

@@
 The beam position $x$ is discretized at $N=31$ equidistant points, and
-the vertical deflection $u(x)$ when the beam is subject to an external load $f(x)$ satisfies the Euler-Bernoulli partial differential equation (PDE):
-$$ \frac{\partial^2}{\partial x^2}\left[ \exp(m(x)) \frac{\partial^2  u(x)}{\partial x^2}\right] = f(x),$$
+the vertical deflection $u(x)$ when the beam is subject to a given external load  $g(x)$ satisfies the Euler-Bernoulli partial differential equation (PDE):
+$$ \frac{\partial^2}{\partial x^2}\left[ \exp(m(x)) \frac{\partial^2  u(x)}{\partial x^2}\right] = g(x),$$
 %where $m(x) = \log E(x)$ is the log of an effective stiffness $E(x)$.  The  
@@
 The objective is to determine the distributions of $Q(F(m))$.
-QMCPy computes the expected deflection of the beam at the node points for a sequence of error tolerances using lattice rules.
+QMCPy simulates the random Young's modulus and computes the expected deflection of the beam at the node points for a sequence of error tolerances using lattice rules.
 By taking advantage of virtual machines and docker containers, UM-Bridge provides automatic load balancing to facilitate seamless parallel model evaluation.% the evaluation of the beam deflection can be computed in parallel in multiple cores of a single CPU to decrease execution time.

### 2023-06-18 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
SC: change f to g to make it clearer

@@
 \left.\frac{\partial^2 u(x)}{\partial x^2}\right|_{x=L} = 0, \qquad  \left.\frac{\partial^3 u(x)}{\partial x^3}\right|_{x=L} = 0. $$
-This PDE can be discretized with finite differences yielding a linear system $K(m) \hat{u} = \hat{f}$, where $\hat{u}$ and $\hat{f} \in \mathbb{R}^N$. }
+This PDE can be discretized with finite differences yielding a linear system $K(m) \hat{u} = \hat{g}$, where $\hat{u}$ and $\hat{g} \in \mathbb{R}^N$. }
 %approximates $u(x)$ and $m(x)$ at the nodes, respectively. 

### 2023-06-19 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
SC: Response to comment #3

@@
 
-\section{Summary}
+\section{Summary and Future Work}
 QMC has experienced decades of exciting growth in theory since the late 1950s.
@@
 
-\SCComment{Perhaps we could respond to referee's comment \#2 and \#4 here by framing it as future work.}
+\SCComment{Response to comment \#2: In this paper, we have focused predominantly on employing QMC sampling in estimating the expectations of functions.
+However, it can be extended to other important mathematical problems such as function approximation, simulation of Markov chains,  density estimation, and many MC-based randomized algorithms.
+For instance, Array-RQMC has proven invaluable for maintaining and sorting arrays of Markov chains \cite{l2018sorting,puchhammer2021variance}.
+The SSJ library~\cite{l2002ssj} is a comprehensive tool kit that includes a wide range of QMC point sets, sequences, and randomizations in the ``hups'' and ``mcqmctools'' packages.
+It will be of great interest to the QMCPy community to make connections to these and other well established software in order to extend its applicability to an increasingly wider scope of  problems in computational mathematics and statistics.
+}
+
+\SCComment{Perhaps we could respond to referee's comment \#4 here.}
 
@@
 \bibliographystyle{spmpsci}
-\bibliography{FJH23,FJHown23}
+\bibliography{FJH23,FJHown23,sc}
 

### 2023-06-20 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
SC: add response to comment #4.

@@
 QMCPy simulates the random Young's modulus and computes the expected deflection of the beam at the node points for a sequence of error tolerances using lattice rules.
-By taking advantage of virtual machines and docker containers, UM-Bridge provides automatic load balancing to facilitate seamless parallel model evaluation.% the evaluation of the beam deflection can be computed in parallel in multiple cores of a single CPU to decrease execution time.
+By taking advantage of virtual machines and docker containers, UM-Bridge provides automatic load balancing to facilitate seamless parallel model evaluation.
+% the evaluation of the beam deflection can be computed in parallel in multiple cores of a single CPU to decrease execution time.
 
@@
 The SSJ library~\cite{l2002ssj} is a comprehensive tool kit that includes a wide range of QMC point sets, sequences, and randomizations in the ``hups'' and ``mcqmctools'' packages.
-It will be of great interest to the QMCPy community to make connections to these and other well established software in order to extend its applicability to an increasingly wider scope of  problems in computational mathematics and statistics.
+It will be of great interest to the QMCPy community to make connections to these and other well established software in order to become applicable to an increasingly wider scope of problems and applications of computational mathematics and statistics.
 }
 
-\SCComment{Perhaps we could respond to referee's comment \#4 here.}
+\SCComment{Response to comment \#4:  Python is known for its versatility, but it can be slower than highly optimized implementations in languages like Java or C++.
+For QMCPy developers, it would be beneficial to investigate such underlying interfaces or consider other code acceleration techniques by leveraging Numba, Joblib, or Ray in the future.
+}
+
+\SCComment{Question: Do we have only one referee's report for this paper?}
 

### 2023-06-20 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
SC: add non-code contribution in open source

@@
 
+\SCComment{In QMCPy, we recognize the crucial role of non-code contributors to the project~\cite{K23}.
+Besides coding,  we also need documentation, interface design, testing, community management, and release management.
+These non-code contributions help make the project more  accessible.
+Non-coding roles offer opportunities for individuals with less coding experiences to develop technical and non-technical skills over time.
+Additionally, the emergence of no-code or low-code analytical platform such as SAS Studio and SAS Viya offer a chance for individuals who are less experienced with complex coding to learn and contribute in a more user-friendly environment.
+We need and welcome non-code contributors to work alongside code contributors, enhancing the overall health of the project and the diversity of the community.
+}
+
 \section{Summary and Future Work}
@@
 \SCComment{Response to comment \#4:  Python is known for its versatility, but it can be slower than highly optimized implementations in languages like Java or C++.
-For QMCPy developers, it would be beneficial to investigate such underlying interfaces or consider other code acceleration techniques by leveraging Numba, Joblib, or Ray in the future.
+For QMCPy developers, it would be beneficial to investigate such underlying interfaces or consider other code acceleration techniques by leveraging high-performance libraries such as Numba, JAX, Joblib, or Ray in the future.
 }

### 2023-06-20 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
SC: add low-code tools

@@
 \SCComment{In QMCPy, we recognize the crucial role of non-code contributors to the project~\cite{K23}.
-Besides coding,  we also need documentation, interface design, testing, community management, and release management.
+Besides coding, we also need documentation, data visualization,  testing, user interface or experience design, web content development, community management, and release management.
 These non-code contributions help make the project more  accessible.
 Non-coding roles offer opportunities for individuals with less coding experiences to develop technical and non-technical skills over time.
-Additionally, the emergence of no-code or low-code analytical platform such as SAS Studio and SAS Viya offer a chance for individuals who are less experienced with complex coding to learn and contribute in a more user-friendly environment.
+Additionally, the emergence of no-code or low-code analytical software such as RapidMiner, KNIME, Orange, SAS Studio, or SAS Viya offer a chance for individuals who are less experienced with complex coding to learn and contribute in a more user-friendly environment.
 We need and welcome non-code contributors to work alongside code contributors, enhancing the overall health of the project and the diversity of the community.

### 2023-07-03 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
changes before 7/3

@@
 Non-coding roles offer opportunities for individuals with less coding experiences to develop technical and non-technical skills over time.
-Additionally, the emergence of no-code or low-code analytical software such as RapidMiner, KNIME, Orange, SAS Studio, or SAS Viya offer a chance for individuals who are less experienced with complex coding to learn and contribute in a more user-friendly environment.
-We need and welcome non-code contributors to work alongside code contributors, enhancing the overall health of the project and the diversity of the community.
+Additionally, the emergence of no-code or low-code analytical software such as RapidMiner, KNIME, Orange, SAS Studio, or SAS Viya enable individuals less experienced with complex coding to learn and contribute in a user-friendly environment.
+We need and welcome non-code contributors to work alongside code contributors for QMCPy, enhancing the overall health of the project and the diversity of the community.
 }

### 2023-07-04 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
Aleksei: address referee comment #4
SC: Fix references of GAIL and QMCPy MCQMC 2020

@@
 members of the academic QMC community, developers of open-source and commercial scientific software, and users of a wide range of scientific software as research scientists or data scientists.
-Many of the insights that we have gained have come through our development of the Guaranteed Automatic Integration Library (GAIL)~\cite{ChoEtal22a} in MATLAB and the Quasi-Monte Carlo Python library (QMCPy)~ \cite{QMCPy2020a}.
+Many of the insights that we have gained have come through our development of the Guaranteed Automatic Integration Library (GAIL)~\cite{TonEtAl22a} in MATLAB and the Quasi-Monte Carlo Python library (ChoEtal22a)~ \cite{QMCPy2020a}.
 
@@
 
-Great QMC software eliminates the need for researchers to resort to  ``do-it-yourself'' for established algorithms.  The less code we have to write, the fewer errors.
+Great QMC software eliminates the need for researchers to resort to  ``do-it-yourself'' for established algorithms.  The less code we have to write, the fewer the errors.
 
@@
 
-We next discuss the characteristics of great QMC software outlined in the introduction.  For each characteristic, we identify what is lacking and what might be done to remedy the lack.
+We next discuss the characteristics of great QMC software as outlined in the introduction.  For each characteristic, we identify potential deficiencies and suggest possible remedies.
 
@@
 Intensive computations may need to be rewritten in lower-level languages, like C.
-Code may need to be refactored to avoid excessive calls to memory.
+We have found looped randomization and point generation routines to greatly benefit from low level implementations.
+Code may also need to be refactored to avoid excessive calls to memory.
 As new hardware architectures arise, QMC software should adapt to take advantage of them.

### 2023-07-09 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
SC: obtained permission to add an image in Figure 2 and worked in some comments from Linus.

@@
 \newcommand{\JagsComment}[1]{{\color{violet} Jags: #1}}
+\newcommand{\LSComment}[1]{{\color{blue} Linus Seelinger: #1}}
 \DeclareMathOperator{\disc}{disc}
@@
     \includegraphics[width=1\textwidth]{ldparallelbeam.eps}
-    \caption{The expected deflection (upper left, image taken from~\cite{ParSee22a} with permission \SCComment{to be obtained}) of a cantilevered beam  (upper right, \SCComment{what is $n$ for this figure?}) with random Young's modulus.
+    \caption{The expected deflection (upper left, image reproduced from~\cite{ParSee22a} with permission) of a cantilevered beam (upper right, \SCComment{what is $n$ for this figure?}) with random Young's modulus.
 QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.
@@
 Complex integrands might be best evaluated by libraries that have no direct connection to QMC software libraries.
-In the uncertainty quantification (UQ) space, the recently developed UQ and Model Bridge (UM-Bridge) \cite{umbridge}, connects libraries that generate sequences of model parameters, $\bsx_1, \bsx_2, \ldots$ with libraries that evaluate models, $f(\bsx_1), f(\bsx_2), \ldots$.
+In the uncertainty quantification (UQ) space, the recently developed UQ and Model Bridge (UM-Bridge) \LSComment{\cite{UMBridge23}}, connects libraries that generate sequences of model parameters, $\bsx_1, \bsx_2, \ldots$ with libraries that evaluate models, $f(\bsx_1), f(\bsx_2), \ldots$   \LSComment{and optionally, their derivatives.
+UM-Bridge allows coupling of UQ codes to models regardless of the programming languages.}
 
-We illustrate in Figure~\ref{CDHJSfigbeam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam, with a random Young's modulus \SCComment{$m(x)$ of length $L$, assumed to be piecewise constant over three regions, each of which with a uniform distribution of uncertainty~\cite{ParSee22a}, i.e., $m(x) = \left[\begin{matrix} m_1(x), m_2(x), m_3(x) \end{matrix} \right]$ with $m_i(x) \sim \calU[1,1.2]^3$.
+We illustrate in Figure~\ref{CDHJSfigbeam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam \SCComment{of length~$L$}, with a random Young's modulus \SCComment{$m(x)$, assumed to be piecewise constant over three regions, each of which with a uniform distribution of uncertainty~\cite{ParSee22a}, i.e., $m(x) = \left[\begin{matrix} m_1(x), m_2(x), m_3(x) \end{matrix} \right]$ with $m_i(x) \sim \calU[1,1.2]^3$.
 The beam position $x$ is discretized at $N=31$ equidistant points, and
@@
 The objective is to determine the distributions of $Q(F(m))$.
+\LSComment{The beam model was written in C++ using the MUQ's model framework~\cite{MUQ}. It is one of the UM-Bridge's benchmark problem library, available as a container.}
 QMCPy simulates the random Young's modulus and computes the expected deflection of the beam at the node points for a sequence of error tolerances using lattice rules.
-By taking advantage of virtual machines and docker containers, UM-Bridge provides automatic load balancing to facilitate seamless parallel model evaluation.
-% the evaluation of the beam deflection can be computed in parallel in multiple cores of a single CPU to decrease execution time.
+By taking advantage of \LSComment{clusters and containers}, UM-Bridge provides automatic load balancing to facilitate seamless parallel model evaluation.
+\LSComment{As a result, thread-parallel QMCPy can offload costly model evaluations to remote clusters~\cite{seelinger2023lowering}.}
 
@@
 
-\section*{Acknowledgements}\SCComment{Added this section}
-The authors are grateful to the referee's constructive comments.
+\section*{Acknowledgements}
+\SCComment{The authors are grateful to the referee's constructive comments. We also want to thank Linus Seelinger for feedback and discussion. }
 

### 2023-07-09 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
SC: minor changes.

@@
 members of the academic QMC community, developers of open-source and commercial scientific software, and users of a wide range of scientific software as research scientists or data scientists.
-Many of the insights that we have gained have come through our development of the Guaranteed Automatic Integration Library (GAIL)~\cite{TonEtAl22a} in MATLAB and the Quasi-Monte Carlo Python library (ChoEtal22a)~ \cite{QMCPy2020a}.
+Many of the insights that we have gained have come through our development of the Guaranteed Automatic Integration Library (GAIL)~\SCComment{\cite{TonEtAl22a}} in MATLAB and the Quasi-Monte Carlo Python library (ChoEtal22a)~\SCComment{\cite{QMCPy2020a}}.
 
@@
 Complex integrands might be best evaluated by libraries that have no direct connection to QMC software libraries.
-In the uncertainty quantification (UQ) space, the recently developed UQ and Model Bridge (UM-Bridge) \LSComment{\cite{UMBridge23}}, connects libraries that generate sequences of model parameters, $\bsx_1, \bsx_2, \ldots$ with libraries that evaluate models, $f(\bsx_1), f(\bsx_2), \ldots$   \LSComment{and optionally, their derivatives.
+In the uncertainty quantification (UQ) space, the recently developed UQ and Model Bridge (UM-Bridge) \LSComment{\cite{UMBridge23}}, connects libraries that generate sequences of model parameters, $\bsx_1, \bsx_2, \ldots$ with libraries that evaluate models, $f(\bsx_1), f(\bsx_2), \ldots$   \LSComment{, and their derivatives if applicable.
 UM-Bridge allows coupling of UQ codes to models regardless of the programming languages.}

### 2023-07-12 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
SC: fix a bib item
Fred 2023 Mac 14 Updates

@@
 We summarize the  QMC software that has been developed to date, propose some criteria for developing great QMC software, and suggest some steps toward achieving great software.
-We illustrate these criteria and steps in practice with the Quasi-Monte Carlo Python library (QMCPy), an open-source community software framework, extensible by design with common programming interfaces to an increasing number of existing or emerging QMC libraries developed by the greater community of QMC researchers.
+We illustrate these criteria and steps with the Quasi-Monte Carlo Python library (QMCPy), an open-source community software framework, extensible by design with common programming interfaces to an increasing number of existing or emerging QMC libraries developed by the greater community of QMC researchers.
 }
@@
 members of the academic QMC community, developers of open-source and commercial scientific software, and users of a wide range of scientific software as research scientists or data scientists.
-Many of the insights that we have gained have come through our development of the Guaranteed Automatic Integration Library (GAIL)~\SCComment{\cite{TonEtAl22a}} in MATLAB and the Quasi-Monte Carlo Python library (ChoEtal22a)~\SCComment{\cite{QMCPy2020a}}.
+Many of the insights that we have gained have come through our development of the Guaranteed Automatic Integration Library (GAIL)~\cite{TonEtAl22a} in MATLAB and the Quasi-Monte Carlo Python library (QMCPy)~\cite{QMCPy2020a}.
 
@@
 QMC software has been successful in quantitative finance, uncertainty quantification, and image rendering.
-Unexpectedly good or bad computational results lead to open theoretical questions.
+Unexpectedly good or bad computational results prompt theoretical questions.
 For example, the early application of QMC to high dimensional integrals arising in financial risk \cite{PasTra95} led to a wave of theoretical results on the tractability of integration in weighted spaces \cite{Woz99a,DicEtal14a,NovWoz10a}.
@@
 The objective is to determine the distributions of $Q(F(m))$.
-\LSComment{The beam model was written in C++ using the MUQ's model framework~\cite{MUQ}. It is one of the UM-Bridge's benchmark problem library, available as a container.}
+\LSComment{The beam model was written in C++ using the MUQ's model framework~\cite{MUQ2}. It is one of the UM-Bridge's benchmark problem library, available as a container.}
 QMCPy simulates the random Young's modulus and computes the expected deflection of the beam at the node points for a sequence of error tolerances using lattice rules.

### 2023-07-12 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
Fred 2023 Mac 14 Updates

@@
 
+\bigskip
+
 We next discuss the characteristics of great QMC software as outlined in the introduction.  For each characteristic, we identify potential deficiencies and suggest possible remedies.
@@
 During lunch at MCQMC 2022, it was proposed that we  agree on a consistent format for storing the parameters that define good LD sequences so that they can be shared across software libraries.
-\SCComment{Changed LatNet Builder's reference in response to Referee's \#3, but the citation relies on another reference ArnGilKel22a which is not found}
 
@@
 \subsection{Integrands and Variable Transformations} \label{CDHJS:sec:integrands}
-The original integral or expectation arising from an application is defined in terms of the integrand, $g$, and the non-negative weight, $\lambda$.  For example,
+The original integral or expectation arising from an application is defined in terms of the integrand, $g :\mathcal{T} \to \R$, and the non-negative weight, $\lambda$, as described in \eqref{eq:cubSummary}.
+For example,
 \begin{itemize}
@@
 Such a transformation is typically non-unique, and the choice of a good one is equivalent to importance sampling.
-If $\lambda$ is a PDF for a random variable $\bsT$, then $\mu = \EE[g(\bsT)]$.
-In this case, one might choose $\bsPsi$ such that $\bsT \sim \bsPsi(\bsX)$, in which case $\lambda(\bsPsi(\bsx)) \abs{\partial \bsPsi/\partial\bsx} = 1$ and $f(\bsx) = g(\bsPsi(\bsx))$.
+If $\lambda$ is the PDF for a random variable $\bsT$, then $\mu = \EE[g(\bsT)]$.
+In this case, one might choose $\bsPsi$ such that $\bsT \sim \bsPsi(\bsX)$, $\bsX \sim \calU(0,1)^d$, in which case $\lambda(\bsPsi(\bsx)) \abs{\partial \bsPsi/\partial\bsx} = 1$ and $f(\bsx) = g(\bsPsi(\bsx))$.
 Software such as SciPy and QMCPy facilitate such variable transformations by providing default $\bsPsi$ for common $\bsT$.
@@
     \includegraphics[width=1\textwidth]{ldparallelbeam.eps}
-    \caption{The expected deflection (upper left, image reproduced from~\cite{ParSee22a} with permission) of a cantilevered beam (upper right, \SCComment{what is $n$ for this figure?}) with random Young's modulus.
+    \caption{The expected deflection (upper left, image reproduced from~\cite{ParSee22a} with permission) of a cantilevered beam (upper right) with random Young's modulus.
 QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.
@@
 Complex integrands might be best evaluated by libraries that have no direct connection to QMC software libraries.
-In the uncertainty quantification (UQ) space, the recently developed UQ and Model Bridge (UM-Bridge) \LSComment{\cite{UMBridge23}}, connects libraries that generate sequences of model parameters, $\bsx_1, \bsx_2, \ldots$ with libraries that evaluate models, $f(\bsx_1), f(\bsx_2), \ldots$   \LSComment{, and their derivatives if applicable.
-UM-Bridge allows coupling of UQ codes to models regardless of the programming languages.}
-
-We illustrate in Figure~\ref{CDHJSfigbeam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam \SCComment{of length~$L$}, with a random Young's modulus \SCComment{$m(x)$, assumed to be piecewise constant over three regions, each of which with a uniform distribution of uncertainty~\cite{ParSee22a}, i.e., $m(x) = \left[\begin{matrix} m_1(x), m_2(x), m_3(x) \end{matrix} \right]$ with $m_i(x) \sim \calU[1,1.2]^3$.
+In the uncertainty quantification (UQ) space, the recently developed UQ and Model Bridge (UM-Bridge) \cite{UMBridge23}, connects libraries that generate sequences of model parameters, $\bsx_1, \bsx_2, \ldots$ with libraries that evaluate models, $f(\bsx_1), f(\bsx_2), \ldots$, and their derivatives if applicable.
+UM-Bridge allows coupling of UQ codes to models regardless of the programming languages.
+{eq:cubSummary}
+We illustrate in Figure~\ref{CDHJSfigbeam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam of length~$L$, with a random Young's modulus $m(x)$, assumed to be piecewise constant over three regions, each of which with a uniform distribution of uncertainty~\cite{ParSee22a}, i.e., $m(x) = \left[\begin{matrix} m_1(x), m_2(x), m_3(x) \end{matrix} \right]$ with $m_i(x) \sim \calU[1,1.2]^3$.
 The beam position $x$ is discretized at $N=31$ equidistant points, and
@@
 The objective is to determine the distributions of $Q(F(m))$.
-\LSComment{The beam model was written in C++ using the MUQ's model framework~\cite{MUQ2}. It is one of the UM-Bridge's benchmark problem library, available as a container.}
+The beam model was written in C++ using the MUQ's model framework~\cite{MUQ2}. It is one of the UM-Bridge's benchmark problem library, available as a container.
+
 QMCPy simulates the random Young's modulus and computes the expected deflection of the beam at the node points for a sequence of error tolerances using lattice rules.
-By taking advantage of \LSComment{clusters and containers}, UM-Bridge provides automatic load balancing to facilitate seamless parallel model evaluation.
-\LSComment{As a result, thread-parallel QMCPy can offload costly model evaluations to remote clusters~\cite{seelinger2023lowering}.}
+By taking advantage of clusters and containers, UM-Bridge provides automatic load balancing to facilitate seamless parallel model evaluation.
+As a result, thread-parallel QMCPy can offload costly model evaluations to remote clusters~\cite{seelinger2023lowering}.
 
@@
 Some of these stopping criteria for absolute error criteria, \eqref{CDHJS_eq:error_tol},  have been extended to relative error criteria or a combination of absolute and relative error criteria, as well as stopping criteria for approximating functions of several integrals \cite{Jia16a,GilJim16b,HicEtal17a,JagSor23a}.
+
+\bigskip
 
@@
 Trade-offs will be made between efficiency and convenience.
+For example, as noted by a referee, QMCPy is not particularly fast at generating tens of thousands of replicates of randomized QMC sample means in comparison to SSJ, which is based on Java.
+Such cases may point to the need for a Python layer over a faster implementation.
 
@@
 
-\SCComment{In QMCPy, we recognize the crucial role of non-code contributors to the project~\cite{K23}.
+We recognize the crucial role of non-code contributors to the project~\cite{K23}.
 Besides coding, we also need documentation, data visualization, testing, user interface or experience design, web content development, community management, and release management.
@@
 Additionally, the emergence of no-code or low-code analytical software such as RapidMiner, KNIME, Orange, SAS Studio, or SAS Viya enable individuals less experienced with complex coding to learn and contribute in a user-friendly environment.
-We need and welcome non-code contributors to work alongside code contributors for QMCPy, enhancing the overall health of the project and the diversity of the community.
-}
+%We need and welcome non-code contributors to work alongside code contributors for QMCPy, enhancing the overall health of the project and the diversity of the community.
 
@@
 
-\SCComment{Response to comment \#2: In this paper, we have focused predominantly on employing QMC sampling in estimating the expectations of functions.
-However, it can be extended to other important mathematical problems such as function approximation, simulation of Markov chains,  density estimation, and many MC-based randomized algorithms.
-For instance, Array-RQMC has proven invaluable for maintaining and sorting arrays of Markov chains \cite{l2018sorting,puchhammer2021variance}.
+In this paper, we have focused predominantly on employing QMC sampling in estimating the expectations of functions.
+However, it can be extended to other important mathematical problems such as function approximation \cite{Kmmerer2015, PotSch21}, simulation of Markov chains \cite{l2018sorting,puchhammer2021variance},  density estimation \cite{LEcuyer2022b}.
 The SSJ library~\cite{l2002ssj} is a comprehensive tool kit that includes a wide range of QMC point sets, sequences, and randomizations in the ``hups'' and ``mcqmctools'' packages.
 It will be of great interest to the QMCPy community to make connections to these and other well established software in order to become applicable to an increasingly wider scope of problems and applications of computational mathematics and statistics.
-}
 
-\SCComment{Response to comment \#4:  Python is known for its versatility, but it can be slower than highly optimized implementations in languages like Java or C++.
+Python is known for its versatility, but it can be slower than highly optimized implementations in languages like Java or C++.
 For QMCPy developers, it would be beneficial to investigate such underlying interfaces or consider other code acceleration techniques by leveraging high-performance libraries such as Numba, JAX, Joblib, or Ray in the future.
-}
-
-\SCComment{Question: Do we have only one referee's report for this paper?}
 
 \section*{Acknowledgements}
-\SCComment{The authors are grateful to the referee's constructive comments. We also want to thank Linus Seelinger for feedback and discussion. }
+The authors are grateful to the referee's constructive comments.
+We also want to thank Linus Seelinger for feedback and discussion.
+The authors acknowledge many students and collaborators who have provided insight in our journey to develop great QMC software.
 

### 2023-07-12 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
Updates from Overleaf

@@
 \left.\frac{\partial^2 u(x)}{\partial x^2}\right|_{x=L} = 0, \qquad  \left.\frac{\partial^3 u(x)}{\partial x^3}\right|_{x=L} = 0. $$
-This PDE can be discretized with finite differences yielding a linear system $K(m) \hat{u} = \hat{g}$, where $\hat{u}$ and $\hat{g} \in \mathbb{R}^N$. }
+This PDE can be discretized with finite differences yielding a linear system $K(m) \hat{u} = \hat{g}$, where $\hat{u}$ and $\hat{g} \in \mathbb{R}^N$. 
 %approximates $u(x)$ and $m(x)$ at the nodes, respectively.
@@
 Trade-offs will be made between efficiency and convenience.
-For example, as noted by a referee, QMCPy is not particularly fast at generating tens of thousands of replicates of randomized QMC sample means in comparison to SSJ, which is based on Java.
+As noted by a referee, QMCPy is not particularly fast at generating tens of thousands of replicates of randomized QMC sample means in comparison to SSJ, which is based on Java.
 Such cases may point to the need for a Python layer over a faster implementation.
@@
 
-Python is known for its versatility, but it can be slower than highly optimized implementations in languages like Java or C++.
-For QMCPy developers, it would be beneficial to investigate such underlying interfaces or consider other code acceleration techniques by leveraging high-performance libraries such as Numba, JAX, Joblib, or Ray in the future.
+Although Python is popular and versatile, but it can be slower than highly optimized implementations in languages like Java or C++.
+Developers of QMC software in Python  developers would do well to investigate interfaces to optimized code in other languages or consider other code acceleration techniques provided by libraries such as Numba, JAX, Joblib, or Ray.
+With time, languages rise and fall in popularity, and those who wish to provide great QMC software may need to pivot to new software environments.
 

### 2023-07-12 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
updates from today's meeting

@@
 
-%%%%%%% Author sign off on this first submission, last revision 3/19/2023
-% Sou-Cheng Choi signed off on 3/17/20223
-% Yuhan Ding signed off on 3/19/2023
-% Fred J. Hickernell signed off on 3/17/2023
-% Jagadeeswaran Rathinavel signed off on 3/17/2023
-% Aleksei Sorokin signed off on 3/17/2023
+%%%%%%% Author sign off on this revised submission, last revision 7/12/2023
+% Sou-Cheng Choi signed off on 
+% Yuhan Ding signed off on 
+% Fred J. Hickernell signed off on 
+% Jagadeeswaran Rathinavel signed off on
+% Aleksei Sorokin signed off on
 
@@
     \includegraphics[width=1\textwidth]{ldparallelbeam.eps}
-    \caption{The expected deflection (upper left, image reproduced from~\cite{ParSee22a} with permission) of a cantilevered beam (upper right) with random Young's modulus.
+    \caption{The upper left shows the piecewise log effective stiffness (image reproduced from~\cite{ParSee22a} with permission).
+The expected deflection of a cantilevered beam with random log effective stiffness is shown in the upper right.
 QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.
@@
 {eq:cubSummary}
-We illustrate in Figure~\ref{CDHJSfigbeam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam of length~$L$, with a random Young's modulus $m(x)$, assumed to be piecewise constant over three regions, each of which with a uniform distribution of uncertainty~\cite{ParSee22a}, i.e., $m(x) = \left[\begin{matrix} m_1(x), m_2(x), m_3(x) \end{matrix} \right]$ with $m_i(x) \sim \calU[1,1.2]^3$.
+We illustrate in Figure~\ref{CDHJSfigbeam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam of length~$L$, with a random log effective stiffness $m(\cdot)$, assumed to be piecewise constant over three regions, each of which with a uniform distribution of uncertainty~\cite{ParSee22a}, i.e., $m(x) = \left[\begin{matrix} m_1(x), m_2(x), m_3(x) \end{matrix} \right]$ with $m_i(x) \sim \calU[1,1.2]^3$.
 The beam position $x$ is discretized at $N=31$ equidistant points, and

### 2023-07-12 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
aleksei updates and sign off

@@
 % Jagadeeswaran Rathinavel signed off on
-% Aleksei Sorokin signed off on
+% Aleksei Sorokin signed off on 7/12/2023
 
@@
     \includegraphics[width=1\textwidth]{ldparallelbeam.eps}
-    \caption{The upper left shows the piecewise log effective stiffness (image reproduced from~\cite{ParSee22a} with permission).
-The expected deflection of a cantilevered beam with random log effective stiffness is shown in the upper right.
-QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.
-For small error tolerances, the time and sample size is $\calO(\varepsilon^{-1})$ (lower right), and parallel processing gives a speed-up factor of four to five (lower left).
+    \caption{The upper left plot shows a piecewise log effective stiffness (image reproduced from~\cite{ParSee22a} with permission).
+The upper right plot shows the expected deflection of a cantilevered beam with random log effective stiffness.
+In the lower plots, QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.
+For small error tolerances, the time and sample size are $\calO(\varepsilon^{-1})$ and parallel processing gives a speed-up factor of four to five.
 These plots may be reproduced using the Jupyter notebook \cite{MCQMC2022Figs}.
@@
 UM-Bridge allows coupling of UQ codes to models regardless of the programming languages.
-{eq:cubSummary}
-We illustrate in Figure~\ref{CDHJSfigbeam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam of length~$L$, with a random log effective stiffness $m(\cdot)$, assumed to be piecewise constant over three regions, each of which with a uniform distribution of uncertainty~\cite{ParSee22a}, i.e., $m(x) = \left[\begin{matrix} m_1(x), m_2(x), m_3(x) \end{matrix} \right]$ with $m_i(x) \sim \calU[1,1.2]^3$.
+
+We illustrate in Figure~\ref{CDHJSfigbeam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam of length~$L$, with a random log effective stiffness $m(\cdot)$.
+Following ~\cite{ParSee22a}, we assume $m(\cdot)$ is piecewise constant over three regions and the three constants take a $\calU[1,1.2]^3$ distribution.
+%, i.e., $m(x) = \left[\begin{matrix} m_1(x), m_2(x), m_3(x) \end{matrix} \right]$ with $m_i(x) \sim \calU[1,1.2]^3$.  
 The beam position $x$ is discretized at $N=31$ equidistant points, and
@@
 The objective is to determine the distributions of $Q(F(m))$.
-The beam model was written in C++ using the MUQ's model framework~\cite{MUQ2}. It is one of the UM-Bridge's benchmark problem library, available as a container.
+The beam model was written in C++ using the MUQ's model framework~\cite{MUQ2}. %It is one of the UM-Bridge's benchmark problem library, available as a container.
 
@@
 
-Although Python is popular and versatile, but it can be slower than highly optimized implementations in languages like Java or C++.
-Developers of QMC software in Python  developers would do well to investigate interfaces to optimized code in other languages or consider other code acceleration techniques provided by libraries such as Numba, JAX, Joblib, or Ray.
+Although Python is popular and versatile, it can be slower than highly optimized implementations in languages like Java or C++.
+Developers of QMC software in Python would do well to investigate interfaces to optimized code in other languages or consider other code acceleration techniques provided by libraries such as Numba, JAX, Joblib, or Ray.
 With time, languages rise and fall in popularity, and those who wish to provide great QMC software may need to pivot to new software environments.

### 2023-07-15 response_to_referee.tex
SC: Responses to reviewer

\documentclass{article}[12pt]
\usepackage[letterpaper, left=0.8in, top=0.7in, right=0.8in, bottom=1in]{geometry}

\usepackage{url}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=blue}
\usepackage{xcolor}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsthm}

\title{Response to Reviewer Comments for  ``Challenges in Developing Great Quasi-Monte Carlo Software''}
\author{Sou-Cheng T.~Choi \and Yuhan Ding \and Fred J. Hickernell \and Jagadeeswaran  Rathinavel \and Aleksei G. Sorokin}
\date{MCQMC 2022
   \\[1ex]  Authors' response to the reviewer
   \\[1ex] \today}

\begin{document}

\maketitle

We appreciate your insightful comments.  In the following, we recap each comment followed by our response summary. In addition, we have added a few changes in the manuscript.

\begin{enumerate}
    \item \textbf{Recap of comment:} \textit{The example on page 6 is not very clear to me. What is the meaning of the ``dscretize
beam position'' here? Where can we see that? What is $f$? I think there should be a
bit more explanations.}

\textbf{Our response:} We agree that the example on page 6 could have additional clarification. In the revised manuscript, we have added an image in Fig.~2 and  substantially more details on the problem.

    \item \textbf{Recap of comment:} \textit{The paper is essentially all about estimating the expectation of a function by QMC sampling.
But (R)QMC is also used for function approximation, simulating Markov chains,
density estimation, etc. This would required extra (connected) software tools. For example
to maintain and sort and array of Markov chains in Array-RQMC [102, 103].
A lot of this is actually implemented in the SSJ library [101], which also implements
many QMC point sets, sequences, and randomizations in the “hups” and “mcqmctools”
packages. Maybe this could be mentioned briefly somewhere.}

\textbf{Our response:} In Section 8, we have added these additional problems and important tools as future work for any QMC software, including QMCPy.

    \item \textbf{Recap of comment:} \textit{Refer. [9]: Maybe the authors can give the reference to the MCQMC 2020 paper on
LatNet Builder instead, because it better explains what the software is doing.}

       \textbf{Our response:} We have replaced the original reference with the suggested one, which is now listed as reference [37] in the revised manuscript.

    \item \textbf{Recap of comment:} \textit{Section 5: Although Python is very popular and easy to use, perhaps it should be
recognized explicitly that it is not the best tool if we want speed. I made the following
simple experiment to compare the speeds of qmcpy with that of SSJ. We want to
integrate the simple function $f(u_1, u_2) = u_1 + u_2 - 1$ over the unit square, using
RQMC with Sobol points $+$ a linear scrambling $+$ a random digital shift, say with
$n = 214$ points. I computed $20,000$ independent replicates of the RQMC estimator,
with independent randomizations. With SSJ, this took 4.7 seconds on my Thinkpad
laptop. With qmcpy, it took about two minutes, which is about 25 times more. I
am not asking that the authors report this experiment in their paper, but I think
they should mention that Python can be significantly slower than a good Java or
\texttt{C++} implementation. One possibility would be to put a Python layer over a fast implementation.
}

    	    \textbf{Our response:} We have added encouragement to rewrite looped point generating or randomized routines into low level languages for increased efficiency. Your example of SSJ outperforming QMCPy when getting thousands of scrambles is also mentioned. Additional encouragement has also been added to the concluding section.

    \item  \textbf{Other major change in the manuscript:}
    %\begin{enumerate}
    %\item  
    At the end of Section 7, we have highlighted the important role of non-code contributors in high-quality QMC software.
    %\end{enumerate}

\end{enumerate}

\end{document}

### 2023-07-15 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
SC: Responses to reviewer

@@
 % Fred J. Hickernell signed off on 
-% Jagadeeswaran Rathinavel signed off on
+% Jagadeeswaran Rathinavel signed off on 7/13/2023
 % Aleksei Sorokin signed off on 7/12/2023
@@
 Although Python is popular and versatile, it can be slower than highly optimized implementations in languages like Java or C++.
-Developers of QMC software in Python would do well to investigate interfaces to optimized code in other languages or consider other code acceleration techniques provided by libraries such as Numba, JAX, Joblib, or Ray.
+Developers of QMC software in Python would do well to investigate interfaces to optimize code in other languages or consider other code acceleration techniques provided by libraries such as Numba, JAX, Joblib, or Ray.
 With time, languages rise and fall in popularity, and those who wish to provide great QMC software may need to pivot to new software environments.

### 2023-07-15 response_to_referee.tex
SC: fix some minor issues in responses.

@@
 \begin{enumerate}
-    \item \textbf{Recap of comment:} \textit{The example on page 6 is not very clear to me. What is the meaning of the ``dscretize
+    \item \textbf{Comment:} \textit{The example on page 6 is not very clear to me. What is the meaning of the ``dscretize
 beam position'' here? Where can we see that? What is $f$? I think there should be a
@@
 
-    \item \textbf{Recap of comment:} \textit{The paper is essentially all about estimating the expectation of a function by QMC sampling.
+    \item \textbf{Comment:} \textit{The paper is essentially all about estimating the expectation of a function by QMC sampling.
 But (R)QMC is also used for function approximation, simulating Markov chains,
@@
 
-    \item \textbf{Recap of comment:} \textit{Refer. [9]: Maybe the authors can give the reference to the MCQMC 2020 paper on
+    \item \textbf{Comment:} \textit{Refer.\ [9]: Maybe the authors can give the reference to the MCQMC 2020 paper on
 LatNet Builder instead, because it better explains what the software is doing.}
@@
 
-    \item \textbf{Recap of comment:} \textit{Section 5: Although Python is very popular and easy to use, perhaps it should be
+    \item \textbf{Comment:} \textit{Section 5: Although Python is very popular and easy to use, perhaps it should be
 recognized explicitly that it is not the best tool if we want speed. I made the following
@@
 they should mention that Python can be significantly slower than a good Java or
-\texttt{C++} implementation. One possibility would be to put a Python layer over a fast implementation.
+\text{C++} implementation. One possibility would be to put a Python layer over a fast implementation.
 }

### 2023-07-17 response_to_referee.tex
Updates from Overleaf

@@
 
-\textbf{Our response:} In Section 8, we have added these additional problems and important tools as future work for any QMC software, including QMCPy.
+\textbf{Our response:} In Section 8, we have added these additional problems, summarized some results so far, and proposed future work in this direction for  QMC software, including QMCPy.
 

### 2023-07-17 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
SC: Signed off
Updates from Overleaf

@@
 %%%%%%% Author sign off on this revised submission, last revision 7/12/2023
-% Sou-Cheng Choi signed off on 
+% Sou-Cheng Choi signed off on 7/15/2023
 % Yuhan Ding signed off on 
@@
 \begin{enumerate}
-    \item Draw a sequence of $n$ low-discrepancy (LD) \cite{Nie92, SloJoe94, DicEtal14a, DicEtal22a} nodes,
+    \item Draw a sequence of $n$ low discrepancy (LD) \cite{Nie92, SloJoe94, DicEtal14a, DicEtal22a} nodes,
     $\bsx_1,\dots,\bsx_n$ that mimic $\calU[0,1]^d$.
@@
 \end{enumerate}
+
 However, the practice of QMC is often more complicated.
@@
 \section{Sustainable} \label{CDHJS_sec:sustainable}
-Great QMC software libraries should be alive.
+Great QMC software libraries should be awake, not dormant.
 They should be under active development or merged into other libraries that are actively maintained.

### 2023-10-19 MCQMC2022ChallengesQMCSoftwareArticle_revision_2_toeditors.tex
Fred 2023 Mac 14 Updates

--- 2023-07-17 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
@@
 %\input FJHDef.tex
-\newcommand{\AGSComment}[1]{{\color{red} #1}}
-\newcommand{\FredComment}[1]{{\color{blue} Fred: #1}}
-\newcommand{\SCComment}[1]{{\color{green} Sou-Cheng: #1}}
-\newcommand{\YDComment}[1]{{\color{orange} Yuhan: #1}}
-\newcommand{\JagsComment}[1]{{\color{violet} Jags: #1}}
-\newcommand{\LSComment}[1]{{\color{blue} Linus Seelinger: #1}}
-\DeclareMathOperator{\disc}{disc}
-\newcommand{\FJHNote}[1]{{\color{blue}Fred: #1}}
+%\newcommand{\AGSComment}[1]{{\color{red} #1}}
+%\newcommand{\FredComment}[1]{{\color{blue} Fred: #1}}
+%\newcommand{\SCComment}[1]{{\color{green} Sou-Cheng: #1}}
+%\newcommand{\YDComment}[1]{{\color{orange} Yuhan: #1}}
+%\newcommand{\JagsComment}[1]{{\color{violet} Jags: #1}}
+%\newcommand{\LSComment}[1]{{\color{blue} Linus Seelinger: #1}}
+%\DeclareMathOperator{\disc}{disc}
+%\newcommand{\FJHNote}[1]{{\color{blue}Fred: #1}}
 
@@
     \item Estimate the true mean, $\mu$, by the sample mean,
-    \begin{equation}
+    \begin{equation*}
         \hat{\mu}_n := \frac{1}{n} \sum_{i=1}^n f(\bsx_i).
-        \label{eq:mc_average}
-    \end{equation}
+        %\label{CDHJSeq:mc_average}
+    \end{equation*}
 \end{enumerate}
@@
 \begin{gather}
-\label{eq:cubSummary}
+\label{CDHJSeq:cubSummary}
 	\mu : = \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  = \bbE[f(\bsX)] = \int_{[0,1]^d} f(\bsx)  \, \D \bsx \approx \frac 1n \sum_{i=1}^{n} f(\bsx_i) =: \hat{\mu}_n, \\
@@
 \subsection{Integrands and Variable Transformations} \label{CDHJS:sec:integrands}
-The original integral or expectation arising from an application is defined in terms of the integrand, $g :\mathcal{T} \to \R$, and the non-negative weight, $\lambda$, as described in \eqref{eq:cubSummary}.
+The original integral or expectation arising from an application is defined in terms of the integrand, $g :\mathcal{T} \to \R$, and the non-negative weight, $\lambda$, as described in \eqref{CDHJSeq:cubSummary}.
 For example,

### 2023-10-19 HickernellToTheMCQMC2022EditorsVersion2/MCQMC2022ChallengesQMCSoftwareArticle_revision_2_toeditors.tex
Fred 2023 Mac M2 Updates

--- 2023-10-19 MCQMC2022ChallengesQMCSoftwareArticle_revision_2_toeditors.tex
@@
 \usepackage[hyphens]{url}
-\providecommand{\HickernellFJ}{Hickernell\xspace}
-\graphicspath{{ProgramsImages/}}
+%\providecommand{\HickernellFJ}{Hickernell\xspace}
+%\graphicspath{{ProgramsImages/}}
 
@@
 
-\bibliographystyle{spmpsci}
-\bibliography{FJH23,FJHown23,sc}
+%\bibliographystyle{spmpsci}
+%\bibliography{FJH23,FJHown23,sc}
+
+%\def\Ignore#1{}\def\notesupp#1{}\providecommand{\HickernellFJ}{Hickernell\xspace}\def\Ignore#1{}\def\notesupp#1{}\providecommand{\HickernellFJ}{Hickernell\xspace}
+\begin{thebibliography}{10}
+	\providecommand{\url}[1]{{#1}}
+	\providecommand{\urlprefix}{URL }
+	\expandafter\ifx\csname urlstyle\endcsname\relax
+	\providecommand{\doi}[1]{DOI~\discretionary{}{}{}#1}\else
+	\providecommand{\doi}{DOI~\discretionary{}{}{}\begingroup
+		\urlstyle{rm}\Url}\fi
+
+	\bibitem{QMCJulia}
+	{SciML QuasiMonteCarlo.jl} (2022).
+	\newblock \urlprefix\url{https://github.com/SciML/QuasiMonteCarlo.jl}
+
+	\bibitem{DakotaUsersManual}
+	Adams, B.M., Bohnhoff, W.J., Dalbey, K.R., Ebeida, M.S., Eddy, J.P., Eldred,
+	M.S., Hooper, R.W., Hough, P.D., Hu, K.T., Jakeman, J.D., Khalil, M., Maupin,
+	K.A., Monschke, J.A., Ridgway, E.M., Rushdi, A.A., Seidl, D.T., Stephens,
+	J.A., Swiler, L.P., Tran, A., Winokur, J.G.: Dakota, A Multilevel Parallel
+	Object-Oriented Framework for Design Optimization, Parameter Estimation,
+	Uncertainty Quantification, and Sensitivity Analysis: Version 6.16 User's
+	Manual.
+	\newblock Sandia National Laboratories (2022)
+
+	\bibitem{osti_1846008}
+	Bernholdt, D.E., Cary, J., Heroux, M., McInnes, L.C.: The science of
+	scientific-software development and use  (2022).
+	\newblock \doi{10.2172/1846008}.
+	\newblock \urlprefix\url{https://www.osti.gov/biblio/1846008}
+
+	\bibitem{BraFox88}
+	Bratley, P., Fox, B.L.: Algorithm 659: Implementing {Sobol's} quasirandom
+	sequence generator.
+	\newblock ACM Trans.\ Math.\ Software \textbf{14}, 88--100 (1988)
+
+	\bibitem{BraFoxNie92}
+	Bratley, P., Fox, B.L., Niederreiter, H.: Implementation and tests of
+	low-discrepancy sequences.
+	\newblock ACM Trans.\ Model.\ Comput.\ Simul. \textbf{2}, 195--213 (1992)
+
+	\bibitem{MCQMC2022Figs}
+	Choi, S.C.T., Ding, Y., Hickernell, F.J., Jagadeeswaran, R., Sorokin, A.G.:
+	{MCQMC} 2022 figures.
+	\newblock
+	\urlprefix\url{https://github.com/QMCSoftware/QMCSoftware/tree/develop/demos/talk_paper_demos/MCQMC2022_Article_Figures}
+
+	\bibitem{QMCPy2020a}
+	Choi, S.C.T., Hickernell, F.J., Jagadeeswaran, R., McCourt, M., Sorokin, A.:
+	{QMCPy}: A quasi-{M}onte {C}arlo {P}ython library (versions 1--1.4) (2023).
+	\newblock \doi{10.5281/zenodo.3964489}.
+	\newblock \urlprefix\url{https://qmcsoftware.github.io/QMCSoftware/}
+
+	\bibitem{ChoEtal22a}
+	Choi, S.C.T., Hickernell, F.J., Jagadeeswaran, R., McCourt, M.J., Sorokin,
+	A.G.: Quasi-{M}onte {C}arlo software.
+	\newblock In: A.~Keller (ed.) {M}onte {C}arlo and Quasi-{M}onte {C}arlo
+	Methods: {MCQMC}, {O}xford, England, {A}ugust 2020, Springer Proceedings in
+	Mathematics and Statistics, pp. 23--50. Springer, Cham (2022).
+	\newblock \url{https://arxiv.org/abs/2102.07833}
+
+	\bibitem{DicEtal22a}
+	Dick, J., Kritzer, P., Pillichshammer, F.: Lattice Rules: {N}umerical
+	Integration, Approximation, and Discrepancy.
+	\newblock Springer Series in Computational Mathematics. Springer Cham (2022).
+	\newblock \doi{https://doi.org/10.1007/978-3-031-09951-9}
+
+	\bibitem{DicEtal14a}
+	Dick, J., Kuo, F., Sloan, I.H.: High dimensional integration --- the
+	{Q}uasi-{M}onte {C}arlo way.
+	\newblock Acta Numer. \textbf{22}, 133--288 (2013).
+	\newblock \doi{10.1017/S0962492913000044}
+
+	\bibitem{K23}
+	Finley, K.: Non-code contributions are the secret to open source success.
+	\newblock The ReadME Project, GitHub (2023).
+	\newblock
+	\urlprefix\url{https://github.com/readme/featured/open-source-non-code-contributions}
+
+	\bibitem{FriKelweb}
+	Friedel, I., Keller, A.: libseq -- low discrepancy sequence library (2001).
+	\newblock \urlprefix\url{http://www.multires.caltech.edu/software/libseq/}
+
+	\bibitem{FriKel02}
+	Friedel, I., Keller, A.: Fast generation of randomized low-discrepancy point
+	sets.
+	\newblock In: K.T. Fang, F.J. Hickernell, H.~Niederreiter (eds.) {M}onte
+	{C}arlo and Quasi-{M}onte {C}arlo Methods 2000, pp. 257--273.
+	Springer-Verlag, Berlin (2002)
+
+	\bibitem{GelEtal13}
+	Gelman, A., Carlin, J.B., Stern, H.S., Dunson, D.B., Vehtari, A., Rubin, D.B.:
+	Bayesian Data Analysis.
+	\newblock CRC Texts in Statistical Science. Chapman \& Hall (2013)
+
+	\bibitem{Gla03}
+	Glasserman, P.: {M}onte {C}arlo Methods in Financial Engineering,
+	\emph{Applications of Mathematics}, vol.~53.
+	\newblock Springer-Verlag, New York (2004)
+
+	\bibitem{GruWeb}
+	Gr\"unschlo{\ss}, L.: Leonhard gr\"unschlo{\ss}'s webpage.
+	\newblock \urlprefix\url{https://gruenschloss.org}
+
+	\bibitem{HicEtal18a}
+	Hickernell, F.J., Choi, S.C.T., Jiang, L., {Jim\'enez Rugama}, {\relax
+		Ll}.A.: {M}onte {C}arlo simulation, automatic stopping criteria for.
+	\newblock In: M.~Davidian, B.~Everitt, R.S. Kenett, G.~Molenberghs,
+	W.~Piegorsch, F.~Ruggeri (eds.) Wiley StatsRef-Statistics Reference Online.
+	John Wiley \& Sons Ltd. (2018).
+	\newblock \doi{10.1002/9781118445112.stat08035}
+
+	\bibitem{HicEtal14a}
+	Hickernell, F.J., Jiang, L., Liu, Y., Owen, A.B.: Guaranteed conservative
+	fixed width confidence intervals via {M}onte {C}arlo sampling.
+	\newblock In: J.~Dick, F.Y. Kuo, G.W. Peters, I.H. Sloan (eds.) {M}onte {C}arlo
+	and Quasi-{M}onte {C}arlo Methods 2012, \emph{Springer Proceedings in
+		Mathematics and Statistics}, vol.~65, pp. 105--128. Springer-Verlag, Berlin
+	(2013).
+	\newblock \doi{10.1007/978-3-642-41095-6}
+
+	\bibitem{HicJim16a}
+	Hickernell, F.J., {Jim\'enez Rugama}, {\relax Ll}.A.: Reliable adaptive
+	cubature using digital sequences.
+	\newblock In: R.~Cools, D.~Nuyens (eds.) {M}onte {C}arlo and Quasi-{M}onte
+	{C}arlo Methods: {MCQMC}, {L}euven, {B}elgium, {A}pril 2014, \emph{Springer
+		Proceedings in Mathematics and Statistics}, vol. 163, pp. 367--383.
+	Springer-Verlag, Berlin (2016).
+	\newblock ArXiv:1410.8615 [math.NA]
+
+	\bibitem{HicEtal17a}
+	Hickernell, F.J., {Jim\'enez Rugama}, {\relax Ll}.A., Li, D.: Adaptive
+	quasi-{M}onte {C}arlo methods for cubature.
+	\newblock In: J.~Dick, F.Y. Kuo, H.~Wo\'zniakowski (eds.) Contemporary
+	Computational Mathematics --- a celebration of the 80th birthday of {I}an
+	{S}loan, pp. 597--619. Springer-Verlag (2018).
+	\newblock \doi{10.1007/978-3-319-72456-0}
+
+	\bibitem{QRNG2020}
+	Hofert, M., Lemieux, C.: qrng {R} package (2017).
+	\newblock \urlprefix\url{https://cran.r-project.org/web/packages/qrng/qrng.pdf}
+
+	\bibitem{HonHic00a}
+	Hong, H.S., Hickernell, F.J.: Algorithm 823: Implementing scrambled digital
+	nets.
+	\newblock ACM Trans.\ Math.\ Software \textbf{29}, 95--109 (2003).
+	\newblock \doi{10.1145/779359.779360}
+
+	\bibitem{RatHic19a}
+	Jagadeeswaran, R., Hickernell, F.J.: Fast automatic {B}ayesian cubature
+	using lattice sampling.
+	\newblock Stat.\ Comput. \textbf{29}, 1215--1229 (2019).
+	\newblock \doi{10.1007/s11222-019-09895-9}
+
+	\bibitem{JagHic22a}
+	Jagadeeswaran, R., Hickernell, F.J.: Fast automatic {B}ayesian cubature
+	using {S}obol' sampling.
+	\newblock In: Z.~Botev, A.~Keller, C.~Lemieux, B.~Tuffin (eds.) Advances in
+	Modeling and Simulation: {F}estschrift in Honour of {P}ierre {L'E}cuyer, pp.
+	301--318. Springer, Cham (2022).
+	\newblock \doi{10.1007/978-3-031-10193-9\_15}
+
+	\bibitem{Jia16a}
+	Jiang, L.: Guaranteed adaptive {M}onte {C}arlo methods for estimating means of
+	random variables.
+	\newblock Ph.D. thesis, Illinois Institute of Technology (2016)
+
+	\bibitem{GilJim16b}
+	{Jim\'enez Rugama}, {\relax Ll}.A., Gilquin, L.: Reliable error estimation for
+	{S}obol' indices.
+	\newblock Statistics and Computing \textbf{28}, 725--738 (2018).
+	\newblock \doi{10.1007/s11222-017-9759-1}
+
+	\bibitem{JimHic16a}
+	{Jim\'enez Rugama}, {\relax Ll}.A., Hickernell, F.J.: Adaptive
+	multidimensional integration based on rank-1 lattices.
+	\newblock In: R.~Cools, D.~Nuyens (eds.) {M}onte {C}arlo and Quasi-{M}onte
+	{C}arlo Methods: {MCQMC}, {L}euven, {B}elgium, {A}pril 2014, \emph{Springer
+		Proceedings in Mathematics and Statistics}, vol. 163, pp. 407--422.
+	Springer-Verlag, Berlin (2016).
+	\newblock ArXiv:1411.1966
+
+	\bibitem{SobolDirection}
+	Joe, S., Kuo, F.Y.: Sobol sequence generator.
+	\newblock \urlprefix\url{https://web.maths.unsw.edu.au/~fkuo/sobol/}
+
+	\bibitem{JoeKuo03}
+	Joe, S., Kuo, F.Y.: Remark on algorithm 659: Implementing {S}obol's quasirandom
+	sequence generator.
+	\newblock ACM Trans.\ Math.\ Software \textbf{29}, 49--57 (2003)
+
+	\bibitem{KuoJoe08a}
+	Joe, S., Kuo, F.Y.: Constructing {S}obol sequences with better two-dimensional
+	projections.
+	\newblock SIAM J.\ Sci.\ Comput. \textbf{30}, 2635--2654 (2008)
+
+	\bibitem{Kmmerer2015}
+	K\"{a}mmerer, L., Potts, D., Volkmer, T.: Approximation of multivariate
+	periodic functions by trigonometric polynomials based on rank-1 lattice
+	sampling.
+	\newblock J.\ Complexity \textbf{31}(4), 543--576 (2015).
+	\newblock \doi{10.1016/j.jco.2015.02.004}.
+	\newblock \urlprefix\url{https://doi.org/10.1016/j.jco.2015.02.004}
+
+	\bibitem{SamplePack}
+	Kollig, T., Keller, A.: Samplepack (2002).
+	\newblock \urlprefix\url{https://www.uni-kl.de/AG-Heinrich/SamplePack.html}
+
+	\bibitem{BRODA20a}
+	Kucherenko, S.: {BRODA} (2022).
+	\newblock \urlprefix\url{https://www.broda.co.uk/index.html}
+
+	\bibitem{KuoNuy16a}
+	Kuo, F.Y., Nuyens, D.: Application of quasi-{M}onte {C}arlo methods to elliptic
+	{PDE}s with random diffusion coefficients -- a survey of analysis and
+	implementation.
+	\newblock Found.\ Comput.\ Math. \textbf{16}, 1631--1696 (2016).
+	\newblock \urlprefix\url{https://people.cs.kuleuven.be/~dirk.nuyens/qmc4pde/}
+
+	\bibitem{LEcEtal22a}
+	L'Ecuyer, P., Marion, P., Godin, M., Puchhammer, F.: A tool for custom
+	construction of {QMC} and {RQMC} point sets.
+	\newblock In: A.~Keller (ed.) {M}onte {C}arlo and Quasi-{M}onte {C}arlo
+	Methods: {MCQMC}, {O}xford, 2020, Springer Proceedings in Mathematics and
+	Statistics. Springer, Cham (2022)
+
+	\bibitem{l2002ssj}
+	L'Ecuyer, P., Meliani, L., Vaucher, J.: Ssj: A framework for stochastic
+	simulation in {Java}.
+	\newblock In: Proceedings of the Winter Simulation Conference, vol.~1, pp.
+	234--242. IEEE (2002)
+
+	\bibitem{l2018sorting}
+	L'Ecuyer, P., Munger, D., L{\'e}cot, C., Tuffin, B.: Sorting methods and
+	convergence rates for {Array-RQMC}: Some empirical comparisons.
+	\newblock Mathematics and Computers in Simulation \textbf{143}, 191--201 (2018)
+
+	\bibitem{LEcuyer2022b}
+	L'Ecuyer, P., Puchhammer, F., {Ben Abdellah}, A.: Monte carlo and
+	quasi{\textendash}monte carlo density estimation via conditioning.
+	\newblock INFORMS J.\ Comput. \textbf{34}(3), 1729--1748 (2022).
+	\newblock \doi{10.1287/ijoc.2021.1135}.
+	\newblock \urlprefix\url{https://doi.org/10.1287/ijoc.2021.1135}
+
+	\bibitem{Nie92}
+	Niederreiter, H.: Random Number Generation and Quasi-{M}onte {C}arlo Methods.
+	\newblock CBMS-NSF Regional Conference Series in Applied Mathematics. SIAM,
+	Philadelphia (1992)
+
+	\bibitem{NovWoz10a}
+	Novak, E., Wo{\'{z}}niakowski, H.: Tractability of Multivariate Problems
+	{V}olume {II}: {S}tandard Information for Functionals.
+	\newblock No.~12 in EMS Tracts in Mathematics. European Mathematical Society,
+	Z\"urich (2010)
+
+	\bibitem{NuyWeb}
+	Nuyens, D.: Dirk's homepage.
+	\newblock \urlprefix\url{https://people.cs.kuleuven.be/~dirk.nuyens/}
+
+	\bibitem{Owe22a}
+	Owen, A.B.: On dropping the first {Sobol'} point.
+	\newblock In: A.~Keller (ed.) {M}onte {C}arlo and Quasi-{M}onte {C}arlo
+	Methods: {MCQMC}, {O}xford, England, {A}ugust 2020, Springer Proceedings in
+	Mathematics and Statistics, pp. 71--86. Springer, Cham (2022).
+	\newblock \urlprefix\url{https://arxiv.org/pdf/2008.08051.pdf}
+
+	\bibitem{FinDer}
+	Papageorgiou, A.: Finder.
+	\newblock \urlprefix\url{http://www.cs.columbia.edu/~ap/html/finder.html}
+
+	\bibitem{MUQ2}
+	Parno, M., Davis, A., Seelinger, L.: {MUQ}: The {MIT} {U}ncertainty
+	{Q}uantification library.
+	\newblock Journal of Open Source Software \textbf{6}(68), 3076 (2021).
+	\newblock \doi{10.21105/joss.03076}.
+	\newblock \urlprefix\url{https://doi.org/10.21105/joss.03076}
+
+	\bibitem{ParSee22a}
+	Parno, M., Seelinger, L.: Uncertainty propagation of material properties of a
+	cantilevered beam (2022).
+	\newblock
+	\urlprefix\url{https://um-bridge-benchmarks.readthedocs.io/en/docs/forward-benchmarks/muq-beam-propagation.html}
+
+	\bibitem{PasTra95}
+	Paskov, S., Traub, J.: Faster valuation of financial derivatives.
+	\newblock J. Portfolio Management \textbf{22}, 113--120 (1995)
+
+	\bibitem{paszke2019pytorch}
+	Paszke, A., Gross, S., Massa, F., Lerer, A., Bradbury, J., Chanan, G., Killeen,
+	T., Lin, Z., Gimelshein, N., Antiga, L., Desmaison, A., K{\"o}pf, A., Yang,
+	E., DeVito, Z., Raison, M., Tejani, A., Chilamkurthy, S., Steiner, B., Fang,
+	L., Bai, J., Chintala, S.: Py{T}orch: An imperative style, high-performance
+	deep learning library.
+	\newblock Advances in neural information processing systems \textbf{32},
+	8026--8037 (2019)
+
+	\bibitem{paulin2022}
+	Paulin, L., Bonneel, N., Coeurjolly, D., Iehl, J.C., Keller, A., Ostromoukhov,
+	V.: Matbuilder: Mastering sampling uniformity over projections.
+	\newblock ACM Transactions on Graphics (Proceedings of SIGGRAPH) \textbf{41}(4)
+	(2022).
+	\newblock \doi{https://doi.org/10.1145/3528223.3530063}
+
+	\bibitem{PotSch21}
+	Potts, D., Schmischke, M.: Approximation of high-dimensional periodic functions
+	with fourier-based methods.
+	\newblock SIAM J.\ Numer.\ Anal. \textbf{59}(5), 2393--2429 (2021).
+	\newblock \doi{10.1137/20M1354921}.
+	\newblock \urlprefix\url{https://doi.org/10.1137/20M1354921}
+
+	\bibitem{puchhammer2021variance}
+	Puchhammer, F., Ben~Abdellah, A., L'Ecuyer, P.: Variance reduction with
+	{Array-RQMC} for tau-leaping simulation of stochastic biological and chemical
+	reaction networks.
+	\newblock Bulletin of Mathematical Biology \textbf{83}(8), 91 (2021)
+
+	\bibitem{pytorchSoboldiscussion2020a}
+	{PyTorch Developers}: {S}obol point implementation (2020).
+	\newblock \urlprefix\url{https://github.com/pytorch/pytorch/issues/32047}
+
+	\bibitem{SAS_LD}
+	{SAS Institute}: Sas documentation for the model procedure (2023).
+	\newblock
+	\urlprefix\url{https://documentation.sas.com/doc/en/pgmsascdc/9.4_3.4/etsug/etsug_model_sect197.htm}
+
+	\bibitem{scipySobol2020a}
+	{SciPy Developers}: {SciPy} discussion of {S}obol' sequence implementation
+	(2020).
+	\newblock \urlprefix\url{https://github.com/scipy/scipy/pull/10844}
+
+	\bibitem{UMBridge23}
+	Seelinger, L., Cheng-Seelinger, V., Davis, A., Parno, M., Reinarz, A.:
+	{UM-Bridge}: {U}ncertainty quantification and modeling bridge.
+	\newblock Journal of Open Source Software \textbf{8}(83), 4748 (2023).
+	\newblock \doi{10.21105/joss.04748}.
+	\newblock \urlprefix\url{https://doi.org/10.21105/joss.04748}
+
+	\bibitem{seelinger2023lowering}
+	Seelinger, L., Reinarz, A., Benezech, J., Lykkegaard, M.B., Tamellini, L.,
+	Scheichl, R.: Lowering the entry bar to {HPC}-scale uncertainty
+	quantification (2023)
+
+	\bibitem{SloJoe94}
+	Sloan, I.H., Joe, S.: Lattice Methods for Multiple Integration.
+	\newblock Oxford University Press, Oxford (1994)
+
+	\bibitem{smith2016software}
+	Smith, A.M., Katz, D.S., Niemeyer, K.E.: Software citation principles.
+	\newblock PeerJ Computer Science \textbf{2}, e86 (2016)
+
+	\bibitem{JagSor23a}
+	Sorokin, A.G., Jagadeeswaran, R.: Monte {C}arlo for vector functions of
+	integrals.
+	\newblock In: A.~Hinrichs, P.~Kritzer, F.~Pillichshammer (eds.) {M}onte {C}arlo
+	and Quasi-{M}onte {C}arlo Methods: {MCQMC}, {L}inz, Austria, {J}uly 2022,
+	Springer Proceedings in Mathematics and Statistics. Springer, Cham (2023+).
+	\newblock Submitted for publication
+
+	\bibitem{tfqfQMC2021a}
+	{TF Quant Finance Contributors}: Quasi {M}onte-{C}arlo methods (2021).
+	\newblock \urlprefix\url{https://github.com/google/tf-quant-finance/math/qmc}
+
+	\bibitem{MAT9.13}
+	{The MathWorks, Inc.}: {MATLAB} R2022b.
+	\newblock Natick, MA (2022)
+
+	\bibitem{NAG27}
+	{The Numerical Algorithms Group}: The {NAG} Library.
+	\newblock Oxford, {M}ark 27 edn. (2021)
+
+	\bibitem{TonEtAl22a}
+	Tong, X., Choi, S.C.T., Ding, Y., Hickernell, F.J., Jiang, L., {Jim\'enez
+		Rugama}, {\relax Ll}.A., Jagadeeswaran, R., Zhang, K., Zhang, Y., Zhou, X.:
+	Guaranteed automatic integration library {(GAIL)}: An open-source {MATLAB}
+	library for function approximation, minimization, and integration.
+	\newblock J. Open Res.\ Software \textbf{10}, 7 (2022).
+	\newblock \doi{10.5334/jors.381}
+
+	\bibitem{ASCR-SSSDU}
+	{U.S. Department of Energy, Office of Advanced Scientific Computing Research}:
+	Workshop on the science of scientific-software development and use (2021).
+	\newblock
+	\urlprefix\url{https://web.cvent.com/event/1b7d7c3a-e9b4-409d-ae2b-284779cfe72f/summary}
+
+	\bibitem{virtanen2020scipy}
+	Virtanen, P., Gommers, R., Oliphant, T.E., Haberland, M., Reddy, T.,
+	Cournapeau, D., Burovski, E., Peterson, P., Weckesser, W., Bright, J.,
+	van~der Walt, S.J., Brett, M., Wilson, J., Millman, K.J., Mayorov, N.,
+	Nelson, A.R.J., Jones, E., Kern, R., Larson, E., Carey, C.J., Polat, I.,
+	Feng, Y., Moore, E.W., VanderPlas, J., Laxalde, D., Perktold, J., Cimrman,
+	R., Henriksen, I., Quintero, E.A., Harris, C.R., Archibald, A.M., Ribeiro,
+	A.H., Pedregosa, F., van Mulbregt, P., Vijaykumar, A., Bardelli, A.P.,
+	Rothberg, A., Hilboll, A., Kloeckner, A., Scopatz, A., Lee, A., Rokem, A.,
+	Woods, C.N., Fulton, C., Masson, C., H{\"a}ggstr{\"o}m, C., Fitzgerald, C.,
+	Nicholson, D.A., Hagen, D.R., Pasechnik, D.V., Olivetti, E., Martin, E.,
+	Wieser, E., Silva, F., Lenders, F., Wilhelm, F., Young, G., Price, G.A.,
+	Ingold, G.L., Allen, G.E., Lee, G.R., Audren, H., Probst, I., Dietrich, J.P.,
+	Silterra, J., Webber, J.T., Slavi{\v c}, J., Nothman, J., Buchner, J.,
+	Kulick, J., Sch{\"o}nberger, J.L., de~Miranda~Cardoso, J., Reimer, J.,
+	Harrington, J., Rodr{\'\i}guez, J.L.C., Nunez-Iglesias, J., Kuczynski, J.,
+	Tritz, K., Thoma, M., Newville, M., K{\"u}mmerer, M., Bolingbroke, M.,
+	Tartre, M., Pak, M., Smith, N.J., Nowaczyk, N., Shebanov, N., Pavlyk, O.,
+	Brodtkorb, P.A., Lee, P., McGibbon, R.T., Feldbauer, R., Lewis, S., Tygier,
+	S., Sievert, S., Vigna, S., Peterson, S., More, S., Pudlik, T., Oshima, T.,
+	Pingel, T.J., Robitaille, T.P., Spura, T., Jones, T.R., Cera, T., Leslie, T.,
+	Zito, T., Krauss, T., Upadhyay, U., Halchenko, Y.O., V{\'a}zquez-Baeza, Y.,
+	Contributors: Scipy 1.0: fundamental algorithms for scientific computing in
+	python.
+	\newblock Nature Methods \textbf{17}(3), 261--272 (2020)
+
+	\bibitem{Woz99a}
+	Wo{\'z}niakowski, H.: Efficiency of quasi-{M}onte {C}arlo algorithms for high
+	dimensions.
+	\newblock In: H.~Niederreiter, J.~Spanier (eds.) {M}onte {C}arlo and
+	Quasi-{M}onte {C}arlo Methods 1998, pp. 114--136. Springer-Verlag, Berlin
+	(2000)
+
+\end{thebibliography}
 
